### 2020-04-19 doc/preliminary report/Alpine.Rmd
preliminary report rmarkdown

---
title: "Alpine seed germination: global synthesis, preliminary report"
author: "Eduardo Fernández Pascual, University of Oviedo (fernandezpeduardo@uniovi.es)"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output: 
  pdf_document:
      df_print: kable
documentclass: article
geometry: margin=1in
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
load("AlpineOutput.RData")
library(tidyverse); library(kableExtra)
```

```{r, echo = FALSE, fig.align = "center"}
knitr::include_graphics("2018062333 H.ou de Los Boches reducida.jpg")
```

# Dataset

````{r echo = FALSE, fig.pos = "H", fig.cap = "Map of the alpine regions with records in the dataset. The size of the circles is proportional to the number of species with records.", fig.align = "center", fig.width = 5, fig.height = 3}
Mapa
```

The current dataset includes seed germination data (germination proportions at given combinations of temperature, temperature alternation, light, stratification, GA3 and scarification) for `r length(unique(TableSpecies$Species))` species from `r length(unique(TableSpecies$Region))` mountain regions and it is based on a selection of the data originally sent by contributors. We standarized the species names provided by the contributors using [The Plant List](http://www.theplantlist.org/). For each species, we recorded the elevation range and the regional treeline elevation from local sources. Based on this information, we removed the species that were not recorded to reach at least the treeline elevation. We also removed the records that had been provided by [ENSCOBASE](http://enscobase.maich.gr/) and [Kew's Millennium Seed Bank database](https://www.kew.org/wakehurst/whats-at-wakehurst/millennium-seed-bank/) because of the impossibility to confirm that they had been collected in an alpine region, and also records from China that came from field instead of laboratory experiments. For the remaining species, we classified them as generalist alpine species (when their minimum elevation reaches below 70% of the regional treeline) or strictly alpine species (when it does not). The actual list of species, with the classification, is provided in a separate Excel file (**please check the classification and recorded treeline elevation for your own region**).

# Sources

Below is a list of sources with the number of species contributed by each. Tudela and Liu are datasets extracted from articles, the rest are original data provided by the contributors.

```{r warning = FALSE, echo = FALSE}
TableSources %>% kable(caption = "List of sources and number of species per source.", booktabs = TRUE) %>% kable_styling(latex_options = c("striped", "hold_position"))
```

# Meta-analysis of the dataset

To meta-analyse the data we followed the recommendations of [Koricheva et al.](https://press.princeton.edu/books/hardcover/9780691137285/handbook-of-meta-analysis-in-ecology-and-evolution) and the [metafor package website](http://www.metafor-project.org/doku.php/todo#extending_rmaglmm_to_multilevel_multivariate_models) for proportion data. To be able to handle random-effect terms (in our case, species and mountain region) we used GLMM. At the moment, the analysis does not take into account species phylogeny because of the complexity of including that into a GLM framework.

Because of the sparse and unbalanced nature of the dataset, we subsetted the dataset and fitted several relatively simple models (with one fixed effect of interest) to analyse different aspects. For example, to study the effect of temperature we selected a subset of the records that had the same level for the rest of the experimental variables (the most common combination of unstratified seeds, alternating temperatures, light, no GA3 and no scarification).

We present results for germination temperature, temperature alternation and stratification. Ligth (germination in darkness) was only available for 2 species. We repeated the analysis for the full dataset, for the generalists and for the stricly alpine species.

# Germination temperature

```{r warning = FALSE, echo = FALSE}
tableTemperatures %>% kable(caption = "Number of species with germination temperature data included in the GLMM.", booktabs = TRUE) %>% kable_styling(latex_options = c("striped", "hold_position"))
```

````{r echo = FALSE, fig.pos = "H", fig.cap = "Germination proportions at different temperatures. The data are for unstratified seeds with alternating temperatures (the germination temperature is the weighted average of the two phases, rounded up to 5ºC intervals), in light, without GA3 or scarification.", fig.align = "center"}
plotTemperatures
```

# Alternating temperatures

```{r warning = FALSE, echo = FALSE}
tableAlternating %>% kable(caption = "Number of species with alternating temperature data included in the GLMM.", booktabs = TRUE) %>% kable_styling(latex_options = c("striped", "hold_position"))
```

````{r echo = FALSE, fig.pos = "H", fig.cap = "Germination proportions at constant and alternating temperatures. The data are for unstratified seeds with wighted average germination temperatures of 20ºC, in light, without GA3 or scarification.", fig.align = "center"}
plotAlternating
```

# Cold stratification

```{r warning = FALSE, echo = FALSE}
tableStratification %>% kable(caption = "Number of species with cold stratification data included in the GLMM.", booktabs = TRUE) %>% kable_styling(latex_options = c("striped", "hold_position"))
```

````{r echo = FALSE, fig.pos = "H", fig.cap = "Germination proportions in unstratified and cold-stratified seeds. The data are for experiments with alternating temperatures and weighted average germination temperatures of 20ºC, in light, without GA3 or scarification.", fig.align = "center"}
plotStratification
```

### 2020-04-19 src/preliminary report analysis.R

@@ -1,0 +1,458 @@
+library(tidyverse); library(lme4); library(cowplot)
+
+respuesta <- function(x) 1/(1+1/(exp(x)))
+
+# ALL SPECIES
+
+load("C:/EFP/Trabayu/Activos/2018 FICYT3/GT1 Traits/Joined DB/AlpineDB.RData")
+
+# TEMPERATURES
+
+d <- subset(AlpineDB, Scarification == "N" & GA3 == "N" & Light == "Y" & Stratification == "N" & Alternating == "Y")
+
+m <- glmer(cbind(Germinated, Germinable - Germinated) ~ factor(Temperature) + (1|Species) + (1|Region), family = "binomial", data = d)
+
+summary(m)
+
+ci <- confint(m, method = "Wald")
+
+p <- predict(m, expand.grid(Temperature = c(5, 10, 15, 20, 25)), re.form = NA)
+
+mt <- cbind(p, 
+            c(ci[3, 1], ci[4, 1] + ci[3, 1], ci[5, 1] + ci[3, 1], ci[6, 1] + ci[3, 1], ci[7, 1] + ci[3, 1]), 
+            c(ci[3, 2], ci[4, 2] + ci[3, 2], ci[5, 2] + ci[3, 2], ci[6, 2] + ci[3, 2], ci[7, 2] + ci[3, 2]))
+
+r <- respuesta(mt)
+
+e <- data.frame(Stratification = "Unstratified", Alternating = "Alternating",
+                 Temperature = c(5, 10, 15, 20, 25),
+                 Mean = r[, 1],
+                 Lower = r[, 2],
+                 Upper = r[, 3])
+
+tableTemperatures<- d %>% select(Region, Alpine, Species) %>% unique %>% group_by(Region, Alpine) %>% tally() %>% spread(Alpine, n)  
+
+table1b <- summary(m)
+
+plot1 <- ggplot(e, aes(x = Temperature, y = Mean, ymin = Lower, ymax = Upper, 
+                               fill = Temperature)) +
+  geom_bar(stat = "identity", position = "dodge", alpha = 0.7, show.legend = FALSE) + 
+  geom_errorbar(aes(color = Temperature), width = 1, size = 1, 
+                position = position_dodge(.9), show.legend = FALSE) +
+  coord_cartesian(ylim = c(0, 1)) +
+  labs(title = "All species", subtitle = "Germination peaks at 20ºC") +
[…]

### 2020-04-22 src/germination.R
corrected Chinese from original data

@@ -1,150 +1,10 @@
-library(tidyverse); library(here)
+library(tidyverse); library(here); library(readxl)
 
-# Join germination data
+# Format Russian data (Sergey Rosbakh)
+# Read from original files
 
-## Format Australian data (Annisa Satyanti, Susanna Venn, Karen Sommerville)
-
-read.csv(here("data", "germination", "australian germination.csv")) %>%
-  mutate(Region = "Australian Alps",
-         Tmin = ifelse(is.na(Tmin), Tmax, Tmin),
-         Tmean = (Tmax * Photoperiod + Tmin * (24 - Photoperiod)) / 24, 
-         Tdif = Tmax - Tmin, 
-         Alternating = as.factor(ifelse(Tdif != 0, "Y", "N"))) %>%
-  select(Taxon,
-         Source, 
-         Country, 
-         Region, 
-         Accession, 
-         Latitude, 
-         Longitude, 
-         Elevation, 
-         Dish, 
-         Scarification,
-         GA3,
-         Stratification_temperature, 
-         Stratification_days, 
-         Stratification, 
-         Light,
-         Photoperiod, 
-         Alternating,
-         Tmean,
-         Tmax, 
-         Tmin,
-         Tdif,
-         Sown, 
-         Germinated, 
-         Normal) ->
-  Australia
-
-## Format Spanish data (Eduardo Fernández-Pascual)
-
-read.csv(here("data", "germination", "spanish germination.csv")) %>%
-  mutate(Region = "Cantabrian Mountains",
-         Tmin = ifelse(is.na(Tmin), Tmax, Tmin),
-         Tmean = (Tmax * Photoperiod + Tmin * (24 - Photoperiod)) / 24, 
-         Tdif = Tmax - Tmin, 
-         Alternating = as.factor(ifelse(Tdif != 0, "Y", "N"))) %>%
-  select(Taxon,
-         Source, 
-         Country, 
-         Region, 
-         Accession, 
-         Latitude, 
-         Longitude, 
-         Elevation, 
-         Dish, 
-         Scarification,
-         GA3,
-         Stratification_temperature, 
-         Stratification_days, 
-         Stratification, 
-         Light,
-         Photoperiod, 
-         Alternating,
-         Tmean,
-         Tmax, 
-         Tmin,
-         Tdif,
-         Sown, 
-         Germinated, 
-         Normal) ->
-  Spain
-
-## Format Italian data (Andrea Mondoni)
-
[…]

### 2020-04-22 src/traits.R

@@ -1,12 +1,22 @@
-library(tidyverse); library(here)
+library(tidyverse); library(here); library(openxlsx)
 
-# Read and prepare seed mass data
+# Read and prepare seed mass data from A Carta
 
-read.csv(here("data", "seed mass", "mass.alpine.csv")) %>%
+read.csv(here("data", "seeds", "original", "A Carta", "mass.alpine.csv")) %>%
   mutate(TPLName = paste(New.Genus, New.Species)) %>%
   select(TPLName, mass) %>%
   group_by(TPLName) %>%
   summarise(Seed.mass = mean(mass)) -> smass
+
+# Read and prepare seed mass data from Bu
+
+read.xlsx(here("data", "seed mass", "species information and seed mass.xlsx")) %>%
+  select(species, `seed.mass.(g,100seeds)-sample1`, 
+         `seed.mass.(g,100seeds)-sample2`, `seed.mass.(g,100seeds)-sample3`) %>%
+  gather(Trait, Value, -species) %>%
+  group_by(species) %>%  summarise(Seed.mass = mean(Value))
+
+# Read and prepare seed mass data from Liu
 
 # Prepare traits object
 

### 2020-05-16 data/elevations/elevations.R
added germination indices

@@ -1,0 +1,84 @@
+library(tidyverse); library(here); library(openxlsx)
+
+# This is an atempt to clean the mess of the many elevation files..
+
+# Create a merged file with the spp elevations
+
+read.csv(here("data", "elevations", "elevations_original.csv")) %>% # Original file, with more info for lowland species
+  merge(read.xlsx(here("data", "elevations", "Species Cavieres.xlsx")), # Merge with Cavieres, who altered some elevations
+        by = c("Species", "Region"), all = TRUE) %>%
+  mutate(Min.elevation = ifelse(is.na(Min.elevation.y), Min.elevation.x, Min.elevation.y),
+         Max.elevation = ifelse(is.na(Max.elevation.y), Max.elevation.x, Max.elevation.y)) %>%
+  select(Species, Region, Alpine.x, Min.elevation, Max.elevation) -> sppelevations
+
+# Read Poschlod file with updated Alpine trait
+
+read.xlsx(here("data", "elevations", "Species_rev_Peter_2020_04_15.xlsx")) %>% 
+  select(Species, Alpine) %>%
+  unique %>%
+  group_by(Species) %>%
+  filter(length(Species) >1) -> spp # Species with different classification in different regions
+
+read.xlsx(here("data", "elevations", "Species_rev_Peter_2020_04_15.xlsx")) %>% 
+  select(Species, Alpine) %>%
+  filter(! (Species %in% spp$Species & 
+              Alpine == "Generalist")) %>%
+  unique -> # I should do the opposite but I trust more the European classification by Peter...
+  alpine
+
+# List lowland spp
+
+sppelevations %>%
+  filter(Alpine.x == "Lowland") %>%
+  select(Species, Alpine.x) %>%
+  rename(Alpine = Alpine.x) %>%
+  unique -> lowland1 # Manually classified as lowland since beginning
+  
+sppelevations %>%
+  filter(Alpine.x != "Lowland") %>%
+  merge(read.csv(here("data", "elevations", "regions.csv"))) %>%
+  filter(Max.elevation <= Treeline) %>%
+  mutate(Alpine = "Lowland") %>%
+  select(Species, Alpine) %>%
+  unique -> lowland2 # Classified as lowland because max elevation in the treeline or below
+
+rbind(lowland1, lowland2) -> lowland # Merged list of lowland spp
+  
+# Merge lowland and alpine, with preference to alpine classification
+
+lowland %>%
[…]

### 2020-05-16 data/elevations/original1/elevations.R

@@ -1,0 +1,107 @@
+library(tidyverse); library(Taxonstand); library(here)
+
+## Australia
+
+read.csv(here("data", "elevations", "australian spp.csv")) %>%
+  rename(Min.elevation = Min.Elevation,
+         Max.elevation = Max.Elevation) %>%
+  rename(Taxon = NameDB) %>%
+  mutate(Alpine = "Alpine",
+         Region = "Australian Alps") %>%
+  select(Taxon, Region, Alpine, Min.elevation, Max.elevation) ->
+  AustralianSpp
+
+## Spain
+
+read.csv(here("data", "elevations", "spanish spp.csv")) %>%
+  separate(Habitat, c("Alpine", "Soil", "Frost"), sep = ",") %>%
+  mutate(Region = "Cantabrian Mountains") %>%
+  select(Taxon, Region, Alpine, Min.elevation, Max.elevation) ->
+  SpanishSpp
+
+## Flora alpina
+
+read.csv(here("data", "elevations", "flora alpina.csv")) %>%
+  rename(Taxon = Species) %>%
+  mutate(Region = "Southern Alps") %>%
+  select(Taxon, Region, Alpine, Min.elevation, Max.elevation) ->
+  SAlpsSpp
+
+SAlpsSpp %>%
+  mutate(Region = "Northern Alps") ->
+  NAlpsSpp
+
+## Russia
+
+read.csv(here("data", "elevations", "russian spp.csv")) %>%
+  filter(! Taxon %in% "") %>%
+  mutate(Region = "Caucasus",
+         Max.elevation = as.character(Max.elevation), Min.elevation = as.character(Min.elevation),
+         Max.elevation = ifelse(Max.elevation == "alpine", 2900, Max.elevation),
+         Max.elevation = ifelse(Max.elevation == "subalpine", 2400, Max.elevation),
+         Min.elevation = ifelse(Min.elevation == "alpine", 2400, Min.elevation),
+         Min.elevation = ifelse(Min.elevation == "subalpine", 1700, Min.elevation),
+         Min.elevation = ifelse(Min.elevation == "montane", 1000, Min.elevation),
+         Min.elevation = ifelse(Min.elevation == "lowlands", 0, Min.elevation),
+         Max.elevation = as.numeric(Max.elevation), Min.elevation = as.numeric(Min.elevation)) %>%
+  select(Taxon, Region, Alpine, Min.elevation, Max.elevation) ->
+  RussianSpp
[…]

### 2020-05-16 data/taxonomy/original/taxonomy.R

@@ -1,47 +1,0 @@
-library(tidyverse); library(Taxonstand); library(here); library(openxlsx)
-
-# Read names, TPL them
-
-read.csv(here("results", "database", "germination.raw.csv")) %>%
-  pull(Taxon) %>%
-#  merge(read.csv(here("data", "taxonomy", "tpl.csv"))) %>%
-  #  pull(Species) %>%
-  
-  unique %>%
-  TPL() %>%
-  write.csv(here("data", "taxonomy", "TPLNames.csv"))
-
-read.csv(here("data", "taxonomy", "TPLNames.csv")) %>%
-  filter(! New.Taxonomic.status %in% c("Accepted", "Unresolved")) %>%
-  select(Taxon) %>%
-  merge(read.csv(here("data", "taxonomy", "tpl.csv"))) %>%
-  pull(Species) %>%
-  unique %>%
-  TPL() %>%
-  write.csv(here("data", "taxonomy", "TPLNames_spelling.csv"))
-
-
-read.csv(here("data", "taxonomy", "TPLNames_spelling.csv")) %>%
-  filter(! New.Taxonomic.status %in% c("Accepted", "Unresolved"))
-
-read.csv(here("data", "taxonomy", "TPLNames.csv")) %>%
-  filter(New.Taxonomic.status %in% c("Accepted", "Unresolved")) -> TPLfound
-read.csv(here("data", "taxonomy", "TPLNames_spelling.csv")) %>%
-  filter(New.Taxonomic.status %in% c("Accepted", "Unresolved")) -> TPLspelling
-read.csv(here("data", "taxonomy", "TPLNames_spelling.csv")) %>%
-  filter(! New.Taxonomic.status %in% c("Accepted", "Unresolved")) -> TPLmissing
-
-rbind(TPLfound, TPLspelling, TPLmissing) -> tpl
-
-read.csv(here("data", "taxonomy", "tpl.csv")) %>%
-  mutate(Original = as.character(Taxon),
-         Spelling = ifelse(Taxon %in% TPLfound$Taxon, Original, as.character(Species)),
-         Familia = Family) %>%
-  select(Original, Spelling, Familia) %>%
-  merge(tpl, by.x = "Spelling", by.y = "Taxon") %>%
-  select(-X) %>%
-  mutate(Original.clean = paste(Genus, Species),
-         TPLName = ifelse(New.Taxonomic.status %in% c("Accepted", "Unresolved"),
[…]

### 2020-05-16 src/cleaning.R

@@ -1,0 +1,1165 @@
+library(tidyverse); library(here); library(readxl)
+
+# WITH GERMINATION INDICES
+
+# Format Australian data (susanna Venn)
+# Read from original files with germination indices (S Rosbakh)
+
+read_excel(here("data",
+                "Sergey Rosbakh (Indices)", "VENN hand joined.xlsx"),
+           sheet = 1, skip = 0) %>%
+  mutate(Region = "Australian Alps",
+         Tmin = ifelse(is.na(Tmin), Tmax, Tmin),
+         Tmean = (Tmax * Photoperiod + Tmin * (24 - Photoperiod)) / 24, 
+         Tdif = Tmax - Tmin, 
+         Alternating = as.factor(ifelse(Tdif != 0, "Y", "N"))) %>%
+  select(Taxon,
+         Source, 
+         Country, 
+         Region, 
+         Accession, 
+         Latitude, 
+         Longitude, 
+         Elevation, 
+         Dish, 
+         Scarification,
+         GA3,
+         Stratification_temperature, 
+         Stratification_days, 
+         Stratification, 
+         Light,
+         Photoperiod, 
+         Alternating,
+         Tmean,
+         Tmax, 
+         Tmin,
+         Tdif,
+         Sown, 
+         Germinated, 
+         Normal,
+         GRS:CVG) ->
+  AustraliaVenn
+
+# Format Australian data (Annisa Satyanti)
+# Read from original files with germination indices (S Rosbakh)
+
+read_excel(here("data",
+                "Sergey Rosbakh (Indices)", "Aust Alpine Germination timing 11MAY2020.xlsx"),
+           sheet = 4, skip = 0) %>%
+  select(Species, Accession, Elevation, Treatment, Rep, seeds, GRS:CVG) -> satyanti1
+read_excel(here("data",
+                "Sergey Rosbakh (Indices)", "Aust Alpine Germination timing 11MAY2020.xlsx"),
+           sheet = 5, skip = 0)  %>%
+  select(Species, Accession, Elevation, Treatment, Rep, seeds, GRS:CVG) -> satyanti2
+read_excel(here("data",
+                "Sergey Rosbakh (Indices)", "Aust Alpine Germination timing 11MAY2020.xlsx"),
+           sheet = 6, skip = 0)  %>%
+  select(Species, Accession, Elevation, Treatment, Rep, seeds, GRS:CVG) -> satyanti3
+read_excel(here("data",
[…]

### 2020-05-16 src/germination.R

@@ -1,706 +1,0 @@
-library(tidyverse); library(here); library(readxl)
-
-# Format Russian data (Sergey Rosbakh)
-# Read from original files
-
-read.csv(here("data", "seeds", "original",
-                "Sergey Rosbakh (Caucasus)", "NCaucasus_Rosbakh.csv")) %>%
-  mutate(Region = "Caucasus", 
-         GA3 = ifelse(GA3 == 0, "N", "Y"),
-         Scarification = ifelse(Scarification == 0, "N", "Y"),
-         Stratification = ifelse(Stratification == 0, "N", "Y"),
-         Stratification_temperature = ifelse(Stratification == "N", NA, 0),
-         Tmin = ifelse(is.na(Tmin), Tmax, Tmin),
-         Tmean = (Tmax * Photoperiod + Tmin * (24 - Photoperiod)) / 24, 
-         Tdif = Tmax - Tmin, 
-         Alternating = as.factor(ifelse(Tdif != 0, "Y", "N"))) %>%
-  select(Taxon,
-         Source, 
-         Country, 
-         Region, 
-         Accession, 
-         Latitude, 
-         Longitude, 
-         Elevation, 
-         Dish, 
-         Scarification,
-         GA3,
-         Stratification_temperature, 
-         Stratification_days, 
-         Stratification, 
-         Light,
-         Photoperiod, 
-         Alternating,
-         Tmean,
-         Tmax, 
-         Tmin,
-         Tdif,
-         Sown, 
-         Germinated, 
-         Normal) ->
-  Russia
-
-# Format German data (Sergey Rosbakh)
-# Read from original files
-
-read_excel(here("data", "seeds", "original",
-                "Sergey Rosbakh (Alps)", "Alps_Rosbakh.xlsx")) %>%
-  mutate(Region = "Northern Alps",
-         Stratification_temperature = ifelse(Stratification_days == 0, NA, Stratification_temperature),
-         Stratification = ifelse(Stratification == "0", "N", Stratification),
-         Scarification = ifelse(Scarification == "0", "N", Scarification),
-         GA3 = "N",
-         Tmin = ifelse(is.na(Tmin), Tmax, Tmin),
-         Tmean = (Tmax * Photoperiod + Tmin * (24 - Photoperiod)) / 24, 
-         Tdif = Tmax - Tmin, 
-         Alternating = as.factor(ifelse(Tdif != 0, "Y", "N")),
[…]

### 2020-05-16 src/traits.R

@@ -1,112 +1,0 @@
-library(tidyverse); library(here); library(openxlsx); library(readxl)
-
-
-# Read and prepare seed mass data from A Carta
-
-read.csv(here("data", "seeds", "original", "A Carta", "mass.alpine.csv")) %>%
-  mutate(TPLName = paste(New.Genus, New.Species)) %>%
-  mutate(mass = mass/10) %>% # Kew if for 1000 seeds
-  select(TPLName, mass) %>%
-  group_by(TPLName) %>%
-  summarise(Seed.mass = mean(mass)) -> smassCarta
-
-# Read and prepare seed mass data from Bu
-
-read.xlsx(here("data", "seeds", "original", "Haiyan Bu", 
-               "germination in light -2010", "species information and seed mass.xlsx")) %>%
-  select(species, `seed.mass.(g,100seeds)-sample1`, 
-         `seed.mass.(g,100seeds)-sample2`, `seed.mass.(g,100seeds)-sample3`) %>%
-  gather(Trait, Value, -species) %>%
-  group_by(species) %>%
-  rename(Taxon = species) %>%
-  summarise(Seed.mass = mean(Value, na.rm = TRUE)) -> SMBu2010
-
-read_excel(path = here("data", "seeds", "original", "Haiyan Bu", 
-                       "germination in darkness-2005", 
-                       "species information-2005.xlsx"), 
-           sheet = 1, skip = 0, col_types = "text") %>%
-  select(`serial number`, Species) -> Bu2005spp
-
-read_excel(path = here("data", "seeds", "original", "Haiyan Bu", 
-                       "germination in darkness-2005", 
-                       "seed mass-2005.xls"), 
-           sheet = 1, skip = 0, col_types = "text") %>%
-  mutate(`serial number` = round(as.numeric(`serial number`), 2)) %>%
-  merge(Bu2005spp) %>%
-  mutate(`seed weight (g,100 seeds)` = as.numeric(`seed weight (g,100 seeds)`)) %>%
-  rename(Taxon = Species) %>%
-  group_by(Taxon) %>%
-  summarise(Seed.mass = mean(`seed weight (g,100 seeds)`, na.rm = TRUE)) -> SMBu2005
-
-# Liu paper (Lydia)
-
-read_excel(path = here("data", "seeds", "original", "Lydia Guja (Liu et al.)", 
-                       "Liu et al 2013 Table S2.xlsx"), 
[…]

### 2020-05-18 doc/indices report/indices report.Rmd
removed too long scoring

---
title: "Alpine indices"
author: "Eduardo Fernández Pascual"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(tidyverse); library(here); library(FactoMineR); library(cowplot)
library(ggthemes); library(ggrepel); library(flextable); library(ggpubr)

pcaEFP <- function(df, id = 1, header = NULL)
{
  z <- list()
  z$pca <- PCA(df[, -id], scale.unit = TRUE, graph = FALSE)
  if(is.null(header)) z$inds <- data.frame(Plot = df[, id], z$pca$ind$coord[, 1:2])
  if(! is.null(header)) z$inds <- merge(header, data.frame(Plot = df[, id], z$pca$ind$coord[, 1:2]))
  z$vars <- rownames_to_column(data.frame(z$pca$var$coord[, 1:2]), var = "Variable")
  z$vars$Variable <- gsub(".", " ", z$vars$Variable, fixed = TRUE)
  z$contrib <- rownames_to_column(data.frame(z$pca$var$contrib[, 1:2]), var = "Variable")
  z$contrib$Variable <- gsub(".", " ", z$contrib$Variable, fixed = TRUE)
  class(z) <- "mvEFP"
  z
}

summary.mvEFP <- function(pca)
{
  z <- data.frame(pca$pca$var$contrib[, 1:2], pca$pca$var$cor[, 1:2])
  is.num <- sapply(z, is.numeric)
  z[is.num] <- lapply(z[is.num], round, 2)
  colnames(z) <- c("cont.Dim.1", "cont.Dim.2", "corr.Dim.1", "corr.Dim.2")
  z
}

read.csv(here("results", "database", "germination.csv")) %>%
  filter(! is.na(GRS)) %>% # Remove cases without indices
  filter(GRP >= 90) %>% # At least 50% germination
  filter_all(all_vars(!is.infinite(.))) %>% # Remove infinites
  merge(read.csv(here("results", "database", "traits.csv")), by = "TPLName") %>% # Merge with spp traits
  select(TPLName, Region,
         Alpine, Life.form, Lifespan, Seed.mass, Dormancy, Embryo,
         Scarification, GA3, Stratification, Light, Alternating, Temperature,
         GRP:VGT) ->
  indi

indi %>%
  pcaEFP(id = c(1:14)) -> pca

pca$inds %>% 
  filter(Dim.2 < 10) %>% # Outlier
  ggplot(aes(x = Dim.1, y = Dim.2)) +
  labs(title = "Region") +
  theme_wsj() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_point(aes(color = Plot.Region), size = 3, show.legend = T, alpha = 0.6, shape = 15) +
  theme(legend.position = "top", legend.title = element_blank(),
        legend.text = element_text(size = 7.5),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm"),
        plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"),
        text = element_text(size = 8)) +
  scale_x_continuous(name = paste("Axis 1 (", round(pca$pca$eig[1, 2], 0),
                                  "% variance explained)", sep = "")) + 
  scale_y_continuous(name = paste("Axis 2 (", round(pca$pca$eig[2, 2], 0), 
                                  "% variance explained)", sep = "")) +
  geom_label_repel(data = pca$vars, aes(x = 4*Dim.1, y = 4*Dim.2, label = Variable), 
                  alpha = 0.8, size = 2.5) +
  geom_segment(data = pca$vars, aes(x = 0, y = 0, xend = 4*Dim.1, yend = 4*Dim.2)) -> p1

pca$inds %>% 
  filter(Dim.2 < 10) %>% # Outlier
  ggplot(aes(x = Dim.1, y = Dim.2)) +
  labs(title = "Alpine") +
  theme_wsj() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_point(aes(color = Plot.Alpine), size = 3, show.legend = T, alpha = 0.6, shape = 15) +
  theme(legend.position = "top", legend.title = element_blank(),
        legend.text = element_text(size = 7.5),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm"),
        plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"),
        text = element_text(size = 8)) +
  scale_x_continuous(name = paste("Axis 1 (", round(pca$pca$eig[1, 2], 0),
                                  "% variance explained)", sep = "")) + 
  scale_y_continuous(name = paste("Axis 2 (", round(pca$pca$eig[2, 2], 0), 
                                  "% variance explained)", sep = "")) +
  geom_label_repel(data = pca$vars, aes(x = 4*Dim.1, y = 4*Dim.2, label = Variable), 
                   alpha = 0.8, size = 2.5) +
  geom_segment(data = pca$vars, aes(x = 0, y = 0, xend = 4*Dim.1, yend = 4*Dim.2)) -> p2

pca$inds %>% 
  filter(Dim.2 < 10) %>% # Outlier
  ggplot(aes(x = Dim.1, y = Dim.2)) +
  labs(title = "Germ. temperature") +
  theme_wsj() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_point(aes(color = Plot.Temperature), size = 3, show.legend = T, alpha = 0.6, shape = 15) +
  theme(legend.position = "top", legend.title = element_blank(),
        legend.text = element_text(size = 7.5),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm"),
        plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"),
        text = element_text(size = 8)) +
  scale_x_continuous(name = paste("Axis 1 (", round(pca$pca$eig[1, 2], 0),
                                  "% variance explained)", sep = "")) + 
  scale_y_continuous(name = paste("Axis 2 (", round(pca$pca$eig[2, 2], 0), 
                                  "% variance explained)", sep = "")) +
  geom_label_repel(data = pca$vars, aes(x = 4*Dim.1, y = 4*Dim.2, label = Variable), 
                   alpha = 0.8, size = 2.5) +
  geom_segment(data = pca$vars, aes(x = 0, y = 0, xend = 4*Dim.1, yend = 4*Dim.2)) -> p3

pca$inds %>% 
  filter(Dim.2 < 10) %>% # Outlier
  ggplot(aes(x = Dim.1, y = Dim.2)) +
  labs(title = "Life form") +
  theme_wsj() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_point(aes(color = Plot.Life.form), size = 3, show.legend = T, alpha = 0.6, shape = 15) +
  theme(legend.position = "top", legend.title = element_blank(),
        legend.text = element_text(size = 7.5),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm"),
        plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"),
        text = element_text(size = 8)) +
  scale_x_continuous(name = paste("Axis 1 (", round(pca$pca$eig[1, 2], 0),
                                  "% variance explained)", sep = "")) + 
  scale_y_continuous(name = paste("Axis 2 (", round(pca$pca$eig[2, 2], 0), 
                                  "% variance explained)", sep = "")) +
  geom_label_repel(data = pca$vars, aes(x = 4*Dim.1, y = 4*Dim.2, label = Variable), 
                   alpha = 0.8, size = 2.5) +
  geom_segment(data = pca$vars, aes(x = 0, y = 0, xend = 4*Dim.1, yend = 4*Dim.2)) -> p4
```

# Explanation of the indices by S Rosbakh

Here come some explanations about the indices.

The abbreviations are read as follows:

* GRS - number of germination seeds
* GRP - germination percentage (from 0 to 100%)
* MGT - mean germination time (time units)
* MGR - mean germination rate (time units)
* GSP - germination speed (%)
* UNC - uncertanty index (bits)
* SYN - syncronization index (from 0 to 1)
* VGT - germination variance
* SDG - germination standard deviation
* CVG - coefficient of variation

GRS and GRP are easy and you have already these data. Basically, MGT, MGR and GSP describe seed germination speed; these three indices are tightly positively (r>0.8) correlated with each other. I would do descriptive statistics with all of them first to see how they differ in your data set. These three indices are strongly temperature-dependent - in my data the relationship is linear and almost perfect (r2=0.8). Thus, you might consider analysing germination speed for one most optimal temperature of germination (around 20 C).

UNC and SYN characterise germination synchrony and tightly positively correlated with each other. I personally prefer the SYN, because it is more intuitive, but I would try both of them.

I have not had time to explore VGT, SDG and CVG, but I think they are not very informative (though I might be wrong). You might consider exploring them as well.

IMPORTANT: I have noticed that calculation of all indices for the cases with a very low number of germinated seeds (GRS) result in spurious results. For example, if only one single seed germinated, its germination timing (e.g. day 7 or day 14) considerably influenced seed germination time and synchrony. May be it makes sense from statistical point of you, but as an ecologist I would not trust them. Thus, I recommend to use only the cases with a comparatively high proportion of seeds germinated (e.g. GRP >30%). Furthermore, in some cases calculations resulted in zero or infinite values and NAs, because of no germination at all or very low number of germinated seeds.

I used the GerminaR package for calculations; more details on the indices are in the paper (https://www.researchgate.net/publication/331073880_GerminaR_An_R_package_for_germination_analysis_with_the_interactive_web_application_GerminaQuant_for_R).

# A note by Sergey on scoring length and intensity

I took a closer look at the experiment duration and scoring intensity of the data I processed and, came to the conclusion that the majority of the germination experiments were conducted in a more or less the same way. The only problematic data set are experiments from Anissa, in which seeds were incubated for several months.

Source	Duration	Scoring intensity
Andrea Mondoni	42 days	6 times
Bu	40-60 days	40-60 times
Rosbakh_Alps	42 days	8 times
Rosbakh_Caucaus	42 days	8 times
Anissa	49-140	12-22 times
Eduardo	28 days	4 times
Susanne Venn	40-60 days	11-21 times
Veronica	24-48 days	9-17 times

I re-calculated the indices for shorter timer series (ca. 60 days; orange tabs; attached) and hope that the PCA will look better now. BTW, Eduardo, did you scale the data before the analysis? The indices scales have very different scale lengths that should be accounted for.

One further improvement would be to reduce the time series in Anissa's, Venn's and Bu's data to 40 days, to make the differences even smaller, but it would require additional data re-formatting.

# Number of records with >80% germination and germination indices, by region.

```{r table1, echo = FALSE}
indi %>% group_by(Region) %>% tally
```

# Number of records with >80% germination and germination indices, by alpine strict vs. generalist.

```{r table2, echo = FALSE}
indi %>% group_by(Alpine) %>% tally
```

# PCA of the indices (see figures below)

```{r table3, echo = FALSE}
summary(pca)
```

# Some conclusions (see figures below)

* MGR and SYN seem to capture the most variance, and are the most intuitive
* Main differences seem to be among regions, BUT the Australian (highest mean germination time) were also the longest running experiments, I bet this is an artifact of experiment lenght (please kill me)
* Indices are not normal

```{r fig1, echo = FALSE, fig.cap = "PCA of the germination indices colored by Region"}
p1
```

```{r fig2, echo = FALSE, fig.cap = "PCA of the germination indices colored by Alpine"}
p2
```

```{r fig3, echo = FALSE, fig.cap = "PCA of the germination indices colored by germination temperature"}
p3
```

```{r fig4, echo = FALSE, fig.cap = "PCA of the germination indices colored by life form"}
p4
```

```{r fig5, echo = FALSE, fig.cap = "Normality of the indices"}
plot_grid(
# Density plot
ggdensity(indi$MGR, fill = "lightgray") + 
  theme_wsj() +
  labs(title = "MGR density plot"),
# QQ plot
ggqqplot(indi$MGR) +
    theme_wsj() +
  labs(title = "QQplot"),
# Density plot
ggdensity(indi$SYN, fill = "lightgray") + 
    theme_wsj() +
  labs(title = "SYN density plot"),
# QQ plot
ggqqplot(indi$SYN) +
    theme_wsj() +
  labs(title = "QQ plot"))
```

### 2020-05-18 src/cleaning.R

@@ -44,19 +44,19 @@ read_excel(here("data",
 # Read from original files with germination indices (S Rosbakh)
 
 read_excel(here("data",
-                "Sergey Rosbakh (Indices)", "Aust Alpine Germination timing 11MAY2020.xlsx"),
+                "Sergey Rosbakh (Indices)", "UPDATE_Aust Alpine Germination timing 11MAY2020.xlsx"),
            sheet = 4, skip = 0) %>%
   select(Species, Accession, Elevation, Treatment, Rep, seeds, GRS:CVG) -> satyanti1
 read_excel(here("data",
-                "Sergey Rosbakh (Indices)", "Aust Alpine Germination timing 11MAY2020.xlsx"),
-           sheet = 5, skip = 0)  %>%
+                "Sergey Rosbakh (Indices)", "UPDATE_Aust Alpine Germination timing 11MAY2020.xlsx"),
+           sheet = 8, skip = 0)  %>%
   select(Species, Accession, Elevation, Treatment, Rep, seeds, GRS:CVG) -> satyanti2
 read_excel(here("data",
-                "Sergey Rosbakh (Indices)", "Aust Alpine Germination timing 11MAY2020.xlsx"),
-           sheet = 6, skip = 0)  %>%
+                "Sergey Rosbakh (Indices)", "UPDATE_Aust Alpine Germination timing 11MAY2020.xlsx"),
+           sheet = 9, skip = 0)  %>%
   select(Species, Accession, Elevation, Treatment, Rep, seeds, GRS:CVG) -> satyanti3
 read_excel(here("data",
-                "Sergey Rosbakh (Indices)", "Aust Alpine Germination timing 11MAY2020.xlsx"),
+                "Sergey Rosbakh (Indices)", "UPDATE_Aust Alpine Germination timing 11MAY2020.xlsx"),
            sheet = 7, skip = 0)  %>%
   select(Species, Accession, Elevation, Treatment, Rep, seeds, GRS:CVG) -> satyanti4
 

### 2020-05-18 src/species names list.R

@@ -1,0 +1,23 @@
+
+read.csv(here("results", "database", "traits.csv")) %>% 
+  pull(TPLName) -> spp
+
+read.csv("../tpl/results/TPLNames.csv") %>%
+  mutate(TPLName = paste(New.Genus, New.Species)) %>%
+  select(TPLName, Family, New.Genus, New.Species) %>%
+  unique %>%
+  filter(TPLName %in% spp) %>%
+  write.csv("Species for tree.csv", row.names = FALSE)
+
+
+read.csv(here("results", "database", "traits.csv")) %>% 
+  filter(is.na(Embryo)) %>%
+  pull(TPLName) -> spp
+
+
+read.csv("../tpl/results/TPLNames.csv") %>%
+  mutate(TPLName = paste(New.Genus, New.Species)) %>%
+  select(TPLName, Family, New.Genus, New.Species) %>%
+  unique %>%
+  filter(TPLName %in% spp) %>%
+  write.csv("Species without embryo.csv", row.names = FALSE)

### 2020-05-19 doc/indices report/indices report.Rmd
full filled seed mass embryo

@@
 
-# Number of records with >80% germination and germination indices, by region.
+# Number of records with >90% germination and germination indices, by region.
 
@@
 
-# Number of records with >80% germination and germination indices, by alpine strict vs. generalist.
+# Number of records with >90% germination and germination indices, by alpine strict vs. generalist.
 

### 2020-05-19 doc/traits report/traits report.Rmd
full filled seed mass embryo

---
title: "Alpine traits report"
author: "Eduardo Fernández Pascual"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(tidyverse); library(here); library(FactoMineR); library(cowplot)
library(ggthemes); library(ggrepel); library(flextable); library(ggpubr)

read.csv(here("results", "database", "traits.csv")) %>%
  merge(read.csv(here("results", "database", "germination.csv")), by = "TPLName") %>% 
  merge(read.csv(here("data", "elevations", "regions.csv")), by = "Region", all.x = TRUE) %>%
  mutate(Macroregion = as.character(Macroregion),
         Macroregion = ifelse(is.na(Macroregion), "Europe", Macroregion),
         Macroregion = as.factor(Macroregion)) -> master

read.csv(here("results", "database", "traits.csv")) %>%
  select(TPLName, Seed.mass, Embryo) -> tt

```

```{r fig1, echo = FALSE, fig.cap = "Embryo by region and alpine character (brackets are SD)"}
master %>%
  group_by(TPLName, Macroregion, Alpine) %>%
  summarise(rate = mean(MGR, na.rm = TRUE), 
            syn = mean(SYN, na.rm = TRUE)) %>%
  merge(tt, by = "TPLName") %>%
  select(TPLName, Macroregion, Alpine, rate, syn, Seed.mass, Embryo) %>%
  gather(Trait, Value, rate:Embryo) %>%
  filter(is.finite(Value)) %>%
  filter(Trait == "Embryo") %>%
  group_by(Trait, Alpine, Macroregion) %>%
  summarise(y = mean(Value), SD = sd(Value)) %>%
  ggplot(aes(Alpine, y, ymin = y - SD, ymax = y + SD, fill = Alpine)) +
    theme_wsj() +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.4) + 
  facet_wrap(~ Macroregion, scales = "fixed") +
  geom_errorbar(aes(color = Alpine), width=.2,
                position=position_dodge(.9))
```

```{r fig2, echo = FALSE, fig.cap = "Seed mass by region and alpine character (brackets are SD)"}
master %>%
  group_by(TPLName, Macroregion, Alpine) %>%
  summarise(rate = mean(MGR, na.rm = TRUE), 
            syn = mean(SYN, na.rm = TRUE)) %>%
  merge(tt, by = "TPLName") %>%
  select(TPLName, Macroregion, Alpine, rate, syn, Seed.mass, Embryo) %>%
  gather(Trait, Value, rate:Embryo) %>%
  filter(is.finite(Value)) %>%
  filter(Trait == "Seed.mass") %>%
  group_by(Trait, Alpine, Macroregion) %>%
  summarise(y = mean(Value), SD = sd(Value)) %>%
  ggplot(aes(Alpine, y, ymin = y - SD, ymax = y + SD, fill = Alpine)) +
    theme_wsj() +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.4) + 
  facet_wrap(~ Macroregion, scales = "fixed") +
  geom_errorbar(aes(color = Alpine), width=.2,
                position=position_dodge(.9))
```

```{r fig3, echo = FALSE,fig.cap = "Mean germination rate by region and alpine character (brackets are SD)"}
master %>%
  group_by(TPLName, Macroregion, Alpine) %>%
  summarise(rate = mean(MGR, na.rm = TRUE), 
            syn = mean(SYN, na.rm = TRUE)) %>%
  merge(tt, by = "TPLName") %>%
  select(TPLName, Macroregion, Alpine, rate, syn, Seed.mass, Embryo) %>%
  gather(Trait, Value, rate:Embryo) %>%
  filter(is.finite(Value)) %>%
  filter(Trait == "rate") %>%
  group_by(Trait, Alpine, Macroregion) %>%
  summarise(y = mean(Value), SD = sd(Value)) %>%
  ggplot(aes(Alpine, y, ymin = y - SD, ymax = y + SD, fill = Alpine)) +
    theme_wsj() +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.4) + 
  facet_wrap(~ Macroregion, scales = "fixed") +
  geom_errorbar(aes(color = Alpine), width=.2,
                position=position_dodge(.9))
```

```{r fig4, echo = FALSE, fig.cap = "Germination synchrony by region and alpine character (brackets are SD)"}
master %>%
  group_by(TPLName, Macroregion, Alpine) %>%
  summarise(rate = mean(MGR, na.rm = TRUE), 
            syn = mean(SYN, na.rm = TRUE)) %>%
  merge(tt, by = "TPLName") %>%
  select(TPLName, Macroregion, Alpine, rate, syn, Seed.mass, Embryo) %>%
  gather(Trait, Value, rate:Embryo) %>%
  filter(is.finite(Value)) %>%
  filter(Trait == "syn") %>%
  group_by(Trait, Alpine, Macroregion) %>%
  summarise(y = mean(Value), SD = sd(Value)) %>%
  ggplot(aes(Alpine, y, ymin = y - SD, ymax = y + SD, fill = Alpine)) +
    theme_wsj() +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.4) + 
  facet_wrap(~ Macroregion, scales = "fixed") +
  geom_errorbar(aes(color = Alpine), width=.2,
                position=position_dodge(.9))
```

```{r fig5, echo = FALSE, fig.cap = "Proportion of full out of sown seeds by region and alpine character (brackets are binomial ci)"}
master %>%
  group_by(Macroregion, Alpine) %>%
  summarise(Sown = sum(Sown, na.rm = TRUE), 
            Germinable = sum(Germinable, na.rm = TRUE)) %>%
  filter(Germinable < Sown) %>% data.frame -> props

binom::binom.confint(props$Germinable, props$Sown, method = "wilson") %>% 
  cbind(props) %>%
  ggplot(aes(Alpine, mean, ymin = lower, ymax = upper, fill = Alpine)) +
    theme_wsj() +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.4) + 
  facet_wrap(~ Macroregion, scales = "fixed") +
  geom_errorbar(aes(color = Alpine), width=.2,
                position=position_dodge(.9))
```

### 2020-05-19 src/cleaning.R

@@ -1137,12 +1137,66 @@ rbind(SMBu2005, SMBu2010, LiuSM) %>%
   group_by(TPLName) %>%
   summarise(Seed.mass = mean(Seed.mass, na.rm = TRUE)) %>%
   mutate(Seed.mass = 10*Seed.mass) -> # Convert to mg (= g per 1000 seeds)
-  seed.mass
+  seed.mass1
+
+read_excel(here("data", "A Carta", "germ&traits.xlsx"), sheet = 2, skip = 0) %>%
+  select(TPLName, mass.new) %>%
+  filter(! TPLName %in% seed.mass1$TPLName) %>%
+  filter(mass.new != "NA") %>%
+  rename(Seed.mass = mass.new) ->
+  seed.mass2 # Values calculated for the genus
+
+# Missing: Helictochloa versicolor: Use SID average of genus Helictotrichon
+# rbind(c("Helictotrichon canescens", 0.4908),
+#       c("Helictotrichon desertorum", 3.4804),
+#       c("Helictotrichon elongatum",  1.3792),
+#       c("Helictotrichon fluitans", 0.7612),
+#       c("Helictotrichon hookeri", 1.7928),
+#       c("Helictotrichon parlatorei", 1.789),
+#       c("Helictotrichon planiculme", 4.7016),
+#       c("Helictotrichon pratense", 2.32),
+#       c("Helictotrichon pubescens", 1.93),
+#       c("Helictotrichon sedenense", 2.6692)) %>%
+#   data.frame-> Helictotrichon
+
+# Missing Neopaxia australasica: SID Data available for synonym
+# "Montia australasica", 0.418
+
+# Missing Nastanthus scapigerus: There is SID data for a related (I think) species
+# "Calycera eryngioides", 7.1784
+
+# Missing Dichosciadium ranunculaceum: There is SID data for related (I think) species
+#rbind(c("Azorella compacta", 2.6322),
+#      c("Azorella filamentosa", 0.7352),
+#      c("Azorella madreporica", 1.3),
+#      c("Azorella spinosa", 3.032)) %>%
+#  data.frame -> Azorella
+
+data.frame(TPLName = c("Helictochloa versicolor",
+                       "Neopaxia australasica",
+                       "Nastanthus scapigerus",
+                       "Dichosciadium ranunculaceum"),
[…]

### 2020-05-19 src/species names list.R

@@ -21,3 +21,4 @@ read.csv("../tpl/results/TPLNames.csv") %>%
   unique %>%
   filter(TPLName %in% spp) %>%
   write.csv("Species without embryo.csv", row.names = FALSE)
+

### 2020-05-19 src/cleaning.R
single table

@@ -1213,7 +1213,31 @@ germination %>%
   merge(filip, all.x = TRUE) -> # Merge embryo
   traits
 
+# Table for Angelino
+
+traits %>%
+  merge(germination, by = "TPLName") %>%
+  merge(read.csv(here("data", "elevations", "regions.csv")), by = "Region", all.x = TRUE) %>%
+  mutate(Macroregion = as.character(Macroregion),
+         Macroregion = ifelse(is.na(Macroregion), "Europe", Macroregion),
+         Stratification = as.character(Stratification),
+         Stratification_Y_N = ifelse(Stratification == "None", "N", "Y")) %>%
+  select(TPLName, 
+         Alpine, `Plant category`, `Life form`, Lifespan, `Reproduction frequency`,
+         Dormancy, Embryo, Seed.mass,
+         Source, Macroregion, Region, Accession,
+         Dish,
+         Scarification, GA3, 
+         Stratification,
+         Stratification_Y_N,
+         Light, Alternating,
+         Tmean, Tdif, Temperature,
+         Sown, Germinated, Germinable,
+         MGT, MGR,
+         UNC, SYN) -> angelino
+
 # Save
 
 write.csv(germination, here("results", "database", "germination.csv"), row.names = FALSE)
 write.csv(traits, here("results", "database", "traits.csv"), row.names = FALSE)
+write.csv(angelino, here("results", "database", "angelino.csv"), row.names = FALSE)

### 2020-06-04 doc/mcmc report/mcm-report.Rmd
Neopaxia

---
title: "MCMC report"
author: "Eduardo Fernández Pascual"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(tidyverse); library(here); library(readxl)

read_excel(here("results", "analysis", "results.preliminary.xlsx")) -> sheet1

sheet1[8:19, ]  %>%
  select(variable, post.mean, l.95..CI, u.95..CI, pMCMC) %>%
  separate(variable, into = c("Alpine", "Treatment"), sep = ":") %>%
  mutate(Alpine = gsub("Alpine", "", Alpine)) %>%
  rename(Mean = post.mean,
         Lower = l.95..CI,
         Upper = u.95..CI,
         p = pMCMC) -> df1

ggplot(df1, aes(y = Alpine, x = Mean, xmin = Lower, xmax = Upper, color = Alpine)) +
  geom_point() +
  labs(title = "Final germination proportion by alpine") + 
  geom_errorbarh(height = .1) +
  facet_grid(~ Treatment, scales = "free") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme(legend.position = "none", axis.title = element_blank()) -> fig1

sheet1[21:57, ]  %>%
  select(variable, post.mean, l.95..CI, u.95..CI, pMCMC) %>%
  separate(variable, into = c("Macroregion", "Alpine", "Treatment"), sep = ":") %>%
  mutate(Alpine = gsub("Alpine", "", Alpine),
         Macroregion = gsub("Macroregion", "", Macroregion)) %>%
  rename(Mean = post.mean,
         Lower = l.95..CI,
         Upper = u.95..CI,
         p = pMCMC) -> df2

ggplot(df2, aes(y = Alpine, x = Mean, xmin = Lower, xmax = Upper, color = Alpine)) +
  geom_point() +
  labs(title = "Final germination proportion by alpine and region") + 
  geom_errorbarh(height = .1) +
  facet_grid(Macroregion ~ Treatment, scales = "free") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme(legend.position = "none", axis.title = element_blank()) -> fig2
```

```{r echo = FALSE, fig.height = 1.5}
fig1
```

```{r echo = FALSE, fig.height = 3}
fig2
```

### 2020-06-04 src/analysis.R

@@ -1,0 +1,76 @@
+library(tidyverse); library(here)
+
+read.csv(here("results", "analysis", "alpR.germ.type.csv")) %>%
+  select(variable, post.mean, l.95..CI, u.95..CI, pMCMC) %>%
+  separate(variable, into = c("Alpine", "Treatment"), sep = ":") %>%
+  mutate(Alpine = gsub("Alpine", "", Alpine)) %>%
+  rename(Mean = post.mean,
+         Lower = l.95..CI,
+         Upper = u.95..CI,
+         p = pMCMC) %>%
+  ggplot(aes(y = Alpine, x = Mean, 
+             xmin = Lower, xmax = Upper, 
+             color = Alpine)) +
+  geom_point() +
+  labs(title = "Final germination proportion") + 
+  geom_errorbarh(height = .1) +
+  facet_grid(~ Treatment, scales = "free") +
+  geom_vline(xintercept = 0, linetype = "dashed") +
+  theme(legend.position = "none", axis.title = element_blank())
+
+read.csv(here("results", "analysis", "alpR.MGT.type.csv")) %>%
+  select(variable, post.mean, l.95..CI, u.95..CI, pMCMC) %>%
+  separate(variable, into = c("Alpine", "Treatment"), sep = ":") %>%
+  mutate(Alpine = gsub("Alpine", "", Alpine)) %>%
+  rename(Mean = post.mean,
+         Lower = l.95..CI,
+         Upper = u.95..CI,
+         p = pMCMC) %>%
+  ggplot(aes(y = Alpine, x = Mean, 
+             xmin = Lower, xmax = Upper, 
+             color = Alpine)) +
+  geom_point() +
+  labs(title = "Mean germination time") + 
+  geom_errorbarh(height = .1) +
+  facet_grid(~ Treatment, scales = "free") +
+  geom_vline(xintercept = 0, linetype = "dashed") +
+  theme(legend.position = "none", axis.title = element_blank())
+
+read.csv(here("results", "analysis", "alpR.UNC.type.csv")) %>%
+  select(variable, post.mean, l.95..CI, u.95..CI, pMCMC) %>%
+  separate(variable, into = c("Alpine", "Treatment"), sep = ":") %>%
+  mutate(Alpine = gsub("Alpine", "", Alpine)) %>%
+  rename(Mean = post.mean,
+         Lower = l.95..CI,
+         Upper = u.95..CI,
+         p = pMCMC) %>%
[…]

### 2020-06-04 src/cleaning.R

@@ -1176,11 +1176,11 @@ read_excel(here("data", "A Carta", "germ&traits.xlsx"), sheet = 2, skip = 0) %>%
 #  data.frame -> Azorella
 
 data.frame(TPLName = c("Helictochloa versicolor",
-                       "Neopaxia australasica",
+                       "Neopaxia australasica", # Told by S Venn
                        "Nastanthus scapigerus",
                        "Dichosciadium ranunculaceum"),
            Seed.mass = c(mean(c(0.4908, 3.4804,  1.3792, 0.7612, 1.7928, 1.789, 4.7016, 2.32, 1.93, 2.6692)),
-                         0.418,
+                         -0.7522,
                          7.1784,
                          mean(c(2.6322, 0.7352, 1.3, 3.032)))) -> seed.mass3
 

### 2020-06-05 src/cleaning.R
corrected Neopaxia negative mass

@@ -1180,7 +1180,7 @@ data.frame(TPLName = c("Helictochloa versicolor",
                        "Nastanthus scapigerus",
                        "Dichosciadium ranunculaceum"),
            Seed.mass = c(mean(c(0.4908, 3.4804,  1.3792, 0.7612, 1.7928, 1.789, 4.7016, 2.32, 1.93, 2.6692)),
-                         -0.7522,
+                         0.7522,
                          7.1784,
                          mean(c(2.6322, 0.7352, 1.3, 3.032)))) -> seed.mass3
 
@@ -1221,21 +1221,28 @@ germination %>%
 traits %>%
   merge(germination, by = "TPLName") %>%
   merge(read.csv(here("data", "elevations", "regions.csv")), by = "Region", all.x = TRUE) %>%
-  mutate(Macroregion = as.character(Macroregion),
+  mutate(TPLName = gsub(" ", "_", TPLName),
+         Macroregion = as.character(Macroregion),
          Macroregion = ifelse(is.na(Macroregion), "Europe", Macroregion),
-         Stratification = as.character(Stratification),
-         Stratification_Y_N = ifelse(Stratification == "None", "N", "Y")) %>%
+         Stratification2 = as.character(Stratification),
+         Scarification = ifelse(Scarification == "Y", 1, 0),
+         Stratification = ifelse(Stratification == "None", 0, 1),
+         GA3 = ifelse(GA3 == "Y", 1, 0),
+         Light = ifelse(Light == "Y", 1, 0),
+         Alternating = ifelse(Alternating == "Y", 1, 0),
+         percent = Germinated / Germinable) %>%
+  rename(embryo = Embryo, mass = Seed.mass) %>%
   select(TPLName, 
          Alpine, `Plant category`, `Life form`, Lifespan, `Reproduction frequency`,
-         Dormancy, Embryo, Seed.mass,
+         Dormancy, embryo, mass,
          Source, Macroregion, Region, Accession,
-         Dish,
+         Dish, Stratification2,
          Scarification, GA3, 
          Stratification,
-         Stratification_Y_N,
          Light, Alternating,
          Tmean, Tdif, Temperature,
          Sown, Germinated, Germinable,
+         percent,
          MGT, MGR,
          UNC, SYN) -> angelino
 
@@ -1243,4 +1250,4 @@ traits %>%
 
 write.csv(germination, here("results", "database", "germination.csv"), row.names = FALSE)
[…]

### 2020-06-06 src/species names list.R
cleaned html

@@ -1,24 +1,0 @@
-
-read.csv(here("results", "database", "traits.csv")) %>% 
-  pull(TPLName) -> spp
-
-read.csv("../tpl/results/TPLNames.csv") %>%
-  mutate(TPLName = paste(New.Genus, New.Species)) %>%
-  select(TPLName, Family, New.Genus, New.Species) %>%
-  unique %>%
-  filter(TPLName %in% spp) %>%
-  write.csv("Species for tree.csv", row.names = FALSE)
-
-
-read.csv(here("results", "database", "traits.csv")) %>% 
-  filter(is.na(Embryo)) %>%
-  pull(TPLName) -> spp
-
-
-read.csv("../tpl/results/TPLNames.csv") %>%
-  mutate(TPLName = paste(New.Genus, New.Species)) %>%
-  select(TPLName, Family, New.Genus, New.Species) %>%
-  unique %>%
-  filter(TPLName %in% spp) %>%
-  write.csv("Species without embryo.csv", row.names = FALSE)
-

### 2020-06-18 doc/mcmc report/mcm-report.Rmd
map figure

@@
 title: "MCMC report"
-author: "Eduardo Fernández Pascual"
 date: "`r format(Sys.time(), '%d/%m/%Y')`"
@@
 ```{r message = FALSE, echo = FALSE, warning = FALSE}
-library(tidyverse); library(here); library(readxl)
+library(tidyverse); library(here)
 
-read_excel(here("results", "analysis", "results.preliminary.xlsx")) -> sheet1
+source(here("src", "analysis.R"))
 
-sheet1[8:19, ]  %>%
-  select(variable, post.mean, l.95..CI, u.95..CI, pMCMC) %>%
-  separate(variable, into = c("Alpine", "Treatment"), sep = ":") %>%
-  mutate(Alpine = gsub("Alpine", "", Alpine)) %>%
-  rename(Mean = post.mean,
-         Lower = l.95..CI,
-         Upper = u.95..CI,
-         p = pMCMC) -> df1
-
-ggplot(df1, aes(y = Alpine, x = Mean, xmin = Lower, xmax = Upper, color = Alpine)) +
-  geom_point() +
-  labs(title = "Final germination proportion by alpine") + 
-  geom_errorbarh(height = .1) +
-  facet_grid(~ Treatment, scales = "free") +
-  geom_vline(xintercept = 0, linetype = "dashed") +
-  theme(legend.position = "none", axis.title = element_blank()) -> fig1
-
-sheet1[21:57, ]  %>%
-  select(variable, post.mean, l.95..CI, u.95..CI, pMCMC) %>%
-  separate(variable, into = c("Macroregion", "Alpine", "Treatment"), sep = ":") %>%
-  mutate(Alpine = gsub("Alpine", "", Alpine),
-         Macroregion = gsub("Macroregion", "", Macroregion)) %>%
-  rename(Mean = post.mean,
-         Lower = l.95..CI,
-         Upper = u.95..CI,
-         p = pMCMC) -> df2
-
-ggplot(df2, aes(y = Alpine, x = Mean, xmin = Lower, xmax = Upper, color = Alpine)) +
-  geom_point() +
-  labs(title = "Final germination proportion by alpine and region") + 
-  geom_errorbarh(height = .1) +
-  facet_grid(Macroregion ~ Treatment, scales = "free") +
-  geom_vline(xintercept = 0, linetype = "dashed") +
-  theme(legend.position = "none", axis.title = element_blank()) -> fig2
 ```
 
-```{r echo = FALSE, fig.height = 1.5}
+# Manuscript figures
+
+```{r echo = FALSE, fig.height = 5.5}
 fig1
@@
 
+# Final germination proportions
+
 ```{r echo = FALSE, fig.height = 3}
-fig2
+fgFull
 ```
+
+```{r echo = FALSE, fig.height = 3}
+fgAlpine
+```
+
+```{r echo = FALSE, fig.height = 3}
+fgRegion
+```
+
+```{r echo = FALSE, fig.height = 3}
+fgDormancy
+```
+
+```{r echo = FALSE, fig.height = 3}
+fgPlanttype
+```
+```{r echo = FALSE, fig.height = 3}
+fgLifeform
+```
+
+```{r echo = FALSE, fig.height = 3}
+fgLifespan
+```
+
+# Mean germination time
+
+```{r echo = FALSE, fig.height = 3}
+mgtFull
+```
+
+```{r echo = FALSE, fig.height = 3}
+mgtAlpine
+```
+
+```{r echo = FALSE, fig.height = 3}
+mgtRegion
+```
+
+```{r echo = FALSE, fig.height = 3}
+mgtDormancy
+```
+
+```{r echo = FALSE, fig.height = 3}
+mgtPlanttype
+```
+```{r echo = FALSE, fig.height = 3}
+mgtLifeform
+```
+
+```{r echo = FALSE, fig.height = 3}
+mgtLifespan
+```
+
+# Uncertainty
+
+```{r echo = FALSE, fig.height = 3}
+uncFull
+```
+
+```{r echo = FALSE, fig.height = 3}
+uncAlpine
+```
+
+```{r echo = FALSE, fig.height = 3}
+uncRegion
+```
+
+```{r echo = FALSE, fig.height = 3}
+uncDormancy
+```
+
+```{r echo = FALSE, fig.height = 3}
+uncPlanttype
+```
+```{r echo = FALSE, fig.height = 3}
+uncLifeform
+```
+
+```{r echo = FALSE, fig.height = 3}
+uncLifespan
+```

### 2020-06-18 doc/ms/ms.Rmd
map figure

---
title: "Alpine Seeds MS"
date: "15/6/2020"
output: 
  bookdown::html_document2:
    reference_docx: gebtemplate.docx
bibliography: refs.bib
csl: global-ecology-and-biogeography.csl
link-citations: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",")
})
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(tidyverse); library(here)
source(here("src", "msSource.R"))
```

# Introduction

Alpine environments are those that occur above the maximum elevation at which trees can grow, and are characterized by low temperatures, strong winds, unstable substrates, and short growing seasons [@RN2392]. The short alpine growing season is a major barrier for plant recruitment, as it constrains the period favorable for seedling establishment [@RN4714; @RN4717]. Alpine seedlings must survive extreme environmental stresses and yet grow to a critical biomass by the end of the growing season, to withstand the harsh and long-lasting winter [@RN2868]. The physiological process of seed germination is the essential, yet most vulnerable, start of seedling establishment [@RN3397]. Because it is an irreversible process, germination must be timed to occur when the environment is favourable for subsequent seedling survival and growth [@RN4691]. Thus, the mechanisms regulating the timing of the seed-to-seedling transition are expected to be under strong selective pressure [@RN3375; @RN2968]. Natural selection should favour seed germination requirements that reduce the probability of facing environmental conditions that are not appropriate for seedlings. The main physiological drivers of germination, water availability and temperature [@RN3368], are also two main features of climate. In addition, many plant species have developed seed dormancy, by which germination is prevented during periods that are only ephemerally favourable, ensuring germination in the right season [@RN3063]. Different degrees of dormancy also ensure the distribution of offspring emergence across time, as bet-hedging against unpredictable environments [@RN3065]. Other germination cues allow for a detection of fine-scale germination niches and safe sites [@RN4719], such as the response to light [@RN4656] and alternating temperatures [@RN1380]. In the alpine environment, the diversity of microhabitats has apparently resulted in variable germination ecologies across species, making it difficult to define a common *alpine germination syndrome* [@RN2392; @RN2943].

Early studies on the germination ecology of alpine plants demonstrated that freshly collected seeds of most arctic and alpine species required relatively high temperatures for germination [@RN3258; @RN3213; @RN4712]. High-temperature germination has been considered to be an adaptation to prevent seed germination at the timing of seed dispersal (autumn), when temperatures are low and there is a high risk of frost [@RN3393]. Indeed, seed germination of alpine species tends to occur after exposure to winter in early summer [@RN2392; @RN2382]; at this stage the temperature window for germination usually widens towards lower values [@RN1021]. Nevertheless, there is an increasing number of studies highlighting the germination of freshly collected seeds at cool incubation temperatures or even during cold stratification treatments [@RN2943; @RN3285; @RN2371; @RN4713]. Low-temperature germination could be an adaptation to germination under snow or at snowmelt, which could presumably allow seedlings to develop a deeper root system before topsoil desiccation in summer [@RN4720], as well as to attain an optimal size for overwintering [@RN4712].

Many alpine species have been described as having deep physiological dormancy [@RN2943; @RN3703; @RN3214]. Dormancy in freshly-dispersed seeds would prevent precocious germination under autumn conditions, when appropriate soil moisture and temperature are not likely to persist for more than a few weeks or days [@RN3330]. A requirement for cold stratification to break dormancy would allow seeds to sense the snow season, thereby timing germination to a posterior period, more appropriate for seedling survival and growth. Thus, it can be expected that a cold stratification period would be a common requirement for seed germination in alpine plant species. Nonetheless, in an early review, Amen [-@RN3213] concluded that cold stratification was not a common requirement for the seed germination of several alpine plant species from different mountains in the USA (see also Sayers and Ward [-@RN1484]; Marchand and Roach [-@RN2995]; Kaye [-@RN4721]). More recently, Sommerville *et al.* [-@RN3703] and Hoyle *et al.* [-@RN3285] showed that in 19 and 54 Australian alpine plant species, respectively, a cold stratification period significantly increased seed germination only in half of the species tested. In contrast, Söyrinki (1938, cited in Körner [-@RN2392]), experimenting with 91 alpine species from the Alps, found that storage at winter temperatures (i.e. cold stratification) increased seed germination in the great majority of cases. Shimono and Kudo [-@RN707] reported that cold stratification increased seed germination over a range of temperatures in most of the 27 plant species studied in the alpine zone of Japan. Cavieres and Sierra-Almeida [-@RN4713] reported that, along an elevational gradient in the central Chilean Andes, cold stratification was an important requirement for seed germination in species from lower elevations, while species from higher sites still did not germinate after cold stratification. Therefore, how important cold stratification is for seed germination in alpine plant species remains unclear and requires a global assessment.

The response of alpine seeds to light and alternating temperatures can provide further information on their capacity to detect fine-scale environmental variation. The preference of some species to germinate in light [@RN4730; @RN4729] could favour the creation of a persistent soil seed bank [@RN4718], adventageous in temporally and spatially unpredictable alpine environments [@RN4731; @RN4732]. Some species seem to prefer dark germination [@RN2943], a behaviour interpreted by some as detecting safe sites in rock crevices [@RN3008]. Regarding alternating temperatures, a study of 445 species from the Qinhai-Tibet plateau detected that species from high elevation alpine meadows did not show a significant response to diurnal temperature oscillations [@RN3376].

# Materials and methods

## Data collection

We have conducted a meta-analysis of primary data [@RN4734]. To gather data, we contacted research groups who have performed experiments on the germination of alpine seeds, asking them to send raw data from laboratory germination experiments. The submitted data had to report the studied species, the germination treatments, the number of seeds sown and the number of seeds germinated. We received data from the Cantabrian Mountains [@RN2371], the Alps [@RN4729; @RN3029; @RN4727; Rosbakh, unpublished], the Apennines (Mondoni, unpublished), the Caucasus [Rosbakh, unpublished], the Qinhai-Tibet Plateau [@RN4728; @RN4735; @RN3376], the Andes [@RN3393; @RN4713; Briceño, unpublished] and the Australian Alps [@RN4737; @RN4736; @RN3703; Satyanti, unpublished]. We also received data on European alpine species from *Enscobase* (http://enscobase.maich.gr/index.tml). Once received, we standardized the species names using *The Plant List* [@RN2321] with the package *Taxonstand* [@RN4660] in *R* [@RN2315]. For each species, we checked its elevation range in local floras [@RN4741; @RN4739; @RN3243; @RN4742; @RN4740], catalogues [@RN4738] and herbaria (*Australian National Herbarium*). We used to elevation ranges to (i) remove from the dataset species that do not grow above the treeline; and (ii) classify the rest of the species as *strict alpine* (i.e. species that only grow above the subalpine) or *generalist* (i.e. species that can grow above the subalpine but also in the lowlands). For each species, we also compiled the following traits: life form and life span from the references used for the elevation, plus some extra references [@RN4747; @RN4744; RN4745; @RN4746; @RN4748; @RN4749] and online sources (https://www.infoflora.ch/de/; http://plantnet.rbgsyd.nsw.gov.au/search/simple.htm; http://buscador.floraargentina.edu.ar/; https://patagoniawildflowers.org/); seed dormancy class [@RN3214; @RN4743]; seed mass [@RN4728; @RN4735; @RN3376; @RN2987] and embryo to endosperm ratio (*Vandelook, ???*). For the species missing seed mass or embryo values, we used genus or family means.

## Description of the dataset

The final dataset contains `r MSrecords` records (germination proportions for a given seed lot of a species, recorded in a set of experimental conditions) contributed by `r MSgroups` research groups, representing `r MSfamilies` seed plant families and `r MSspecies` species (`r MSstrict` strict alpine species and `r MSgeneral` generalists). Seed lots had been collected in `r MSregions` alpine regions of the world (**Fig. \@ref(fig:fig1)**), in both hemispheres. The total number of seeds used in the experiments is `r MSseeds`. The range of germination temperatures (weighted average of the daily thermoperiod) goes from `r MSminT` to `r MSmaxT` ºC, with `r MSaltY` records of constant temperatures and `r MSaltN` of alternating temperatures. Light was used in `r MSlightY` records, darkness in `r MSlightN`. The experiments were performed with unstratified seeds in `r MSstratN` records, and of the rest, the majority (`r MSstratCold`) went through cold stratification. Scarification was applied to `r MSscarY` records, and GA~3~ to `r MSga3`.

## Data analysis

Across species, germination and dormancy patterns are affected by a combination of factors such as phylogeny [@RN4723; @RN4724; @RN4715] and seed traits [@RN3061; @RN4728].

# Discussion

The influence of environmental conditions as drivers of these contrasting germination responses to temperature is gaining recognition. For example, alpine species show higher optimal temperatures for germination than either subalpine species (i.e. species that live close to the treeline, Fernández-Pascual *et al.* [-@RN2371]) or congeneric counterparts from below the treeline [@RN3266]. Differences in germination traits have been attributed also to species successional niche, with pioneer species germinating better at colder temperatures than later successional species [@RN3229]. Species from calcareous and siliceous alpine grasslands show a slow overwinter germination or high germination under all conditions, respectively [@RN4727; @RN4726]. Other factors related to species germination characteristics include slope [@RN4733] and biogeographical origin [@RN698]. Nevertheless, no habitat-related germination strategies were identified when comparing a major divide in alpine microhabitats, as is the one between fellfields and snowbeds [@RN707].

# Figures

```{r fig1, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Geographical origin of the germination data with the number of records, species and strict alpine species per alpine region. Ensconet is not plotted because its records come from alpine regions throughout Europe. The purple areas correspond to the global extension of alpine environments according to Testolin *et al.* [-@RN4750].", fig.align = "center", fig.width=7, fig.height=3}
cowplot::ggdraw(fig1)
```

# References

### 2020-06-18 src/analysis.R

@@ -1,6 +1,187 @@
 library(tidyverse); library(here)
 
-read.csv(here("results", "analysis", "alpR.germ.type.csv")) %>%
+# Final germination proportion
+
+## Figure for ms
+
+rbind(
+  rbind(
+    read.csv(here("results", "analysis", "proportion", "alpR.germ.general.csv")) %>%
+      select(model, post.mean, l.95..CI, u.95..CI, pMCMC) %>%
+      mutate(Moderator = "Alone"),
+    read.csv(here("results", "analysis", "proportion", "alpR.germ.general.embryo.csv")) %>%
+      select(model, post.mean, l.95..CI, u.95..CI, pMCMC) %>% 
+      mutate(Moderator = "Embryo"),
+    read.csv(here("results", "analysis", "proportion", "alpR.germ.general.mass.csv")) %>%
+      select(model, post.mean, l.95..CI, u.95..CI, pMCMC) %>% 
+      mutate(Moderator = "Mass")) %>%
+    rename(Mean = post.mean,
+           Lower = l.95..CI,
+           Upper = u.95..CI,
+           p = pMCMC,
+           Treatment = model) %>%
+    select(Moderator, Treatment, Mean, Lower, Upper, p),
+  read.csv(here("results", "analysis", "proportion", "alpR.germ.type.csv")) %>%
+    select(variable, post.mean, l.95..CI, u.95..CI, pMCMC) %>%
+    separate(variable, into = c("Moderator", "Treatment"), sep = ":") %>%
+    mutate(Moderator = gsub("Alpine", "", Moderator)) %>%
+    rename(Mean = post.mean,
+           Lower = l.95..CI,
+           Upper = u.95..CI,
+           p = pMCMC)) %>%
+  mutate(Moderator = recode(Moderator, Strict = "Strict alpine spp.",
+                            Generalist = "Generalist spp.",
+                            Mass = "Seed mass",
+                            Embryo = "Embryo:endosperm",
+                            Alone = "Main effect"),
+         Moderator = fct_relevel(Moderator, c("Embryo:endosperm", "Seed mass",  "Generalist spp.", "Strict alpine spp.", "Main effect")),
+         Treatment = recode(Treatment, Tmean = "Temperature", GA3 = "GA[3]"),
[…]

### 2020-06-18 src/cleaning.R

@@ -1027,7 +1027,7 @@ rbind(AustraliaVenn, AustraliaSommerville, AustraliaSat, ChileBriceño, ChileCav
          Sown = as.numeric(Sown)) %>%
   rename(Germinable = Normal) %>%
   filter(! is.na(Germinated)) %>%
-  merge(read.csv("../tpl/results/TPLNames.csv"), by = "Taxon") %>%
+  merge(read.csv("../#tpl/results/TPLNames.csv"), by = "Taxon") %>%
   mutate(TPLName = paste(New.Genus, New.Species)) %>%
   #unique %>% # Remove duplicates (may remove non duplicates with same info)
   merge(read_excel(here("data", "Peter Poschlod", "Traits_to_complete_PP_20200510.xlsx")), 
@@ -1132,7 +1132,7 @@ read_excel(path = here("data", "Lydia Guja (Liu et al.)",
 # Merge seed mass files
 
 rbind(SMBu2005, SMBu2010, LiuSM) %>%
-  merge(read.csv("../tpl/results/TPLNames.csv")) %>%
+  merge(read.csv("../#tpl/results/TPLNames.csv")) %>%
   mutate(TPLName = paste(New.Genus, New.Species)) %>%
   select(TPLName, Seed.mass) %>%
   rbind(smassCarta) %>%
@@ -1212,7 +1212,7 @@ germination %>%
   unique %>%
   merge(life.form, all.x = TRUE) %>%
   merge(seed.mass, all.x = TRUE) %>% # Merge seed mass
-  merge(read.csv("../baskin/results/dormancy.csv"), all.x = TRUE) %>%
+  merge(read.csv("../#baskin/results/dormancy.csv"), all.x = TRUE) %>%
   merge(filip, all.x = TRUE) -> # Merge embryo
   traits
 

### 2020-06-18 src/msSource.R

@@ -1,0 +1,94 @@
+library(tidyverse); library(here)
+
+read.csv(here("data", "Elevations", "Regions.csv")) %>%
+  merge(read.csv(here("results", "database", "germination.csv")), by = "Region", all = TRUE) %>%
+  merge(read.csv(here("results", "database", "traits.csv")), by = "TPLName") %>%
+  mutate(Region = as.character(Region),
+         Region = ifelse(is.na(Macroregion), "Enscobase", Region),
+         Region = ifelse(Region %in% c("Northern Alps", "Southern Alps"), "Alps", Region),
+         Region = ifelse(Region %in% c("Central Andes", "Southern Andes"), "Andes", Region),
+         Region = ifelse(Region %in% c("Qinghai-Tibet Plateau"), "Qinghai-Tibet", Region),
+         Region = ifelse(Region %in% c("Cantabrian Mountains"), "Cantabrian Mts", Region),     
+         Region = as.factor(Region)) -> dataset
+
+# Numbers for the main text
+
+dataset %>% summary
+dataset %>% pull(Region) %>% unique %>% length -> MSregions
+dataset %>% tally() -> MSrecords
+dataset %>% pull(Source) %>% unique %>% length -> MSgroups
+read.csv("../../../#tpl/results/TPLNames.csv") %>% 
+  select(Taxon, Family) %>%
+  unique %>%
+  merge(dataset, by = "Taxon", all.y = TRUE) %>%
+  pull(Family) %>% unique %>% length -> MSfamilies
+dataset %>% pull(TPLName) %>% unique %>% length -> MSspecies
+dataset %>% filter(Alpine == "Strict") %>% pull(TPLName) %>% unique %>% length -> MSstrict
+dataset %>% filter(Alpine == "Generalist") %>% pull(TPLName) %>% unique %>% length -> MSgeneral
+dataset$Germinable %>% sum -> MSseeds
+dataset %>% pull(Tmean) %>% min -> MSminT
+dataset %>% pull(Tmean) %>% max -> MSmaxT
+dataset %>% filter(Alternating == "Y") %>% tally -> MSaltY
+dataset %>% filter(Alternating == "N") %>% tally -> MSaltN
+dataset %>% filter(Light == "Y") %>% tally -> MSlightY
+dataset %>% filter(Light == "N") %>% tally -> MSlightN
+dataset %>% filter(Stratification == "None") %>% tally -> MSstratN
+dataset %>% filter(Stratification == "Cold") %>% tally -> MSstratCold
[…]

### 2020-06-30 doc/ms/ms.Rmd
1st draft sent

@@
 ---
-title: "Alpine Seeds MS"
-date: "15/6/2020"
+title: "Seed germination ecology of alpine plants: a global meta-analysis of primary data"
+date: "`r format(Sys.time(), '%d/%m/%Y')`"
 output: 
-  bookdown::html_document2:
-    reference_docx: gebtemplate.docx
+  bookdown::word_document2:
+    reference_docx: template.docx
 bibliography: refs.bib
-csl: global-ecology-and-biogeography.csl
+csl: new-phytologist.csl
 link-citations: yes
@@
 
-Alpine environments are those that occur above the maximum elevation at which trees can grow, and are characterized by low temperatures, strong winds, unstable substrates, and short growing seasons [@RN2392].
-The short alpine growing season is a major barrier for plant recruitment, as it constrains the period favorable for seedling establishment [@RN4714; @RN4717].
-Alpine seedlings must survive extreme environmental stresses and yet grow to a critical biomass by the end of the growing season, to withstand the harsh and long-lasting winter [@RN2868].
-The physiological process of seed germination is the essential, yet most vulnerable, start of seedling establishment [@RN3397].
+Apine environments are those that occur above the maximum elevation at which trees can grow, and are characterized by low temperatures, strong winds, unstable substrates, and short growing seasons [@RN2392].
+The short alpine summer is a major barrier for plant recruitment, as it constrains the period favorable for seedling establishment [@RN4714; @RN4717].
+Alpine seedlings must survive extreme environmental stress and yet grow to a critical biomass by the end of the growing season, to withstand the harsh and long-lasting winter [@RN2868].
+The physiological process of seed germination is the essential, yet most vulnerable, start of seedling establishment.
 Because it is an irreversible process, germination must be timed to occur when the environment is favourable for subsequent seedling survival and growth [@RN4691].
-Thus, the mechanisms regulating the timing of the seed-to-seedling transition are expected to be under strong selective pressure [@RN3375; @RN2968].
+Thus, the mechanisms regulating the timing of the seed-to-seedling transition are expected to be under strong selective pressure [@RN3375].
 Natural selection should favour seed germination requirements that reduce the probability of facing environmental conditions that are not appropriate for seedlings.
 The main physiological drivers of germination, water availability and temperature [@RN3368], are also two main features of climate.
-In addition, many plant species have developed seed dormancy, by which germination is prevented during periods that are only ephemerally favourable, ensuring germination in the right season [@RN3063].
+In addition, many plant species have developed seed dormancy, by which germination is prevented during periods that are only ephemerally favourable, ensuring germination in the right season [@RN3214].
 Different degrees of dormancy also ensure the distribution of offspring emergence across time, as bet-hedging against unpredictable environments [@RN3065].
-Other germination cues allow for a detection of fine-scale germination niches and safe sites [@RN4719], such as the response to light [@RN4656] and alternating temperatures [@RN1380].
+Other germination cues, such as the response to light [@RN4656] and alternating temperatures [@RN1380], allow for a fine-scale detection of germination niches and safe sites [@RN4719].
 In the alpine environment, the diversity of microhabitats has apparently resulted in variable germination ecologies across species, making it difficult to define a common *alpine germination syndrome* [@RN2392; @RN2943].
@@
 High-temperature germination has been considered to be an adaptation to prevent seed germination at the timing of seed dispersal (autumn), when temperatures are low and there is a high risk of frost [@RN3393].
-Indeed, seed germination of alpine species tends to occur after exposure to winter in early summer [@RN2392; @RN2382]; at this stage the temperature window for germination usually widens towards lower values [@RN1021].
+Indeed, germination of alpine seeds tends to occur after exposure to winter in early summer [@RN2392; @RN2382]; at this stage the temperature window for germination usually widens towards lower values [@RN1021].
 Nevertheless, there is an increasing number of studies highlighting the germination of freshly collected seeds at cool incubation temperatures or even during cold stratification treatments [@RN2943; @RN3285; @RN2371; @RN4713].
@@
 A requirement for cold stratification to break dormancy would allow seeds to sense the snow season, thereby timing germination to a posterior period, more appropriate for seedling survival and growth.
-Thus, it can be expected that a cold stratification period would be a common requirement for seed germination in alpine plant species.
-Nonetheless, in an early review, Amen [-@RN3213] concluded that cold stratification was not a common requirement for the seed germination of several alpine plant species from different mountains in the USA (see also Sayers and Ward [-@RN1484]; Marchand and Roach [-@RN2995]; Kaye [-@RN4721]).
+Thus, it can be expected that a cold stratification period would be a common requirement for seed germination in alpine plants.
+Nonetheless, in an early review, Amen [-@RN3213] concluded that cold stratification was not a common requirement for the seed germination of several alpine species from different mountains in the USA (see also Sayers and Ward [-@RN1484]; Marchand and Roach [-@RN2995]; Kaye [-@RN4721]).
 More recently, Sommerville *et al.* [-@RN3703] and Hoyle *et al.* [-@RN3285] showed that in 19 and 54 Australian alpine plant species, respectively, a cold stratification period significantly increased seed germination only in half of the species tested.
@@
 For each species, we checked its elevation range in local floras [@RN4741; @RN4739; @RN3243; @RN4742; @RN4740], catalogues [@RN4738] and herbaria (*Australian National Herbarium*).
-We used to elevation ranges to (i) remove from the dataset species that do not grow above the treeline; and (ii) classify the rest of the species as *strict alpine* (i.e.
+We used the elevation ranges to (i) remove from the dataset species that do not grow above the treeline; and (ii) classify the rest of the species as *strict alpine* (i.e.
 species that only grow above the subalpine) or *generalist* (i.e.
 species that can grow above the subalpine but also in the lowlands).
-For each species, we also compiled the following traits: life form and life span from the references used for the elevation, plus some extra references [@RN4747; @RN4744; RN4745; @RN4746; @RN4748; @RN4749] and online sources (https://www.infoflora.ch/de/; http://plantnet.rbgsyd.nsw.gov.au/search/simple.htm; http://buscador.floraargentina.edu.ar/; https://patagoniawildflowers.org/); seed dormancy class [@RN3214; @RN4743]; seed mass [@RN4728; @RN4735; @RN3376; @RN2987] and embryo to endosperm ratio (*Vandelook, ???*).
+For each species, we also compiled the following traits: life form and life span from the references used for the elevation, plus some extra references [@RN4747; @RN4744; RN4745; @RN4746; @RN4748; @RN4749] and online sources (https://www.infoflora.ch/de/; http://plantnet.rbgsyd.nsw.gov.au/search/simple.htm; http://buscador.floraargentina.edu.ar/; https://patagoniawildflowers.org/); seed dormancy class [@RN3214; @RN4743]; seed mass [@RN4728; @RN4735; @RN3376; @RN2987] and embryo to endosperm ratio (Vandelook, unpublished).
 For the species missing seed mass or embryo values, we used genus or family means.
@@
 Scarification was applied to `r MSscarY` records, and GA~3~ to `r MSga3`.
+Germination incubations had ran for different periods of time, we established a cut-off time of 8 weeks to calculate the final germination percentages, in order to have an homogeneous comparison of germination.
 
-## Data analysis
+## Germination indices
 
-Across species, germination and dormancy patterns are affected by a combination of factors such as phylogeny [@RN4723; @RN4724; @RN4715] and seed traits [@RN3061; @RN4728].
+As a response variable for the germination treatments, we used the final germination proportion and, in addition, we calculated the mean time to germination (MGT) and the germination uncertainty index (UNC) with the package *GerminaR* [@RN4752].
+MGT indicates the time it takes for half of the seed lot to germinate, while UNC estimates how scattered germination is through time.
+To calculate these indices, we used only records that had (i) more than 50% final germination *check with Angelino*; and (ii) less than 4 weeks of MGT; we established these limits to prevent the results being altered by a modification of seed dormancy status during the germination incubation.
+
+## MCMC analysis
+
+We meta-analysed germination data by fitting generalized mixed models with Bayesian estimation (Markov Chain Monte Carlo generalized linear mixed models, MCMCglmms) using the R package *MCMCglmm* [@RN4755].
+To model final germination proportion we used binomial MCMCglmms, while for MGT and UNC we used gaussian MCMCglmms.
+Models had, as fixed effects, the experimental conditions (temperature, alternating temperature, ligth, scarification, stratification and GA~3~), plus their interaction with seed mass and embryo:endosperm, and the strict alpine or generalist character of the species.
+Random effects included a reconstructed phylogenetic tree for the `r MSspecies`, species identity, seed lot, lab and alpine region.
+To create the phylogeny, we used the R package *V.PhyloMaker* [@RN4753].
+*V.PhyloMaker* contains a mega-tree which relies on GBOTB for the seed plants [@RN4754], with updates, corrections and expansion.
+We binded taxa absent from the mega-tree to the genus-level basal node.
+In all models, all variables were scaled.
+We used weakly informative priors in all models, with parameter-expanded priors for the random effects.
+Each model was run for 500,000 MCMC steps, with an initial burn-in phase of 50,000 and a thinning interval of 50 [@RN4756], resulting, on average, in 9,000 posterior distributions.
+From the resulting posterior distributions, we calculated mean parameter estimates and 95% Highest Posterior Density (HPD) and Credible Intervals (CI).
+We estimated the significance of model parameters by examining CIs, considering parameters with CIs overlapping with zero as non-significant.
+To estimate the phylogenetic signal of seed germination over all variables we used Pagel’s lambda (λ) [@RN4757], estimated simultaneously with the models by calculating the mean of the posterior distribution and the 95% CI of λ as indicated by [-@RN4756].
+When λ = 0, related taxa are no more similar than expected by chance, while when λ = 1, the trait is evolving following a constant variance random walk or Brownian motion model; intermediate values of λ indicate a phylogenetic correlation in traits evolution that does not fully follow a Brownian motion model [@RN4757].
+
+## FAMD ordination
+
+To visualize the alpine seed germination spectrum, we performed a Factorial Analysis of Mixed Data (FAMD) as implemented in the package *FactoMineR* [@RN3166].
+Roughly, FAMD ordination is a mix between PCA and MCA that allows to include continuous and categorical variables.
+We calculated the ordination at the species level, i.e.
+for each species we calculated a series of continuous and categorical traits.
+We created a continuous variable for each germination cue (i.e.
+temperature, alternating temperatures, light, scarification, stratification and GA~3~).
+To do so, for each cue and species, we calculated a weighted average of the cue levels (cue levels were the temperature treatments in the case of temperature; for the other cues the levels were 0: absence and 1: presence), weighting by the germination propotion at each level.
+This approach underrepresents the importance of the levels that were not tested for a given species, but can serve as a proxy of germination cues when visualized across the whole dataset; it must be stressed that this stage of the analysis serves only for visualization and not hypothesis testing.
+For each species, we also calculated a single MGT (the min, i.e.
+the time taken at the most favourable treatment) and UNC (the mean).
+We conducted the FAMD only with the `r MSfamdspp` species for which we had MGT and UNC values.
+We also included the continuous variables seed mass and embryo:endosperm, as well as the categorical variables dormancy class, distribution (alpine vs.
+generalist), life form and life span.
+
+# Results
+
+## Dormancy classes
+
+Most of the species had some type of seed dormancy (**Fig.
+\@ref(fig:fig2)**a), the most common one being physiological (PD).
+A minority of species were non-dormant (ND), and ND was slightly more frequent in generalist species.
+The most noticeable difference between generalist and strict species was in the morphological (MD) and morphophysiological (MPD) classes, which were more frequent in the strict alpine species.
+Physical dormancy (PY) was rare in both strict alpines and generalists.
+
+## Seed mass and embryo:endosperm
+
+Both generalist and strict alpine species had a similar range of values for seed mass (**Fig.
+\@ref(fig:fig2)**b) and embryo to endosperm ratio (**Fig.
+\@ref(fig:fig2)**c).
+The values of seed mass ranged from `r MSminMass` to `r MSmaxMass` mg, with a median of `r MSmedianMass`.
+Embryo:endosperm encompassed the full range of potential values, from endospermic species with very small embryos (`r MSminEmbryo`) to non-endospermic seeds (`r MSmaxEmbryo`).
+The median value was `r MSmedianEmbryo`.
+
+## Germination temperature
+
+Temperature had a positive effect on the final germination proportion (**Fig.
+\@ref(fig:fig3)**a), indicating a preference for warmer temperatures.
+Increasing temperatures also decreased germination time (**Fig.
+\@ref(fig:fig3)**b) and uncertainty (**Fig.
+\@ref(fig:fig3)**c).
+The effect of warmer temperatures was more intense in strict alpine than in generalist species.
+Seed mass and embryo:endosperm reduced the response to temperature: heavier and less endospermic seeds increased less their germination proportion and decreased less their germination time in warmer temperatures, although uncertainty did not show an interaction with mass or embryo:endosperm.
+
+## Alternating temperatures
+
+Alternating temperatures had a positive effect on the germination proportion (**Fig.
+\@ref(fig:fig3)**a), but not in germination time (**Fig.
+\@ref(fig:fig3)**b) or uncertainty (**Fig.
+\@ref(fig:fig3)**c).
+The posive effect on the germination proportion was slightly higher in the generalist species.
+Again, higher seed mass and embryo:endosperm reduced the positive effect of alternating temperatures.
+
+## Light
+
+Light enhanced the germination proportion (**Fig.
+\@ref(fig:fig3)**a), with no differences between strict alpine and generalist species.
+It did not affect the germination time (**Fig.
+\@ref(fig:fig3)**b), but it did increase the germination uncertainty of generalist species (**Fig.
+\@ref(fig:fig3)**c).
+Heavier and less endospermic seeds responded less to light.
+
+## Scarification
+
+Scarification promoted germination (**Fig.
+\@ref(fig:fig3)**a) and reduced germination time (**Fig.
+\@ref(fig:fig3)**b), but it did not affect germination uncertainty (**Fig.
+\@ref(fig:fig3)**c).
+Strict alpine species responded slightly more to scarification.
+Heavier seeds reponded more to stratification, while seeds with smaller embryos responded less.
+
+## Stratification and GA~3~
+
+Stratification increased the germination proportion (**Fig.
+\@ref(fig:fig3)**a), had a small negative effect on germination time (**Fig.
+\@ref(fig:fig3)**b) and reduced germination uncertainty (**Fig.
+\@ref(fig:fig3)**c).
+The positive effect on the germination proportion was higher on the strict alpine species.
+Seed mass did not interact with stratification, but species with smaller embryos had a smaller response to stratification in their germination proportion, and reduced less their germination time and uncertainty.
+GA~3~ had similar effects to stratification, with the main difference being than seed mass did reduce the response to GA~3~.
+
+## Random effects and phylogenetic signal
+
+Random factors had an effect on the germination response to the environment (**Fig.
+\@ref(fig:fig4)**).
+The final germination proportion and the mean germination time were affected more strongly, on average, by the seed lot and the species identity.
+Alpine region and lab had a lower average effect, but more variability.
+Germination uncertainty showed the opposite pattern, with a stronger effect of lab and alpine region, which might be representing different germination scoring schedules among research groups.
+The phylogenetic signal in the response to all germination parameters (**Fig.
+\@ref(fig:fig5)**) was positive for the final germination proportion and the mean germination time.
+For the germination uncertainty it was closer to zero, indicating a weak effect of phylogeny, which perhaps is also caused by a stronger effect of different scoring schedules.
+
+## FAMD ordination
+
+The first FADM axis explained `r round(pcaAlpine$eig[1, 2], 0)`% of the variation.
+The quantitative variables with the largest contribution to this first axis were germination temperature, germination uncertainty, GA~3~ and alternating temperatures.
+Among the qualitative factors, strict, generalist, physiological dormancy and annual contributed to axis 1.
+In summary, the axis separated (i, left) generalist and annual species with physiological dormancy, a positive response to alternating temperatures and more germination uncertainty from (ii, right) strict alpine species with a preferences for warmer temperatures and a positive response to GA~3~ (**Fig.
+\@ref(fig:fig6)**).
+Axis 2 explained `r round(pcaAlpine$eig[2, 2], 0)`% of the variability.
+The main contributing variables were embryo:endosperm, mean germination time, seed mass and scarification; and the main categorical factors were physical dormancy, graminoid life form and morphophysiological dormancy.
+This axis separated (iii, bottom) morphophysiologically dormant and graminoid species with longer mean time to germination from physically dormant species that responded to scarification and had heavier, less endospermic seeds.
 
@@
 
+```{r fig2, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Seed dormancy classes, seed mass and embryo to endosperm ratio in generalist and strict alpine species.
+Seed dormancy classes are abbreviated as follows: MD (morphological), MPD (morphophysiological), ND (non dormant), PD (physiological), PY (physical).", fig.align = "center", fig.width=7, fig.height=3}
+cowplot::ggdraw(fig2)
+```
+
+```{r fig3, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the germination environment on germination, according to the MCMC meta-analysis of the primary data.", fig.align = "center", fig.width=7, fig.height=5.5}
+cowplot::ggdraw(fig3)
+```
+
+```{r fig4, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the random factors on germination, according to the MCMC meta-analysis of the primary data.", fig.align = "center", fig.width=7, fig.height=5}
+cowplot::ggdraw(fig4)
+```
+
+```{r fig5, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Phylogenetic signal in the response to the germination environment.", fig.align = "center", fig.width=7, fig.height=1.5}
+cowplot::ggdraw(fig5)
+```
+
+```{r fig6, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "FADM ordination of the seed variables and the species distribution, life form and life span.", fig.align = "center", fig.width=7, fig.height=7}
+cowplot::ggdraw(fig6)
+```
+
 # References

### 2020-06-30 doc/ordination report/ordination-report.Rmd
1st draft sent

---
title: "Ordinations PCA"
author: "Eduardo Fernández Pascual"
date: "19/6/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(tidyverse); library(here)

read.csv(here("data", "Elevations", "Regions.csv")) %>%
  merge(read.csv(here("results", "database", "germination.csv")), by = "Region", all = TRUE) %>%
  merge(read.csv(here("results", "database", "traits.csv")), by = "TPLName") %>%
  mutate(Region = as.character(Region),
         Region = ifelse(is.na(Macroregion), "Enscobase", Region),
         Region = ifelse(Region %in% c("Northern Alps", "Southern Alps"), "Alps", Region),
         Region = ifelse(Region %in% c("Central Andes", "Southern Andes"), "Andes", Region),
         Region = ifelse(Region %in% c("Qinghai-Tibet Plateau"), "Qinghai-Tibet", Region),
         Region = ifelse(Region %in% c("Cantabrian Mountains"), "Cantabrian Mts", Region),     
         Region = as.factor(Region)) -> dataset

# Numbers for the main text

dataset %>% summary
dataset %>% pull(Region) %>% unique %>% length -> MSregions
dataset %>% tally() -> MSrecords
dataset %>% pull(Source) %>% unique %>% length -> MSgroups
read.csv("../../../#tpl/results/TPLNames.csv") %>% 
  select(Taxon, Family) %>%
  unique %>%
  merge(dataset, by = "Taxon", all.y = TRUE) %>%
  pull(Family) %>% unique %>% length -> MSfamilies
dataset %>% pull(TPLName) %>% unique %>% length -> MSspecies
dataset %>% filter(Alpine == "Strict") %>% pull(TPLName) %>% unique %>% length -> MSstrict
dataset %>% filter(Alpine == "Generalist") %>% pull(TPLName) %>% unique %>% length -> MSgeneral
dataset$Germinable %>% sum -> MSseeds
dataset %>% pull(Tmean) %>% min -> MSminT
dataset %>% pull(Tmean) %>% max -> MSmaxT
dataset %>% filter(Alternating == "Y") %>% tally -> MSaltY
dataset %>% filter(Alternating == "N") %>% tally -> MSaltN
dataset %>% filter(Light == "Y") %>% tally -> MSlightY
dataset %>% filter(Light == "N") %>% tally -> MSlightN
dataset %>% filter(Stratification == "None") %>% tally -> MSstratN
dataset %>% filter(Stratification == "Cold") %>% tally -> MSstratCold
dataset %>% filter(Scarification == "Y") %>% tally -> MSscarY
dataset %>% filter(GA3 == "Y") %>% tally -> MSga3

# Map figure

dataset %>%
  group_by(Region) %>%
  summarise(Longitude = mean(Longitude.x), Latitude = mean(Latitude.x),
            Records =length(TPLName),
            Species = length(unique(TPLName))) -> dfRecords

dataset %>%
  filter(Alpine == "Strict") %>%
  group_by(Region) %>%
  summarise(Strict = length(unique(TPLName))) -> dfStricts

dfRecords %>%
  merge(dfStricts) %>%
  data.frame -> dfNumbers

library(maps); library(mapdata); library(extrafont)

load("../../../#wwfmap/alpinemap/results/alpinemap.RData")

dfNumbers %>%
  ggplot(aes(Longitude, Latitude)) +
  geom_polygon(data = map_data("world"), aes(x = long, y = lat, group = group),
               color = "grey96", fill = "grey96") +
  geom_polygon(data = testolin_simple, aes(x = long, y = lat, group = group),
               color = "purple", fill = "purple") +
  geom_label(aes(label = rownames(dfNumbers)), size = 2.5,
             label.padding = unit(0.1, "lines"),
             label.r = unit(0.25, "lines"),
             alpha = 0.6, fill = "gold") +
  guides(size = F) +
  coord_cartesian(xlim = c(-180, 180), ylim = c(-49, 80)) +
  ggthemes::theme_tufte() +
  theme(axis.line=element_blank(), axis.text.x=element_blank(),
        axis.text.y=element_blank(), axis.ticks=element_blank(),
        axis.title.x=element_blank(), axis.title.y=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(), panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(), plot.background=element_blank(),
        legend.title = element_blank(),
        legend.position = "none",
        legend.text = element_text(margin = margin(r = 11, unit = "pt"))) -> p1

dfNumbers %>%
  select(Region, Records, Species, Strict) %>%
  gridExtra::tableGrob(theme = gridExtra::ttheme_minimal(
    core = list(fg_params = list(cex = .6, hjust = 1, x = 0.9, fontfamily = "serif")),
    colhead = list(fg_params = list(cex = .6, hjust = 1, x = 0.9, fontfamily = "serif")),
    rowhead = list(fg_params = list(cex = .6), fontfamily = "serif"))) -> TableMap

gridExtra::arrangeGrob(p1, TableMap, ncol = 2, widths = 3:2) -> fig1

# Figure 2: description of seed dormancy, mass, embryo

dataset %>%
  select(TPLName, Alpine, Dormancy) %>%
  na.omit %>%
  unique %>%
  group_by(Alpine, Dormancy) %>%
  tally %>%
  ggplot(aes(Alpine, n, fill = Dormancy)) +
  geom_col(position = "fill") +
  labs(y = "Frequency", title = "(A) Seed dormancy classes") +
  ggthemes::theme_tufte() +
  theme(legend.position = "right", axis.title.x = element_blank(),
        panel.background = element_rect(color = "grey96", fill = "grey96"),
        axis.text.x = element_text(size = 7)) +
  scale_fill_manual(values = c("olivedrab", "yellowgreen", "turquoise4", "gold", "purple")) -> fig2a

dataset %>%
  select(TPLName, Alpine, Seed.mass) %>%
  na.omit %>%
  unique %>%
  ggplot(aes(Alpine, log(Seed.mass), fill = Alpine)) +
  geom_boxplot() +
  labs(y = "Seed mass (log)", title = "(B) Seed mass") +
  ggthemes::theme_tufte() +
  theme(legend.position = "right", axis.title.x = element_blank(),
        panel.background = element_rect(color = "grey96", fill = "grey96"),
        axis.text.x = element_text(size = 7)) +
  scale_fill_manual(values = c("yellowgreen", "gold")) -> fig2b

dataset %>%
  select(TPLName, Alpine, Embryo) %>%
  na.omit %>%
  unique %>%
  ggplot(aes(Alpine, Embryo, fill = Alpine)) +
  geom_boxplot() +
  labs(y = "Embryo:endosperm", title = "(C) Embryo to endosperm ratio") +
  ggthemes::theme_tufte() +
  theme(legend.position = "right", axis.title.x = element_blank(),
        panel.background = element_rect(color = "grey96", fill = "grey96"),
        axis.text.x = element_text(size = 7)) +
  scale_fill_manual(values = c("yellowgreen", "gold")) -> fig2c

gridExtra::arrangeGrob(fig2a, fig2b, fig2c, ncol = 1) -> fig2

# Figure 3 - MCMC

rbind(
  rbind(
    read.csv(here("results", "analysis", "proportion", "alpR.germ.general.csv")) %>%
      select(model, post.mean, l.95..CI, u.95..CI, pMCMC) %>%
      mutate(Moderator = "Alone"),
    read.csv(here("results", "analysis", "proportion", "alpR.germ.general.embryo.csv")) %>%
      select(model, post.mean, l.95..CI, u.95..CI, pMCMC) %>% 
      mutate(Moderator = "Embryo"),
    read.csv(here("results", "analysis", "proportion", "alpR.germ.general.mass.csv")) %>%
      select(model, post.mean, l.95..CI, u.95..CI, pMCMC) %>% 
      mutate(Moderator = "Mass")) %>%
    rename(Mean = post.mean,
           Lower = l.95..CI,
           Upper = u.95..CI,
           p = pMCMC,
           Treatment = model) %>%
    select(Moderator, Treatment, Mean, Lower, Upper, p),
  read.csv(here("results", "analysis", "proportion", "alpR.germ.type.csv")) %>%
    select(variable, post.mean, l.95..CI, u.95..CI, pMCMC) %>%
    separate(variable, into = c("Moderator", "Treatment"), sep = ":") %>%
    mutate(Moderator = gsub("Alpine", "", Moderator)) %>%
    rename(Mean = post.mean,
           Lower = l.95..CI,
           Upper = u.95..CI,
           p = pMCMC)) %>%
  mutate(Moderator = recode(Moderator, Strict = "Strict alpine spp.",
                            Generalist = "Generalist spp.",
                            Mass = "Seed mass",
                            Embryo = "Embryo:endosperm",
                            Alone = "Main effect"),
         Moderator = fct_relevel(Moderator, c("Embryo:endosperm", "Seed mass",  "Generalist spp.", "Strict alpine spp.", "Main effect")),
         Treatment = recode(Treatment, Tmean = "Temperature", GA3 = "GA[3]"),
         Treatment = fct_relevel(Treatment, c("Temperature", "Alternating", "Light", "Scarification", "Stratification", "GA[3]"))) %>%
  ggplot(aes(y = Moderator, x = Mean, 
             xmin = Lower, xmax = Upper,
             color = Moderator)) +
  facet_wrap(~ Treatment, labeller = label_parsed, nrow = 1) +
  geom_point(size = 1, alpha = 0.7) +
  labs(x = "Effect size", title = "(A) Effect on final germination proportions") +
  geom_errorbarh(height = .3) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  ggthemes::theme_tufte() +
  theme(legend.position = "none", axis.title.y = element_blank(),
        panel.background = element_rect(color = "grey96", fill = "grey96"),
        axis.text.y = element_text(color = c("olivedrab",  "turquoise4", "yellowgreen", "gold", "purple")),
        axis.text.x = element_text(size = 7)) +
  scale_color_manual(values = c("olivedrab",  "turquoise4", "yellowgreen", "gold", "purple")) -> fig1a

rbind(
  rbind(
    read.csv(here("results", "analysis", "mgt", "alpR.MGT.general.csv")) %>%
      select(model, post.mean, l.95..CI, u.95..CI, pMCMC) %>%
      mutate(Moderator = "Alone"),
    read.csv(here("results", "analysis", "mgt", "alpR.MGT.general.embryo.csv")) %>%
      select(model, post.mean, l.95..CI, u.95..CI, pMCMC) %>% 
      mutate(Moderator = "Embryo"),
    read.csv(here("results", "analysis", "mgt", "alpR.MGT.general.mass.csv")) %>%
      select(model, post.mean, l.95..CI, u.95..CI, pMCMC) %>% 
      mutate(Moderator = "Mass")) %>%
    rename(Mean = post.mean,
           Lower = l.95..CI,
           Upper = u.95..CI,
           p = pMCMC,
           Treatment = model) %>%
    select(Moderator, Treatment, Mean, Lower, Upper, p),
  read.csv(here("results", "analysis", "mgt", "alpR.MGT.type.csv")) %>%
    select(variable, post.mean, l.95..CI, u.95..CI, pMCMC) %>%
    separate(variable, into = c("Moderator", "Treatment"), sep = ":") %>%
    mutate(Moderator = gsub("Alpine", "", Moderator)) %>%
    rename(Mean = post.mean,
           Lower = l.95..CI,
           Upper = u.95..CI,
           p = pMCMC)) %>%
  filter(Mean < 0.5) %>%
  mutate(Moderator = recode(Moderator, Strict = "Strict alpine spp.",
                            Generalist = "Generalist spp.",
                            Mass = "Seed mass",
                            Embryo = "Embryo:endosperm",
                            Alone = "Main effect"),
         Moderator = fct_relevel(Moderator, c("Embryo:endosperm", "Seed mass",  "Generalist spp.", "Strict alpine spp.", "Main effect")),
         Treatment = recode(Treatment, Tmean = "Temperature", GA3 = "GA[3]"),
         Treatment = fct_relevel(Treatment, c("Temperature", "Alternating", "Light", "Scarification", "Stratification", "GA[3]"))) %>%
  ggplot(aes(y = Moderator, x = Mean, 
             xmin = Lower, xmax = Upper,
             color = Moderator)) +
  facet_wrap(~ Treatment, labeller = label_parsed, nrow = 1) +
  geom_point(size = 1, alpha = 0.7) +
  labs(x = "Effect size", title = "(B) Effect on mean germination time") +
  geom_errorbarh(height = .3) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  ggthemes::theme_tufte() +
  theme(legend.position = "none", axis.title.y = element_blank(),
        panel.background = element_rect(color = "grey96", fill = "grey96"),
        axis.text.y = element_text(color = c("olivedrab",  "turquoise4", "yellowgreen", "gold", "purple")),
        axis.text.x = element_text(size = 7)) +
  scale_color_manual(values = c("olivedrab",  "turquoise4", "yellowgreen", "gold", "purple")) -> fig1b

rbind(
  rbind(
    read.csv(here("results", "analysis", "unc", "alpR.UNC.general.csv")) %>%
      select(model, post.mean, l.95..CI, u.95..CI, pMCMC) %>%
      mutate(Moderator = "Alone"),
    read.csv(here("results", "analysis", "unc", "alpR.UNC.general.embryo.csv")) %>%
      select(model, post.mean, l.95..CI, u.95..CI, pMCMC) %>% 
      mutate(Moderator = "Embryo"),
    read.csv(here("results", "analysis", "unc", "alpR.UNC.general.mass.csv")) %>%
      select(model, post.mean, l.95..CI, u.95..CI, pMCMC) %>% 
      mutate(Moderator = "Mass")) %>%
    rename(Mean = post.mean,
           Lower = l.95..CI,
           Upper = u.95..CI,
           p = pMCMC,
           Treatment = model) %>%
    select(Moderator, Treatment, Mean, Lower, Upper, p),
  read.csv(here("results", "analysis", "unc", "alpR.UNC.type.csv")) %>%
    select(variable, post.mean, l.95..CI, u.95..CI, pMCMC) %>%
    separate(variable, into = c("Moderator", "Treatment"), sep = ":") %>%
    mutate(Moderator = gsub("Alpine", "", Moderator)) %>%
    rename(Mean = post.mean,
           Lower = l.95..CI,
           Upper = u.95..CI,
           p = pMCMC)) %>%
  filter(Mean < 0.5) %>%
  mutate(Moderator = recode(Moderator, Strict = "Strict alpine spp.",
                            Generalist = "Generalist spp.",
                            Mass = "Seed mass",
                            Embryo = "Embryo:endosperm",
                            Alone = "Main effect"),
         Moderator = fct_relevel(Moderator, c("Embryo:endosperm", "Seed mass",  "Generalist spp.", "Strict alpine spp.", "Main effect")),
         Treatment = recode(Treatment, Tmean = "Temperature", GA3 = "GA[3]"),
         Treatment = fct_relevel(Treatment, c("Temperature", "Alternating", "Light", "Scarification", "Stratification", "GA[3]"))) %>%
  ggplot(aes(y = Moderator, x = Mean, 
             xmin = Lower, xmax = Upper,
             color = Moderator)) +
  facet_wrap(~ Treatment, labeller = label_parsed, nrow = 1) +
  geom_point(size = 1, alpha = 0.7) +
  labs(x = "Effect size", title = "(C) Effect on germination uncertainty") +
  geom_errorbarh(height = .3) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  ggthemes::theme_tufte() +
  theme(legend.position = "none", axis.title.y = element_blank(),
        panel.background = element_rect(color = "grey96", fill = "grey96"),
        axis.text.y = element_text(color = c("olivedrab",  "turquoise4", "yellowgreen", "gold", "purple")),
        axis.text.x = element_text(size = 7)) +
  scale_color_manual(values = c("olivedrab",  "turquoise4", "yellowgreen", "gold", "purple")) -> fig1c

cowplot::plot_grid(fig1a, fig1b, fig1c, ncol = 1) -> fig3

# Figure 4 - Ordination

dataset %>%
  mutate(Macroregion = as.character(Macroregion),
         Macroregion = ifelse(is.na(Macroregion), "Europe", Macroregion),
         Germination = Germinated / Germinable,
         Scarification = ifelse(Scarification == "Y", 1, 0),
         Stratification = ifelse(Stratification == "None", 0, 1),
         GA3 = ifelse(GA3 == "Y", 1, 0),
         Light = ifelse(Light == "Y", 1, 0),
         Alternating = ifelse(Alternating == "Y", 1, 0)) %>%
  group_by(TPLName, Macroregion) %>%
  summarise(Temperature = weighted.mean(Tmean, w = Germination),
            Alternating = weighted.mean(Alternating, w = Germination),
            Light = weighted.mean(Light, w = Germination),
            Scarification = weighted.mean(Scarification, w = Germination),
            Stratification = weighted.mean(Stratification, w = Germination),
            GA3 = weighted.mean(GA3, w = Germination)) %>%
  merge(read.csv(here("results", "database", "traits.csv")), by = "TPLName") %>%
  select(TPLName, Macroregion, Alpine, Plant.category, Life.form, Lifespan, Reproduction.frequency, Dormancy, Seed.mass, Embryo, Temperature:GA3) %>%
  filter(!is.nan(Temperature)) %>% 
  mutate(Plant.category = fct_recode(Plant.category, Woody = "Herb, Woody", Woody = "Woody, Subshrub?"),
         Life.form = fct_recode(Life.form, T = "T, H", G = "H, G", CH = "H, CH", T = "T, Hydrophyt", CH = "CH, H", CH = "CH, NP", G = "G, H", H = "H, Hydrophyt"),
         Lifespan = fct_recode(Lifespan, annual = "annual, biennial", annual = "annual, perennial", perennial = "biennial", perennial = "biennial, perennial"),
         Reproduction.frequency = fct_recode(Reproduction.frequency, monocarp = "monocarp, ?polycarp", monocarp = "monocarp, polycarp")) -> traits

traits %>%
  select(Seed.mass:GA3) %>%
  FactoMineR::PCA(graph = FALSE) -> pcaAlpine

traits %>%
  select(-c(Seed.mass:GA3)) %>%
  cbind(pcaAlpine$ind$coord[, 1:2]) -> pcaInds

pcaAlpine$var$coord[, 1:2] %>%
  data.frame %>%
  rownames_to_column(var = "Variable") %>%
  mutate(Variable = fct_recode(Variable, `Seed mass` = "Seed.mass", 
                               `Embryo:endosperm` = "Embryo")) -> pcaVars

ggplot(pcaInds, aes(x = Dim.1, y = Dim.2)) +
  geom_text(aes(label = TPLName, color = Macroregion), size = 3, show.legend = T, alpha = 0.6, shape = 15) +
  ggthemes::theme_tufte() + 
  theme(panel.background = element_rect(color = "grey96", fill = "grey96"),
        legend.position = "top", legend.title = element_blank(),
        legend.text = element_text(size = 7.5),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm"),
        plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"),
        text = element_text(size = 8)) +
  scale_color_manual(values = c("olivedrab",  "turquoise4", "gold", "purple")) +
  scale_x_continuous(name = paste("Axis 1 (", round(pcaAlpine$eig[1, 2], 0),
                                  "% variance explained)", sep = "")) + 
  scale_y_continuous(name = paste("Axis 2 (", round(pcaAlpine$eig[2, 2], 0), 
                                  "% variance explained)", sep = "")) +
  geom_segment(data = pcaVars, aes(x = 0, y = 0, xend = 4*Dim.1, yend = 4*Dim.2)) +
  geom_label(data = pcaVars, aes(x = 4*Dim.1, y = 4*Dim.2, label = Variable), 
                  size = 2.5)

ggplot(pcaInds, aes(x = Dim.1, y = Dim.2)) +
  geom_point(aes(color = Alpine), size = 3, show.legend = T, alpha = 0.6, shape = 15) +
  ggthemes::theme_tufte() + 
  theme(panel.background = element_rect(color = "grey96", fill = "grey96"),
        legend.position = "top", legend.title = element_blank(),
        legend.text = element_text(size = 7.5),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm"),
        plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"),
        text = element_text(size = 8)) +
  scale_color_manual(values = c("yellowgreen", "gold")) +
  scale_x_continuous(name = paste("Axis 1 (", round(pcaAlpine$eig[1, 2], 0),
                                  "% variance explained)", sep = "")) + 
  scale_y_continuous(name = paste("Axis 2 (", round(pcaAlpine$eig[2, 2], 0), 
                                  "% variance explained)", sep = "")) +
  geom_segment(data = pcaVars, aes(x = 0, y = 0, xend = 4*Dim.1, yend = 4*Dim.2)) +
  geom_label(data = pcaVars, aes(x = 4*Dim.1, y = 4*Dim.2, label = Variable), 
             size = 2.5)

ggplot(pcaInds, aes(x = Dim.1, y = Dim.2)) +
  geom_point(aes(color = Life.form), size = 3, show.legend = T, alpha = 0.6, shape = 15) +
  ggthemes::theme_tufte() + 
  theme(panel.background = element_rect(color = "grey96", fill = "grey96"),
        legend.position = "top", legend.title = element_blank(),
        legend.text = element_text(size = 7.5),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm"),
        plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"),
        text = element_text(size = 8)) +
  scale_color_manual(values = c("olivedrab",  "turquoise4", "gold", "purple", "yellowgreen", "indianred")) +
  scale_x_continuous(name = paste("Axis 1 (", round(pcaAlpine$eig[1, 2], 0),
                                  "% variance explained)", sep = "")) + 
  scale_y_continuous(name = paste("Axis 2 (", round(pcaAlpine$eig[2, 2], 0), 
                                  "% variance explained)", sep = "")) +
  geom_segment(data = pcaVars, aes(x = 0, y = 0, xend = 4*Dim.1, yend = 4*Dim.2)) +
  geom_label(data = pcaVars, aes(x = 4*Dim.1, y = 4*Dim.2, label = Variable), 
             size = 2.5)

ggplot(pcaInds, aes(x = Dim.1, y = Dim.2)) +
  geom_point(aes(color = Lifespan), size = 3, show.legend = T, alpha = 0.6, shape = 15) +
  ggthemes::theme_tufte() + 
  theme(panel.background = element_rect(color = "grey96", fill = "grey96"),
        legend.position = "top", legend.title = element_blank(),
        legend.text = element_text(size = 7.5),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm"),
        plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"),
        text = element_text(size = 8)) +
  scale_color_manual(values = c("olivedrab",  "turquoise4", "gold", "purple", "yellowgreen", "indianred")) +
  scale_x_continuous(name = paste("Axis 1 (", round(pcaAlpine$eig[1, 2], 0),
                                  "% variance explained)", sep = "")) + 
  scale_y_continuous(name = paste("Axis 2 (", round(pcaAlpine$eig[2, 2], 0), 
                                  "% variance explained)", sep = "")) +
  geom_segment(data = pcaVars, aes(x = 0, y = 0, xend = 4*Dim.1, yend = 4*Dim.2)) +
  geom_label(data = pcaVars, aes(x = 4*Dim.1, y = 4*Dim.2, label = Variable), 
             size = 2.5)

ggplot(pcaInds, aes(x = Dim.1, y = Dim.2)) +
  geom_point(aes(color = Plant.category), size = 3, show.legend = T, alpha = 0.6, shape = 15) +
  ggthemes::theme_tufte() + 
  theme(panel.background = element_rect(color = "grey96", fill = "grey96"),
        legend.position = "top", legend.title = element_blank(),
        legend.text = element_text(size = 7.5),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm"),
        plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"),
        text = element_text(size = 8)) +
  scale_color_manual(values = c("olivedrab",  "turquoise4", "gold", "purple", "yellowgreen", "indianred")) +
  scale_x_continuous(name = paste("Axis 1 (", round(pcaAlpine$eig[1, 2], 0),
                                  "% variance explained)", sep = "")) + 
  scale_y_continuous(name = paste("Axis 2 (", round(pcaAlpine$eig[2, 2], 0), 
                                  "% variance explained)", sep = "")) +
  geom_segment(data = pcaVars, aes(x = 0, y = 0, xend = 4*Dim.1, yend = 4*Dim.2)) +
  geom_label(data = pcaVars, aes(x = 4*Dim.1, y = 4*Dim.2, label = Variable), 
             size = 2.5)

ggplot(pcaInds, aes(x = Dim.1, y = Dim.2)) +
  geom_point(aes(color = Dormancy), size = 3, show.legend = T, alpha = 0.6, shape = 15) +
  ggthemes::theme_tufte() + 
  theme(panel.background = element_rect(color = "grey96", fill = "grey96"),
        legend.position = "top", legend.title = element_blank(),
        legend.text = element_text(size = 7.5),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm"),
        plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"),
        text = element_text(size = 8)) +
  scale_color_manual(values = c("olivedrab",  "turquoise4", "gold", "purple", "yellowgreen", "indianred")) +
  scale_x_continuous(name = paste("Axis 1 (", round(pcaAlpine$eig[1, 2], 0),
                                  "% variance explained)", sep = "")) + 
  scale_y_continuous(name = paste("Axis 2 (", round(pcaAlpine$eig[2, 2], 0), 
                                  "% variance explained)", sep = "")) +
  geom_segment(data = pcaVars, aes(x = 0, y = 0, xend = 4*Dim.1, yend = 4*Dim.2)) +
  geom_label(data = pcaVars, aes(x = 4*Dim.1, y = 4*Dim.2, label = Variable), 
             size = 2.5)

traits %>%
  select(Plant.category, Lifespan, Reproduction.frequency, Dormancy, Seed.mass:GA3) %>%
  FactoMineR::FAMD(graph = FALSE) -> pcaAlpine

traits %>%
  select(-c(Plant.category, Lifespan, Reproduction.frequency, Dormancy, Seed.mass:GA3)) %>%
  cbind(pcaAlpine$ind$coord[, 1:2]) -> pcaInds

cent <- aggregate(cbind(Dim.1, Dim.2) ~ Alpine, data = pcaInds, FUN = mean)
segs <- merge(pcaInds, setNames(cent, c("Alpine", "oDim.1", "oDim.2")), by = "Alpine", sort = FALSE)

pcaAlpine$quanti.var$coord[, 1:2] %>%
  data.frame %>%
  rownames_to_column(var = "Variable") %>%
  mutate(Variable = fct_recode(Variable, `Seed mass` = "Seed.mass", 
                               `Embryo:endosperm` = "Embryo")) -> pcaVars

pcaAlpine$quali.var$coord[, 1:2] %>%
  data.frame %>%
  rownames_to_column(var = "Variable") %>%
  mutate(Variable = fct_recode(Variable, Annual = "annual", Perennial = "perennial",
                               Monocarpic =  "monocarp", Polycarpic = "polycarp")) -> pcaVars2

ggplot(pcaInds, aes(x = Dim.1, y = Dim.2)) +
  geom_point(aes(color = Alpine), alpha = 0.5, size = 2.5, shape = 15) +
  ggthemes::theme_tufte() + 
  theme(panel.background = element_rect(color = "grey96", fill = "grey96"),
        legend.position = "top", legend.title = element_blank(),
        legend.text = element_text(size = 7.5),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm"),
        plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"),
        text = element_text(size = 8)) +
  scale_color_manual(values = c("yellowgreen", "gold")) +
  scale_x_continuous(name = paste("Axis 1 (", round(pcaAlpine$eig[1, 2], 0),
                                  "% variance explained)", sep = "")) + 
  scale_y_continuous(name = paste("Axis 2 (", round(pcaAlpine$eig[2, 2], 0), 
                                  "% variance explained)", sep = "")) +
  geom_segment(data = pcaVars, aes(x = 0, y = 0, xend = 4*Dim.1, yend = 4*Dim.2)) +
  geom_label(data = pcaVars, aes(x = 4*Dim.1, y = 4*Dim.2, label = Variable), 
                  size = 2.5) +
  geom_label(data = pcaVars2, aes(x = 4*Dim.1, y = 4*Dim.2, label = Variable), 
             size = 2.5)

```

### 2020-08-09 doc/ms/ms.Rmd
unified ms

--- 2020-06-18 doc/ms/ms.Rmd
@@
 ---
-title: "Alpine Seeds MS"
-date: "15/6/2020"
+title: "Seed germination ecology of alpine plants: a global meta-analysis of primary data"
 output: 
-  bookdown::html_document2:
-    reference_docx: gebtemplate.docx
+  bookdown::word_document2:
+    reference_docx: template.docx
 bibliography: refs.bib
-csl: global-ecology-and-biogeography.csl
+csl: new-phytologist.csl
 link-citations: yes
@@
 
+Eduardo Fernández-Pascual^1^ ^+^, Elías Afif Khouri^1^, Joaquín Ascaso^2^, Beatriz Morais^1^, Ramón Reiné^2^, Madalena Vaz^3^
+
+^1^ Departamento de Biología de Organismos y Sistemas, Universidad de Oviedo, Oviedo/Uviéu, Spain; ^2^ Departamento de Ciencias Agrarias y Medio Natural, Universidad de Zaragoza, Huesca, Spain; ^3^ Banco Português de Germoplasma Vegetal, Instituto Nacional de Investigação Agrária e Veterinária (INIAV), Braga, Portugal
+
+^+^ Correspondence: Departamento de Biología de Organismos y Sistemas, Universidad de Oviedo, C/ Catedrático Rodrigo Uría, 33006 Oviedo/Uviéu, Spain.
+Email: efernandezpascual@gmail.com.
+Telephone: +34985104787.
+
+# Summary (200 words)
+
+# Keywords (8)
+
 # Introduction
 
-Alpine environments are those that occur above the maximum elevation at which trees can grow, and are characterized by low temperatures, strong winds, unstable substrates, and short growing seasons [@RN2392].
-The short alpine growing season is a major barrier for plant recruitment, as it constrains the period favorable for seedling establishment [@RN4714; @RN4717].
-Alpine seedlings must survive extreme environmental stresses and yet grow to a critical biomass by the end of the growing season, to withstand the harsh and long-lasting winter [@RN2868].
-The physiological process of seed germination is the essential, yet most vulnerable, start of seedling establishment [@RN3397].
-Because it is an irreversible process, germination must be timed to occur when the environment is favourable for subsequent seedling survival and growth [@RN4691].
-Thus, the mechanisms regulating the timing of the seed-to-seedling transition are expected to be under strong selective pressure [@RN3375; @RN2968].
-Natural selection should favour seed germination requirements that reduce the probability of facing environmental conditions that are not appropriate for seedlings.
-The main physiological drivers of germination, water availability and temperature [@RN3368], are also two main features of climate.
-In addition, many plant species have developed seed dormancy, by which germination is prevented during periods that are only ephemerally favourable, ensuring germination in the right season [@RN3063].
-Different degrees of dormancy also ensure the distribution of offspring emergence across time, as bet-hedging against unpredictable environments [@RN3065].
-Other germination cues allow for a detection of fine-scale germination niches and safe sites [@RN4719], such as the response to light [@RN4656] and alternating temperatures [@RN1380].
-In the alpine environment, the diversity of microhabitats has apparently resulted in variable germination ecologies across species, making it difficult to define a common *alpine germination syndrome* [@RN2392; @RN2943].
+Alpine environments occur worldwide above the maximum elevation at which trees can grow [@RN4750].
+These treeless habitats are characterized by low temperatures, strong winds, unstable substrates, and short growing seasons [@RN2392].
+Although some alpine plants may reproduce clonally, sexual reproduction is the main strategy to maintain genetic diversity and to colonize suitable habitats in response to environmental changes.
+The short reproductive season of alpine habitats constrains the phenological timing of flowering, mating, seed development and seed dispersal.
+Further, dispersed seeds will need to germinate in the most appropriate period to ensure survival of seedlings under extreme environmental stress [@RN4714; @RN2868; @RN4717].
+The physiological process of seed germination is therefore an essential life stage that must be timed to occur when the environment is favourable for subsequent seedling survival and growth [@RN4691].
+In alpine habitats, germination by seeds will be the final test for plant populations to cope with climate change and to determine whether they will persist or go into extinction.
 
-Early studies on the germination ecology of alpine plants demonstrated that freshly collected seeds of most arctic and alpine species required relatively high temperatures for germination [@RN3258; @RN3213; @RN4712].
-High-temperature germination has been considered to be an adaptation to prevent seed germination at the timing of seed dispersal (autumn), when temperatures are low and there is a high risk of frost [@RN3393].
-Indeed, seed germination of alpine species tends to occur after exposure to winter in early summer [@RN2392; @RN2382]; at this stage the temperature window for germination usually widens towards lower values [@RN1021].
-Nevertheless, there is an increasing number of studies highlighting the germination of freshly collected seeds at cool incubation temperatures or even during cold stratification treatments [@RN2943; @RN3285; @RN2371; @RN4713].
-Low-temperature germination could be an adaptation to germination under snow or at snowmelt, which could presumably allow seedlings to develop a deeper root system before topsoil desiccation in summer [@RN4720], as well as to attain an optimal size for overwintering [@RN4712].
+Natural selection should favour seed germination requirements that reduce the probability of facing environmental conditions that are not appropriate for seedlings [@RN3375].
+The main physiological drivers of germination, water availability and temperature [@RN3368], are also the main ecological signals of climatic variation.
+Many plant species have also developed seed dormancy, by which germination is prevented during periods that are only ephemerally favourable, postponing germination to the right season [@RN3214].
+Different degrees of dormancy also ensure the distribution of offspring emergence across time, a bet-hedge effect against unpredictable environments [@RN3065].
+Other germination cues, such as the response to light [@RN4656] and alternating temperatures [@RN1380], allow for a fine-scale detection of germination micro-niches and safe sites [@RN4719].
+Given the heterogeneity of alpine climates and species lineages adapted to different regions, and the wealth of local studies on alpine seed ecology (reviewed in the following paragraphs), a current challenge is to synthesize, at the global level, the seed germination spectrum of alpine plants [@RN4655].
+Such a global synthesis would allow us to (i) better understand how climatic changes will affect alpine plant diversity; and (ii) manage and eventually counteract regression of restricted alpine species.
 
-Many alpine species have been described as having deep physiological dormancy [@RN2943; @RN3703; @RN3214].
-Dormancy in freshly-dispersed seeds would prevent precocious germination under autumn conditions, when appropriate soil moisture and temperature are not likely to persist for more than a few weeks or days [@RN3330].
-A requirement for cold stratification to break dormancy would allow seeds to sense the snow season, thereby timing germination to a posterior period, more appropriate for seedling survival and growth.
-Thus, it can be expected that a cold stratification period would be a common requirement for seed germination in alpine plant species.
-Nonetheless, in an early review, Amen [-@RN3213] concluded that cold stratification was not a common requirement for the seed germination of several alpine plant species from different mountains in the USA (see also Sayers and Ward [-@RN1484]; Marchand and Roach [-@RN2995]; Kaye [-@RN4721]).
-More recently, Sommerville *et al.* [-@RN3703] and Hoyle *et al.* [-@RN3285] showed that in 19 and 54 Australian alpine plant species, respectively, a cold stratification period significantly increased seed germination only in half of the species tested.
-In contrast, Söyrinki (1938, cited in Körner [-@RN2392]), experimenting with 91 alpine species from the Alps, found that storage at winter temperatures (i.e.
-cold stratification) increased seed germination in the great majority of cases.
-Shimono and Kudo [-@RN707] reported that cold stratification increased seed germination over a range of temperatures in most of the 27 plant species studied in the alpine zone of Japan.
-Cavieres and Sierra-Almeida [-@RN4713] reported that, along an elevational gradient in the central Chilean Andes, cold stratification was an important requirement for seed germination in species from lower elevations, while species from higher sites still did not germinate after cold stratification.
-Therefore, how important cold stratification is for seed germination in alpine plant species remains unclear and requires a global assessment.
+Early studies on the germination ecology of alpine plants demonstrated that, in most species, recently dispersed seeds require relatively high temperatures for germination [@RN3258; @RN3213; @RN4712].
+Warm-cued germination has been considered as an adaptation to prevent seed germination at the time of seed dispersal (autumn) when temperatures are low and there is a high risk of frost [@RN3393].
+Indeed, germination of alpine seeds tends to occur after winter, mainly in early summer [@RN2392; @RN2382].
+In recent times, an increasing number of studies reported that fresh seeds of some alpine plants also germinate at cool incubation temperatures or during cold stratification treatments [@RN2943; @RN3285; @RN2371; @RN4713].
+This cold-cued germination could be an adaptation to germinate under snow or during snowmelt, which could presumably allow seedlings to develop a deeper root system before topsoil desiccation in summer [@RN4720], or to attain an optimal size for overwintering [@RN4712].
+The main question that remains to be answered is the generality of warm-cued germination across alpine micro-climatic niches.
 
-The response of alpine seeds to light and alternating temperatures can provide further information on their capacity to detect fine-scale environmental variation.
-The preference of some species to germinate in light [@RN4730; @RN4729] could favour the creation of a persistent soil seed bank [@RN4718], adventageous in temporally and spatially unpredictable alpine environments [@RN4731; @RN4732].
-Some species seem to prefer dark germination [@RN2943], a behaviour interpreted by some as detecting safe sites in rock crevices [@RN3008].
-Regarding alternating temperatures, a study of 445 species from the Qinhai-Tibet plateau detected that species from high elevation alpine meadows did not show a significant response to diurnal temperature oscillations [@RN3376].
+Many alpine species have also been described as having seeds with deep physiological dormancy [@RN2943; @RN3703; @RN3214].
+Much like warm-cued germination, dormancy in recently dispersed seeds would prevent precocious germination under autumn cold conditions, when appropriate soil moisture and temperature are not likely to persist for more than a few weeks or days [@RN3330].
+A requirement for cold stratification to break dormancy would allow seeds to sense the snow season, thereby postponing germination to a better period for seedling survival and development.
+Thus, it can be expected that a cold stratification period is a common requirement for seed germination in alpine plants.
+Nonetheless, an early review [@RN3213] suggested that cold stratification was not a requirement for the seed germination of several alpine species from different mountains in the USA, an idea that was repeated in subsequent articles [@RN1484; @RN2995; @RN4721].
+More recently, research on Australian alpine plants found that a cold stratification period only increased germination in half of the species tested [@RN3703; @RN3285].
+In contrast, cold stratification increased seed germination over a range of temperatures in most of the species studied in the alpine zone of Japan [@RN707].
+Along an elevational gradient in the central Chilean Andes, cold stratification was an important requirement for seed germination in species from lower elevations, while species from higher sites did not germinate even after cold stratification [@RN4713].
+Therefore, how important cold stratification is for seed germination in alpine plant species remains unclear.
 
-# Materials and methods
+Besides germination temperature and patterns of seed dormancy, the response of alpine seeds to light and alternating temperatures can provide further information on their capacity to detect fine-scale environmental cues.
+The preference of alpine species to germinate in light [@RN4730; @RN4729] could favour the creation of a persistent soil seed bank [@RN4718], which is advantageous in temporally and spatially unpredictable alpine environments [@RN4731; @RN4732].
+Indeed, alpine soil seed banks are more frequent than previously supposed, as it has been consistently reported in the last years [@RN4929; @RN4718].
+Nevertheless, other alpine species seem to prefer dark germination [@RN2943], which may be interpreted as a strategy for detecting safe sites in rock crevices [@RN3008].
+Regarding alternating temperatures, a study of 445 species from the Qinghai-Tibet plateau found that species from high elevation alpine meadows did not show a significant response to diurnal temperature oscillations [@RN3376].
+This is intriguing given that alpine habitats typically have a strong day-night climatic variation during the growing season.
+Although some germination studies report laboratory experiments conducted for constant and alternate temperatures, or light/dark conditions, they are very much case-dependent, therefore it is difficult to evaluate whether these factors show a consistent pattern across global alpine ecosystems.
+
+Here, we collect primary germination data from eight alpine regions and four continents to investigate the germination response of alpine plant species to key environmental factors.
+We used raw data obtained from laboratory experiments to test the response of seed germination to temperature, cold stratification, light and alternating temperatures, and their relationships with dormancy classes, seed mass, and embryo:endosperm ratio obtained from bibliographic data.
+Our central question was to investigate the seed germination spectrum of alpine plants to test the consistency at the global scale of general assumptions about their seed germination ecology: in short, that alpine seeds (1) have physiological dormancy and (2) require cold stratification to break dormancy, (3) need relatively warm temperatures to germinate, (4) response positively to light and (5) do not require alternating temperatures for germination.
+To understand the ecological constraints of our meta-analysis on alpine species pools, we further analyse how the germination responses differ between strict alpine species which occur exclusively above the treeline, and generalist species that also occur at lower elevations.
+
+# Materials and Methods
 
@@
 
-We have conducted a meta-analysis of primary data [@RN4734].
-To gather data, we contacted research groups who have performed experiments on the germination of alpine seeds, asking them to send raw data from laboratory germination experiments.
-The submitted data had to report the studied species, the germination treatments, the number of seeds sown and the number of seeds germinated.
-We received data from the Cantabrian Mountains [@RN2371], the Alps [@RN4729; @RN3029; @RN4727; Rosbakh, unpublished], the Apennines (Mondoni, unpublished), the Caucasus [Rosbakh, unpublished], the Qinhai-Tibet Plateau [@RN4728; @RN4735; @RN3376], the Andes [@RN3393; @RN4713; Briceño, unpublished] and the Australian Alps [@RN4737; @RN4736; @RN3703; Satyanti, unpublished].
-We also received data on European alpine species from *Enscobase* (http://enscobase.maich.gr/index.tml).
-Once received, we standardized the species names using *The Plant List* [@RN2321] with the package *Taxonstand* [@RN4660] in *R* [@RN2315].
+We conducted a meta-analysis of primary data [@RN4734], which consists in pooling together original data from different local studies to synthetize their conclusions at the global level.
+The original data files, as well as R code for data cleaning, analysis and manuscript production are available at https://github.com/efernandezpascual/alpineseeds.
+To gather data, we contacted research groups who have performed experiments on seed germination of alpine plants, asking them to deliver raw data from laboratory germination experiments.
+The submitted data had to report the studied species, the germination treatments, the number of seeds sown, and the number of seeds germinated.
+We collected data from the Cantabrian Mountains [@RN2371], the European Alps [@RN4729; @RN3029; @RN4727; Rosbakh, unpublished], the Apennines (Mondoni, unpublished), the Caucasus [Rosbakh, unpublished], the Qinghai-Tibet Plateau [@RN4728; @RN4735; @RN3376], the Andes [@RN3393; @RN4713; Briceño, unpublished] and the Australian Alps [@RN4737; @RN4736; @RN3703; Satyanti, unpublished].
+We also collected data on European alpine species from *Enscobase*, the database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml).
+
+Once received, we standardized the species names using *The Plant List* [@RN2321] and the package *Taxonstand* [@RN4660] in *R* [@RN2315].
 For each species, we checked its elevation range in local floras [@RN4741; @RN4739; @RN3243; @RN4742; @RN4740], catalogues [@RN4738] and herbaria (*Australian National Herbarium*).
-We used to elevation ranges to (i) remove from the dataset species that do not grow above the treeline; and (ii) classify the rest of the species as *strict alpine* (i.e.
-species that only grow above the subalpine) or *generalist* (i.e.
-species that can grow above the subalpine but also in the lowlands).
-For each species, we also compiled the following traits: life form and life span from the references used for the elevation, plus some extra references [@RN4747; @RN4744; RN4745; @RN4746; @RN4748; @RN4749] and online sources (https://www.infoflora.ch/de/; http://plantnet.rbgsyd.nsw.gov.au/search/simple.htm; http://buscador.floraargentina.edu.ar/; https://patagoniawildflowers.org/); seed dormancy class [@RN3214; @RN4743]; seed mass [@RN4728; @RN4735; @RN3376; @RN2987] and embryo to endosperm ratio (*Vandelook, ???*).
-For the species missing seed mass or embryo values, we used genus or family means.
+We used the elevation ranges (i) to remove from the dataset species that do not grow above the treeline; and (ii) to classify the rest of the species as *strict alpine* (i.e.
+species that only grow above the treeline) or *generalist* (i.e.
+species that can grow above the treeline but also in the lowlands).
+For each species, we also compiled the following traits: life form and life span from the references used for the elevation, plus some extra references [@RN4747; @RN4744; @RN4745; @RN4746; @RN4748; @RN4749] and online resources (https://www.infoflora.ch/de/; http://plantnet.rbgsyd.nsw.gov.au/search/simple.htm; http://buscador.floraargentina.edu.ar/; https://patagoniawildflowers.org/); seed dormancy class [@RN3214; @RN4743]; seed mass [@RN4728; @RN4735; @RN3376; @RN2987] and embryo to endosperm ratio (Vandelook, unpublished).
+For the species missing seed mass or embryo values, we used genus or family averages.
 
@@
 
-The final dataset contains `r MSrecords` records (germination proportions for a given seed lot of a species, recorded in a set of experimental conditions) contributed by `r MSgroups` research groups, representing `r MSfamilies` seed plant families and `r MSspecies` species (`r MSstrict` strict alpine species and `r MSgeneral` generalists).
-Seed lots had been collected in `r MSregions` alpine regions of the world (**Fig.
-\@ref(fig:fig1)**), in both hemispheres.
-The total number of seeds used in the experiments is `r MSseeds`.
-The range of germination temperatures (weighted average of the daily thermoperiod) goes from `r MSminT` to `r MSmaxT` ºC, with `r MSaltY` records of constant temperatures and `r MSaltN` of alternating temperatures.
-Light was used in `r MSlightY` records, darkness in `r MSlightN`.
+The final dataset used for meta-analysis contained `r MSrecords` records (i.e.
+germination proportions for a given seed lot of a species, recorded in a set of experimental conditions) contributed by `r MSgroups` research groups, representing `r MSfamilies` seed plant families and `r MSspecies` species (`r MSstrict` strict alpine species and `r MSgeneral` generalists).
+Seed lots were originally sampled in `r MSregions` alpine regions of the world (Fig.
+\@ref(fig:fig1)), in both hemispheres.
+
+Most species in the dataset (> 90 %) were perennial, and either graminoids or forbs.
+Most were reported to have dormant seeds (Fig.
+\@ref(fig:fig2)a), but the frequency of their dormancy classes did not differ between generalist and strict alpine species.
+Among species with dormant seeds, physiological dormancy was the most common dormancy class, followed by morpho-physiological and physical.
+Both strict alpine and generalist species had a similar range of values for seed mass (Fig.
+\@ref(fig:fig2)b) and embryo to endosperm ratio (Fig.
+\@ref(fig:fig2)c).
+The values of seed mass ranged from `r MSminMass` to `r MSmaxMass` mg, with a median of `r MSmedianMass`.
+Embryo:endosperm ratio encompassed the full range of potential values, from endospermic species with very small embryos (`r MSminEmbryo`) to non-endospermic seeds (`r MSmaxEmbryo`).
+The median value was `r MSmedianEmbryo`.
+
+The total number of seeds used in the experiments was `r MSseeds`.
+Germination temperatures (weighted average of the daily thermoperiod) ranged from `r MSminT` to `r MSmaxT` ºC, with `r MSaltY` records of constant temperatures and `r MSaltN` of alternating temperatures.
+Light was used in `r MSlightY` records, and darkness in `r MSlightN`.
 The experiments were performed with unstratified seeds in `r MSstratN` records, and of the rest, the majority (`r MSstratCold`) went through cold stratification.
-Scarification was applied to `r MSscarY` records, and GA~3~ to `r MSga3`.
+Scarification was applied in `r MSscarY` records, and GA~3~ in `r MSga3`.
+Since germination incubations ran for different periods of time, we established a cut-off time of 8 weeks to calculate the final germination percentages, to have a homogeneous comparison.
 
-## Data analysis
+## Germination response variables
 
-Across species, germination and dormancy patterns are affected by a combination of factors such as phylogeny [@RN4723; @RN4724; @RN4715] and seed traits [@RN3061; @RN4728].
+We used the final germination proportion as the main response variable for the germination treatments.
+We further calculated the mean time to germination and the germination uncertainty index with the package *GerminaR* [@RN4752].
+Mean germination time indicates the time it takes for half of the seed lot to germinate, while the germination uncertainty index estimates how scattered germination is through time.
+To calculate these indices, we used only records that had (i) more than 50% final germination; and (ii) less than 4 weeks of mean germination time; we established these limits to prevent the results being altered by a modification of seed dormancy status during the germination incubation.
+
+## MCMCglmm models
+
+We meta-analysed germination data by fitting generalized mixed models with Bayesian estimation (Markov Chain Monte Carlo generalized linear mixed models, MCMCglmms) using the R package *MCMCglmm* [@RN4755].
+To model final germination proportion, we used binomial MCMCglmms, while for mean germination time and germination uncertainty index we used gaussian MCMCglmms.
+Models had, as fixed effects, the experimental conditions (temperature, alternating temperature, light, scarification, stratification and GA~3~), plus their interaction with seed mass and embryo:endosperm, and the strict alpine or generalist character of the species.
+Random effects included a reconstructed phylogenetic tree for the `r MSspecies`, species identity, seed lot, lab, and alpine region.
+To create the phylogeny, we used the R package *V.PhyloMaker* [@RN4753].
+*V.PhyloMaker* contains a mega-tree which is an updated version of GBOTB for the seed plants [@RN4754].
+We placed taxa absent from the mega-tree at the genus-level basal node.
+In all models, all variables were scaled.
+We used weakly informative priors in all models, with parameter-expanded priors for the random effects.
+Each model was run for 500,000 MCMC steps, with an initial burn-in phase of 50,000 and a thinning interval of 50 [@RN4756], resulting, on average, in 9,000 posterior distributions.
+From the resulting posterior distributions, we calculated mean parameter estimates and 95% Highest Posterior Density (HPD) and Credible Intervals (CI).
+We estimated the significance of model parameters by examining CIs, considering parameters with CIs overlapping with zero as non-significant.
+To estimate the phylogenetic signal of seed germination over all variables we used Pagels’s lambda (λ) [@RN4757], estimated simultaneously with the models by calculating the mean of the posterior distribution and the 95% CI of λ as indicated by [-@RN4756].
+When λ = 0, related taxa are no more similar than expected by chance, while when λ = 1, the trait is evolving following a constant variance random walk or Brownian motion model; intermediate values of λ indicate a phylogenetic correlation in traits evolution that does not fully follow a Brownian motion model [@RN4757].
+
+## FAMD ordination
+
+To visualize the alpine seed germination spectrum, we performed a Factorial Analysis of Mixed Data (FAMD) as implemented in the package *FactoMineR* [@RN3166].
+FAMD ordination combines the properties of PCA and MCA to jointly measure the variation of continuous and categorical variables.
+The ordination was performed at the species level, i.e.
+calculating a series of continuous and categorical traits for each species.
+We created a continuous variable for each germination cue (i.e.
+temperature, alternating temperatures, light, scarification, stratification, and GA~3~).
+To do so, for each cue and species, we calculated a weighted average of the cue levels (cue levels were the temperature treatments in the case of temperature; for the other cues the levels were 0 = absence and 1 = presence), weighting by the germination proportion at each level.
+This approach underrepresents the importance of the levels that were not tested for a given species, but can serve as a proxy of germination cues when visualized across the whole dataset; it must be stressed that this stage of the analysis serves only for visualization and not hypothesis testing.
+For each species, we also calculated a single mean germination time (the minimum, i.e.
+the time taken at the most favourable treatment) and germination uncertainty index (the mean).
+We conducted the FAMD only with the `r MSfamdspp` species for which we had mean germination time and germination uncertainty index values.
+We also included the continuous variables seed mass and embryo:endosperm, as well as the categorical variables dormancy class, distribution (alpine vs.
+generalist), life form and life span.
+
+# Results
+
+## Germination temperature
+
+Temperature had a positive effect on the final germination proportion (Fig.
+\@ref(fig:fig3)a), indicating a preference for warmer temperatures.
+Increasing temperatures also decreased mean germination time (Fig.
+\@ref(fig:fig3)b) and the germination uncertainty index (Fig.
+\@ref(fig:fig3)c).
+The effect of warmer temperatures was more intense in strict alpine than in generalist species.
+Seed mass and embryo:endosperm reduced the response to temperature: heavier and less endospermic seeds increased less their germination proportion and decreased less their germination time in warmer temperatures, although uncertainty did not show an interaction with mass or embryo:endosperm.
+
+## Alternating temperatures
+
+Alternating temperatures had a positive effect on the germination proportion (Fig.
+\@ref(fig:fig3)a), but not in germination time (Fig.
+\@ref(fig:fig3)b) or uncertainty (Fig.
+\@ref(fig:fig3)c).
+The positive effect on the germination proportion was slightly higher in the generalist species.
+Again, higher seed mass and embryo:endosperm reduced the positive effect of alternating temperatures.
+
+## Light
+
+Light enhanced the germination proportion (Fig.
+\@ref(fig:fig3)a), with no differences between strict alpine and generalist species.
+It did not affect the germination time (Fig.
+\@ref(fig:fig3)b), but it did increase the germination uncertainty of generalist species (Fig.
+\@ref(fig:fig3)c).
+Heavier and less endospermic seeds responded less to light.
+
+## Scarification
+
+Scarification promoted germination (Fig.
+\@ref(fig:fig3)a) and reduced germination time (Fig.
+\@ref(fig:fig3)b), but it did not affect germination uncertainty (Fig.
+\@ref(fig:fig3)c).
+Strict alpine species responded slightly more to scarification.
+Heavier seeds were more responsive to scarification, while seeds with smaller embryos responded less.
+
+## Stratification and GA~3~
+
+In general, stratification increased the germination proportion (Fig.
+\@ref(fig:fig3)a).
+This positive effect occurred in both generalist and strict alpine species, although the effect was higher on the latter group.
+Stratification slightly reduced mean germination time (Fig.
+\@ref(fig:fig3)b), which was evident on the generalist species, but not on the strict alpines.
+Further, stratification reduced germination uncertainty as a general response (Fig.
+\@ref(fig:fig3)c) as well as on generalist and strict alpine species in particular.
+Seed mass did not interact with stratification, but species with smaller embryos had a smaller response to stratification in their germination proportion and reduced their germination time and uncertainty.
+GA~3~ had similar effects to stratification, with the main difference being that a higher seed mass generally reduced the response to GA~3~.
+
+## Random effects and phylogenetic signal
+
+Random factors influenced the germination response to the environment (Fig.
+\@ref(fig:fig4)).
+The final germination proportion and the mean germination time were affected more strongly, on average, by the seed lot and the species identity.
+Alpine region and lab had a lower average effect, but more variability.
+Germination uncertainty showed the opposite pattern, with a stronger effect of lab and alpine region.
+The phylogenetic signal in the response to all germination parameters (Fig.
+\@ref(fig:fig5)) was positive for the final germination proportion and the mean germination time.
+For the germination uncertainty it was close to zero, indicating a weak effect of phylogeny.
+
+## FAMD ordination
+
+The first FADM axis explained `r round(pcaAlpine$eig[1, 2], 0)`% of the variation.
+The quantitative variables with the largest contribution to this first axis were germination temperature, germination uncertainty, GA~3~ and alternating temperatures.
+Among the qualitative factors, the strict alpine / generalist categories contributed to axis 1.
+In summary, the horizontal axis separated (i, left) generalist with a positive response to alternating temperatures and more germination uncertainty from (ii, right) strict alpine species with a preference for warmer temperatures and a positive response to GA~3~ (Fig.
+\@ref(fig:fig6)).
+Axis 2 explained `r round(pcaAlpine$eig[2, 2], 0)`% of the variability.
+The main contributing variables were embryo:endosperm, mean germination time, seed mass and scarification; and the main categorical factors were physical dormancy, graminoid life form and morphophysiological dormancy.
+This axis separated (iii, bottom) morphophysiologically dormant species and graminoids with longer mean time to germination from physically dormant species that responded to scarification and had heavier, less endospermic seeds.
 
@@
 
-The influence of environmental conditions as drivers of these contrasting germination responses to temperature is gaining recognition.
-For example, alpine species show higher optimal temperatures for germination than either subalpine species (i.e.
-species that live close to the treeline, Fernández-Pascual *et al.* [-@RN2371]) or congeneric counterparts from below the treeline [@RN3266].
-Differences in germination traits have been attributed also to species successional niche, with pioneer species germinating better at colder temperatures than later successional species [@RN3229].
-Species from calcareous and siliceous alpine grasslands show a slow overwinter germination or high germination under all conditions, respectively [@RN4727; @RN4726].
-Other factors related to species germination characteristics include slope [@RN4733] and biogeographical origin [@RN698].
-Nevertheless, no habitat-related germination strategies were identified when comparing a major divide in alpine microhabitats, as is the one between fellfields and snowbeds [@RN707].
+## Seed dormancy
+
+Our dataset indicates that seed dormancy is frequent across many alpine regions of the world and plays an important role in plant recruitment in the harsh alpine environment.
+However, dormancy was similarly frequent in both strict alpine and generalist species, suggesting that this is a general strategy of temperate species but not restricted to alpine lineages.
+Seed dormancy is a common adaptation in seasonal climates where germination can be risky during certain seasons [@RN3214].
+The dominance of alpine species with dormant seeds suggests that germination after dispersal, usually in late summer or autumn, is delayed to favourable seasons for seedling establishment and survival such as spring and early summer [@RN3285].
+There may also be a phylogenetic component to the frequency of some dormancy classes as, for example, morphological dormancy is common in the Campanulaceae, Gentianaceae and Ranunculaceae families [@RN3214], which are well represented in alpine floras.
+
+Seeds with the two most frequent dormancy classes, physiological and morphophysiological, require a stratification period during which they experience cold and wet conditions over a period of months [@RN3214; @RN4743].
+Our results confirm that cold stratification is important for promoting seed germination in species inhabiting alpine habitats, especially for strict alpine species.
+Our results also indicate that the hormone gibberellic acid (GA~3~) can be used as a substitute of cold stratification to overcome inherent physiological dormancy.
+Cold stratification decreases time to germination and increases overall percent germination.
+These results concur with several studies that have shown that cold-stratification is important for seed germination in different alpine regions of the world [@RN3393; @RN2868; @RN707; @RN698; @RN3703; @RN2355; @RN3285; @RN2371; @RN4713].
+Under natural field conditions, cold stratification occurs over winter when seeds are covered by snow.
+Hence, the role of cold stratification on the different germination parameters assessed suggests that this allows seeds to sense the snow season, and thereby promotes germination to occur in the spring and summer snow-free season, when conditions are more favourable for seedling survival and growth.
+Since global climate change is causing major changes in snow cover duration [@RN4917; @ RN4918], it could disrupt natural cold stratification cycles in alpine environments, compromising population viability and indirectly favouring species with no stratification requirement [@RN3703].
+This risk will be partly alleviated by the know plasticity of physiological seed dormancy in response to the seed maturation temperature, a mechanism of “seed memory” that facilitates acclimatisation to changing environments [@RN2249], and which could reduce the stratification requirements of seeds produced in a warmer climate.
+
+Scarification removes mechanical barriers in species with physical dormancy, an alternative dormancy mechanism based on a water-impermeable seed coat [@RN3214].
+We found physical dormancy to be infrequent in alpine habitats, but when it does occur it could be broken by mechanical scarification via freezing-thawing cycles in spring and/or high diurnal temperature fluctuations in summer.
+
+A small fraction of species in our dataset were classified as non-dormant.
+The response to high germination temperatures that we found suggests that, under a warming climate, non-dormant seeds could germinate readily in autumn, and this would provide a longer time for seedlings to establish.
+Non-dormant seeds could possibly have an advantage as the growing seasons become longer, if seedlings become large enough to survive over winter.
+This raises a question on whether having a non-dormant or less dormant seed is a better strategy for alpine plants to thrive under future climate [@RN3317; @RN2382].
+However, we also might expect that seedlings from these species will be exposed to frost if snow starts to melt earlier and insulation during winter disappears, and therefore a mechanism that enables diverse germination timing (i.e.
+high germination uncertainty) should to be favoured.
+
+## Temperature regulation of germination
+
+Our results show an increase of seed germination with incubation temperature, more pronounced in strict alpine compared to generalist species, confirming the findings of local studies [@RN3266; @RN2371].
+Considering that risk reduction is an important selective pressure for the evolution of germination traits [@RN4915], a plausible explanation for these results is that alpine plants received selection pressure from damaging spring frost and evolved an avoidance mechanism by germinating at high temperatures.
+Frost avoidance is an important survival strategy in alpine plants [@RN4908; @RN4906; @RN4722].
+Complementarily, strict alpine species might have been less subjected to drought damage during establishment compared to generalist species.
+Supporting this view, seedlings of alpine pioneer species have a low frost resistance [@RN4722], yet high heat tolerance up to 40–50°C [@RN4902].
+Plants in high-elevation habitats generally do not experience drought limitation because precipitation increases while evapotranspiration decreases along elevational gradients [@RN4914].
+
+Our results also show that germination is faster with increasing incubation temperature, especially in small seeded, endospermic species.
+This is not surprising, as alpine seed germination must be fast to match the short growing season.
+Small-seeded species have a lower rate of seeding survival under drought [@RN4530] but may require warmer germination temperatures than large-seeded species, especially in seasonal climates [@RN2399].
+Accordingly, this pattern was stronger in strict alpine plants which, being restricted to higher elevations, presumably experience less unpredictable drought but more frost risk, compared with generalist species: precipitation and the probability of early or late season frosts generally increase with elevation [@RN2392; @RN4897].
+Taken together, our results indicate that germination patterns in alpine species are driven by an interplay of seed size and germination temperature to escape either drought or frost stress: small seeds selecting for fast germination at warm temperatures in order to escape unfavourable early-spring frost and larger seeds selecting for slower germination at cooler conditions, which should maximize emergence during snowmelt (when water availability is highest) and a well-developed root system to cope with desiccation risk in summer.
+
+Finally, increasing temperatures also decreased germination uncertainty, showing that germination synchrony is a plastic trait driven by temperature, at least in alpine species.
+Indeed, while asynchronous germination is thought to be a form of adaptation to unpredictable alpine environments [@RN3734; @RN3690], our results indicate that staggered germination occurs when temperatures are still cool, likely as a bet-edging strategy against the risk of early-spring frost.
+On the other hand, a fast and synchronised germination at warm conditions (i.e.
+in late spring) may be another important strategy that increases alpine seedling survival by avoiding summer drought.
+The extreme germination uncertainty syndrome, i.e.
+staggering germination so that emergence occurs both before and after winter, is known in alpine species [@RN2392; @RN3285; @RN4907].
+Staggered germination is usually exhibited as a rapid first wave of germination followed by second wave several weeks later but can also take place in two seasons separated by a whole year [@RN2392].
+If autumn seedlings grow faster than spring seedlings [@RN4907], this will likely have cascading effects on individual fitness along the life of the plant [@RN3036; @RN2384].
+Delving further into the staggering germination syndrome, such as examining whether the source of variation lies within individuals, across individuals within a population, or across populations; or whether such syndrome is conserved; will help us to understand further the ecology of the unique alpine flora and its fate under a changing climate.
+
+## Alternating temperatures and light
+
+We found a positive germination response to alternating temperatures and light, conditions that mimic the environment in the upper soil, during snowmelt at the end of winter [@RN4712; @RN2392; @RN3029].
+Triggering seed germination at the very beginning of the spring season ensures that seedlings have time to establish during the short growing seasons [@RN2392] and grow to a critical biomass before the next winter [@RN2868].
+Alternating temperatures had a slightly lower effect on strict alpine species, compared with generalist species, but still had an effect.
+This partly contradicts the findings of Liu et al.
+[-@RN3376], who reported a lack of a response to temperature fluctuation for the species distributed only at high elevations of the Qinghai-Tibet Plateau.
+Liu and co-workers argued that, in that area, high temperature fluctuations may occur in all months of the year and therefore fluctuating temperature alone may not be a reliable indicator of suitable conditions for seedling establishment and growth, while warm daily temperatures could be a cue that the short summer growing season has begun [@RN3376].
+When analysed at the global level, alternating temperatures seem to increase germination of alpine plants, which corresponds well with the fact that most alpine regions have a strong day-night climatic variation during the growing season.
+
+Solar irradiance in physiologically significant quantities penetrates only the first millimetres of the soil [@RN4909] and fluctuations of diurnal temperature decrease with increasing burial depths: below 10 cm they might be too small to trigger germination of species requiring temperature variation [@RN4910].
+Therefore, it is not surprising that we detected a negative effect of seed mass on final germination values in the light and under alternating temperature regimes.
+The depth of seed burial in the soil is crucial for seedling emergence [@RN4898], as seed mass  - or better said the seed kernel, i.e.
+embryo and endosperm [@RN4899] - may represent a constraint for seedling emergence of small-seeded species.
+Therefore, small seeds are more likely to require light and alternating temperatures for germination, which ensures that germination does not occur too deep in the soil for seedling emergence [@RN4904].
+
+## Seeds, embryos, and phylogenetic signal
+
+Seed mass is a relatively constant trait in our data, with no differences between strict alpine and generalist species.
+This is in line with the general ambiguity of relationships between seed mass and elevation.
+In alpine species, the evidence of a relationship between seed mass and elevation is mixed, being either negative [@RN4912], positive [@RN4903] or absent [@RN3685; @RN4896] when looking at the species level.
+Similarly, at the population level, the relationship between seed mass and elevation is often absent [@RN3442; @RN4896].
+The length of the growing season might influence seed size.
+For instance, at higher elevations where snow duration is longer and thus the growing season is shorter, the production of smaller seeds might be favoured [@RN4895; @RN4912].
+On the other hand, natural selection may also favour the production of larger seeds in species at higher elevations because larger seeds enable plants to cope better with stressful environments [@RN4913; @RN4903].
+
+Much less information is available concerning a possible selective advantage of embryo:endosperm size in alpine species.
+However, the similarity in embryo:endosperm size between generalist and strict alpine species confirms the absence of a correlation between embryo-seed size and elevation found previously in Apiaceae species [@RN3685].
+Although seed size and embryo:endosperm ratio are not different for generalist and strict alpine species in this study, these traits have been found to influence germination strategies in alpine species.
+For instance, previous studies have shown that non-endospermic seeds are smaller and are likely to be non-dormant at the time of dispersal and thus germinate quicker than alpine endospermic seeds [@RN3703; @RN3285; @RN4907].
+Perhaps accumulation of reserves in cotyledons rather than in the endosperm may improve early growth rate and establishment in immediate germinating species [@RN3285].
+
+A general assumption of seed trait ecology is that seed mass and embryo:endosperm ratios are relatively homogeneous across related species [@RN4916; @RN3685].
+In contrast, physiological responses to temperature and dormancy patterns are understood as more plastic traits that can quickly acclimatize in new situations [@RN2249].
+Although our study is not focused on trait evolution, we detected close relationships between seed size and physiological responses of seeds, suggesting that both traits are connected and may be subjected to co-adaptation.
+While phylogenetic signal in seed germination has been found often, especially in highly selective environments [@RN3001; @RN2399; @RN2865], seed germination patterns can also be shared between phylogenetically distant species [@RN4911] with low or absent phylogenetic signal [@RN2365; @RN4900; @RN2371].
+However, the phylogenetic signal found here indicates that evolutionary history cannot be neglected when studying seed germination patterns in the alpine environment.
+Further studies are required to assess whether the degree of the shared germination patterns among phylogenetically related species in the alpine environment is the result of trait conservatism, convergence, or both [@RN4901; @RN3235].
+
+## The alpine seed germination spectrum
+
+The factorial analysis of mixed data (FAMD) separated generalist and strict alpine plants according to a “fast-slow” gradient of regeneration strategies.
+On one end of the gradient, strict alpine species tend to have slow and synchronous germination, require warm temperatures and light to germinate, and need stratification or GA~3~ to break physiological dormancy.
+On the other end, generalist species tend to have fast germination, show a bet-hedging strategy (high germination uncertainty), and show germination traits suggesting possible fast germination after specific cues (PY, alternating temperatures).
+The dormancy classes without a physiological component (non-dormant, physical, and morphological) appear associated with generalist species.
+While most species are perennial, the generalist group includes some annuals.
+Life forms are evenly distributed across regeneration strategies.
+This “fast-slow” gradient of regenerative strategies is similar to prominent gradients of fast or slow population dynamics [@RN4922; @RN4921], fast or slow developing leaf traits [@RN4923], or large sets of plant ecological features [@RN2279].
+“Fast” regeneration can be understood as an opportunistic strategy to cope with frequently disturbed habitats such as avalanche ways, steep eroding slopes, regressing glaciers, and riverbeds [@RN4926; @RN4925].
+The “slow” regeneration of strict alpine species, on the other hand, seems to be a more specialized syndrome of those species that are truly restricted - and adapted - to the alpine belt.
+
+## Conclusions
+
+Our meta-analysis of primary data has shown that strict alpine species from different mountain regions of the world tend to show a specialized seed germination strategy which shows evidence of phylogenetic signal and is characterized by:
+
+* Physiological seed dormancy.
+* A strong need for cold stratification to break dormancy, which can be substituted by GA~3~. 
+* A positive response to light.
+* A positive response to alternating temperatures, although not so prominent as in generalist species.
+* Slow and relatively synchronous germination.
+* An interplay with seed and embryo size, with smaller and more endospermic seeds being more responsive to warmth, light, and alternating temperatures.
+
+# Acknowledgements
+
+E.F.P. received financial support from the Government of Asturias and the FP7 – Marie Curie - COFUND programme of the European Commission (Grant ‘Clarín’ ACB17-19).
+
+# Author Contribution
+
+# Data Availability Statement
+
+The original data, R code for the analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/alpineseeds.
+Upon publication, a version of record of the repository will be deposited in Zenodo.
+
+# References
+
+::: {#refs}
+:::
 
@@
 
-# References
+```{r fig2, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Seed dormancy classes, seed mass and embryo to endosperm ratio in the subsets of strict alpine and generalist species analysed in this study.
+Seed dormancy classes are abbreviated as follows: MD (morphological), MPD (morphophysiological), ND (non-dormant), PD (physiological), PY (physical).", fig.align = "center", fig.width=7, fig.height=3}
+cowplot::ggdraw(fig2)
+```
+
+```{r fig3, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the germination environment on germination, according to the MCMC meta-analysis of the primary data.
+The dots indicate the posterior mean of the effect size, and the brackets its 95 % credible interval.
+The line of zero effect is shown.
+When the credible intervals overlap with the zero-effect line, the effect is not significant.
+The figure shows first the main effect, then the effects for the strict alpine and generalist groups, and then the interaction with seed mass and embryo:endosperm.
+A negative interaction (e.g.
+seed mass and light, in the case of the germination proportion) indicates that, as the interacting factor increases (e.g.
+seed mass) the response to the germination cue (e.g.
+light) decreases, in comparison with the main effect.", fig.align = "center", fig.width=7, fig.height=5.5}
+cowplot::ggdraw(fig3)
+```
+
+```{r fig4, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the random factors on germination, according to the MCMC meta-analysis of the primary data.
+The dots indicate the posterior mean of the effect size, and the brackets its 95 % credible interval.
+The line of zero effect is shown.", fig.align = "center", fig.width=7, fig.height=5}
+cowplot::ggdraw(fig4)
+```
+
+```{r fig5, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Phylogenetic signal in the response to the germination environment.
+When lambda = 0, related taxa are no more similar than expected by chance; when λ = 1, the trait is evolving following a constant variance random walk or Brownian motion model", fig.align = "center", fig.width=7, fig.height=1.5}
+cowplot::ggdraw(fig5)
+```
+
+```{r fig6, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "FADM ordination of the seed variables and the species distribution, life form and life span.
+Each square is a species.
+Labels indicate the contribution of the variables to the axes.", fig.align = "center", fig.width=7, fig.height=7}
+cowplot::ggdraw(fig6)
+```

### 2020-08-09 doc/ms/ms.Rmd
ms summarised

@@
 Other germination cues, such as the response to light [@RN4656] and alternating temperatures [@RN1380], allow for a fine-scale detection of germination micro-niches and safe sites [@RN4719].
-Given the heterogeneity of alpine climates and species lineages adapted to different regions, and the wealth of local studies on alpine seed ecology (reviewed in the following paragraphs), a current challenge is to synthesize, at the global level, the seed germination spectrum of alpine plants [@RN4655].
-Such a global synthesis would allow us to (i) better understand how climatic changes will affect alpine plant diversity; and (ii) manage and eventually counteract regression of restricted alpine species.
+Given the heterogeneity of alpine climates and species lineages adapted to different regions, and the wealth of local studies on alpine seed ecology (reviewed in the following paragraphs), a current challenge is to synthesize, at the global level, the *seed germination spectrum* of alpine plants [@RN4655].
+Saatkamp et al.
+[-@RN4655] define the seed ecological spectrum as the set of trait-based ecological responses of seeds that determines their ability to disperse, persist, germinate, and establish.
+A global synthesis of the alpine germination spectrum would allow us to (i) better understand how climatic changes will affect alpine plant diversity; and (ii) manage and eventually counteract regression of restricted alpine species.
 
@@
 Along an elevational gradient in the central Chilean Andes, cold stratification was an important requirement for seed germination in species from lower elevations, while species from higher sites did not germinate even after cold stratification [@RN4713].
+In a comparative study of species from different habitats of the Austrian Alps which assessed the germination of unstratified seeds, species from alpine grasslands had the lowest final germination proportion [@RN4930].
 Therefore, how important cold stratification is for seed germination in alpine plant species remains unclear.
 
+The underlying question is whether seed dormancy plays a relevant role in postponing alpine germination during autumn and winter, or the warm-cued germination is enough to prevent germination under cool autumn temperatures.
+The response to this question becomes non-trivial under a warming climate, by which non-dormant seeds could germinate readily in autumn, and this would provide a longer time for seedlings to establish.
+Non-dormant seeds could possibly have an advantage as the growing seasons become longer, if seedlings become large enough to survive over winter.
+Potentially, having a non-dormant or less dormant seed could be a better strategy for alpine plants to thrive under future climate [@RN3317; @RN2382].
+However, we also might expect that seedlings from these species will be exposed to frost if snow starts to melt earlier and insulation during winter disappears, and therefore a mechanism that enables diverse germination timing (i.e.
+high germination uncertainty) should to be preferred.
+
 Besides germination temperature and patterns of seed dormancy, the response of alpine seeds to light and alternating temperatures can provide further information on their capacity to detect fine-scale environmental cues.
+Solar irradiance in physiologically significant quantities penetrates only the first millimetres of the soil [@RN4909] and fluctuations of diurnal temperature decrease with increasing burial depths: below 10 cm they might be too small to trigger germination of species requiring temperature variation [@RN4910].
+Therefore, these cues can indicate whether seeds are buried deep in the soil, in rock crevices or under snow.
+The depth of seed burial is crucial for seedling emergence, as small seeds lack the resources to survive until they reach the soil surface [@RN4898].
 The preference of alpine species to germinate in light [@RN4730; @RN4729] could favour the creation of a persistent soil seed bank [@RN4718], which is advantageous in temporally and spatially unpredictable alpine environments [@RN4731; @RN4732].
@@
 The submitted data had to report the studied species, the germination treatments, the number of seeds sown, and the number of seeds germinated.
-We collected data from the Cantabrian Mountains [@RN2371], the European Alps [@RN4729; @RN3029; @RN4727; Rosbakh, unpublished], the Apennines (Mondoni, unpublished), the Caucasus [Rosbakh, unpublished], the Qinghai-Tibet Plateau [@RN4728; @RN4735; @RN3376], the Andes [@RN3393; @RN4713; Briceño, unpublished] and the Australian Alps [@RN4737; @RN4736; @RN3703; Satyanti, unpublished].
+We collected data from the Cantabrian Mountains of Spain [@RN2371], the European Alps [@RN4729; @RN3029; @RN4727; Rosbakh, unpublished], the Italian Apennines (Mondoni, unpublished), the Russian Caucasus [Rosbakh, unpublished], the Qinghai-Tibet Plateau of China [@RN4728; @RN4735; @RN3376], the Chilean Andes [@RN3393; @RN4713; Briceño, unpublished] and the Australian Alps [@RN4737; @RN4736; @RN3703; Satyanti, unpublished].
 We also collected data on European alpine species from *Enscobase*, the database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml).
 
-Once received, we standardized the species names using *The Plant List* [@RN2321] and the package *Taxonstand* [@RN4660] in *R* [@RN2315].
-For each species, we checked its elevation range in local floras [@RN4741; @RN4739; @RN3243; @RN4742; @RN4740], catalogues [@RN4738] and herbaria (*Australian National Herbarium*).
-We used the elevation ranges (i) to remove from the dataset species that do not grow above the treeline; and (ii) to classify the rest of the species as *strict alpine* (i.e.
+Once received, we standardized the species names according to *The Plant List* [@RN2321] using the package *Taxonstand* [@RN4660] in *R* [@RN2315].
+For each species, we recorded its elevation range from local floras [@RN4741; @RN4739; @RN3243; @RN4742; @RN4740], catalogues [@RN4738] and herbaria (*Australian National Herbarium*).
+We used the elevation ranges (i) to remove from the dataset outlier species that do not grow above the treeline; and (ii) to classify the rest of the species as *strict alpine* (i.e.
 species that only grow above the treeline) or *generalist* (i.e.
 species that can grow above the treeline but also in the lowlands).
-For each species, we also compiled the following traits: life form and life span from the references used for the elevation, plus some extra references [@RN4747; @RN4744; @RN4745; @RN4746; @RN4748; @RN4749] and online resources (https://www.infoflora.ch/de/; http://plantnet.rbgsyd.nsw.gov.au/search/simple.htm; http://buscador.floraargentina.edu.ar/; https://patagoniawildflowers.org/); seed dormancy class [@RN3214; @RN4743]; seed mass [@RN4728; @RN4735; @RN3376; @RN2987] and embryo to endosperm ratio (Vandelook, unpublished).
-For the species missing seed mass or embryo values, we used genus or family averages.
+For each species, we also compiled the following traits: (1) life form (forb, graminoid or woody) and (2) life span (annual or perennial) from the references used for the elevation, plus some extra references [@RN4747; @RN4744; @RN4745; @RN4746; @RN4748; @RN4749] and online resources (https://www.infoflora.ch/de/; http://plantnet.rbgsyd.nsw.gov.au/search/simple.htm; http://buscador.floraargentina.edu.ar/; https://patagoniawildflowers.org/); (3) seed dormancy class following the classification system of Baskin and Baskin: physiological (germination prevented by the seed’s internal balance of hormones), morphological (germination prevented by an underdeveloped embryo at the time of dispersal), morphophysiological (both physiological and morphological dormancy present), physical (germination prevented by a water-impermeable seed coat) or non-dormant [@RN3214; @RN4743]; (4) seed mass [@RN4728; @RN4735; @RN3376; @RN2987] and (5) embryo to endosperm ratio (Vandelook, unpublished).
+For the species missing seed mass or embryo values, we calculated genus or family averages.
 
@@
 
-The final dataset used for meta-analysis contained `r MSrecords` records (i.e.
+The final dataset used in our meta-analysis contained `r MSrecords` records (i.e.
 germination proportions for a given seed lot of a species, recorded in a set of experimental conditions) contributed by `r MSgroups` research groups, representing `r MSfamilies` seed plant families and `r MSspecies` species (`r MSstrict` strict alpine species and `r MSgeneral` generalists).
@@
 Most were reported to have dormant seeds (Fig.
-\@ref(fig:fig2)a), but the frequency of their dormancy classes did not differ between generalist and strict alpine species.
-Among species with dormant seeds, physiological dormancy was the most common dormancy class, followed by morpho-physiological and physical.
+\@ref(fig:fig2)a), but the frequency of the dormancy classes did not differ between generalist and strict alpine species.
+Among those species with dormant seeds, physiological dormancy was the most common dormancy class, followed by morpho-physiological and physical.
 Both strict alpine and generalist species had a similar range of values for seed mass (Fig.
@@
 \@ref(fig:fig2)c).
-The values of seed mass ranged from `r MSminMass` to `r MSmaxMass` mg, with a median of `r MSmedianMass`.
+The values of seed mass ranged from `r MSminMass` to `r MSmaxMass` mg, with a median value of `r MSmedianMass` mg.
 Embryo:endosperm ratio encompassed the full range of potential values, from endospermic species with very small embryos (`r MSminEmbryo`) to non-endospermic seeds (`r MSmaxEmbryo`).
@@
 The total number of seeds used in the experiments was `r MSseeds`.
-Germination temperatures (weighted average of the daily thermoperiod) ranged from `r MSminT` to `r MSmaxT` ºC, with `r MSaltY` records of constant temperatures and `r MSaltN` of alternating temperatures.
-Light was used in `r MSlightY` records, and darkness in `r MSlightN`.
-The experiments were performed with unstratified seeds in `r MSstratN` records, and of the rest, the majority (`r MSstratCold`) went through cold stratification.
-Scarification was applied in `r MSscarY` records, and GA~3~ in `r MSga3`.
-Since germination incubations ran for different periods of time, we established a cut-off time of 8 weeks to calculate the final germination percentages, to have a homogeneous comparison.
+Germination temperatures (i.e., the weighted average of the day and night temperatures, weighted by the duration of each phase) ranged from `r MSminT` to `r MSmaxT` ºC, with `r MSaltY` records of constant temperatures (i.e.
+experiments that used the same temperature during all their duration) and `r MSaltN` of alternating temperatures (i.e.
+experiments where different temperatures were applied during the day and the night, in diurnal cycles).
+Seeds were exposed to light during some part of the diurnal cycle in `r MSlightY` records, and kept in total darkness in `r MSlightN` records.
+The experiments were performed with untreated (i.e.
+fresh) seeds in `r MSstratN` records, and of the rest, the majority (`r MSstratCold`) went through cold stratification (< 5 ºC, in darkness).
+Scarification (i.e.
+an abrassion of the seed coat to allow water imbibition) was performed in `r MSscarY` records, and GA~3~ (a plant hormone that promotes embryo growth and germination) in `r MSga3` records.
+Since germination incubations ran for different periods of time, we established a cut-off time of 8 weeks to calculate the final germination percentages, to have a homogeneous comparison across records.
 
@@
 
-We used the final germination proportion as the main response variable for the germination treatments.
-We further calculated the mean time to germination and the germination uncertainty index with the package *GerminaR* [@RN4752].
-Mean germination time indicates the time it takes for half of the seed lot to germinate, while the germination uncertainty index estimates how scattered germination is through time.
-To calculate these indices, we used only records that had (i) more than 50% final germination; and (ii) less than 4 weeks of mean germination time; we established these limits to prevent the results being altered by a modification of seed dormancy status during the germination incubation.
+We used the final germination proportion (i.e.
+the number of seeds germinated out of seeds sown) as the main response variable for the germination treatments.
+We further calculated the mean germination time and the germination uncertainty index with the package *GerminaR* [@RN4752].
+Mean germination time is the time that it takes for half of the seed lot to germinate, indicating the germination speed or rate in a given experimental condition.
+The germination uncertainty index estimates how scattered germination is through time, and gives an idea of whether germination is synchronous (i.e.
+most seeds germinate around the mean germination time) or asynchronous (i.e.
+germination events are separated through time during the germination experiment).
+To calculate these indices, we used only records that met two conditions: (i) to have achieved more than 50% final germination percentage; and (ii) to have a mean germination time under 4 weeks.
+We established these limits to prevent the results being altered by a modification of seed dormancy status during the germination incubation (e.g.
+a fraction of the seed lot is non-dormant at the beginin of the incubation and germinates shortly after the start of the experiment; while another fraction is originally dormant, loses dormancy in response to the incubation conditions, and germinates in a second wave).
 
@@
 Models had, as fixed effects, the experimental conditions (temperature, alternating temperature, light, scarification, stratification and GA~3~), plus their interaction with seed mass and embryo:endosperm, and the strict alpine or generalist character of the species.
-Random effects included a reconstructed phylogenetic tree for the `r MSspecies`, species identity, seed lot, lab, and alpine region.
+Random effects included a reconstructed phylogenetic tree for the `r MSspecies`, and also species identity, seed lot, lab, and alpine region.
 To create the phylogeny, we used the R package *V.PhyloMaker* [@RN4753].
@@
 To estimate the phylogenetic signal of seed germination over all variables we used Pagels’s lambda (λ) [@RN4757], estimated simultaneously with the models by calculating the mean of the posterior distribution and the 95% CI of λ as indicated by [-@RN4756].
-When λ = 0, related taxa are no more similar than expected by chance, while when λ = 1, the trait is evolving following a constant variance random walk or Brownian motion model; intermediate values of λ indicate a phylogenetic correlation in traits evolution that does not fully follow a Brownian motion model [@RN4757].
+When λ = 0, related taxa are no more similar than expected by chance, while when λ = 1, the trait is evolving following a constant variance random walk or Brownian motion model; intermediate values of λ indicate a phylogenetic correlation in trait evolution that does not fully follow a Brownian motion model [@RN4757].
 
@@
 temperature, alternating temperatures, light, scarification, stratification, and GA~3~).
-To do so, for each cue and species, we calculated a weighted average of the cue levels (cue levels were the temperature treatments in the case of temperature; for the other cues the levels were 0 = absence and 1 = presence), weighting by the germination proportion at each level.
-This approach underrepresents the importance of the levels that were not tested for a given species, but can serve as a proxy of germination cues when visualized across the whole dataset; it must be stressed that this stage of the analysis serves only for visualization and not hypothesis testing.
+To do so, for each cue and species, we calculated a weighted average of the cue levels (in the case of temperature, cue levels were the temperature treatments; for the other cues the levels were 0 = absence and 1 = presence), weighting by the germination proportion at each level.
+This approach underrepresents the importance of the levels that were not tested for a given species, but can serve as a proxy of the response to the germination cues when visualized across the whole dataset; it must be stressed that this stage of the analysis serves only for visualization and not hypothesis testing.
 For each species, we also calculated a single mean germination time (the minimum, i.e.
 the time taken at the most favourable treatment) and germination uncertainty index (the mean).
-We conducted the FAMD only with the `r MSfamdspp` species for which we had mean germination time and germination uncertainty index values.
-We also included the continuous variables seed mass and embryo:endosperm, as well as the categorical variables dormancy class, distribution (alpine vs.
-generalist), life form and life span.
+We conducted the FAMD with only the `r MSfamdspp` species for which we had mean germination time and germination uncertainty index values.
+We also included the continuous variables seed mass and embryo:endosperm, as well as the categorical variables dormancy class (physiological, morphophysiological, morphological, physical or non-dormant), distribution (strict alpine vs.
+generalist species), life form (forb, graminoid or woody) and life span (annual or perennial).
 
@@
 
-Temperature had a positive effect on the final germination proportion (Fig.
+Temperature had a positive effect on the final germination proportion as shown by a credible interval not overlapping zero (Fig.
 \@ref(fig:fig3)a), indicating a preference for warmer temperatures.
@@
 \@ref(fig:fig3)c).
-The effect of warmer temperatures was more intense in strict alpine than in generalist species.
-Seed mass and embryo:endosperm reduced the response to temperature: heavier and less endospermic seeds increased less their germination proportion and decreased less their germination time in warmer temperatures, although uncertainty did not show an interaction with mass or embryo:endosperm.
+The effect of warmer temperatures was slightly more intense in strict alpine than in generalist species, although the credible intervals of the two groups of species partially overlap.
+Seed mass and embryo:endosperm had a negative interaction with temperature, in other words, they reduced the response to temperature: heavier and less endospermic seeds increased less their germination proportion and decreased less their mean germination time in warmer temperatures.
+The germination uncertainty index did not show an interaction with seed mass or embryo:endosperm, as shown by the credible interval overlapping zero.
 
@@
 
-Alternating temperatures had a positive effect on the germination proportion (Fig.
-\@ref(fig:fig3)a), but not in germination time (Fig.
-\@ref(fig:fig3)b) or uncertainty (Fig.
+Alternating temperatures had a positive effect on the final germination proportion (Fig.
+\@ref(fig:fig3)a), but not in the mean germination time (Fig.
+\@ref(fig:fig3)b) or the germination uncertainty index (Fig.
 \@ref(fig:fig3)c).
-The positive effect on the germination proportion was slightly higher in the generalist species.
-Again, higher seed mass and embryo:endosperm reduced the positive effect of alternating temperatures.
+The positive effect on the germination proportion was marginally higher in the generalist species.
+As in the case of the mean germination temperature, higher seed mass and embryo:endosperm values reduced the positive effect of alternating temperatures on the final germination proportion.
+Alternating temperatures reduced the mean germination time more in seeds with low values of embryo:endosperm.
 
@@
 
-Light enhanced the germination proportion (Fig.
+Light increased the final germination proportion (Fig.
 \@ref(fig:fig3)a), with no differences between strict alpine and generalist species.
-It did not affect the germination time (Fig.
-\@ref(fig:fig3)b), but it did increase the germination uncertainty of generalist species (Fig.
+It did not affect the mean germination time (Fig.
+\@ref(fig:fig3)b), but it did increase the germination uncertainty index of generalist species (Fig.
 \@ref(fig:fig3)c).
@@
 
-Scarification promoted germination (Fig.
-\@ref(fig:fig3)a) and reduced germination time (Fig.
-\@ref(fig:fig3)b), but it did not affect germination uncertainty (Fig.
+Scarification promoted the germination proportion (Fig.
+\@ref(fig:fig3)a) and reduced the mean germination time (Fig.
+\@ref(fig:fig3)b), but it did not affect the germination uncertainty index (Fig.
 \@ref(fig:fig3)c).
-Strict alpine species responded slightly more to scarification.
+Strict alpine species were slightly more responsive to scarification than generalist species.
 Heavier seeds were more responsive to scarification, while seeds with smaller embryos responded less.
@@
 
-In general, stratification increased the germination proportion (Fig.
+In general, stratification increased the final germination proportion (Fig.
 \@ref(fig:fig3)a).
-This positive effect occurred in both generalist and strict alpine species, although the effect was higher on the latter group.
-Stratification slightly reduced mean germination time (Fig.
-\@ref(fig:fig3)b), which was evident on the generalist species, but not on the strict alpines.
-Further, stratification reduced germination uncertainty as a general response (Fig.
-\@ref(fig:fig3)c) as well as on generalist and strict alpine species in particular.
-Seed mass did not interact with stratification, but species with smaller embryos had a smaller response to stratification in their germination proportion and reduced their germination time and uncertainty.
-GA~3~ had similar effects to stratification, with the main difference being that a higher seed mass generally reduced the response to GA~3~.
+This positive effect occurred in both generalist and strict alpine species, although the effect was higher in the strict alpine group.
+Stratification slightly reduced the mean germination time (Fig.
+\@ref(fig:fig3)b), which was evident in the generalist species, but not in the strict alpine species.
+Further, stratification reduced the germination uncertainty index as a general response (Fig.
+\@ref(fig:fig3)c), and slightly more so in generalist species.
+Seed mass did not interact with stratification, but species with smaller embryos had a smaller response to stratification in their final germination proportion and reduced less markedly their mean germination time and germination uncertainty index.
+GA~3~ had similar effects to stratification, with the main difference being that seed mass affected the response to GA~3~, whereas it did not interact with the response to stratification.
 
@@
 
-Random factors influenced the germination response to the environment (Fig.
+Random factors had a significant effect on the MCMCglmm models (Fig.
 \@ref(fig:fig4)).
 The final germination proportion and the mean germination time were affected more strongly, on average, by the seed lot and the species identity.
-Alpine region and lab had a lower average effect, but more variability.
-Germination uncertainty showed the opposite pattern, with a stronger effect of lab and alpine region.
+Alpine region and lab had a lower average effect, but more variability as shown by their credible intervals.
+The germination uncertainty index showed the opposite pattern, with a stronger effect of lab and alpine region.
 The phylogenetic signal in the response to all germination parameters (Fig.
 \@ref(fig:fig5)) was positive for the final germination proportion and the mean germination time.
-For the germination uncertainty it was close to zero, indicating a weak effect of phylogeny.
+For the germination uncertainty index, it overlapped with zero, indicating a weak effect of phylogeny.
 
@@
 
-Our dataset indicates that seed dormancy is frequent across many alpine regions of the world and plays an important role in plant recruitment in the harsh alpine environment.
-However, dormancy was similarly frequent in both strict alpine and generalist species, suggesting that this is a general strategy of temperate species but not restricted to alpine lineages.
-Seed dormancy is a common adaptation in seasonal climates where germination can be risky during certain seasons [@RN3214].
-The dominance of alpine species with dormant seeds suggests that germination after dispersal, usually in late summer or autumn, is delayed to favourable seasons for seedling establishment and survival such as spring and early summer [@RN3285].
-There may also be a phylogenetic component to the frequency of some dormancy classes as, for example, morphological dormancy is common in the Campanulaceae, Gentianaceae and Ranunculaceae families [@RN3214], which are well represented in alpine floras.
-
+Our dataset indicates that seed dormancy is frequent across many alpine regions of the world and postponing germination plays an important role in plant recruitment in the harsh alpine environment, although the frequency of dormant species is similar in strict alpine and generalist species.
 Seeds with the two most frequent dormancy classes, physiological and morphophysiological, require a stratification period during which they experience cold and wet conditions over a period of months [@RN3214; @RN4743].
-Our results confirm that cold stratification is important for promoting seed germination in species inhabiting alpine habitats, especially for strict alpine species.
+Our results show that cold stratification decreases time to germination and increases overall percent germination.
 Our results also indicate that the hormone gibberellic acid (GA~3~) can be used as a substitute of cold stratification to overcome inherent physiological dormancy.
-Cold stratification decreases time to germination and increases overall percent germination.
-These results concur with several studies that have shown that cold-stratification is important for seed germination in different alpine regions of the world [@RN3393; @RN2868; @RN707; @RN698; @RN3703; @RN2355; @RN3285; @RN2371; @RN4713].
-Under natural field conditions, cold stratification occurs over winter when seeds are covered by snow.
-Hence, the role of cold stratification on the different germination parameters assessed suggests that this allows seeds to sense the snow season, and thereby promotes germination to occur in the spring and summer snow-free season, when conditions are more favourable for seedling survival and growth.
-Since global climate change is causing major changes in snow cover duration [@RN4917; @ RN4918], it could disrupt natural cold stratification cycles in alpine environments, compromising population viability and indirectly favouring species with no stratification requirement [@RN3703].
+Physical dormancy, infrequent in our dataset, is broken by mechanical scarification, a process that could also occur naturally via freezing-thawing cycles in spring.
+Our results thus confirm that overwintering is essential for promoting seed germination in species inhabiting alpine habitats, especially for strict alpine species.
+This concurs with several studies that have shown that cold-stratification is important for seed germination in different alpine regions of the world [@RN3393; @RN2868; @RN707; @RN698; @RN3703; @RN2355; @RN3285; @RN2371; @RN4713].
+Under natural field conditions, cold stratification occurs over winter when seeds are covered by snow, allowing seeds to sense the snow season, and thereby promoting germination in the spring and summer, when conditions are more favourable for seedling survival and growth.
+Since global climate change is causing major changes in snow cover duration [@RN4917; @RN4918], it could disrupt natural cold stratification cycles in alpine environments, compromising population viability and indirectly favouring species with no stratification requirement [@RN3703].
 This risk will be partly alleviated by the know plasticity of physiological seed dormancy in response to the seed maturation temperature, a mechanism of “seed memory” that facilitates acclimatisation to changing environments [@RN2249], and which could reduce the stratification requirements of seeds produced in a warmer climate.
-
-Scarification removes mechanical barriers in species with physical dormancy, an alternative dormancy mechanism based on a water-impermeable seed coat [@RN3214].
-We found physical dormancy to be infrequent in alpine habitats, but when it does occur it could be broken by mechanical scarification via freezing-thawing cycles in spring and/or high diurnal temperature fluctuations in summer.
-
-A small fraction of species in our dataset were classified as non-dormant.
-The response to high germination temperatures that we found suggests that, under a warming climate, non-dormant seeds could germinate readily in autumn, and this would provide a longer time for seedlings to establish.
-Non-dormant seeds could possibly have an advantage as the growing seasons become longer, if seedlings become large enough to survive over winter.
-This raises a question on whether having a non-dormant or less dormant seed is a better strategy for alpine plants to thrive under future climate [@RN3317; @RN2382].
-However, we also might expect that seedlings from these species will be exposed to frost if snow starts to melt earlier and insulation during winter disappears, and therefore a mechanism that enables diverse germination timing (i.e.
-high germination uncertainty) should to be favoured.
 
@@
 Our results show an increase of seed germination with incubation temperature, more pronounced in strict alpine compared to generalist species, confirming the findings of local studies [@RN3266; @RN2371].
-Considering that risk reduction is an important selective pressure for the evolution of germination traits [@RN4915], a plausible explanation for these results is that alpine plants received selection pressure from damaging spring frost and evolved an avoidance mechanism by germinating at high temperatures.
+Our results also show that germination is faster with increasing incubation temperature, especially in small seeded, endospermic species.
+Small-seeded species have a lower rate of seeding survival under drought [@RN4530] but may require warmer germination temperatures than large-seeded species, especially in seasonal climates [@RN2399].
+Considering that risk reduction is an important selective pressure for the evolution of germination traits [@RN4915], a plausible explanation for these patterns is that alpine plants received selection pressure from damaging spring frost and evolved an avoidance mechanism by germinating at high temperatures.
 Frost avoidance is an important survival strategy in alpine plants [@RN4908; @RN4906; @RN4722].
-Complementarily, strict alpine species might have been less subjected to drought damage during establishment compared to generalist species.
+Complementarily, strict alpine species might have been less subjected to drought damage during establishment compared to generalist species: as elevation increases, plants face less drought risk because precipitation increases and evapotranspiration decreases, but at the same there is a higher probability of early or late season frosts [@RN4897; @RN2392; @RN4914].
+Taken together, our results indicate that germination patterns in alpine species are driven by an interplay of seed size and germination temperature to escape either drought or frost stress: small seeds selecting for fast germination at warm temperatures in order to escape unfavourable early-spring frost; and large seeds selecting for slower germination at cooler conditions, which should maximize emergence during snowmelt (when water availability is highest) and a well-developed root system to cope with desiccation risk in summer.
 Supporting this view, seedlings of alpine pioneer species have a low frost resistance [@RN4722], yet high heat tolerance up to 40–50°C [@RN4902].
-Plants in high-elevation habitats generally do not experience drought limitation because precipitation increases while evapotranspiration decreases along elevational gradients [@RN4914].
 
-Our results also show that germination is faster with increasing incubation temperature, especially in small seeded, endospermic species.
-This is not surprising, as alpine seed germination must be fast to match the short growing season.
-Small-seeded species have a lower rate of seeding survival under drought [@RN4530] but may require warmer germination temperatures than large-seeded species, especially in seasonal climates [@RN2399].
-Accordingly, this pattern was stronger in strict alpine plants which, being restricted to higher elevations, presumably experience less unpredictable drought but more frost risk, compared with generalist species: precipitation and the probability of early or late season frosts generally increase with elevation [@RN2392; @RN4897].
-Taken together, our results indicate that germination patterns in alpine species are driven by an interplay of seed size and germination temperature to escape either drought or frost stress: small seeds selecting for fast germination at warm temperatures in order to escape unfavourable early-spring frost and larger seeds selecting for slower germination at cooler conditions, which should maximize emergence during snowmelt (when water availability is highest) and a well-developed root system to cope with desiccation risk in summer.
-
-Finally, increasing temperatures also decreased germination uncertainty, showing that germination synchrony is a plastic trait driven by temperature, at least in alpine species.
+Increasing temperatures also decreased germination uncertainty, showing that germination synchrony is a plastic trait driven by temperature, at least in alpine species.
 Indeed, while asynchronous germination is thought to be a form of adaptation to unpredictable alpine environments [@RN3734; @RN3690], our results indicate that staggered germination occurs when temperatures are still cool, likely as a bet-edging strategy against the risk of early-spring frost.
@@
 staggering germination so that emergence occurs both before and after winter, is known in alpine species [@RN2392; @RN3285; @RN4907].
-Staggered germination is usually exhibited as a rapid first wave of germination followed by second wave several weeks later but can also take place in two seasons separated by a whole year [@RN2392].
+Staggered germination is usually exhibited as a rapid first wave of germination followed by second wave several weeks or months later [@RN2392].
 If autumn seedlings grow faster than spring seedlings [@RN4907], this will likely have cascading effects on individual fitness along the life of the plant [@RN3036; @RN2384].
-Delving further into the staggering germination syndrome, such as examining whether the source of variation lies within individuals, across individuals within a population, or across populations; or whether such syndrome is conserved; will help us to understand further the ecology of the unique alpine flora and its fate under a changing climate.
+Delving further into the staggering germination syndrome, such as examining whether the source of variation lies at the individual or population level; or whether such syndrome is conserved; will help us to understand further the ecology of the unique alpine flora and its fate under a changing climate.
 
@@
 
-We found a positive germination response to alternating temperatures and light, conditions that mimic the environment in the upper soil, during snowmelt at the end of winter [@RN4712; @RN2392; @RN3029].
-Triggering seed germination at the very beginning of the spring season ensures that seedlings have time to establish during the short growing seasons [@RN2392] and grow to a critical biomass before the next winter [@RN2868].
+We found a positive germination response to alternating temperatures and light, conditions that mimic the environment in the upper soil during the snow-free season [@RN4712; @RN2392; @RN3029].
 Alternating temperatures had a slightly lower effect on strict alpine species, compared with generalist species, but still had an effect.
@@
 When analysed at the global level, alternating temperatures seem to increase germination of alpine plants, which corresponds well with the fact that most alpine regions have a strong day-night climatic variation during the growing season.
-
-Solar irradiance in physiologically significant quantities penetrates only the first millimetres of the soil [@RN4909] and fluctuations of diurnal temperature decrease with increasing burial depths: below 10 cm they might be too small to trigger germination of species requiring temperature variation [@RN4910].
-Therefore, it is not surprising that we detected a negative effect of seed mass on final germination values in the light and under alternating temperature regimes.
-The depth of seed burial in the soil is crucial for seedling emergence [@RN4898], as seed mass  - or better said the seed kernel, i.e.
-embryo and endosperm [@RN4899] - may represent a constraint for seedling emergence of small-seeded species.
-Therefore, small seeds are more likely to require light and alternating temperatures for germination, which ensures that germination does not occur too deep in the soil for seedling emergence [@RN4904].
-
-## Seeds, embryos, and phylogenetic signal
-
-Seed mass is a relatively constant trait in our data, with no differences between strict alpine and generalist species.
-This is in line with the general ambiguity of relationships between seed mass and elevation.
-In alpine species, the evidence of a relationship between seed mass and elevation is mixed, being either negative [@RN4912], positive [@RN4903] or absent [@RN3685; @RN4896] when looking at the species level.
-Similarly, at the population level, the relationship between seed mass and elevation is often absent [@RN3442; @RN4896].
-The length of the growing season might influence seed size.
-For instance, at higher elevations where snow duration is longer and thus the growing season is shorter, the production of smaller seeds might be favoured [@RN4895; @RN4912].
-On the other hand, natural selection may also favour the production of larger seeds in species at higher elevations because larger seeds enable plants to cope better with stressful environments [@RN4913; @RN4903].
-
-Much less information is available concerning a possible selective advantage of embryo:endosperm size in alpine species.
-However, the similarity in embryo:endosperm size between generalist and strict alpine species confirms the absence of a correlation between embryo-seed size and elevation found previously in Apiaceae species [@RN3685].
-Although seed size and embryo:endosperm ratio are not different for generalist and strict alpine species in this study, these traits have been found to influence germination strategies in alpine species.
-For instance, previous studies have shown that non-endospermic seeds are smaller and are likely to be non-dormant at the time of dispersal and thus germinate quicker than alpine endospermic seeds [@RN3703; @RN3285; @RN4907].
-Perhaps accumulation of reserves in cotyledons rather than in the endosperm may improve early growth rate and establishment in immediate germinating species [@RN3285].
-
-A general assumption of seed trait ecology is that seed mass and embryo:endosperm ratios are relatively homogeneous across related species [@RN4916; @RN3685].
-In contrast, physiological responses to temperature and dormancy patterns are understood as more plastic traits that can quickly acclimatize in new situations [@RN2249].
-Although our study is not focused on trait evolution, we detected close relationships between seed size and physiological responses of seeds, suggesting that both traits are connected and may be subjected to co-adaptation.
-While phylogenetic signal in seed germination has been found often, especially in highly selective environments [@RN3001; @RN2399; @RN2865], seed germination patterns can also be shared between phylogenetically distant species [@RN4911] with low or absent phylogenetic signal [@RN2365; @RN4900; @RN2371].
-However, the phylogenetic signal found here indicates that evolutionary history cannot be neglected when studying seed germination patterns in the alpine environment.
-Further studies are required to assess whether the degree of the shared germination patterns among phylogenetically related species in the alpine environment is the result of trait conservatism, convergence, or both [@RN4901; @RN3235].
+It is also not surprising that we detected a negative interaction between seed mass and the germination response to light and under alternating temperature regimes, since large seeds can afford to germinate at greater depths (where they are more protected from the alpine environment) and still reach the soil surface [@RN4904].
 
@@
 
-## Conclusions
+A general assumption of seed trait ecology is that seed mass and embryo:endosperm ratios are relatively conserved across related species [@RN4916; @RN3685].
+In our alpine dataset, both traits were relatively constant, with no differences between strict alpine and generalist species.
+This is in line with the general ambiguity of relationships between seed mass and elevation that have been found for alpine species, including examples of negative [@RN4912], positive [@RN4903] or absent [@RN3685; @RN4896] relationships when looking at the species level.
+The similarity in embryo:endosperm size between generalist and strict alpine species confirms the absence of a correlation between embryo-seed size and elevation found in the Apiaceae [@RN3685].
+In contrast with the supposed phylogenetic conservatism of these morphological and anatomical traits, physiological responses to temperature and dormancy patterns are understood as more plastic traits that can quickly acclimatize to new situations [@RN2249].
+While phylogenetic signal in seed germination has been found, especially in highly selective environments [@RN3001; @RN2399; @RN2865], seed germination patterns can also be shared between phylogenetically distant species [@RN4911] and have low or absent phylogenetic signal [@RN2365; @RN4900; @RN2371].
+However, the phylogenetic signal found here indicates that evolutionary history cannot be neglected when studying seed germination patterns.
+Furthermore, we detected close relationships between seed-embryo size and physiological responses of seeds, suggesting that both sets of traits are connected and may be subjected to co-adaptation.
+Previous studies have shown that non-endospermic seeds are smaller and are likely to be non-dormant at the time of dispersal and thus germinate quicker than alpine endospermic seeds [@RN3703; @RN3285; @RN4907].
+Taken together, these results indicate that seed morphology and physiology are connected aspects of the seed ecological spectrum and must be studied in the light of this connection and of their shared phylogenetic history.
 
-Our meta-analysis of primary data has shown that strict alpine species from different mountain regions of the world tend to show a specialized seed germination strategy which shows evidence of phylogenetic signal and is characterized by:
-
-* Physiological seed dormancy.
-* A strong need for cold stratification to break dormancy, which can be substituted by GA~3~. 
-* A positive response to light.
-* A positive response to alternating temperatures, although not so prominent as in generalist species.
-* Slow and relatively synchronous germination.
-* An interplay with seed and embryo size, with smaller and more endospermic seeds being more responsive to warmth, light, and alternating temperatures.
+In conclusion, our meta-analysis of primary data has shown that strict alpine species from different mountain regions of the world tend to show a specialized seed germination strategy which shows evidence of phylogenetic signal and is characterized by (1) physiological seed dormancy; (2) a strong need for cold stratification or GA~3~ to break dormancy; (3) a positive response to light; (4) a positive response to alternating temperatures, although not so prominent as in generalist species; (5) slow and relatively synchronous germination, and (3) an interplay with seed and embryo size, with smaller and more endospermic seeds being more responsive to warmth, light, and alternating temperatures.
 
@@
 ```{r fig5, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Phylogenetic signal in the response to the germination environment.
-When lambda = 0, related taxa are no more similar than expected by chance; when λ = 1, the trait is evolving following a constant variance random walk or Brownian motion model", fig.align = "center", fig.width=7, fig.height=1.5}
+When lambda = 0, related taxa are no more similar than expected by chance; when λ = 1, the trait is evolving following a constant variance random walk or Brownian motion model.", fig.align = "center", fig.width=7, fig.height=1.5}
 cowplot::ggdraw(fig5)
@@
 Each square is a species.
-Labels indicate the contribution of the variables to the axes.", fig.align = "center", fig.width=7, fig.height=7}
+Labels indicate the contribution of the variables to the axes.
+Seed dormancy classes are abbreviated as follows: MD (morphological), MPD (morphophysiological), ND (non-dormant), PD (physiological), PY (physical).", fig.align = "center", fig.width=7, fig.height=7}
 cowplot::ggdraw(fig6)

### 2020-08-13 doc/ms/ms.Rmd
first full draft

@@
 ---
-title: "Seed germination ecology of alpine plants: a global meta-analysis of primary data"
+title: "Seed germination ecology of alpine plants: a global meta-analysis"
 output: 
@@
 ```{r message = FALSE, echo = FALSE, warning = FALSE}
-library(tidyverse); library(here)
-source(here("src", "msSource.R"))
+library(tidyverse); library(here); library(flextable)
+source(here("src", "msSource.R"), encoding = "UTF-8")
 ```
 
-Eduardo Fernández-Pascual^1^ ^+^, Elías Afif Khouri^1^, Joaquín Ascaso^2^, Beatriz Morais^1^, Ramón Reiné^2^, Madalena Vaz^3^
+Eduardo Fernández-Pascual^1^ ^+^, Angelino Carta^2^, Andrea Mondoni^3^, Lohengrin Cavieres^4^, Sergey Rosbakh^5^, Susanna Venn^6^, Annisa Satyanti^7^, Lydia Guja^8^, Verónica Briceño^7^, Filip Vandelook^9^, Efisio Mattana^10^, Arne Saatkamp^11^, Haiyan Bu^12^, Karen Sommerville^13^, Peter Poschlod^14^, Kun Liu^12^, Adrienne Nicotra^7^, Borja Jiménez-Alfaro^15^
 
-^1^ Departamento de Biología de Organismos y Sistemas, Universidad de Oviedo, Oviedo/Uviéu, Spain; ^2^ Departamento de Ciencias Agrarias y Medio Natural, Universidad de Zaragoza, Huesca, Spain; ^3^ Banco Português de Germoplasma Vegetal, Instituto Nacional de Investigação Agrária e Veterinária (INIAV), Braga, Portugal
+^1^ Departamento de Biología de Organismos y Sistemas, Universidad de Oviedo, Oviedo/Uviéu, Spain; ^2^ Dipartimento di Biologia, Botany Unit, University of Pisa, Pisa, Italy; ^3^ Department of Earth and Environmental Sciences, University of Pavia, Pavia, Italy; ^4^ Departamento de Botánica, Facultad de Ciencias Naturales y Oceanográficas, Universidad de Concepción, Concepción, Chile; ^5^ Chair of Ecology and Conservation Biology, University of Regensburg, Regensburg, Germany; ^6^ Centre for Integrative Ecology, School of Life and Environmental Sciences, Deakin University, Burwood, Victoria, Australia; ^7^ Division of Ecology & Evolution, Research School of Biology, The Australian National University, Canberra, ACT, Australia; ^8^ Centre for Australian National Biodiversity Research, CSIRO, Canberra, ACT, Australia; ^9^ Meise Botanic Garden, Meise, Belgium; ^10^ Natural Capital and Plant Health Department, Royal Botanic Gardens, Kew, Ardingly, UK; ^11^ Aix Marseille Université, Université d'Avignon, CNRS, IRD, IMBE, Marseille, France; ^12^ State Key Laboratory of Grassland and Agro-Ecosystems, Lanzhou University, Lanzhou, PR China; ^13^ The Australian Plant Bank, Royal Botanic Gardens and Domain Trust, Mount Annan, NSW, Australia; ^14^ Department of Ecology and Conservation Biology, Institute of Plant Sciences, University of Regensburg, Regensburg, Germany; ^15^ Research Unit of Biodiversity (CSUC/UO/PA), University of Oviedo, Mieres, Spain
 
@@
 
-# Summary (200 words)
+*Total word count for the main body of the text (Introduction, Materials and Methods, Results, Discussion), word counts for each section plus the number of figures (indicating which figures should be published in colour), tables and supporting information must also be stated on the front page.*
 
-# Keywords (8)
+# Summary
+
+* We present a meta-analysis of the alpine seed ecological spectrum based on primary data from germination experiments conducted in four continents with `r MSspecies` plant species, with the aim of testing the generality of alpine germination patterns at the global level.
+* We use Markov Chain Monte Carlo Generalized Linear Mixed Models (MCMCglmm) to model the influence of six environmental cues on germination proportion, mean germination time and the uncertainty of germination; controlling the effects of seed mass, embryo:endosperm ratio and phylogeny.
+* Alpine plants are characterized by physiological seed dormancy; strong need for cold stratification; warm-cued germination; and positive germination responses to light and alternating temperatures.
+Strict alpine species have a higher preference for warm temperature and a stronger response to cold stratification than generalist species, and their germination is slower and more synchronous.
+Seed mass, embryo size and phylogeny have strong effects on germination responses.
+* Physiological dormancy and overwintering are key drivers of germination in alpine habitats.
+The interplay of warm temperatures and seed size further explains mechanisms to avoid frost or drought stress.
+These results support the generality, at the global level, of germination patterns expected for alpine species.
+
+# Keywords
+
+alpine; alternating temperature; cold stratification; embryo endosperm ratio; light germination; seed dormancy; seed germination; seed mass
 
@@
 These treeless habitats are characterized by low temperatures, strong winds, unstable substrates, and short growing seasons [@RN2392].
-Although some alpine plants may reproduce clonally, sexual reproduction is the main strategy to maintain genetic diversity and to colonize suitable habitats in response to environmental changes.
-The short reproductive season of alpine habitats constrains the phenological timing of flowering, mating, seed development and seed dispersal.
-Further, dispersed seeds will need to germinate in the most appropriate period to ensure survival of seedlings under extreme environmental stress [@RN4714; @RN2868; @RN4717].
+Although some alpine plants may reproduce clonally, sexual reproduction is the main strategy to maintain genetic diversity and to colonize suitable habitats by seeds in response to environmental changes.
+However, the short reproductive season of alpine habitats constrains the phenological timing of flowering, mating, seed development and seed dispersal.
+Regeneration by seeds requires dispersed propagules to germinate in the most appropriate period to ensure survival of seedlings under environmental stress [@RN4714; @RN2868; @RN4717].
 The physiological process of seed germination is therefore an essential life stage that must be timed to occur when the environment is favourable for subsequent seedling survival and growth [@RN4691].
@@
 
-Natural selection should favour seed germination requirements that reduce the probability of facing environmental conditions that are not appropriate for seedlings [@RN3375].
-The main physiological drivers of germination, water availability and temperature [@RN3368], are also the main ecological signals of climatic variation.
+Natural selection is expected to favour seed germination requirements by reducing the probability of facing environmental conditions which are not appropriate for seedlings [@RN3375].
+This is mostly regulated by climatic signals related to water availability and soil temperature as the two general drivers of seed germination [@RN3368].
 Many plant species have also developed seed dormancy, by which germination is prevented during periods that are only ephemerally favourable, postponing germination to the right season [@RN3214].
@@
 Other germination cues, such as the response to light [@RN4656] and alternating temperatures [@RN1380], allow for a fine-scale detection of germination micro-niches and safe sites [@RN4719].
-Given the heterogeneity of alpine climates and species lineages adapted to different regions, and the wealth of local studies on alpine seed ecology (reviewed in the following paragraphs), a current challenge is to synthesize, at the global level, the *seed germination spectrum* of alpine plants [@RN4655].
-Saatkamp et al.
-[-@RN4655] define the seed ecological spectrum as the set of trait-based ecological responses of seeds that determines their ability to disperse, persist, germinate, and establish.
-A global synthesis of the alpine germination spectrum would allow us to (i) better understand how climatic changes will affect alpine plant diversity; and (ii) manage and eventually counteract regression of restricted alpine species.
+Given the heterogeneity of alpine climates and species lineages adapted to different regions, and the wealth of local studies on alpine seed ecology (reviewed in the following paragraphs), a current challenge is to synthesize, at the global level, the seed ecological spectrum [@RN4655] of alpine plants.
+The seed ecological spectrum is a set of seed traits that determines the ability of plants to disperse, persist, germinate, and establish in different habitats.
 
@@
 This cold-cued germination could be an adaptation to germinate under snow or during snowmelt, which could presumably allow seedlings to develop a deeper root system before topsoil desiccation in summer [@RN4720], or to attain an optimal size for overwintering [@RN4712].
-The main question that remains to be answered is the generality of warm-cued germination across alpine micro-climatic niches.
+A major question that remains to be answered is the generality of warm-cued germination across different alpine regions and habitats.
 
-Many alpine species have also been described as having seeds with deep physiological dormancy [@RN2943; @RN3703; @RN3214].
-Much like warm-cued germination, dormancy in recently dispersed seeds would prevent precocious germination under autumn cold conditions, when appropriate soil moisture and temperature are not likely to persist for more than a few weeks or days [@RN3330].
+Many alpine species have been described as having seeds with deep physiological dormancy [@RN2943; @RN3703; @RN3214].
+Much like warm-cued germination, dormancy would prevent precocious germination of recently dispersed seeds under autumn cold conditions, when appropriate soil moisture and temperature are not likely to persist for more than a few weeks or days [@RN3330].
 A requirement for cold stratification to break dormancy would allow seeds to sense the snow season, thereby postponing germination to a better period for seedling survival and development.
 Thus, it can be expected that a cold stratification period is a common requirement for seed germination in alpine plants.
-Nonetheless, an early review [@RN3213] suggested that cold stratification was not a requirement for the seed germination of several alpine species from different mountains in the USA, an idea that was repeated in subsequent articles [@RN1484; @RN2995; @RN4721].
+Nonetheless, an early review [@RN3213] suggested that cold stratification was not a requirement for the seed germination of several alpine species from different mountains in the USA, an idea that was repeated in subsequent studies [@RN1484; @RN2995; @RN4721].
 More recently, research on Australian alpine plants found that a cold stratification period only increased germination in half of the species tested [@RN3703; @RN3285].
@@
 In a comparative study of species from different habitats of the Austrian Alps which assessed the germination of unstratified seeds, species from alpine grasslands had the lowest final germination proportion [@RN4930].
-Therefore, how important cold stratification is for seed germination in alpine plant species remains unclear.
-
-The underlying question is whether seed dormancy plays a relevant role in postponing alpine germination during autumn and winter, or the warm-cued germination is enough to prevent germination under cool autumn temperatures.
+The importance of cold stratification as a general pre-requisite for seed germination in alpine plants remains unclear.
+An underlying question is whether physiological seed dormancy plays a relevant role in postponing alpine germination during autumn and winter, or the warm-cued germination is enough to prevent germination under cool autumn temperatures.
 The response to this question becomes non-trivial under a warming climate, by which non-dormant seeds could germinate readily in autumn, and this would provide a longer time for seedlings to establish.
@@
 high germination uncertainty) should to be preferred.
+The question turns even more complex if we consider other types of dormancy that might also occur in alpine environments (e.g.
+physical dormancy), and the fact that dormancy patterns may vary across populations of the same species.
 
@@
 Here, we collect primary germination data from eight alpine regions and four continents to investigate the germination response of alpine plant species to key environmental factors.
-We used raw data obtained from laboratory experiments to test the response of seed germination to temperature, cold stratification, light and alternating temperatures, and their relationships with dormancy classes, seed mass, and embryo:endosperm ratio obtained from bibliographic data.
-Our central question was to investigate the seed germination spectrum of alpine plants to test the consistency at the global scale of general assumptions about their seed germination ecology: in short, that alpine seeds (1) have physiological dormancy and (2) require cold stratification to break dormancy, (3) need relatively warm temperatures to germinate, (4) response positively to light and (5) do not require alternating temperatures for germination.
-To understand the ecological constraints of our meta-analysis on alpine species pools, we further analyse how the germination responses differ between strict alpine species which occur exclusively above the treeline, and generalist species that also occur at lower elevations.
+We collated raw data obtained from laboratory experiments to test the general response of alpine seed germination to temperature, cold stratification, light and alternating temperatures, and their relationships with dormancy classes, seed mass, and embryo:endosperm ratio obtained from bibliographic data.
+Our main aim was to test the consistency of general assumptions about the influence of major environmental cues on seed germination of alpine plants: in short, we investigate whether, at the global scale, alpine seeds (1) have physiological dormancy and (2) require cold stratification to break dormancy, (3) need relatively warm temperatures to germinate, (4) response positively to light and (5) do not require alternating temperatures for germination.
+To understand the ecological constraints of our meta-analysis on alpine species pools, we analyse how the germination responses differ between strict alpine species which occur exclusively above the treeline, and generalist species that also occur at lower elevations.
 
@@
 
-We conducted a meta-analysis of primary data [@RN4734], which consists in pooling together original data from different local studies to synthetize their conclusions at the global level.
+We conducted a meta-analysis of primary data [@RN4734], which consists of pooling together original data from different local studies to synthetize their conclusions at the global level.
 The original data files, as well as R code for data cleaning, analysis and manuscript production are available at https://github.com/efernandezpascual/alpineseeds.
@@
 The submitted data had to report the studied species, the germination treatments, the number of seeds sown, and the number of seeds germinated.
-We collected data from the Cantabrian Mountains of Spain [@RN2371], the European Alps [@RN4729; @RN3029; @RN4727; Rosbakh, unpublished], the Italian Apennines (Mondoni, unpublished), the Russian Caucasus [Rosbakh, unpublished], the Qinghai-Tibet Plateau of China [@RN4728; @RN4735; @RN3376], the Chilean Andes [@RN3393; @RN4713; Briceño, unpublished] and the Australian Alps [@RN4737; @RN4736; @RN3703; Satyanti, unpublished].
-We also collected data on European alpine species from *Enscobase*, the database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml).
+We received data (Table 1) from the Cantabrian Mountains of Spain, the European Alps, the Italian Apennines, the Russian Caucasus, the Qinghai-Tibet Plateau of China, the Chilean Andes, and the Australian Alps.
+We also received data on European alpine species from *Enscobase*, the database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml).
+
+---
+nocite: |
+  @RN2371, @RN4729, @RN3029, @RN4727, @RN4728, @RN4735, @RN3376, @RN3393, @RN4713, @RN4737, @RN4736, @RN3703
+...
 
@@
 species that can grow above the treeline but also in the lowlands).
-For each species, we also compiled the following traits: (1) life form (forb, graminoid or woody) and (2) life span (annual or perennial) from the references used for the elevation, plus some extra references [@RN4747; @RN4744; @RN4745; @RN4746; @RN4748; @RN4749] and online resources (https://www.infoflora.ch/de/; http://plantnet.rbgsyd.nsw.gov.au/search/simple.htm; http://buscador.floraargentina.edu.ar/; https://patagoniawildflowers.org/); (3) seed dormancy class following the classification system of Baskin and Baskin: physiological (germination prevented by the seed’s internal balance of hormones), morphological (germination prevented by an underdeveloped embryo at the time of dispersal), morphophysiological (both physiological and morphological dormancy present), physical (germination prevented by a water-impermeable seed coat) or non-dormant [@RN3214; @RN4743]; (4) seed mass [@RN4728; @RN4735; @RN3376; @RN2987] and (5) embryo to endosperm ratio (Vandelook, unpublished).
+For each species, we compiled (1) life form (forb, graminoid or woody) and (2) life span (annual or perennial) from the references used for the elevation, plus some extra references [@RN4747; @RN4744; @RN4745; @RN4746; @RN4748; @RN4749] and online resources (https://www.infoflora.ch/de/; http://plantnet.rbgsyd.nsw.gov.au/search/simple.htm; http://buscador.floraargentina.edu.ar/; https://patagoniawildflowers.org/); (3) seed dormancy class following the classification of Baskin and Baskin: physiological (germination prevented by the seed’s internal balance of hormones), morphological (germination prevented by an underdeveloped embryo at the time of dispersal), morphophysiological (both physiological and morphological dormancy present), physical (germination prevented by a water-impermeable seed coat) or non-dormant [@RN3214; @RN4743]; (4) seed mass [@RN4728; @RN4735; @RN3376; @RN2987] and (5) embryo to endosperm ratio (Vandelook, unpublished).
 For the species missing seed mass or embryo values, we calculated genus or family averages.
@@
 germination proportions for a given seed lot of a species, recorded in a set of experimental conditions) contributed by `r MSgroups` research groups, representing `r MSfamilies` seed plant families and `r MSspecies` species (`r MSstrict` strict alpine species and `r MSgeneral` generalists).
-Seed lots were originally sampled in `r MSregions` alpine regions of the world (Fig.
-\@ref(fig:fig1)), in both hemispheres.
+Seed lots were originally sampled in `r MSregions` alpine regions of the world (Table 1), in both hemispheres.
 
-Most species in the dataset (> 90 %) were perennial, and either graminoids or forbs.
-Most were reported to have dormant seeds (Fig.
-\@ref(fig:fig2)a), but the frequency of the dormancy classes did not differ between generalist and strict alpine species.
+Most species in the dataset (> 90 %) were perennial graminoids or forbs.
+They were mostly reported to have dormant seeds (Fig.
+\@ref(fig:fig1)a), but the frequency of the dormancy classes did not differ between generalist and strict alpine species.
 Among those species with dormant seeds, physiological dormancy was the most common dormancy class, followed by morpho-physiological and physical.
 Both strict alpine and generalist species had a similar range of values for seed mass (Fig.
-\@ref(fig:fig2)b) and embryo to endosperm ratio (Fig.
-\@ref(fig:fig2)c).
+\@ref(fig:fig1)b) and embryo to endosperm ratio (Fig.
+\@ref(fig:fig1)c).
 The values of seed mass ranged from `r MSminMass` to `r MSmaxMass` mg, with a median value of `r MSmedianMass` mg.
@@
 Scarification (i.e.
-an abrassion of the seed coat to allow water imbibition) was performed in `r MSscarY` records, and GA~3~ (a plant hormone that promotes embryo growth and germination) in `r MSga3` records.
+an abrasion of the seed coat to allow water imbibition) was performed in `r MSscarY` records, and GA~3~ (a plant hormone that promotes embryo growth and germination) in `r MSga3` records.
 Since germination incubations ran for different periods of time, we established a cut-off time of 8 weeks to calculate the final germination percentages, to have a homogeneous comparison across records.
@@
 We further calculated the mean germination time and the germination uncertainty index with the package *GerminaR* [@RN4752].
-Mean germination time is the time that it takes for half of the seed lot to germinate, indicating the germination speed or rate in a given experimental condition.
+Mean germination time is the time that it takes for half of the seed lot to germinate, indicating the germination speed or rate in each experimental condition.
 The germination uncertainty index estimates how scattered germination is through time, and gives an idea of whether germination is synchronous (i.e.
@@
 We established these limits to prevent the results being altered by a modification of seed dormancy status during the germination incubation (e.g.
-a fraction of the seed lot is non-dormant at the beginin of the incubation and germinates shortly after the start of the experiment; while another fraction is originally dormant, loses dormancy in response to the incubation conditions, and germinates in a second wave).
+a fraction of the seed lot is non-dormant at the beginning of the incubation and germinates shortly after the start of the experiment; while another fraction is originally dormant, loses dormancy in response to the incubation conditions, and germinates in a second wave).
 
@@
 Models had, as fixed effects, the experimental conditions (temperature, alternating temperature, light, scarification, stratification and GA~3~), plus their interaction with seed mass and embryo:endosperm, and the strict alpine or generalist character of the species.
-Random effects included a reconstructed phylogenetic tree for the `r MSspecies`, and also species identity, seed lot, lab, and alpine region.
+Random effects included a reconstructed phylogenetic tree for the `r MSspecies`, and species identity, seed lot, lab (i.e.
+data source), and alpine region (Table 1).
 To create the phylogeny, we used the R package *V.PhyloMaker* [@RN4753].
@@
 
-To visualize the alpine seed germination spectrum, we performed a Factorial Analysis of Mixed Data (FAMD) as implemented in the package *FactoMineR* [@RN3166].
+To visualize the alpine seed ecological spectrum, we performed a Factorial Analysis of Mixed Data (FAMD) as implemented in the package *FactoMineR* [@RN3166].
 FAMD ordination combines the properties of PCA and MCA to jointly measure the variation of continuous and categorical variables.
-The ordination was performed at the species level, i.e.
+The ordination was performed for all traits available in this study at the species level, i.e.
 calculating a series of continuous and categorical traits for each species.
-We created a continuous variable for each germination cue (i.e.
+We transformed the final germination proportions to create a continuous variable for each germination cue (i.e.
 temperature, alternating temperatures, light, scarification, stratification, and GA~3~).
@@
 This approach underrepresents the importance of the levels that were not tested for a given species, but can serve as a proxy of the response to the germination cues when visualized across the whole dataset; it must be stressed that this stage of the analysis serves only for visualization and not hypothesis testing.
-For each species, we also calculated a single mean germination time (the minimum, i.e.
-the time taken at the most favourable treatment) and germination uncertainty index (the mean).
+We also included germination timing traits by calculating single values for the mean germination time (the minimum, i.e.
+the time taken at the most favourable treatment) and the germination uncertainty index (the mean).
 We conducted the FAMD with only the `r MSfamdspp` species for which we had mean germination time and germination uncertainty index values.
-We also included the continuous variables seed mass and embryo:endosperm, as well as the categorical variables dormancy class (physiological, morphophysiological, morphological, physical or non-dormant), distribution (strict alpine vs.
+We finally included the continuous variables seed mass and embryo:endosperm, as well as the categorical variables dormancy class (physiological, morphophysiological, morphological, physical or non-dormant), distribution (strict alpine vs.
 generalist species), life form (forb, graminoid or woody) and life span (annual or perennial).
@@
 Temperature had a positive effect on the final germination proportion as shown by a credible interval not overlapping zero (Fig.
-\@ref(fig:fig3)a), indicating a preference for warmer temperatures.
+\@ref(fig:fig2)a), indicating a preference for warmer temperatures.
 Increasing temperatures also decreased mean germination time (Fig.
-\@ref(fig:fig3)b) and the germination uncertainty index (Fig.
-\@ref(fig:fig3)c).
+\@ref(fig:fig2)b) and the germination uncertainty index (Fig.
+\@ref(fig:fig2)c).
 The effect of warmer temperatures was slightly more intense in strict alpine than in generalist species, although the credible intervals of the two groups of species partially overlap.
-Seed mass and embryo:endosperm had a negative interaction with temperature, in other words, they reduced the response to temperature: heavier and less endospermic seeds increased less their germination proportion and decreased less their mean germination time in warmer temperatures.
+Seed mass and embryo:endosperm had a negative interaction with temperature, reducing the response to temperature.
+In other words, heavier and less endospermic seeds increased less their germination proportion and decreased less their mean germination time in warmer temperatures.
 The germination uncertainty index did not show an interaction with seed mass or embryo:endosperm, as shown by the credible interval overlapping zero.
@@
 Alternating temperatures had a positive effect on the final germination proportion (Fig.
-\@ref(fig:fig3)a), but not in the mean germination time (Fig.
-\@ref(fig:fig3)b) or the germination uncertainty index (Fig.
-\@ref(fig:fig3)c).
-The positive effect on the germination proportion was marginally higher in the generalist species.
+\@ref(fig:fig2)a), but not in the mean germination time (Fig.
+\@ref(fig:fig2)b) or the germination uncertainty index (Fig.
+\@ref(fig:fig2)c).
+The positive effect on the germination proportion was slightly higher in the generalist species.
 As in the case of the mean germination temperature, higher seed mass and embryo:endosperm values reduced the positive effect of alternating temperatures on the final germination proportion.
@@
 Light increased the final germination proportion (Fig.
-\@ref(fig:fig3)a), with no differences between strict alpine and generalist species.
+\@ref(fig:fig2)a), with no differences between strict alpine and generalist species.
 It did not affect the mean germination time (Fig.
-\@ref(fig:fig3)b), but it did increase the germination uncertainty index of generalist species (Fig.
-\@ref(fig:fig3)c).
+\@ref(fig:fig2)b), but it did increase the germination uncertainty index of generalist species (Fig.
+\@ref(fig:fig2)c).
 Heavier and less endospermic seeds responded less to light.
@@
 Scarification promoted the germination proportion (Fig.
-\@ref(fig:fig3)a) and reduced the mean germination time (Fig.
-\@ref(fig:fig3)b), but it did not affect the germination uncertainty index (Fig.
-\@ref(fig:fig3)c).
+\@ref(fig:fig2)a) and reduced the mean germination time (Fig.
+\@ref(fig:fig2)b), but it did not affect the germination uncertainty index (Fig.
+\@ref(fig:fig2)c).
 Strict alpine species were slightly more responsive to scarification than generalist species.
@@
 In general, stratification increased the final germination proportion (Fig.
-\@ref(fig:fig3)a).
+\@ref(fig:fig2)a).
 This positive effect occurred in both generalist and strict alpine species, although the effect was higher in the strict alpine group.
 Stratification slightly reduced the mean germination time (Fig.
-\@ref(fig:fig3)b), which was evident in the generalist species, but not in the strict alpine species.
+\@ref(fig:fig2)b), which was evident in the generalist species, but not in the strict alpine species.
 Further, stratification reduced the germination uncertainty index as a general response (Fig.
-\@ref(fig:fig3)c), and slightly more so in generalist species.
+\@ref(fig:fig2)c), and slightly more so in generalist species.
 Seed mass did not interact with stratification, but species with smaller embryos had a smaller response to stratification in their final germination proportion and reduced less markedly their mean germination time and germination uncertainty index.
@@
 Random factors had a significant effect on the MCMCglmm models (Fig.
-\@ref(fig:fig4)).
+\@ref(fig:fig3)).
 The final germination proportion and the mean germination time were affected more strongly, on average, by the seed lot and the species identity.
-Alpine region and lab had a lower average effect, but more variability as shown by their credible intervals.
+Alpine region and lab had a lower average effect and more variability as shown by their credible intervals.
 The germination uncertainty index showed the opposite pattern, with a stronger effect of lab and alpine region.
 The phylogenetic signal in the response to all germination parameters (Fig.
-\@ref(fig:fig5)) was positive for the final germination proportion and the mean germination time.
+\@ref(fig:fig4)) was positive for the final germination proportion and the mean germination time.
 For the germination uncertainty index, it overlapped with zero, indicating a weak effect of phylogeny.
@@
 In summary, the horizontal axis separated (i, left) generalist with a positive response to alternating temperatures and more germination uncertainty from (ii, right) strict alpine species with a preference for warmer temperatures and a positive response to GA~3~ (Fig.
-\@ref(fig:fig6)).
+\@ref(fig:fig5)).
 Axis 2 explained `r round(pcaAlpine$eig[2, 2], 0)`% of the variability.
@@
 
-Our dataset indicates that seed dormancy is frequent across many alpine regions of the world and postponing germination plays an important role in plant recruitment in the harsh alpine environment, although the frequency of dormant species is similar in strict alpine and generalist species.
-Seeds with the two most frequent dormancy classes, physiological and morphophysiological, require a stratification period during which they experience cold and wet conditions over a period of months [@RN3214; @RN4743].
-Our results show that cold stratification decreases time to germination and increases overall percent germination.
-Our results also indicate that the hormone gibberellic acid (GA~3~) can be used as a substitute of cold stratification to overcome inherent physiological dormancy.
-Physical dormancy, infrequent in our dataset, is broken by mechanical scarification, a process that could also occur naturally via freezing-thawing cycles in spring.
-Our results thus confirm that overwintering is essential for promoting seed germination in species inhabiting alpine habitats, especially for strict alpine species.
-This concurs with several studies that have shown that cold-stratification is important for seed germination in different alpine regions of the world [@RN3393; @RN2868; @RN707; @RN698; @RN3703; @RN2355; @RN3285; @RN2371; @RN4713].
-Under natural field conditions, cold stratification occurs over winter when seeds are covered by snow, allowing seeds to sense the snow season, and thereby promoting germination in the spring and summer, when conditions are more favourable for seedling survival and growth.
+Our dataset indicates that seed dormancy is frequent across many alpine regions of the world and postponing germination plays an important role in plant recruitment in alpine habitats.
+Nevertheless, the frequency of dormant species is similar in strict alpine and generalist species, suggesting that seed dormancy is a general trait in cold-adapted mountain species, either below or above the climatic treeline.
+Seeds with the two most frequent dormancy classes, physiological and morphophysiological, require a stratification period with cold and wet conditions over a period of months [@RN3214; @RN4743].
+Indeed, our results show that cold stratification decreases time to germination and increases overall percent germination.
+We also found that the hormone gibberellic acid (GA~3~) can be used as a general substitute of cold stratification to overcome the inherent physiological dormancy of alpine species.
+Finally, we found a few species with physical dormancy.
+This type of dormancy is broken by mechanical scarification, a process that could also occur naturally via freezing-thawing cycles in spring.
+Our results confirm that overwintering is essential for promoting seed germination in species inhabiting alpine habitats, especially for strict alpine species.
+This concurs with many studies showing that cold-stratification is important for seed germination in different alpine regions [@RN3393; @RN2868; @RN707; @RN698; @RN3703; @RN2355; @RN3285; @RN2371; @RN4713].
 Since global climate change is causing major changes in snow cover duration [@RN4917; @RN4918], it could disrupt natural cold stratification cycles in alpine environments, compromising population viability and indirectly favouring species with no stratification requirement [@RN3703].
-This risk will be partly alleviated by the know plasticity of physiological seed dormancy in response to the seed maturation temperature, a mechanism of “seed memory” that facilitates acclimatisation to changing environments [@RN2249], and which could reduce the stratification requirements of seeds produced in a warmer climate.
+This risk will be partly alleviated by the plasticity of physiological seed dormancy in response to the seed maturation temperature, a mechanism of “seed memory” that facilitates acclimatisation to changing environments [@RN2249], and which could reduce the stratification requirements of seeds produced in a warmer climate.
 
@@
 
-Our results show an increase of seed germination with incubation temperature, more pronounced in strict alpine compared to generalist species, confirming the findings of local studies [@RN3266; @RN2371].
-Our results also show that germination is faster with increasing incubation temperature, especially in small seeded, endospermic species.
+We found a general increase of seed germination with incubation temperature.
+This effect was more pronounced in strict alpine compared to generalist species, confirming the findings of local studies [@RN3266; @RN2371].
+It was also more pronounced in small seeded and endospermic species.
 Small-seeded species have a lower rate of seeding survival under drought [@RN4530] but may require warmer germination temperatures than large-seeded species, especially in seasonal climates [@RN2399].
-Considering that risk reduction is an important selective pressure for the evolution of germination traits [@RN4915], a plausible explanation for these patterns is that alpine plants received selection pressure from damaging spring frost and evolved an avoidance mechanism by germinating at high temperatures.
-Frost avoidance is an important survival strategy in alpine plants [@RN4908; @RN4906; @RN4722].
-Complementarily, strict alpine species might have been less subjected to drought damage during establishment compared to generalist species: as elevation increases, plants face less drought risk because precipitation increases and evapotranspiration decreases, but at the same there is a higher probability of early or late season frosts [@RN4897; @RN2392; @RN4914].
-Taken together, our results indicate that germination patterns in alpine species are driven by an interplay of seed size and germination temperature to escape either drought or frost stress: small seeds selecting for fast germination at warm temperatures in order to escape unfavourable early-spring frost; and large seeds selecting for slower germination at cooler conditions, which should maximize emergence during snowmelt (when water availability is highest) and a well-developed root system to cope with desiccation risk in summer.
+Considering that risk reduction is an important selective pressure for the evolution of germination traits [@RN4915], a plausible explanation for these patterns is that alpine plants received selection pressure from damaging spring frost and evolved a frost avoidance mechanism by germinating at high temperatures [@RN4908; @RN4906; @RN4722].
+Complementarily, strict alpine species might have been less subjected to drought damage during establishment compared to generalist species: as elevation increases, plants face less drought risk because precipitation increases (except in Mediterranean and semi-arid climates), but at the same time there is higher probability of early or late season frosts [@RN4897; @RN2392; @RN4914].
+Taken together, these results indicate that germination patterns in alpine species are driven by an interplay of seed size and germination temperature to escape either frost or drought stress: small seeds selecting for fast germination at warm temperatures in order to escape unfavourable early-spring frost; and large seeds selecting for slower germination at cooler conditions, which should maximize emergence during snowmelt (when water availability is highest) and a well-developed root system to cope with desiccation risk in summer.
 Supporting this view, seedlings of alpine pioneer species have a low frost resistance [@RN4722], yet high heat tolerance up to 40–50°C [@RN4902].
 
-Increasing temperatures also decreased germination uncertainty, showing that germination synchrony is a plastic trait driven by temperature, at least in alpine species.
-Indeed, while asynchronous germination is thought to be a form of adaptation to unpredictable alpine environments [@RN3734; @RN3690], our results indicate that staggered germination occurs when temperatures are still cool, likely as a bet-edging strategy against the risk of early-spring frost.
+Increasing temperatures also decreased germination uncertainty, showing that germination synchrony is a plastic trait driven by temperature.
+While asynchronous germination is thought to be a form of adaptation to unpredictable alpine environments [@RN3734; @RN3690], our results indicate that staggered germination occurs when temperatures are still cool, likely as a bet-edging strategy against the risk of early-spring frost.
 On the other hand, a fast and synchronised germination at warm conditions (i.e.
-in late spring) may be another important strategy that increases alpine seedling survival by avoiding summer drought.
+in late spring) may be another important strategy that increases alpine seedling survival by avoiding drier conditions in summer.
 The extreme germination uncertainty syndrome, i.e.
@@
 If autumn seedlings grow faster than spring seedlings [@RN4907], this will likely have cascading effects on individual fitness along the life of the plant [@RN3036; @RN2384].
-Delving further into the staggering germination syndrome, such as examining whether the source of variation lies at the individual or population level; or whether such syndrome is conserved; will help us to understand further the ecology of the unique alpine flora and its fate under a changing climate.
+Delving further into the staggering germination syndrome, e.g.
+by examining whether the source of variation lies at the individual or population level, will help us to understand the ecology of alpine flora and its fate under climate change.
 
@@
 
-## The alpine seed germination spectrum
+## The alpine seed ecological spectrum
 
-The factorial analysis of mixed data (FAMD) separated generalist and strict alpine plants according to a “fast-slow” gradient of regeneration strategies.
-On one end of the gradient, strict alpine species tend to have slow and synchronous germination, require warm temperatures and light to germinate, and need stratification or GA~3~ to break physiological dormancy.
-On the other end, generalist species tend to have fast germination, show a bet-hedging strategy (high germination uncertainty), and show germination traits suggesting possible fast germination after specific cues (PY, alternating temperatures).
+Using the whole set of traits compiled in this study, we visualize how the variation in seed germination links to other sources of ecological variation across alpine plants.
+On one end of the gradient, strict alpine species tend to have slow and synchronous germination, requiring warm temperatures and light to germinate, and needing stratification or GA~3~ to break physiological dormancy.
+On the other end, generalist species tend to have fast germination, showing a bet-hedging strategy (high germination uncertainty), and germination traits suggesting possible fast germination after specific cues (PY, alternating temperatures).
 The dormancy classes without a physiological component (non-dormant, physical, and morphological) appear associated with generalist species.
-While most species are perennial, the generalist group includes some annuals.
+While most species are perennial, the generalist group also includes some annuals.
 Life forms are evenly distributed across regeneration strategies.
-This “fast-slow” gradient of regenerative strategies is similar to prominent gradients of fast or slow population dynamics [@RN4922; @RN4921], fast or slow developing leaf traits [@RN4923], or large sets of plant ecological features [@RN2279].
+This gradient of regenerative strategies is similar to prominent gradients of fast or slow population dynamics [@RN4922; @RN4921], fast or slow developing leaf traits [@RN4923], or large sets of plant ecological features [@RN2279].
 “Fast” regeneration can be understood as an opportunistic strategy to cope with frequently disturbed habitats such as avalanche ways, steep eroding slopes, regressing glaciers, and riverbeds [@RN4926; @RN4925].
-The “slow” regeneration of strict alpine species, on the other hand, seems to be a more specialized syndrome of those species that are truly restricted - and adapted - to the alpine belt.
+The “slow” regeneration of strict alpine species, on the other hand, seems to be a more specialized syndrome of those species that are adapted to the alpine belt.
 
 A general assumption of seed trait ecology is that seed mass and embryo:endosperm ratios are relatively conserved across related species [@RN4916; @RN3685].
-In our alpine dataset, both traits were relatively constant, with no differences between strict alpine and generalist species.
-This is in line with the general ambiguity of relationships between seed mass and elevation that have been found for alpine species, including examples of negative [@RN4912], positive [@RN4903] or absent [@RN3685; @RN4896] relationships when looking at the species level.
+In our alpine dataset, both traits showed no differences between strict alpine and generalist species.
+This is in line with the ambiguity of relationships between seed mass and elevation that have been found for alpine species, including examples of negative [@RN4912], positive [@RN4903] or absent [@RN3685; @RN4896] relationships.
 The similarity in embryo:endosperm size between generalist and strict alpine species confirms the absence of a correlation between embryo-seed size and elevation found in the Apiaceae [@RN3685].
 In contrast with the supposed phylogenetic conservatism of these morphological and anatomical traits, physiological responses to temperature and dormancy patterns are understood as more plastic traits that can quickly acclimatize to new situations [@RN2249].
-While phylogenetic signal in seed germination has been found, especially in highly selective environments [@RN3001; @RN2399; @RN2865], seed germination patterns can also be shared between phylogenetically distant species [@RN4911] and have low or absent phylogenetic signal [@RN2365; @RN4900; @RN2371].
+While phylogenetic signal in seed germination is generally found, especially in highly selective environments [@RN3001; @RN2399; @RN2865], seed germination patterns can also be shared between phylogenetically distant species [@RN4911] and have low or absent phylogenetic signal [@RN2365; @RN4900; @RN2371].
 However, the phylogenetic signal found here indicates that evolutionary history cannot be neglected when studying seed germination patterns.
@@
 
-In conclusion, our meta-analysis of primary data has shown that strict alpine species from different mountain regions of the world tend to show a specialized seed germination strategy which shows evidence of phylogenetic signal and is characterized by (1) physiological seed dormancy; (2) a strong need for cold stratification or GA~3~ to break dormancy; (3) a positive response to light; (4) a positive response to alternating temperatures, although not so prominent as in generalist species; (5) slow and relatively synchronous germination, and (3) an interplay with seed and embryo size, with smaller and more endospermic seeds being more responsive to warmth, light, and alternating temperatures.
+## Conclusions
+
+Our meta-analysis of primary germination data shows that strict alpine species from different mountain regions of the world tend to show a specialized seed germination strategy which is characterized by (1) physiological seed dormancy; (2) a strong need for cold stratification or GA~3~ to break dormancy; (3) a positive response to light; (4) a positive response to alternating temperatures, although not so prominent as in generalist species; (5) slow and relatively synchronous germination, and (6) an interplay with seed and embryo size, with smaller and more endospermic seeds being more responsive to warmth, light, and alternating temperatures.
+These results support the generality of germination patterns expected for alpine species, despite habitat-specific differences that may occur across regions.
+Our study also provides an assessment of the seed ecological spectrum in relation with other plant traits of alpine plants, suggesting potential relationships between regenerative and adult plant traits that will need further research.
+New studies also will need to consider phylogenetic signal as an intrinsic constraint of seed trait variation in alpine habitats.
+Overall, this study provides a global assessment of alpine seed germination that can be used as a reference in further studies focused on spatial or temporal variation on seed germination, their links with other plant traits, and the expected responses of alpine plants to climate change.
 
@@
 # Author Contribution
+
+BJA conceived the original idea.
+EFP, AC, AM, LC, SR, SV, AnS, LG, VB, FV, HB, KS, PP contributed data.
+EFP and AC analysed data.
+EFP led manuscript writing, with contributions from BJA, AC, AM, LC, SR, SV, AnS, LG, VB, FV, EM, ArS.
+All authors revised the manuscript and approved the final version.
 
@@
 
-```{r fig1, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Geographical origin of the germination data with the number of records, species and strict alpine species per alpine region.
-Ensconet is not plotted because its records come from alpine regions throughout Europe.
-The purple areas correspond to the global extension of alpine environments according to Testolin *et al.* [-@RN4750].", fig.align = "center", fig.width=7, fig.height=3}
-cowplot::ggdraw(fig1)
-```
-
-```{r fig2, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Seed dormancy classes, seed mass and embryo to endosperm ratio in the subsets of strict alpine and generalist species analysed in this study.
+```{r fig1, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Seed dormancy classes, seed mass and embryo to endosperm ratio in the subsets of strict alpine and generalist species analysed in this study.
 Seed dormancy classes are abbreviated as follows: MD (morphological), MPD (morphophysiological), ND (non-dormant), PD (physiological), PY (physical).", fig.align = "center", fig.width=7, fig.height=3}
@@
 
-```{r fig3, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the germination environment on germination, according to the MCMC meta-analysis of the primary data.
+```{r fig2, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the germination environment on germination, according to the MCMC meta-analysis of the primary data.
 The dots indicate the posterior mean of the effect size, and the brackets its 95 % credible interval.
@@
 
-```{r fig4, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the random factors on germination, according to the MCMC meta-analysis of the primary data.
+```{r fig3, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the random factors on germination, according to the MCMC meta-analysis of the primary data.
 The dots indicate the posterior mean of the effect size, and the brackets its 95 % credible interval.
@@
 
-```{r fig5, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Phylogenetic signal in the response to the germination environment.
-When lambda = 0, related taxa are no more similar than expected by chance; when λ = 1, the trait is evolving following a constant variance random walk or Brownian motion model.", fig.align = "center", fig.width=7, fig.height=1.5}
+```{r fig4, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Phylogenetic signal in the response to the germination environment.
+When lambda = 0, related taxa are no more similar than expected by chance; when lambda = 1, the trait is evolving following a constant variance random walk or Brownian motion model.", fig.align = "center", fig.width=7, fig.height=1.5}
 cowplot::ggdraw(fig5)
@@
 
-```{r fig6, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "FADM ordination of the seed variables and the species distribution, life form and life span.
+```{r fig5, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "FADM ordination of the seed variables and the species distribution, life form and life span.
 Each square is a species.
@@
 ```
+
+# Tables
+
+```{r table1, echo = FALSE}
+flextable(Table1) %>% width(width = c(3.3, 1, 0.8, 0.7, 0.6)) %>% fontsize(size = 8) %>% padding(padding = 0) %>% set_caption("Table 1 Description of the datasets contributing primary data to this meta-analysis.
+The number of germination records, number of species and number of strict alpine species are given.")
+```

### 2020-08-13 src/msSource.R

@@ -11,6 +11,50 @@ read.csv(here("data", "Elevations", "Regions.csv")) %>%
          Region = ifelse(Region %in% c("Cantabrian Mountains"), "Cantabrian Mts", Region),     
          Region = as.factor(Region)) -> dataset
 
+# Table with the datasets
+
+data.frame(
+merge(
+merge(
+dataset %>%
+  mutate(Source = fct_recode(Source, `Fernández-Pascual` = "Fernández-Pascual Plant Biol")) %>%
+  mutate(Region = fct_recode(Region, `European Alps` = "Alps")) %>%
+  mutate(Region = fct_recode(Region, `Europe` = "Enscobase")) %>%
+  select(Source, Region, Country, TPLName, Alpine) %>%
+  group_by(Source, Region) %>%
+  tally %>%
+  rename(Dataset = Source, 
+         `Coverage` = Region,
+         Records = n),
+dataset %>%
+  mutate(Source = fct_recode(Source, `Fernández-Pascual` = "Fernández-Pascual Plant Biol")) %>%
+  mutate(Region = fct_recode(Region, `European Alps` = "Alps")) %>%
+  mutate(Region = fct_recode(Region, `Europe` = "Enscobase")) %>%
+  select(Source, Region, Country, TPLName, Alpine) %>%
+  unique %>%
+  group_by(Source, Region) %>%
+  tally %>%
+  rename(Dataset = Source, 
+         `Coverage` = Region,
+         Species = n)),
+dataset %>%
+  filter(Alpine == "Strict") %>%
+  mutate(Source = fct_recode(Source, `Fernández-Pascual` = "Fernández-Pascual Plant Biol")) %>%
+  mutate(Region = fct_recode(Region, `European Alps` = "Alps")) %>%
+  mutate(Region = fct_recode(Region, `Europe` = "Enscobase")) %>%
+  select(Source, Region, Country, TPLName, Alpine) %>%
+  unique %>%
+  group_by(Source, Region) %>%
+  tally %>%
+  rename(Dataset = Source, 
+         `Coverage` = Region,
+         `Strict` = n)),
+  Source = c("Briceño, unpublished", "Bu et al. (2007, 2008)", "Cavieres & Arroyo (2000), Cavieres & Sierra-Almeida (2018)", 
+               "enscobase.maich.gr", "Fernández-Pascual et al. (2017a)", "Liu et al. (2013)", "Mondoni, unpublished", 
[…]

### 2020-09-14 README.Rmd
Submission-ready 🌱🏔️

---
title: "README.md"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

![Picos de Europa National Park, Cantabrian Mountains, Spain](data/fernandezpascual/Picos de Europa.JPG)

# The seed germination spectrum of alpine plants: a global meta-analysis

This repository stores all the information related to the publication *The seed germination spectrum of alpine plants: a global meta-analysis*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.

## Contents

This repository is organised following the advice of [Wilson et al. 2017](https://doi.org/10.1371/journal.pcbi.1005510) for recording and storing research projects.

The following materials are available in the folders of this repository:

* `data` Original raw files including primary germination data and species traits.
* `doc` Files to create the manuscript using `Rmarkdown`.
* `results` Output of the `R` scripts, including clean data and MCMCglmm analysis outputs.
* `src` Scripts in `R` language used to clean the raw data and perform the analyses of the manuscript.

## Citation

Please cite the dataset as `CITATION`

### 2020-09-14 doc/manuscript.Rmd
Submission-ready 🌱🏔️

---
title: "The global seed germination spectrum of alpine plants"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
bibliography: refs.bib
csl: new-phytologist.csl
link-citations: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",")
})
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(tidyverse); library(flextable)
source("../src/manuscript-source.R", encoding = "UTF-8")
```

Eduardo Fernández-Pascual^1^ ^+^, Angelino Carta^2^, Andrea Mondoni^3^, Lohengrin Cavieres^4^, Sergey Rosbakh^5^, Susanna Venn^6^, Annisa Satyanti^7^, Lydia Guja^8^, Verónica F. Briceño^7^, Filip Vandelook^10^, Efisio Mattana^11^, Arne Saatkamp^12^, Haiyan Bu^13^, Karen Sommerville^14^, Peter Poschlod^5^, Kun Liu^13^, Adrienne Nicotra^7^, Borja Jiménez-Alfaro^15^

^1^ Departamento de Biología de Organismos y Sistemas, Universidad de Oviedo, Oviedo/Uviéu, Spain; ^2^ Dipartimento di Biologia, Botany Unit, University of Pisa, Pisa, Italy; ^3^ Department of Earth and Environmental Sciences, University of Pavia, Pavia, Italy; ^4^ Departamento de Botánica, Facultad de Ciencias Naturales y Oceanográficas, Universidad de Concepción, Concepción, Chile; ^5^ Chair of Ecology and Conservation Biology, Institute of Plant Sciences, University of Regensburg, Regensburg, Germany; ^6^ Centre for Integrative Ecology, School of Life and Environmental Sciences, Deakin University, Burwood, Victoria, Australia; ^7^ Division of Ecology & Evolution, Research School of Biology, The Australian National University, Canberra, ACT, Australia; ^8^ Centre for Australian National Biodiversity Research, CSIRO, Canberra, ACT, Australia; ^9^ National Seed Bank, Biodiversity Science Section, Australian National Botanic Gardens, Canberra, ACT, Australia; ^10^ Meise Botanic Garden, Meise, Belgium; ^11^ Natural Capital and Plant Health Department, Royal Botanic Gardens, Kew, Ardingly, UK; ^12^ Aix Marseille Université, Université d'Avignon, CNRS, IRD, IMBE, Marseille, France; ^13^ State Key Laboratory of Grassland and Agro-Ecosystems, Lanzhou University, Lanzhou, PR China; ^14^ The Australian Plant Bank, Royal Botanic Gardens and Domain Trust, Mount Annan, NSW, Australia; ^15^ Research Unit of Biodiversity (CSUC/UO/PA), University of Oviedo, Mieres, Spain

^+^ Correspondence: Departamento de Biología de Organismos y Sistemas, Universidad de Oviedo, C/ Catedrático Rodrigo Uría, 33006 Oviedo/Uviéu, Spain. Email: efernandezpascual@gmail.com. Telephone: +34985104787.

**Total word count for the main body of the text** 6,343; **Introduction** 1,968; **Materials and Methods** 1,342; **Results** 1,138; **Discussion** 1,896; **Number of figures** 5 in colour; **Tables** 1; **Supporting information** 0

# Summary

* We present a meta-analysis of the alpine seed germination spectrum based on primary data from germination experiments conducted in four continents with `r MSspecies` plant species, with the aim of testing the generality of alpine germination patterns at the global level.
* We use phylogenetic Generalized Linear Mixed Models to estimate the influence of six environmental cues on seed germination proportion, mean germination time and germination synchrony; accounting for possible effects of seed morphology (mass, embryo:seed ratio) and phylogeny.
* Alpine plants are characterized by physiological seed dormancy, strong need for cold stratification, warm-cued germination and positive germination responses to light and alternating temperatures. Species with a distribution limited to the alpine belt have a higher preference for warm temperatures and a stronger response to cold stratification than species whose distribution extends also to the lowlands. Seed mass, embryo size and phylogeny have strong constraining effects on germination responses to the environment.
* Overwintering and warm-cued germination are key drivers of germination in alpine habitats. The interplay of seed morphological traits and germination physiology reflects pressures to avoid frost or drought stress. These results indicate the convergence, at the global level, of the seed germination spectrum of alpine species.

# Keywords

alpine; alternating temperature; cold stratification; embryo endosperm ratio; light germination; seed dormancy; seed germination; seed mass

# Introduction

Alpine environments occur worldwide, above the maximum elevation at which trees can grow naturally [@RN3000; @RN4750]. These treeless habitats are characterized by low temperatures, strong winds, unstable substrates and short growing seasons [@RN2392]. Their dependence on extreme and strict climatic conditions [@RN3024], coupled with the relatively small terrestrial area they occupy [@RN4750], makes alpine environments disproportionally vulnerable to climate change [@RN4946; @RN4950]. This threatens the survival of the world’s rich and specialized mountain biodiversity [@RN4947]. It also endangers the sustainability of the ecosystem services that alpine environments provide for hundreds of millions worldwide, including the provisioning of natural resources, the regulation of environmental hazards and the maintenance of cultural and recreational activities [@RN4949; @RN4948]. Succesful management of climate change and of its consequences will depend on an adequate understanding of alpine biodiversity.

Specially, plant regeneration by seed is a key process that determines whether species will persist in their current habitat or will be able to migrate to new sites [@RN3055]; and it is also fundamental for active efforts in plant conservation [@RN3763] and seed-based ecological restoration [@RN3507; @RN2241]. Although most alpine plants can reproduce clonally [@RN2392], sexual reproduction by seed is the main strategy to maintain genetic diversity and colonize suitable new sites in response to environmental changes. However, the short reproductive season of the alpine habitat constrains the phenological timing of flowering, pollination, seed development and seed dispersal. For regeneration by seed to be succesful, dispersed propagules must germinate in the most appropriate period to ensure the survival of seedlings under environmental stress [@RN4714; @RN2868; @RN4717]. The physiological process of seed germination is therefore an essential life stage that must be timed to occur when the environment is favourable for subsequent seedling survival and growth [@RN4691]. In alpine habitats, seed germination will be a crucial test for plant populations coping with climate change, determining whether they will persist or go into extinction.

Natural selection is expected to favour seed germination requirements that reduce the probability of facing environmental conditions which are not appropriate for seedling survival and growth [@RN3375]. Thus, germination is mostly regulated by environmental cues related to water availability and soil temperature [@RN3368]. The propagules of many plant species have also developed seed dormancy, by which germination is prevented during periods that are only ephemerally favourable, like a short warm spell in the middle of winter [@RN3214]. Different degrees of dormancy within a seed population work to ensure the distribution of seedling emergence across time, thereby bet hedging offspring survival against unpredictable environments [@RN3065]. Other germination cues, such as the response to diurnal alternating temperatures [@RN1380] and the need for light or darkness [@RN4656], allow for a fine-scale detection of germination micro-niches and safe sites [@RN4719]. Given the comparatively high heterogeneity of alpine climates and plant lineages adapted to different regions, and the wealth of local studies on alpine seed ecology (reviewed in the following paragraphs), a current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants while accounting for the evolutionary relatedness among taxa. The seed ecological spectrum [@RN4655] is defined as a set of seed traits that determines the ability of plants to disperse, persist, germinate, and establish in different habitats. A major goal for trait-based seed ecology is to study the relationships and trade-offs between different aspects of the seed ecological spectrum [@RN4655], such as the physiological drivers of germination versus key morphological traits like seed mass [@RN4916] and the embryo to seed ratio [@RN3685].

In the available literature, many alpine species have been described as having seeds with deep physiological dormancy, i.e. a type of dormancy that is caused by the seed’s internal balance of phytohormones and requires a long exposure to dormancy-breaking treatments to be terminated [@RN2943; @RN3703; @RN3214]. This physiological dormancy would prevent precocious germination of recently dispersed seeds during autumn, when appropriate conditions are not likely to persist for more than a few weeks or days [@RN3330]. A requirement for cold stratification to break dormancy would allow seeds to sense the snow season, thereby postponing germination to a more favourable period for seedling survival and development. Thus, it can be expected that a cold stratification period is a common requirement for seed germination in alpine plants. Nonetheless, an early review [@RN3213] suggested that cold stratification was not a requirement for the seed germination of several alpine species from different mountain ranges in the USA, a finding that has been reported by subsequent studies [@RN1484; @RN2995; @RN4721]. More recently, research on Australian alpine plants found that a cold stratification period only increased germination in half of the species tested [@RN3703; @RN3285]. In contrast, cold stratification increased seed germination over a range of temperatures in most of the species studied in the alpine zone of Japan [@RN707]. Along an elevational gradient in the central Chilean Andes, cold stratification was an important requirement for seed germination in species from lower elevations, while species from higher sites did not germinate even after cold stratification [@RN4713]. Finally, in a comparative study on germination of unstratified seeds from different habitats of the Austrian Alps, species from alpine grasslands had the lowest final germination proportion, suggesting that they had the highest level of dormancy [@RN4930].

Seed dormancy is not the only trait that can prevent, in practice, autumn germination. Early studies on the germination ecology of alpine plants demonstrated that in most species recently dispersed seeds require relatively high temperatures for germination [@RN3258; @RN3213; @RN4712]. Warm-cued germination has been considered as an adaptation to prevent seed germination at the time of seed dispersal (autumn) when temperatures are relatively low and there is a high risk of frost [@RN3393]. Indeed, germination of alpine seeds tends to occur after winter, mainly in early summer [@RN2392; @RN2382]. In recent times, an increasing number of studies reported that fresh seeds of some alpine plants also germinate at cool incubation temperatures or during cold stratification treatments [@RN2943; @RN3285; @RN2371; @RN4713]. This cold-cued germination could be an adaptation to germinate under snow or during snowmelt, which could presumably allow seedlings to develop a deeper root system before topsoil desiccation in summer [@RN4720], or to attain an optimal size for overwintering [@RN4712]. Yet, the generality of warm-cued germination across different alpine regions and habitats has never been studied systematically, and the relative importance of cold stratification versus warm-cued germination remains unclear. A central question is whether physiological seed dormancy plays a relevant role in postponing alpine germination during autumn and winter, or the warm-cued germination is enough to prevent germination under cool autumn temperatures. The response to this question becomes pertinent under a warming climate, which could result in non-dormant seeds germinating readily in autumn. Non-dormant seeds could possibly have an advantage as the growing seasons become longer, if seedlings have enough time to grow and become sufficiently large to survive the winter snow season. Potentially, having a non-dormant or less dormant seed could be a better strategy for alpine plants to thrive under future climate [@RN3317; @RN2382]. However, we might also expect that seedlings from these species will be exposed to frost damage if snow starts to melt earlier and insulation during winter disappears [@RN4944] and therefore a mechanism that enables diverse germination timing (i.e. low germination synchrony) should be preferred.

Besides patterns of seed dormancy and germination temperature, the response of alpine seeds to diurnal alternating temperatures and light can provide further information on their capacity to detect fine-scale environmental cues. Fluctuations of diurnal temperature decrease with increasing burial depths; below 10 cm soil temperature might be too stable to trigger germination in species requiring temperature alternation [@RN4910]. Solar irradiance in physiologically significant quantities penetrates only the first millimetres of the soil [@RN4909]. Therefore, these two cues can indicate whether seeds are buried deep in the soil, in rock crevices or under snow. The depth of seed burial is crucial for seedling emergence, as small seeds lack the resources to survive until they reach the soil surface [@RN4898]. A study of 445 species from the Qinghai-Tibet plateau found that species from high elevation alpine meadows did not show a significant response to diurnal temperature oscillations [@RN3376]. This is intriguing given that alpine habitats typically have a strong day-night climatic variation during the growing season. Regarding light and darkness, some local studies have found alpine species to have a preference for germinating in light [@RN4730; @RN4729]. This trait could favour the creation of a persistent soil seed bank [@RN4718; @RN4929] which is advantageous in the temporally and spatially unpredictable alpine environments [@RN4731; @RN4732]. Indeed, alpine soil seed banks are more frequent than previously supposed, as consistently reported in recent years [@RN4929; @RN4718; @RN4945]. Nevertheless, other alpine species seem to prefer dark germination [@RN2943], which may be interpreted as a strategy for detecting safe sites in rock crevices [@RN3008]. Although some germination studies report laboratory experiments conducted with constant and alternating temperatures, or light and dark conditions, they are very much case-dependent, and therefore it is difficult to evaluate whether these factors show a consistent pattern across global alpine ecosystems.

In this work, we make a global collection of primary germination data from laboratory experiments conducted with seeds of alpine plant species. The assembled dataset represents eight alpine regions belonging to four continents, and we use it to investigate the germination response of alpine plants to key environmental factors. These factors include dormancy-breaking treatments (cold stratification, GA~3~ and scarification), average germination temperature, diurnal alternating temperatures and light [@RN3368; @RN3214]. Given the heterogeneity of alpine species lineages adapted to different regions, we account for the shared evolutionary history using a phylogenetic comparative approach [@RN3081]. We also analyse the relationship between germination ecology and two morphological traits which are key drivers of the seed ecological spectrum: seed mass and the embryo to seed ratio [@RN4655]. Our main aim in conducting this work is to test the consistency of general assumptions about the influence of major environmental cues on seed germination of alpine plants. Specifically, we determine at the global scale whether alpine seeds: (1) have physiological dormancy; (2) require cold stratification to break dormancy; (3) need relatively warm temperatures to germinate; (4) do not require alternating temperatures for germination; and (5) respond positively to light. To understand the ecological constraints of our meta-analysis on alpine species pools, we analyse how the germination responses differ between strict alpine species (that occur exclusively above the treeline) and generalist species (that occur above the treeline but also in the lowlands). We can expect that the former group will show a more convergent alpine germination syndrome, while the latter will be more plastic in their response.

# Materials and Methods

## Data collection

We conducted a meta-analysis of primary data [@RN4734], which consists of pooling together original data from different local studies to synthesize their conclusions at the global level. The original data files, as well as R code for data cleaning, analysis and manuscript production are available at https://github.com/efernandezpascual/alpineseeds. To gather data, we contacted research groups which have performed experiments on seed germination of alpine plants, asking them to provide raw data from laboratory germination experiments. The submitted data had to report the studied species, the germination treatments, the number of seeds sown, and the number of seeds germinated. We received data (Table 1) from the Cantabrian Mountains of Spain, the European Alps, the Italian Apennines, the Northern Caucasus, the Qinghai-Tibet Plateau of China, the Chilean Andes, and the Australian Alps. We also received data on European alpine species from *Enscobase*, the database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml).

---
nocite: |
  @RN2371, @RN4729, @RN3029, @RN4727, @RN4728, @RN4735, @RN3376, @RN3393, @RN4713, @RN4737, @RN4736, @RN3703
...

Once received, we standardized the species names according to *The Plant List* [@RN2321] using the package *Taxonstand* [@RN4660] in *R* [@RN2315]. For each species, we recorded its elevation range from local floras [@RN4741; @RN4739; @RN3243; @RN4742; @RN4740], catalogues [@RN4738] and herbaria (*Australian National Herbarium*, https://www.anbg.gov.au/cgi-bin/anhsir). We used the elevation ranges (i) to remove from the dataset outlier species that do not grow above the treeline; and (ii) to classify the rest of the species as *strict alpine* (i.e. species that only grow above the treeline) or *generalist* (i.e. species that can grow above the treeline but also in the lowlands). For each species, we compiled three additional traits. First, seed dormancy class obtained from Baskin & Baskin [-@RN3214] and Rosbakh et al. [-@RN4743]. We simplified the original dormancy classes to five categories: physiological (germination prevented by the seed’s internal balance of phytohormones), morphological (germination prevented by an embryo which is underdeveloped at the time of dispersal), morphophysiological (both physiological and morphological dormancy present), physical (germination prevented by a water-impermeable seed coat) or non-dormant. Second, seed mass as provided by the Royal Botanic Gardens Kew Seed Information Database [-@RN2987] and some additional sources [@RN4728; @RN4735; @RN3376]. Third, the embryo to seed surface ratio. Embryo to surface area values were based on measurements of the surface area of the embryo and seed from drawings and photographs retrieved mainly from Martin [-@RN3244] and the Royal Botanic Gardens Kew Seed Information Database [-@RN2987], supplemented with own measurements (Vandelook, unpublished). For some species that were missing seed mass or embryo values, we calculated genus or family averages.

## Germination response variables

We used the final germination proportion (i.e. the number of seeds germinated out of seeds sown) at 8 weeks as the main response variable for the germination treatments. We further calculated the mean germination time and the germination uncertainty index with the package *GerminaR* [@RN4752]. Mean germination time is the time that it takes for half of the seed lot to germinate, indicating the germination speed in each experimental condition. The germination uncertainty index estimates how scattered germination is through time, and gives an idea of whether germination is synchronous (i.e. most seeds germinate around the mean germination time) or asynchronous (i.e. germination events are separated through time during the germination experiment). Lower values in the germination uncertainty index indicate higher germination synchrony. The uncertainty index is analogous to the synchrony index and we used it because, unlike the latter, its values are not constrained between 0 and 1 [@RN4752]. To calculate these indices, we used only records that met two conditions: (i) more than 50% final germination percentage; and (ii) a mean germination time under 4 weeks. We established these limits to prevent the results being altered by a modification of the seed’s dormancy status during the germination incubation (e.g. a fraction of the seed lot is non-dormant at the beginning of the incubation and germinates shortly after the start of the experiment; while another fraction is originally dormant, loses dormancy in response to the incubation conditions, and germinates in a second wave).

## Statistical analyses

We meta-analysed germination data by fitting generalized mixed models with Bayesian estimation (Markov Chain Monte Carlo generalized linear mixed models, MCMCglmms) using the R package *MCMCglmm* [@RN4755]. To model final germination proportion, we used binomial MCMCglmms, while for mean germination time and the germination uncertainty index we used gaussian MCMCglmms. Models had, as fixed effects, the experimental conditions (stratification, GA~3~, scarification, temperature, alternating temperature and light), plus their interaction with seed mass and embryo:seed, and the strict alpine or generalist character of the species. Random effects included a reconstructed phylogenetic tree for the `r MSspecies` species, and species identity, seed lot, lab (i.e. data provider), and alpine region (Table 1). To create the phylogeny, we used the R package *V.PhyloMaker* [@RN4753] which contains an updated mega-tree of the seed plants based on Smith & Brown [-@RN4754]. We placed taxa absent from the mega-tree at the genus-level basal node. In all models, all variables were scaled so their contribution to the effect sizes could be compared. We used weakly informative priors in all models, with parameter-expanded priors for the random effects. Each model was run for 500,000 MCMC steps, with an initial burn-in phase of 50,000 and a thinning interval of 50 [@RN4756], resulting, on average, in 9,000 posterior distributions. From the resulting posterior distributions, we calculated mean parameter estimates and 95% Highest Posterior Density (HPD) and Credible Intervals (CI). We estimated the significance of model parameters by examining CIs, considering parameters with CIs overlapping with zero as non-significant. To estimate the phylogenetic signal of seed germination over all variables we used Pagels’s lambda (λ) [@RN4757], estimated simultaneously with the models by calculating the mean of the posterior distribution and the 95% CI of λ as indicated by De Villemereuil et al. [-@RN4756]. When λ = 0, related taxa are no more similar than expected by chance, while when λ = 1, the trait is evolving following a constant variance random walk or Brownian motion model; intermediate values of λ indicate a phylogenetic correlation in trait evolution that does not fully follow a Brownian motion model [@RN4757].

## Visualization of the alpine seed germination spectrum

To visualize the seed germination spectrum of alpine plants, we employed a Factorial Analysis of Mixed Data (FAMD) as implemented in the package *FactoMineR* [@RN3166]. FAMD ordination combines the properties of PCA and MCA to jointly measure the variation of continuous and categorical variables. The ordination was performed at the species level, i.e. calculating a series of continuous and categorical traits for each species. We transformed the final germination proportions to create a continuous variable for each germination cue (i.e. stratification, GA~3~, scarification, temperature, alternating temperatures and light). To do so, for each cue and species, we calculated a weighted average of the cue levels (in the case of temperature, cue levels were the temperature treatments; for the other cues the levels were 0 = absence and 1 = presence), weighting by the germination proportion at each level. This approach underrepresents the importance of the levels that were not tested for a given species, but can serve as a proxy of the response to the germination cues when visualized across the whole dataset; it must be stressed that this stage of the analysis serves only for visualization and not hypothesis testing. We also included germination timing traits by calculating single values for the mean germination time (the minimum, i.e. the time taken at the most favourable treatment) and the germination uncertainty index (the mean). We conducted the FAMD with only the `r MSfamdspp` species for which we had mean germination time and germination uncertainty index values. We finally included the continuous variables seed mass and embryo:seed, as well as the categorical variables dormancy class (physiological, morphophysiological, morphological, physical or non-dormant) and distribution (strict alpine vs. generalist species).

# Results

## Description of the dataset

The final dataset used in our meta-analysis contained `r MSrecords` records (i.e. germination proportions for a given seed lot of a species, recorded in a set of experimental conditions) contributed by `r MSgroups` research groups, representing `r MSfamilies` seed plant families and `r MSspecies` species (`r MSstrict` strict alpine species and `r MSgeneral` generalist species). Seed lots were originally sampled in `r MSregions` alpine regions of the world (Table 1), in both hemispheres.

Most species in the dataset were reported as having dormant seeds (Fig. \@ref(fig:fig1)a), but the frequency of the dormancy classes did not differ between generalist and strict alpine species. Among those species with dormant seeds, physiological dormancy was the most common dormancy class, followed by morpho-physiological and then physical. Both strict alpine and generalist species had a similar range of values for seed mass (Fig. \@ref(fig:fig1)b) and embryo:seed ratio (Fig. \@ref(fig:fig1)c). The values of seed mass ranged from `r MSminMass` mg (*`r MSminMassSpp`*) to `r MSmaxMass` mg (*`r MSmaxMassSpp`*), with a median value of `r MSmedianMass` mg. Embryo:seed ratio encompassed the full range of potential values, from endospermic species with very small embryos (`r MSminEmbryo`) to non-endospermic seeds (`r MSmaxEmbryo`). The median value was `r MSmedianEmbryo`.

The total number of seeds used in the experiments was `r MSseeds`. Germination temperatures (i.e. the weighted average of the day and night temperatures, weighted by the duration of each phase) ranged from `r MSminT` to `r MSmaxT` ºC, with `r MSaltY` records of constant temperatures (i.e. experiments that used the same temperature during all their duration) and `r MSaltN` of alternating temperatures (i.e. experiments where different temperatures were applied during the day and the night, in diurnal cycles). Seeds were exposed to light during some part of the diurnal cycle in `r MSlightY` records and kept in total darkness in `r MSlightN` records. The experiments were performed with untreated seeds in `r MSstratN` records, and of the rest, the majority (`r MSstratCold`) went through cold stratification (< 5 ºC, in darkness). Scarification (i.e. an abrasion of the seed coat to allow water imbibition) was performed in `r MSscarY` records, and GA~3~ (a plant hormone that promotes embryo growth and stimulates germination) in `r MSga3` records. Since germination incubations ran for different periods of time, we established a cut-off time of 8 weeks to calculate the final germination percentages, to have a homogeneous comparison across records.

## Dormancy breaking treatments

In general, stratification significantly increased the final germination proportion (Fig. \@ref(fig:fig2)a), as shown by a credible interval not overlapping zero. This positive effect occurred in both generalist and strict alpine species, although the effect size was larger in the strict alpine group. Stratification slightly reduced the mean germination time (Fig. \@ref(fig:fig2)b), which was evident in the generalist species, but not in the strict alpine species. Further, stratification reduced the germination uncertainty index as a general response (Fig. \@ref(fig:fig2)c), and slightly more so in generalist species. Seed mass did not interact with stratification, but species with larger embryos had a smaller response to stratification in their final germination proportion and reduced less markedly their mean germination time and germination uncertainty index. GA~3~ had similar effects to stratification, with the main difference being that seed mass affected the response to GA~3~, whereas it did not interact with the response to stratification.

Scarification promoted the germination proportion (Fig. \@ref(fig:fig2)a) and reduced the mean germination time (Fig. \@ref(fig:fig2)b), but it did not affect the germination uncertainty index (Fig. \@ref(fig:fig2)c). Strict alpine species were slightly more responsive to scarification than generalist species. Heavier seeds were more responsive to scarification, while seeds with larger embryos responded less.

## Germination temperature

Temperature had a positive significant effect on the final germination proportion as shown by a credible interval not overlapping zero (Fig. \@ref(fig:fig2)a), indicating a preference for warmer temperatures. Warmer temperatures also significantly decreased mean germination time (Fig. \@ref(fig:fig2)b) and the germination uncertainty index (Fig. \@ref(fig:fig2)c). The effect of warmer temperatures was slightly higher in strict alpine than in generalist species, although the credible intervals of the two groups of species do overlap. Seed mass and embryo:seed had a negative interaction with temperature, reducing the response to temperature. In other words, heavier and less endospermic seeds had a smaller increase in germination proportion and a smaller decrease in mean germination time when germinating in warmer temperatures. The germination uncertainty index did not show an interaction with seed mass or embryo:seed, as shown by the credible interval overlapping zero.

## Alternating temperatures

Alternating temperatures had a positive effect on the final germination proportion (Fig. \@ref(fig:fig2)a), but not in the mean germination time (Fig. \@ref(fig:fig2)b) or the germination uncertainty index (Fig. \@ref(fig:fig2)c). The positive effect on the germination proportion was slightly higher in the generalist species. As in the case of the mean germination temperature, higher seed mass and embryo:seed values reduced the positive effect of alternating temperatures on the final germination proportion. Alternating temperatures resulted in faster germination (reduced the mean germination time) for seeds with low values of embryo:seed.

## Light

Light increased the final germination proportion (Fig. \@ref(fig:fig2)a), with no differences between strict alpine and generalist species. It did not affect the mean germination time (Fig. \@ref(fig:fig2)b), but it did result in more variable germination (increased the germination uncertainty index) of generalist species (Fig. \@ref(fig:fig2)c). Heavier and less endospermic seeds responded less to light.

## Random effects and phylogenetic signal

Random factors had a significant effect on the MCMCglmm models (Fig. \@ref(fig:fig3)). The final germination proportion and the mean germination time were affected more strongly, on average, by the seed lot and the species identity, suggesting that these traits are subject to intra- and interspecific variability, respectively. Alpine region and lab had a lower average effect and more variability as shown by their credible intervals. The germination uncertainty index showed the opposite pattern, with a stronger effect of lab and alpine region. The phylogenetic signal in the response to all germination parameters (Fig. \@ref(fig:fig4)) was significantly positive for the final germination proportion and the mean germination time. For the germination uncertainty index, it overlapped with zero, indicating a weak effect of phylogeny.

## FAMD ordination

The first FADM axis explained `r round(pcaAlpine$eig[1, 2], 0)`% of the variation. The quantitative variables with the largest contribution to this first axis were germination temperature, germination uncertainty, GA~3~ and alternating temperatures. Among the qualitative factors, the strict alpine / generalist categories contributed to axis 1. In summary, the horizontal axis separated (i, left) generalist species with a positive response to alternating temperatures and more germination uncertainty from (ii, right) strict alpine species with a preference for warmer temperatures and a positive response to GA~3~ (Fig. \@ref(fig:fig5)). Axis 2 explained `r round(pcaAlpine$eig[2, 2], 0)`% of the variability. The main contributing variables were embryo:seed, mean germination time, seed mass and scarification; and the main categorical factors were physical dormancy and morphophysiological dormancy. This axis separated (iii, bottom) morphophysiologically dormant species with longer mean time to germination from (iv, top) physically dormant species that responded to scarification and had heavier, less endospermic seeds.

# Discussion

Our meta-analysis of primary germination data shows that alpine plants from different mountains of the world tend to show specialized seed germination characteristics. For most alpine species, both dormancy and warm-cued germination combine to ensure germination occurs in the season most conducive to survival: seeds tend to germinate when, after weeks of exposure to conditions that mimic winter, they are moved to mid-summer temperatures. Specifically, the germination ecology of strict alpine species is characterized by (1) physiological seed dormancy; (2) a strong need for cold stratification or GA~3~ to break dormancy; (3) warm-cued germination; (4) a positive response to alternating temperatures, although not so prominent as in generalist species; (5) a positive response to light; (6) slow and relatively synchronous germination; and (7) an interplay with seed and embryo size, with smaller and more endospermic seeds being more responsive to stratification, warmth, alternating temperatures and light. Here we discuss the ecological significance of the different aspects of the alpine seed germination spectrum.

## Seed dormancy

Our dataset indicates that seed dormancy is frequent across alpine species and plays an important role in postponing plant recruitment to favourable times. Furthermore, the frequency of dormancy is similar in strict alpine and generalist species, suggesting that this is a general trait of cold-adapted mountain species. Seeds with the two most frequent dormancy classes, physiological and morphophysiological, require a stratification period during which certain conditions (e.g. cold and wet, although other combinations are possible) are met over a period of months [@RN3214; @RN4743]. Indeed, our results show that cold stratification promotes germination. The phytohormone gibberellic acid (GA~3~) works as a general substitute of cold stratification to overcome the inherent physiological dormancy of alpine species. Additionally, we found some species with physical dormancy. This dormancy type can be broken by mechanical scarification via freezing-thawing cycles in spring or extreme temperature fluctuations during summer, but is also capable of responding to seasonal cycles of temperature and humidity [@RN780].

Overall, these results confirm that overwintering is essential for promoting seed germination in alpine species, concurring with previous local studies [@RN3393; @RN2868; @RN707; @RN698; @RN3703; @RN2355; @RN3285; @RN2371; @RN4713]. Since global climate change is causing major changes in snow cover duration [@RN4917; @RN4918], it could disrupt natural cold stratification cycles in alpine environments, compromising population viability and indirectly favouring species with no stratification requirement [@RN3703]. This risk will be partly alleviated by the plasticity of physiological seed dormancy in response to the seed maturation temperature [@RN2358 ; @RN2899], a mechanism of “seed memory” that facilitates acclimatisation to changing environments [@RN2249], and which could reduce the stratification requirements of seeds produced in a warmer climate.

## Thermal regulation of germination

We found a general increase of seed germination with higher incubation temperatures. This effect was more pronounced in strict alpine compared to generalist species, confirming the findings of local studies [@RN3266; @RN2371]. It was also more pronounced in small seeded and endospermic species. Small-seeded species have a lower proportion of seedling survival under drought [@RN4530] but may require warmer germination temperatures than large-seeded species, especially in seasonal climates [@RN2399]. Considering that risk reduction is an important selective pressure for the evolution of germination traits [@RN4915], a plausible explanation for these patterns is that alpine plants received selection pressure from damaging spring frosts and evolved a frost avoidance mechanism by germinating at high temperatures [@RN4908; @RN4906; @RN4722]. Moreover, strict alpine species may be less subjected to drought damage during establishment compared to generalist species: as elevation increases, plants face less drought risk because precipitation increases (except in Mediterranean and semi-arid mountains), but at the same time there is a higher probability of early or late season frosts [@RN4897; @RN2392; @RN4914]. Together, these results indicate that germination patterns in alpine species are driven by an interplay of seed size and germination temperature to escape either frost or drought stress; small seeds are being selected for fast germination at warm temperatures, in order to escape unfavourable early-spring frost, and large seeds are being selected for slower germination under cooler conditions, which should maximize emergence during snowmelt (when water availability is highest) and gives seedlings adequate time to develop a root system to cope with desiccation risks in summer. Supporting this view, seedlings of alpine pioneer species have a low frost resistance [@RN4722], yet high heat tolerance up to 40–50°C [@RN4902].

In our analysis, increasing temperatures also decreased germination uncertainty, showing that germination synchrony is a plastic trait driven by temperature. While asynchronous germination is thought to be a form of adaptation to unpredictable alpine environments [@RN3734; @RN3690], our results indicate that asynchronous germination occurs when temperatures are relatively cool, likely as a bet-hedging strategy against the risk of early-spring frost. On the other hand, a fast and synchronised germination at warm conditions (i.e. in late spring) may be another important strategy that increases alpine seedling survival by avoiding drier conditions in summer. The extreme germination uncertainty syndrome, i.e. staggering germination so that emergence occurs both before and after winter, is known in alpine species [@RN2392; @RN3285; @RN4907]. Staggered germination is usually exhibited as a rapid first wave of germination followed by second wave several weeks or months later [@RN2392]. If autumn seedlings grow faster than spring seedlings [@RN4907], this will likely have cascading effects on individual fitness along the life of the plant [@RN3036; @RN2384].

## Alternating temperatures and light

We found a positive germination response to alternating temperatures and light, conditions that mimic the environment in the upper soil and certain microsites during the snow-free season [@RN4712; @RN2392; @RN3029]. Alternating temperatures had a slightly lower effect on strict alpine species, compared with generalist species, but still had an effect. This partly contradicts the findings of Liu et al. [-@RN3376], who reported a lack of a response to temperature fluctuation for the species distributed only at high elevations of the Qinghai-Tibet Plateau. Liu and co-workers argued that, in that area, high temperature fluctuations may occur in all months of the year and therefore fluctuating temperature alone may not be a reliable indicator of suitable conditions for seedling establishment and growth, while warm daily temperatures could be a cue that the short summer growing season has begun [@RN3376]. When analysed at the global level, alternating temperatures generally increase germination of alpine plants, which corresponds well with the fact that most alpine regions have a strong day-night climatic variation during the growing season. It is also not surprising that we detected a negative interaction between seed mass and the germination response to light and alternating temperature regimes, since large seeds can afford to germinate at greater depths (where they are more protected from the alpine environment) and still reach the soil surface [@RN4904].

## The alpine seed germination spectrum

Using the whole set of traits compiled in this study we visualize the variation in the seed germination spectrum of alpine plants. On one end of the spectrum, strict alpine species tend to have slow and synchronous germination, require warm temperatures and light to germinate, and need stratification or GA~3~ to break physiological dormancy. On the other end, generalist species tend to have fast germination, show a bet-hedging strategy (high germination uncertainty) and germination traits suggesting possible fast germination after specific cues (scarification, alternating temperatures). The dormancy classes without a physiological component (non-dormant, physical, and morphological) appear associated with generalist species. This gradient of regenerative strategies mirrors prominent gradients of fast or slow population dynamics [@RN4922; @RN4921], fast or slow developing leaf traits [@RN4923], or large sets of plant ecological features [@RN2279]. “Fast” regeneration can be understood as a plastic and opportunistic strategy to cope with frequently disturbed habitats such as avalanche ways, steep eroding slopes, regressing glaciers, and riverbeds [@RN4926; @RN4925]. The “slow” regeneration of strict alpine species, on the other hand, seems to be a more specialized syndrome and might be the result of convergent evolution acting on those species that are truly restricted to the alpine vegetation belt.

A general assumption of seed trait ecology is that seed mass and embryo:seed ratios are relatively conserved across related species [@RN4916; @RN3685]. In our alpine dataset, both traits showed no differences between strict alpine and generalist species. This is in line with the ambiguity of relationships that have been found between seed mass and elevation for alpine species [@RN4912; @RN4903; @RN3685; @RN4896]. The similarity in embryo:seed size between generalist and strict alpine species confirms the absence of a correlation between embryo-seed size and elevation found in the Apiaceae [@RN3685]. In contrast with the supposed phylogenetic conservatism of these morphological traits, physiological germination responses are understood to be more plastic traits that can quickly acclimatize to new situations [@RN2249]. While phylogenetic signal in seed germination is generally found, especially in highly selective environments [@RN3001; @RN2399; @RN2865], germination patterns can also be shared between phylogenetically distant species [@RN4911] and may have low or absent phylogenetic signal [@RN2365; @RN4900; @RN2371]. The phylogenetic signal found here indicates that evolutionary history cannot be neglected when studying alpine seed germination patterns. Furthermore, we detected close relationships between seed-embryo size and physiological responses of seeds, suggesting that both sets of traits are connected and may be subjected to co-adaptation. Previous studies have shown that non-endospermic seeds are smaller and are likely to be non-dormant at the time of dispersal and thus germinate quicker than alpine endospermic seeds [@RN3703; @RN3285; @RN4907]. Together, these results indicate that seed morphology and physiology are connected aspects. This connection, intertwined with their shared phylogenetic history, may constrict the potential variability of the seed ecological spectrum.

## Conclusions

Alpine species from different regions of the world have specialised germination characteristics that support successful regeneration in challenging alpine environments. Our results point to an evolutionary convergence of germination patterns in alpine species. They also provide the first global assessment of the seed germination spectrum in relation with morphological seed traits of alpine plants, suggesting potential relationships and trade-offs between sets of regenerative traits that will need further research. New studies also will need to consider evolutionary history as an intrinsic constraint of the seed ecological spectrum. Overall, this study provides a global assessment of alpine seed germination that can be used as a reference in further studies focused on spatial or temporal variation in seed germination, their links with other plant traits, and the expected responses of alpine plants to climate change.

# Acknowledgements

EFP received financial support from the Government of Asturias and the FP7 – Marie Curie - COFUND programme of the European Commission (Grant ‘Clarín’ ACB17-19). EM is supported by the Future Leaders Fellowship - Diversity and Livelihoods, of the Royal Botanic Gardens, Kew. This research was supported by the NGSeeds lab virtual working group.

# Author Contribution

BJA conceived the original idea. EFP, AC, AM, LC, SR, SV, AnS, LG, VB, FV, HB, KS, PP contributed data. EFP assembled and managed the database. AC and EFP analysed data. EFP led manuscript writing, with contributions from BJA, AC, AM, LC, SR, SV, AnS, LG, VB, FV, EM, ArS. All authors revised the manuscript and approved the final version.

# Data Availability Statement

The original data, R code for the analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/alpineseeds. Upon publication, a version of record of the repository will be deposited in Zenodo.

# References

::: {#refs}
:::

# Figures

```{r fig1, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Seed dormancy classes, seed mass and embryo to seed ratio in the subsets of strict alpine and generalist species analysed in this study. Seed dormancy classes are abbreviated as follows: MD (morphological), MPD (morphophysiological), ND (non-dormant), PD (physiological), PY (physical).", fig.align = "center", fig.width=7, fig.height=3}
cowplot::ggdraw(fig1)
```

```{r fig2, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the germination environment on germination, according to the MCMC meta-analysis of the primary data. The dots indicate the posterior mean of the effect size, and the brackets its 95 % credible interval. The line of zero effect is shown. When the credible intervals overlap with the zero-effect line, the effect is not significant. The figure shows first the main effect, then the effects for the strict alpine and generalist groups, and then the interaction with seed mass and embryo:endosperm. A negative interaction (e.g. seed mass and light, in the case of the germination proportion) indicates that, as the interacting factor increases (e.g. seed mass) the response to the germination cue (e.g. light) decreases, in comparison with the main effect.", fig.align = "center", fig.width=7, fig.height=5.5}
cowplot::ggdraw(fig2)
```

```{r fig3, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the random factors on germination, according to the MCMC meta-analysis of the primary data. The dots indicate the posterior mean of the effect size, and the brackets its 95 % credible interval. The line of zero effect is shown.", fig.align = "center", fig.width=7, fig.height=5}
cowplot::ggdraw(fig3)
```

```{r fig4, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Phylogenetic signal in the response to the germination environment. When lambda = 0, related taxa are no more similar than expected by chance; when lambda = 1, the trait is evolving following a constant variance random walk or Brownian motion model.", fig.align = "center", fig.width=7, fig.height=1.5}
cowplot::ggdraw(fig4)
```

```{r fig5, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "FADM ordination of the alpine seed germination spectrum. Each square is a species. Labels indicate the contribution of the variables to the axes. Seed dormancy classes are abbreviated as follows: MD (morphological), MPD (morphophysiological), ND (non-dormant), PD (physiological), PY (physical).", fig.align = "center", fig.width=7, fig.height=7}
cowplot::ggdraw(fig5)
```

# Tables

```{r table1, echo = FALSE}
flextable(Table1) %>% width(width = c(3.3, 1, 0.8, 0.7, 0.6)) %>% fontsize(size = 8) %>% padding(padding = 0) %>% set_caption("Table 1 Description of the datasets contributing primary data to this meta-analysis. The number of germination records, number of species and number of strict alpine species are given.")
```

### 2020-09-14 data/elevations/elevations.R

@@ -1,84 +1,0 @@
-library(tidyverse); library(here); library(openxlsx)
-
-# This is an atempt to clean the mess of the many elevation files..
-
-# Create a merged file with the spp elevations
-
-read.csv(here("data", "elevations", "elevations_original.csv")) %>% # Original file, with more info for lowland species
-  merge(read.xlsx(here("data", "elevations", "Species Cavieres.xlsx")), # Merge with Cavieres, who altered some elevations
-        by = c("Species", "Region"), all = TRUE) %>%
-  mutate(Min.elevation = ifelse(is.na(Min.elevation.y), Min.elevation.x, Min.elevation.y),
-         Max.elevation = ifelse(is.na(Max.elevation.y), Max.elevation.x, Max.elevation.y)) %>%
-  select(Species, Region, Alpine.x, Min.elevation, Max.elevation) -> sppelevations
-
-# Read Poschlod file with updated Alpine trait
-
-read.xlsx(here("data", "elevations", "Species_rev_Peter_2020_04_15.xlsx")) %>% 
-  select(Species, Alpine) %>%
-  unique %>%
-  group_by(Species) %>%
-  filter(length(Species) >1) -> spp # Species with different classification in different regions
-
-read.xlsx(here("data", "elevations", "Species_rev_Peter_2020_04_15.xlsx")) %>% 
-  select(Species, Alpine) %>%
-  filter(! (Species %in% spp$Species & 
-              Alpine == "Generalist")) %>%
-  unique -> # I should do the opposite but I trust more the European classification by Peter...
-  alpine
-
-# List lowland spp
-
-sppelevations %>%
-  filter(Alpine.x == "Lowland") %>%
-  select(Species, Alpine.x) %>%
-  rename(Alpine = Alpine.x) %>%
-  unique -> lowland1 # Manually classified as lowland since beginning
-  
-sppelevations %>%
-  filter(Alpine.x != "Lowland") %>%
-  merge(read.csv(here("data", "elevations", "regions.csv"))) %>%
-  filter(Max.elevation <= Treeline) %>%
-  mutate(Alpine = "Lowland") %>%
-  select(Species, Alpine) %>%
-  unique -> lowland2 # Classified as lowland because max elevation in the treeline or below
-
-rbind(lowland1, lowland2) -> lowland # Merged list of lowland spp
-  
-# Merge lowland and alpine, with preference to alpine classification
-
-lowland %>%
[…]

### 2020-09-14 data/elevations/original1/elevations.R

@@ -1,107 +1,0 @@
-library(tidyverse); library(Taxonstand); library(here)
-
-## Australia
-
-read.csv(here("data", "elevations", "australian spp.csv")) %>%
-  rename(Min.elevation = Min.Elevation,
-         Max.elevation = Max.Elevation) %>%
-  rename(Taxon = NameDB) %>%
-  mutate(Alpine = "Alpine",
-         Region = "Australian Alps") %>%
-  select(Taxon, Region, Alpine, Min.elevation, Max.elevation) ->
-  AustralianSpp
-
-## Spain
-
-read.csv(here("data", "elevations", "spanish spp.csv")) %>%
-  separate(Habitat, c("Alpine", "Soil", "Frost"), sep = ",") %>%
-  mutate(Region = "Cantabrian Mountains") %>%
-  select(Taxon, Region, Alpine, Min.elevation, Max.elevation) ->
-  SpanishSpp
-
-## Flora alpina
-
-read.csv(here("data", "elevations", "flora alpina.csv")) %>%
-  rename(Taxon = Species) %>%
-  mutate(Region = "Southern Alps") %>%
-  select(Taxon, Region, Alpine, Min.elevation, Max.elevation) ->
-  SAlpsSpp
-
-SAlpsSpp %>%
-  mutate(Region = "Northern Alps") ->
-  NAlpsSpp
-
-## Russia
-
-read.csv(here("data", "elevations", "russian spp.csv")) %>%
-  filter(! Taxon %in% "") %>%
-  mutate(Region = "Caucasus",
-         Max.elevation = as.character(Max.elevation), Min.elevation = as.character(Min.elevation),
-         Max.elevation = ifelse(Max.elevation == "alpine", 2900, Max.elevation),
-         Max.elevation = ifelse(Max.elevation == "subalpine", 2400, Max.elevation),
-         Min.elevation = ifelse(Min.elevation == "alpine", 2400, Min.elevation),
-         Min.elevation = ifelse(Min.elevation == "subalpine", 1700, Min.elevation),
-         Min.elevation = ifelse(Min.elevation == "montane", 1000, Min.elevation),
-         Min.elevation = ifelse(Min.elevation == "lowlands", 0, Min.elevation),
-         Max.elevation = as.numeric(Max.elevation), Min.elevation = as.numeric(Min.elevation)) %>%
-  select(Taxon, Region, Alpine, Min.elevation, Max.elevation) ->
-  RussianSpp
[…]

### 2020-09-14 src/MCMCglmms.R

@@ -1,0 +1,92 @@
+library(MCMCglmm)
+
+# This script shows as a sample the code to fit the MCMCglmms models. Because these models take a long time to be computed,
+# the final output of the definitive models has been saved in the results folder, and is read from there
+# to build the manuscript (by the manuscript-source script).
+
+germination <- read.csv("results/germination.csv") # Read clean germination data
+traits <- read.csv("results/traits.csv") # Read clean species traits
+nnls <- ape::read.tree("data/carta/ALPINE.tree") # Read phylogenetic tree
+
+# MCMCglmm analyses
+
+## Table for MCMCglmm analyses
+
+traits %>%
+  merge(germination, by = "TPLName") %>%
+  merge(read.csv("data/fernandezpascual/regions.csv"), by = "Region", all.x = TRUE) %>%
+  mutate(TPLName = gsub(" ", "_", TPLName),
+         Macroregion = as.character(Macroregion),
+         Macroregion = ifelse(is.na(Macroregion), "Europe", Macroregion),
+         Stratification2 = as.character(Stratification),
+         Scarification = ifelse(Scarification == "Y", 1, 0),
+         Stratification = ifelse(Stratification == "None", 0, 1),
+         GA3 = ifelse(GA3 == "Y", 1, 0),
+         Light = ifelse(Light == "Y", 1, 0),
+         Alternating = ifelse(Alternating == "Y", 1, 0),
+         percent = Germinated / Germinable) %>%
+  rename(embryo = Embryo, mass = Seed.mass) %>%
+  mutate(animal = TPLName, ID = TPLName) %>%
+  select(animal, 
+         ID,
+         Alpine,
+         Dormancy, embryo, mass,
+         Source, Macroregion, Region, Accession,
+         Dish, Stratification2,
+         Scarification, GA3, 
+         Stratification,
+         Light, Alternating,
+         Tmean, Tdif, Temperature,
+         Sown, Germinated, Germinable,
+         percent,
+         MGT, MGR,
+         UNC, SYN) -> alpineDB
+
+## Binomial models
+
+priors <-  list(R = list(V = 1, nu = 50), 
+                G = list(G1 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 500), 
+                         G2 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 500),
[…]

### 2020-09-14 src/analysis.R

@@ -1,594 +1,0 @@
-library(tidyverse); library(here)
-
-# Final germination proportion
-
-## Figure for ms
-
-rbind(
-  rbind(
-    read.csv(here("results", "analysis", "proportion", "alpR.germ.general.csv")) %>%
-      select(model, post.mean, l.95..CI, u.95..CI, pMCMC) %>%
-      mutate(Moderator = "Alone"),
-    read.csv(here("results", "analysis", "proportion", "alpR.germ.general.embryo.csv")) %>%
-      select(model, post.mean, l.95..CI, u.95..CI, pMCMC) %>% 
-      mutate(Moderator = "Embryo"),
-    read.csv(here("results", "analysis", "proportion", "alpR.germ.general.mass.csv")) %>%
-      select(model, post.mean, l.95..CI, u.95..CI, pMCMC) %>% 
-      mutate(Moderator = "Mass")) %>%
-    rename(Mean = post.mean,
-           Lower = l.95..CI,
-           Upper = u.95..CI,
-           p = pMCMC,
-           Treatment = model) %>%
-    select(Moderator, Treatment, Mean, Lower, Upper, p),
-  read.csv(here("results", "analysis", "proportion", "alpR.germ.type.csv")) %>%
-    select(variable, post.mean, l.95..CI, u.95..CI, pMCMC) %>%
-    separate(variable, into = c("Moderator", "Treatment"), sep = ":") %>%
-    mutate(Moderator = gsub("Alpine", "", Moderator)) %>%
-    rename(Mean = post.mean,
-           Lower = l.95..CI,
-           Upper = u.95..CI,
-           p = pMCMC)) %>%
-  mutate(Moderator = recode(Moderator, Strict = "Strict alpine spp.",
-                            Generalist = "Generalist spp.",
-                            Mass = "Seed mass",
-                            Embryo = "Embryo:endosperm",
-                            Alone = "Main effect"),
-         Moderator = fct_relevel(Moderator, c("Embryo:endosperm", "Seed mass",  "Generalist spp.", "Strict alpine spp.", "Main effect")),
-         Treatment = recode(Treatment, Tmean = "Temperature", GA3 = "GA[3]"),
-         Treatment = fct_relevel(Treatment, c("Temperature", "Alternating", "Light", "Scarification", "Stratification", "GA[3]"))) %>%
[…]

### 2020-09-14 src/cleaning.R

@@ -1,12 +1,14 @@
-library(tidyverse); library(here); library(readxl)
+library(tidyverse); library(readxl)
 
-# WITH GERMINATION INDICES
+# This script has the code to read the primary germination data from the different contributors
+# and merge it into a clean germination dataset
 
-# Format Australian data (susanna Venn)
-# Read from original files with germination indices (S Rosbakh)
+# CLEAN GERMINATION DATA
 
-read_excel(here("data",
-                "Sergey Rosbakh (Indices)", "VENN hand joined.xlsx"),
+## Format Australian data (susanna Venn)
+## Read from original files with germination indices (S Rosbakh)
+
+read_excel("data/venn/VENN hand joined.xlsx",
            sheet = 1, skip = 0) %>%
   mutate(Region = "Australian Alps",
          Tmin = ifelse(is.na(Tmin), Tmax, Tmin),
@@ -40,36 +42,31 @@ read_excel(here("data",
          GRS:CVG) ->
   AustraliaVenn
 
-# Format Australian data (Annisa Satyanti)
-# Read from original files with germination indices (S Rosbakh)
+## Format Australian data (Annisa Satyanti)
+## Read from original files with germination indices (S Rosbakh)
 
-read_excel(here("data",
-                "Sergey Rosbakh (Indices)", "UPDATE_Aust Alpine Germination timing 11MAY2020.xlsx"),
+read_excel("data/satyanti/UPDATE_Aust Alpine Germination timing 11MAY2020.xlsx",
            sheet = 4, skip = 0) %>%
   select(Species, Accession, Elevation, Treatment, Rep, seeds, GRS:CVG) -> satyanti1
-read_excel(here("data",
-                "Sergey Rosbakh (Indices)", "UPDATE_Aust Alpine Germination timing 11MAY2020.xlsx"),
+read_excel("data/satyanti/UPDATE_Aust Alpine Germination timing 11MAY2020.xlsx",
            sheet = 8, skip = 0)  %>%
   select(Species, Accession, Elevation, Treatment, Rep, seeds, GRS:CVG) -> satyanti2
-read_excel(here("data",
-                "Sergey Rosbakh (Indices)", "UPDATE_Aust Alpine Germination timing 11MAY2020.xlsx"),
[…]

### 2020-09-14 src/manuscript-source.R

@@ -1,0 +1,563 @@
+library(tidyverse)
+
+# This script is sourced by the Rmarkdown file that creates the manuscript. It contains
+# the code to create the figures, tables and numerical output of the manuscript.
+
+# Prepare dataset object to create tables, figures, etc for the manuscript
+
+read.csv("../data/fernandezpascual/Regions.csv") %>%
+  merge(read.csv("../results/germination.csv"), by = "Region", all = TRUE) %>%
+  merge(read.csv("../results/traits.csv"), by = "TPLName") %>%
+  mutate(Region = as.character(Region),
+         Region = ifelse(is.na(Macroregion), "Enscobase", Region),
+         Region = ifelse(Region %in% c("Northern Alps", "Southern Alps"), "Alps", Region),
+         Region = ifelse(Region %in% c("Central Andes", "Southern Andes"), "Andes", Region),
+         Region = ifelse(Region %in% c("Qinghai-Tibet Plateau"), "Qinghai-Tibet", Region),
+         Region = ifelse(Region %in% c("Cantabrian Mountains"), "Cantabrian Mts", Region),     
+         Region = as.factor(Region)) -> dataset
+
+# Table 1
+
+data.frame(
+merge(
+merge(
+dataset %>%
+  mutate(Source = fct_recode(Source, `Fernández-Pascual` = "Fernández-Pascual Plant Biol")) %>%
+  mutate(Region = fct_recode(Region, `European Alps` = "Alps")) %>%
+  mutate(Region = fct_recode(Region, `Europe` = "Enscobase")) %>%
+  select(Source, Region, Country, TPLName, Alpine) %>%
+  group_by(Source, Region) %>%
+  tally %>%
+  rename(Dataset = Source, 
+         `Coverage` = Region,
+         Records = n),
+dataset %>%
+  mutate(Source = fct_recode(Source, `Fernández-Pascual` = "Fernández-Pascual Plant Biol")) %>%
+  mutate(Region = fct_recode(Region, `European Alps` = "Alps")) %>%
+  mutate(Region = fct_recode(Region, `Europe` = "Enscobase")) %>%
+  select(Source, Region, Country, TPLName, Alpine) %>%
+  unique %>%
+  group_by(Source, Region) %>%
+  tally %>%
+  rename(Dataset = Source, 
+         `Coverage` = Region,
[…]

### 2020-09-14 src/msSource.R

@@ -1,632 +1,0 @@
-library(tidyverse); library(here)
-
-read.csv(here("data", "Elevations", "Regions.csv")) %>%
-  merge(read.csv(here("results", "database", "germination.csv")), by = "Region", all = TRUE) %>%
-  merge(read.csv(here("results", "database", "traits.csv")), by = "TPLName") %>%
-  mutate(Region = as.character(Region),
-         Region = ifelse(is.na(Macroregion), "Enscobase", Region),
-         Region = ifelse(Region %in% c("Northern Alps", "Southern Alps"), "Alps", Region),
-         Region = ifelse(Region %in% c("Central Andes", "Southern Andes"), "Andes", Region),
-         Region = ifelse(Region %in% c("Qinghai-Tibet Plateau"), "Qinghai-Tibet", Region),
-         Region = ifelse(Region %in% c("Cantabrian Mountains"), "Cantabrian Mts", Region),     
-         Region = as.factor(Region)) -> dataset
-
-# Table with the datasets
-
-data.frame(
-merge(
-merge(
-dataset %>%
-  mutate(Source = fct_recode(Source, `Fernández-Pascual` = "Fernández-Pascual Plant Biol")) %>%
-  mutate(Region = fct_recode(Region, `European Alps` = "Alps")) %>%
-  mutate(Region = fct_recode(Region, `Europe` = "Enscobase")) %>%
-  select(Source, Region, Country, TPLName, Alpine) %>%
-  group_by(Source, Region) %>%
-  tally %>%
-  rename(Dataset = Source, 
-         `Coverage` = Region,
-         Records = n),
-dataset %>%
-  mutate(Source = fct_recode(Source, `Fernández-Pascual` = "Fernández-Pascual Plant Biol")) %>%
-  mutate(Region = fct_recode(Region, `European Alps` = "Alps")) %>%
-  mutate(Region = fct_recode(Region, `Europe` = "Enscobase")) %>%
-  select(Source, Region, Country, TPLName, Alpine) %>%
-  unique %>%
-  group_by(Source, Region) %>%
-  tally %>%
-  rename(Dataset = Source, 
-         `Coverage` = Region,
-         Species = n)),
-dataset %>%
-  filter(Alpine == "Strict") %>%
-  mutate(Source = fct_recode(Source, `Fernández-Pascual` = "Fernández-Pascual Plant Biol")) %>%
[…]

### 2020-09-14 src/preliminary report analysis.R

@@ -1,458 +1,0 @@
-library(tidyverse); library(lme4); library(cowplot)
-
-respuesta <- function(x) 1/(1+1/(exp(x)))
-
-# ALL SPECIES
-
-load("C:/EFP/Trabayu/Activos/2018 FICYT3/GT1 Traits/Joined DB/AlpineDB.RData")
-
-# TEMPERATURES
-
-d <- subset(AlpineDB, Scarification == "N" & GA3 == "N" & Light == "Y" & Stratification == "N" & Alternating == "Y")
-
-m <- glmer(cbind(Germinated, Germinable - Germinated) ~ factor(Temperature) + (1|Species) + (1|Region), family = "binomial", data = d)
-
-summary(m)
-
-ci <- confint(m, method = "Wald")
-
-p <- predict(m, expand.grid(Temperature = c(5, 10, 15, 20, 25)), re.form = NA)
-
-mt <- cbind(p, 
-            c(ci[3, 1], ci[4, 1] + ci[3, 1], ci[5, 1] + ci[3, 1], ci[6, 1] + ci[3, 1], ci[7, 1] + ci[3, 1]), 
-            c(ci[3, 2], ci[4, 2] + ci[3, 2], ci[5, 2] + ci[3, 2], ci[6, 2] + ci[3, 2], ci[7, 2] + ci[3, 2]))
-
-r <- respuesta(mt)
-
-e <- data.frame(Stratification = "Unstratified", Alternating = "Alternating",
-                 Temperature = c(5, 10, 15, 20, 25),
-                 Mean = r[, 1],
-                 Lower = r[, 2],
-                 Upper = r[, 3])
-
-tableTemperatures<- d %>% select(Region, Alpine, Species) %>% unique %>% group_by(Region, Alpine) %>% tally() %>% spread(Alpine, n)  
-
-table1b <- summary(m)
-
-plot1 <- ggplot(e, aes(x = Temperature, y = Mean, ymin = Lower, ymax = Upper, 
-                               fill = Temperature)) +
-  geom_bar(stat = "identity", position = "dodge", alpha = 0.7, show.legend = FALSE) + 
-  geom_errorbar(aes(color = Temperature), width = 1, size = 1, 
-                position = position_dodge(.9), show.legend = FALSE) +
-  coord_cartesian(ylim = c(0, 1)) +
-  labs(title = "All species", subtitle = "Germination peaks at 20ºC") +
[…]

### 2020-09-14 doc/cover-letter.Rmd
1st submission 🍋

---
title: "Cover letter"
author: "Eduardo Fernández Pascual"
date: "14/09/2020"
output: word_document
bibliography: refs.bib
csl: new-phytologist.csl
link-citations: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Dear Editorial Board,

I submit the manuscript *The seed germination spectrum of alpine plants: a global meta-analysis* to be considered for publication as a **full paper** in New Phytologist.

## What hypotheses or questions does this work address?

This is a global meta-analysis of primary germination data contributed by 12 research groups from 8 alpine regions in 4 continents. We used 9,799 records representing 62 seed plant families and 661 species to assess whether the germination ecological patterns detected by local studies are generalized at the global level.

## How does this work advance our current understanding of plant science?

We describe a global alpine germination syndrome, information valuable for all scientists working in alpine environments. Our results indicate a convergence of the seed germination spectrum in strict alpine species. We also found interactive effects of phylogeny and seed morphology that point to trade-offs shaping the seed ecological spectrum.

## Why is this work important and timely?

This is the first global evaluation, for a coherent habitat, of the seed germination spectrum and its relationship with other aspects of the seed ecological spectrum. This was one of the objectives in the research agenda for trait-based seed ecology recently published by New Phytologist [@RN4655].

## References

### 2020-09-14 doc/manuscript.Rmd
1st submission 🍋

@@
 ---
-title: "The global seed germination spectrum of alpine plants"
+title: "The seed germination spectrum of alpine plants: a global meta-analysis"
 output: 
@@
 We conducted a meta-analysis of primary data [@RN4734], which consists of pooling together original data from different local studies to synthesize their conclusions at the global level.
-The original data files, as well as R code for data cleaning, analysis and manuscript production are available at https://github.com/efernandezpascual/alpineseeds.
+All data processing, analysis and manuscript production has been performed in *R* [@RN2315].
+The original data, plus R code for the analysis and creation of the manuscript using Rmarkdown, can be accessed at the GitHub repository https://github.com/efernandezpascual/alpineseeds.
+(The repository will be made open when the manuscript is accepted for publication, a zip version has been uploaded to https://drive.google.com/file/d/12Usqi7fOAbOoZ9Uu9wkJt2FANB629LFP/view?usp=sharing for peer-review).
 To gather data, we contacted research groups which have performed experiments on seed germination of alpine plants, asking them to provide raw data from laboratory germination experiments.
@@
 
-Once received, we standardized the species names according to *The Plant List* [@RN2321] using the package *Taxonstand* [@RN4660] in *R* [@RN2315].
+Once received, we standardized the species names according to *The Plant List* [@RN2321] using the package *Taxonstand* [@RN4660] in *R*.
 For each species, we recorded its elevation range from local floras [@RN4741; @RN4739; @RN3243; @RN4742; @RN4740], catalogues [@RN4738] and herbaria (*Australian National Herbarium*, https://www.anbg.gov.au/cgi-bin/anhsir).
@@
 Upon publication, a version of record of the repository will be deposited in Zenodo.
+(The repository will be made open when the manuscript is accepted for publication, a zip version has been uploaded to https://drive.google.com/file/d/12Usqi7fOAbOoZ9Uu9wkJt2FANB629LFP/view?usp=sharing for peer-review)
 
@@
 When the credible intervals overlap with the zero-effect line, the effect is not significant.
-The figure shows first the main effect, then the effects for the strict alpine and generalist groups, and then the interaction with seed mass and embryo:endosperm.
+The figure shows first the main effect, then the effects for the strict alpine and generalist groups, and then the interaction with seed mass and embryo:seed.
 A negative interaction (e.g.

### 2020-10-27 doc/cover-letter.Rmd
major revision v2.1 ⛰️

@@
 ---
-title: "Cover letter"
-author: "Eduardo Fernández Pascual"
-date: "14/09/2020"
+title: "NPH-MS-2020-34477 - Response to reviewers"
+date: "`r format(Sys.time(), '%d/%m/%Y')`"
 output: word_document
@@
 
-Dear Editorial Board,
+# Editor
 
-I submit the manuscript *The seed germination spectrum of alpine plants: a global meta-analysis* to be considered for publication as a **full paper** in New Phytologist.
+> **I am writing you concerning your submitted manuscript NPH-MS-2020-34477 "The seed germination spectrum of alpine plants: a global meta-analysis”, which has now been reviewed.
+We found two reviewers and they both saw a great deal of merit in your manuscript.
+While both reviewers were quite enthusiastic about the topic and the way you’re analyzed your dataset, they also request that additional clarity be introduced at several parts of the manuscript.
+Reviewer #2 has also pointed out several aspects of the methodology that might require a different approach, especially with regard to how to incorporate light and phylogenetic signal.
+Finally, both reviewers also have numerous specific concerns to convey (see below).**
 
-## What hypotheses or questions does this work address?
+We respond below to each of the issues pointed out by the referees.
+We have adopted most of the proposed changes.
+Regarding light, we explain that this is an intrinsic limitation of the meta-analysis approach, and discuss the implications.
+As for the phylogenetic signal, we explain that this is explicitly taken into account by adopting a phylogenetic comparative statistical method such as the MCMCglmm, which considers phylogeny and phylogenetic signal within the analysis.
 
-This is a global meta-analysis of primary germination data contributed by 12 research groups from 8 alpine regions in 4 continents.
-We used 9,799 records representing 62 seed plant families and 661 species to assess whether the germination ecological patterns detected by local studies are generalized at the global level.
+# Referee 1
 
-## How does this work advance our current understanding of plant science?
+> **The introductory bullet point in your abstract is not strong, you may want to change the initial sentence (that rather describes part of the methodology) with one that sets the context for your study.
+Something in line with what you wrote in lines 103-107 would be nice.**
 
-We describe a global alpine germination syndrome, information valuable for all scientists working in alpine environments.
-Our results indicate a convergence of the seed germination spectrum in strict alpine species.
-We also found interactive effects of phylogeny and seed morphology that point to trade-offs shaping the seed ecological spectrum.
+We have changed the introductory point to *Given the micro-climatic heterogeneity of alpine landscapes, and the wealth of local studies on alpine seed ecology, a current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants.*.
 
-## Why is this work important and timely?
+> **Line 277: How many of the species you have included in this study have achieved the final germination proportion by 8 weeks?
+It would be nice to know the reason why you have chosen that timing as a response variable.**
 
-This is the first global evaluation, for a coherent habitat, of the seed germination spectrum and its relationship with other aspects of the seed ecological spectrum.
-This was one of the objectives in the research agenda for trait-based seed ecology recently published by New Phytologist [@RN4655].
+Eight weeks was a compromise deadline to unify datasets.
+There was only a small fraction of trials that had run for longer times, so we cropped these to eight weeks.
+We now indicate this in the main text.
 
-## References
+> **Line 282: Would you provide the equation, citation or more details for how you estimated the uncertainty index please?**
+
+The equation is provided by Lozano-Isla et al. [-@RN4752] in their article about the *GerminaR* package. We now comment this in the manuscript.
+
+> **Lines 289-296: Would the chosen timeframe influence in any way the uncertainty index?**
+
+We think that, rather than the timeframe, the uncertainty index could have been more affected by the germination scoring frequency (i.e.
+daily, every three days, etc.) which was not the same in all datasets.
+We have included this consideration in the *Random effects and phylogenetic signal* section of the *Results*.
+
+> **Line 330: Please check the ‘FAMD’ abbreviation spelling throughout the text.**
+
+We have corrected the spelling of FAMD throughout the text.
+
+> **Lines 402-404: The points for ‘strict alpine species’ and ‘embryo:seed’ data are missing in Figure 2b scarification panel.**
+
+The effects that are missing are those that overlap with zero (i.e.
+no effect) and also have very wide credible intervals (much wider than the other cues).
+They hamper the figure's visibility (i.e.
+they make all other effects appear too compressed), so we removed them from the plots.
+We have included this clarification in the figure's caption.
+
+> **Line 445: Figure 5 should be called earlier in this section. You should also label the axes (axis 1 and 2) and indicate to which panel you refer to in the text.**
+
+We have moved the call to the section's beginning and labelled the axes.
+Indicating to which panel we are referring to in the text is not easy, because the different panels/traits are intertwined in the descriptions, therefore it would require to make many recurrent calls (and, in our opinion, that would burden the text too much.)
+
+> **Legends for figures 4 and 5 should be expanded to make the figures more self-explaining, please include the meaning of abbreviations (UNC, MGT) and axes legends.
+Especially for figure 5 it would be useful to say which panels correspond to quantitative or qualitative data.**
+
+We have included the requested information in the figure captions.
+
+> **Would you enlarge axes legends in figure 1?**
+
+Done accordingly.
+
+> **Line 477: “… the frequency of dormancy…” as in ‘the frequency of dormancy class displayed by…’?**
+
+We have changed the sentence to *the frequency of dormancy classes displayed by strict alpine and generalist species is similar*.
+
+# Referee 2
+
+> **Upfront, the authors should define their object.
+They studied diaspores.
+Some of these are seeds, some are fruits (think of achenes in Asteraceae and caryopses in Poaceae).
+Once defined, you may continue with 'seed' for all practical reasons.**
+
+We have defined "seeds" upon first mention: *diaspores (i.e. seeds, fruits, or fruits with extra-ovarian structures attached; hereafter 'seeds')*.
+
+> **You also need to specify early on that this is for seasonal alpine settings (mostly high latitude).
+It would be a different situation in the humid tropics.
+Had anybody explored this for the tropics?
+[...] 360 please mention/recall the latitude range covered (here or earlier).
+All these are regions with a temperature seasonality.
+Alpine taxa from humid tropical habitats may show very different responses (e.g.
+no chill-requirement?).
+Since the title of the paper should not be lengthened, please specify early on that 'global' excludes tropics and subtropics (abstract, intro).
+Is there a knowledge gap for these latitudes?**
+
+Yes, there is a knowledge gap for tropical mountains.
+We were unable to find any significant amount of data from the tropics (excluding single species studies).
+We have qualified *global* as excluding the tropics throughout the manuscript, and commented on the knowledge gap in the conclusions.
+
+> **Light: It seems, the authors took 'light' during germination trials as a parameter in the sense of 'on or off' (germination in light or dark).
+However light carries substantial information by (a) its spectral composition, that is the R/FR ratio (think of tungsten light, LEDs, daylight) and (b) by photoperiod (the night to day length ratio).
+Both (a) and (b) affect germination.
+In fact, (b) belongs to the environmental cues that also prevent germination at the 'wrong time' (in conjunction with chilling requirement).
+[...] 221 ...respond positively to light.
+Please see my general remark.
+The action of light has many facets [...] 378 you really need to account for photoperiod.
+If the data are too poor, you need to mention at least the two potential quality effects of the light regime on germination.
+[...] 427 same issue [...] 550 'light conditions' is far too vague a phrase, given the multiple possible actions of the light regime [...] 565 again the light issue!**
+
+The referee is right that light is a more complex cue than the binary variable we considered.
+The same can be said about other cues, like alternating temperature (amplitude of the alternation, duration of the cycles...) or cold stratification (actual temperature, length...).
+However, to conduct a meta-analysis, we need to find a compromise between data quality and quantity.
+Detailed descriptions of light spectral composition and photoperiod are only available for a small fraction of the whole dataset.
+Therefore, we must summarise cues into some variables (like light absence/presence) that are routinely recorded in the majority of trials.
+We have added these considerations to the conclusions.
+
+> **Seed storage type.
+There are two major types of storage: endosperm (e.g.
+Poaceae) and (e.g.
+Asteraceae, Fabaceae) cotyledons.
+So these should be specified as such and you need to acknowlege that these seed traits are phylogenetically confounded, hence you cannot separate them from taxonomic relatedness.
+This means the abundance of taxa in your data set that belong to either type influences the outcome of the analysis.
+[...] 370 storage type...
+see my general comment [...] 470 mention/address the phylogenetic bias issue and whether you could solve it.
+[...] 606 seed storage type and phylogeny are confounded [...] 624 are you sure it is phylogeny or is it seed morphology across families?
+[]**
+
+We considered storage type as a continuum (i.e.
+embryo:seed ratio), ranging from species with small embryos with copious endosperm (often referred to as endospermic) to species with no endosperm at all and all nutrient reserves stored in the embryo (often referred to as non-endospermic seeds).
+To improve understanding, we have included a definition of non-endospermic seeds as *non-endospermic seeds that store nutrients in the cotyledons*.
+In any comparative biology analysis, all traits are always phylogenetically confounded.
+This is why we used a phylogenetic comparative analysis such as the MCMCglmm.
+The analysis includes, as a random factor, a matrix of phylogenetic distances among species.
+In this way, traits are analysed in relation to the underlying phylogenetic structure of the dataset.
+In our models, both the trait (seed:endosperm ratio) and phylogeny had significant effects.
+The use of phylogenetic comparative analysis is covered in the analysis section of the methods section.
+
+> **Dormancy: A common distinction in the phenology literature is 'endodormancy' versus 'ecodormancy'.
+Both these dormancies are physiological states, hence there is no non-physiological dormancy.
+Endodormancy means an internal state controlled by genes and hormons.
+The drivers of endodormancy are (a) an (unknown) internal clock that causes germination to occur over a wide time window (in cohorts, staggering) between a few weeks and 100 years (Verbascum) and (b) environmental cues ('switches') that operate at gene level (e.g.
+chilling requirement, passage through an animal gut).
+Ecodormancy indicates an active (turgid) seed, ready to germinate when the environment permits (T, moisture).
+Admittedly, there may be intermediate states, when e.g.
+R/FR comes into play.
+I think, this just needs a bit of better phrasing.
+[...] 264 delayed embryogenesis is quite common.
+See e.g.
+Akhalkatsi M, Wagner J (1996) Reproductive phenology and seed development of Gentianella caucasea in different habitats in the Central Caucasus.
+Flora 191:161-168 [...] 364 dormancy...
+see my general comment [...] 465 dormancy is always a physiological state.
+See general comment**
+
+In our manuscript we use the definition and classification of dormancy that is currently dominant in seed ecology and physiology.
+This framework is based on two main references: (i) the concept by Vleeshouwers et al.
+[-@RN3254] that dormancy is an inner seed property ("endodormancy"), which should be separated from the absence of appropriate germination cues in the environment ("ecodormancy"); and the classification system proposed by Baskin & Baskin [-@RN3261] based on previous work by Nikolaeva et al.
+[-@RN2877] and which considers morphological and physical dormancy separated from physiological.
+There are two practical reasons to use this framework: first, it will be readily understandable to the widest scope of readers; and second, it provides the largest amount of species trait data available [@RN3214; @RN4743].
+We have retained the definition and classification of dormancy originally used in our paper, however, we have also included explanation in the text of how this classification corresponds to the terminology preferred by the reviewer.
+
+> **Conclusion/message: I found the message vague.
+First you mention that alpine seeds are special, but than you do NOT specify what is special, and escape into recommendations for further research (not very concrete ones) and that the data may be a useful reference (they surely are).
+If you wish to add a methods advice, this would best be pointing at the specific limitations of the current data or questions that had not even been addressed.
+If you find that alpine plants are NOT special (this is the impression I get from the data), then why not state this?
+This would also be an important message.
+Don't your findings match what is known from non-ruderal seeds at low elevation?
+If I am wrong, you really need to specify what is special.
+Fenner M (1985) Seed ecology.
+Chapman and Hall, London, might be a source for comparing alpine with other habitats /always excluding (!) ruderals/pioneers (which are opportunistic with regard to germination) and trees, which often employ mass seeding, shade tolerance, and are fighting with predation.
+[...] As I said at the beginning, the conclusions really need upgrading.
+If you think something is special, you need to verify this by comparing with non-alpine data.
+If you arrive at concluding that most responses are not special (similar to lowland herbs/grasses, check Fenner and literature cited therein), but a few are indeed special, this needs to be phrased into a differential take home message.
+As phrased now, the conclusion tells the reader that things varied, you were busy and that more work is needed....
+You would be frustrated reading such conclusion yourself.
+[...] 617 If you think there is something special, this is the place to highlight it specifically.
+[...] 620 Please explain the convergence.
+What exactly do you think converges?
+In which direction?
+Please avoid recalling your activitiesbut rather focus on the main results/take home message.
+Potential relatedness?
+There are always 'potentials' for relatedness.
+Can you name specific traits/responses**
+
+We have revised the entire conclusions paragraph to address these concerns, being more specific about both the conclusions and about the limitations of the data.
+
+> **A language suggestion: better avoid words like change, vary, interplay, interwined, interact.
+All these can be replaced by or combined with directional terms that provide an answer.
+Things always vary, interplay etc.
+That by itself is not worth mentioning.
+If there are interactions you better write them out.
+[...] 523 try to avoid neutral (non-directional) statements.
+See general comment.**
+
+The referee is right and we have screened the text for this class of words.
+*Change* is used to refer to a concept like climate change, or to encompass and summarise changes that can go in different directions (environmental changes, snow cover changes).
+*Interplay* and *intertwined* are used in places like the summary and conclusions, where we must summarise concepts that are detailed elsewhere.
+*Interaction* is used in its statistical meaning, and qualified as *no interaction*, *negative interaction*, etc.
+We could not find *vary*.
+Regarding the specific point at L523, this is not a neutral statement: the direction of *interplay* is explained in the same sentence, after the *;*.
+To make this point clear, we have substituted the *;* with a *:*.
+
+> **In the characterisation of alpine habitats I suggest to remove 'wind'.
+Mountains are in fact less windy than the plains (except for summits and ridges where weather stations often happen to be located) and the most characteristic feature of alpine plants is their aerodynamic decoupling from the free atmosphere.
+I am not saying that wind may not contribute to pollination and seed dispersal, but not in a sense that alpine plants face special wind 'problems'.**
+
+We have removed wind.
+
+> **Line 64 If I may, I would suggest to remove that vulnerability 'soap'.
+Things do not become more true if repeated more often.
+Many authors (including IPCC) confuse the vulnerability of human infrastructure in mountains with that of mountain biota.
+And of all mountain biota, the alpine one is the least vulnerable in terms of fatal environmental influences other than human intervention with bulldozers on ski slopes.
+I consider the topography driven habitat diversity a particular safety factor, by offering alternative habitat conditions at shorter geographical distance than anywhere else (see Scherer and Körner 2011 J Biogeography).
+If climatic warming were such a threat as many argue, we should have lost a great deal of alpine flora in the postglacial 'Atlanticum' period (some 8000 years ago, a period during which most glaciers had gone, the climate was c.
+2 K warmer and the treeline 250 m higher (not sure what happened to the Australian alpine then?).
+If enhanced summer drought comes into play (Chile, Sierra Nevada), then say so.
+GLORIA seems to suggest there is a drought issue in Mediterranean mountain settings.
+I prefer such works to start with scientific motives rather than potential usefulness for society.
+Admittedly, it became a fashion to use such populistic motives.
+At the least, calm down with the alarmistic tone...
+Further, the ecosystem services story is unneeded here.
+Other then securing soils against erosion there are hardly any such 'services' for 'millions', and erosion control by vegetation is always true.
+Körner and Ohsawa talk about the entire mountain terrain and not just the small alpine fraction.
+That makes a big difference...
+73ff reads very well, in contrast.**
+
+We have removed the *soap*, and now the two introductory sentences link directly to former L73.
+
+> **103 heterogeneity of alpine climate?
+Not sure what you are referring to.
+The fact that different alpine regions have different climate?
+Or, you mean microclimatic variation within a given alpine landscape?**
+
+To the latter.
+To be more concrete, we have changed the sentence to *Given the high micro-climatic heterogeneity of alpine landscapes, and the wealth of local studies on alpine seed ecology (reviewed in the following paragraphs), a current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants*.
+
+> **178 -10 cm T seems to be quite irrelevant for germination.
+I guess everything deeper than 1 cm would rarely facilitate germination?
+Hence, solar radiation driven fluctuation of T is the norm.
+I do appreciate language barriers, but the following study is a must to consider.
+This author explored 197 arctic-alpine species and accumulated an enormous wealth of data.
+Peter Poschlod could help with that text.
+You find a brief summary in Alpine Plant Life.
+Soeyrinki N (1938) Studien über die generative und vegetative Vermehrung der Samenpflanzen in der alpinen Vegetation Petsamo-Lappland.
+1.
+Allgemeiner Teil.
+Ann Bot Soc Zool Bot Fenn Vanamo 11:83-93, 124-147  There may be a second paper/part.
+Name spelling could be an issue, try Söyrinki.
+I recall that he explored temperature variation and arrived at concluding that it is essential for many alpine species to properly germinate.
+[...] 555 see Söyrienki above**
+
+We have removed the reference to a hard set depth boundary.
+The depth profile of temperature fluctuations will depend on soil type, vegetation cover, snow cover.
+Specifically, the insulating effect of snow cover means that fluctuation of T is not necessarily the norm.
+We have changed the sentence to *Fluctuations of diurnal temperature decrease with increasing burial depths and the depth profile of this decrease depends on soil type, vegetation cover and snow*.
+We have read Söyrinki's work and cited it where appropriate, however, we did not find that he explicitly tested the effect of fluctuating temperatures.
+
+> **270 not really clear to me what sort of ratios you talk about.
+In a 2D microscopy image you could quantify fraction of area covered or circumference.
+What is the measure?
+length, area, volume?
+A reference is not enough.**
+
+We rewrote more clearly in the methods section *Embryo to seed surface area values were calculated by dividing embryo area by the surface area of the seed, more specifically embryo plus endosperm and perisperm.
+Measurements were made using the ImageJ software (Schneider et al.
+2012) on drawings and photographs of seeds cut in half along the longitudinal axis, retrieved mainly from Martin (1946) and the Royal Botanic Gardens Kew Seed Information Database (2017), supplemented with own photographs (Vandelook, unpublished).*.
+In our opinion this is the most straightforward method to obtain a good perception of the amount of nutrient reserves stored either inside or outside the embryo without going into more advanced 3D methods.
+It was approximately the same methodology as the one applied by Forbis et al.
+[-@RN4958].
+
+> **274 genus or family averages? You really think this works? Within a family the spectrum of trait variation could be far bigger than among families...**
+
+It works as a proxy in as much as these two traits consistently show a strong phylogenetic signal [@RN4958; @RN4916; @RN3685], as we mention in the discussion.
+We must keep in mind that the purpose of the meta-analysis is to detect a pattern across the studied alpine flora, rather than to accurately predict the specific traits of any given species.
+In this sense, we would argue that, if it had not worked, we would have found no pattern or random noise, rather than an effect that is consistent with ecological theory and previous research.
+
+> **296 Did you, and if, how did you standardize trial temperatures for comparison?
+At least this needs a comment.
+It is a pity how little data on actual field temperatures in seedbeds are available (if any) in order to judge the suitability of temperatures during lab trials.
+It is very hard to obtain trustworthy seedbed temperatures.
+Possible you need fine flexible probes (thermocouples, thermistors, Pt100), securely embedded into the seed bed.
+Would be great to install such a trial in the field and observe germination dynamics in relation to actual temperature.
+Because top soils desiccate within a few hours of direct insolation, it may need a very special moisture regime in combination with temperature for successful germination and seedling establishment.
+Given the clonal nature (also treated by Söyrinki), successful seedling establishment needs to occur only once in perhaps half a century to retain a population.
+One would expect micro-evolutionary selection for such coincidental conditions....
+This would be a more useful topic in the introduction than the vulnerability debate.**
+
+Trial temperatures are the ones reported in the original datasets, and they correspond to daily temperature regimes programmed in an incubator/growth chamber.
+For analysis, each trial is represented by two temperature variables: (1) the *average temperature* (i.e.
+the weighted average of the day and night temperatures, weighted by the duration of each phase) and (2) *alternating temperature* (i.e.
+a binary variable describing whether the trial applied constant temperature versus different temperatures during the day and the night, in diurnal cycles).
+These two variables are described later in the text, in the *Description of the dataset* section of the *Results*.
+
+> **531 40 °C would be an exceptionally low heat resistance.
+To my knowledge alpine plants always survive 46 °C and a 48 °C threshold is rather the norm than the exception.
+Dehydration could be a problem with such trials.
+Any evidence that seelings are more sensitive than adult tissue?**
+
+We have removed the adjective *high* and changed the sentence to *yet tolerate heat up to 40–50°C*.
+
+> **535 staggering is true for almost all herbs/graminoids (except for ruderal weeds, but even these perform it for risk mitigation...
+classical works by John Harper I think).
+So this is nothing special with alpine plants.**
+
+Indeed, this is not something exclusive of alpine environments. We have changed the sentence to *a form of adaptation to unpredictable environments*.
+
+# References

### 2020-10-27 doc/manuscript.Rmd
major revision v2.1 ⛰️

@@
   bookdown::word_document2:
+    number_sections: false
     reference_docx: template.docx
@@
 
-* We present a meta-analysis of the alpine seed germination spectrum based on primary data from germination experiments conducted in four continents with `r MSspecies` plant species, with the aim of testing the generality of alpine germination patterns at the global level.
-* We use phylogenetic Generalized Linear Mixed Models to estimate the influence of six environmental cues on seed germination proportion, mean germination time and germination synchrony; accounting for possible effects of seed morphology (mass, embryo:seed ratio) and phylogeny.
-* Alpine plants are characterized by physiological seed dormancy, strong need for cold stratification, warm-cued germination and positive germination responses to light and alternating temperatures.
+* Given the micro-climatic heterogeneity of alpine landscapes, and the wealth of local studies on alpine seed ecology, a current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants.
+* We meta-analyse primary data from laboratory experiments conducted across four continents (excluding the tropics) and 661 species, to estimate the influence of six environmental cues on seed germination proportion, mean germination time and germination synchrony; accounting for possible effects of seed morphology (mass, embryo:seed ratio) and phylogeny.
+* Alpine plants show physiological seed dormancy, a strong need for cold stratification, warm-cued germination and positive germination responses to light and alternating temperatures.
 Species with a distribution limited to the alpine belt have a higher preference for warm temperatures and a stronger response to cold stratification than species whose distribution extends also to the lowlands.
 Seed mass, embryo size and phylogeny have strong constraining effects on germination responses to the environment.
-* Overwintering and warm-cued germination are key drivers of germination in alpine habitats.
-The interplay of seed morphological traits and germination physiology reflects pressures to avoid frost or drought stress.
-These results indicate the convergence, at the global level, of the seed germination spectrum of alpine species.
+* Globally, overwintering and warm-cued germination are key drivers of germination in alpine habitats.
+The interplay between germination physiology and seed morphological traits further reflects pressures to avoid frost or drought stress.
+Our results indicate the convergence, at the global level, of the seed germination spectrum of alpine species.
 
@@
 
-Alpine environments occur worldwide, above the maximum elevation at which trees can grow naturally [@RN3000; @RN4750].
-These treeless habitats are characterized by low temperatures, strong winds, unstable substrates and short growing seasons [@RN2392].
-Their dependence on extreme and strict climatic conditions [@RN3024], coupled with the relatively small terrestrial area they occupy [@RN4750], makes alpine environments disproportionally vulnerable to climate change [@RN4946; @RN4950].
-This threatens the survival of the world’s rich and specialized mountain biodiversity [@RN4947].
-It also endangers the sustainability of the ecosystem services that alpine environments provide for hundreds of millions worldwide, including the provisioning of natural resources, the regulation of environmental hazards and the maintenance of cultural and recreational activities [@RN4949; @RN4948].
-Succesful management of climate change and of its consequences will depend on an adequate understanding of alpine biodiversity.
-
-Specially, plant regeneration by seed is a key process that determines whether species will persist in their current habitat or will be able to migrate to new sites [@RN3055]; and it is also fundamental for active efforts in plant conservation [@RN3763] and seed-based ecological restoration [@RN3507; @RN2241].
-Although most alpine plants can reproduce clonally [@RN2392], sexual reproduction by seed is the main strategy to maintain genetic diversity and colonize suitable new sites in response to environmental changes.
+Alpine environments occur worldwide, above the maximum elevation at which trees can grow naturally [@RN3000; @RN3024; @RN4750].
+These treeless habitats are characterized by low temperatures, unstable substrates and short growing seasons [@RN2392].
+Although most alpine plants can reproduce clonally [@RN2392], sexual reproduction by diaspores (i.e.
+seeds, fruits, or fruits with extra-ovarian structures attached; hereafter 'seeds') is the main strategy to maintain genetic diversity and colonize suitable new sites in response to environmental changes.
+Plant regeneration by seed is a key process that determines whether species will persist in their current habitat or will be able to migrate to new sites [@RN3055]; and it is also fundamental for plant conservation [@RN3763] and seed-based ecological restoration [@RN3507; @RN2241].
 However, the short reproductive season of the alpine habitat constrains the phenological timing of flowering, pollination, seed development and seed dispersal.
-For regeneration by seed to be succesful, dispersed propagules must germinate in the most appropriate period to ensure the survival of seedlings under environmental stress [@RN4714; @RN2868; @RN4717].
+For regeneration by seed to be successful, dispersed propagules must germinate in the most appropriate period to ensure the survival of seedlings [@RN4714; @RN2868; @RN4717].
 The physiological process of seed germination is therefore an essential life stage that must be timed to occur when the environment is favourable for subsequent seedling survival and growth [@RN4691].
@@
 The propagules of many plant species have also developed seed dormancy, by which germination is prevented during periods that are only ephemerally favourable, like a short warm spell in the middle of winter [@RN3214].
-Different degrees of dormancy within a seed population work to ensure the distribution of seedling emergence across time, thereby bet hedging offspring survival against unpredictable environments [@RN3065].
+Different degrees of dormancy within a seed population ensure the distribution of seedling emergence across time, a form of bet-hedging offspring survival against unpredictable environments [@RN3065].
 Other germination cues, such as the response to diurnal alternating temperatures [@RN1380] and the need for light or darkness [@RN4656], allow for a fine-scale detection of germination micro-niches and safe sites [@RN4719].
-Given the comparatively high heterogeneity of alpine climates and plant lineages adapted to different regions, and the wealth of local studies on alpine seed ecology (reviewed in the following paragraphs), a current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants while accounting for the evolutionary relatedness among taxa.
+Given the high micro-climatic heterogeneity of alpine landscapes, and the wealth of local studies on alpine seed ecology (reviewed in the following paragraphs), a current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants.
 The seed ecological spectrum [@RN4655] is defined as a set of seed traits that determines the ability of plants to disperse, persist, germinate, and establish in different habitats.
-A major goal for trait-based seed ecology is to study the relationships and trade-offs between different aspects of the seed ecological spectrum [@RN4655], such as the physiological drivers of germination versus key morphological traits like seed mass [@RN4916] and the embryo to seed ratio [@RN3685].
+A major goal for trait-based seed ecology is therefore to study the relationships and trade-offs between different aspects of the seed ecological spectrum [@RN4655], such as the physiological drivers of germination versus key morphological traits like seed mass [@RN4916] and the embryo to seed ratio [@RN3685].
 
-In the available literature, many alpine species have been described as having seeds with deep physiological dormancy, i.e.
+Traditionally, alpine species have been described as having seeds with physiological dormancy, i.e.
 a type of dormancy that is caused by the seed’s internal balance of phytohormones and requires a long exposure to dormancy-breaking treatments to be terminated [@RN2943; @RN3703; @RN3214].
 This physiological dormancy would prevent precocious germination of recently dispersed seeds during autumn, when appropriate conditions are not likely to persist for more than a few weeks or days [@RN3330].
-A requirement for cold stratification to break dormancy would allow seeds to sense the snow season, thereby postponing germination to a more favourable period for seedling survival and development.
-Thus, it can be expected that a cold stratification period is a common requirement for seed germination in alpine plants.
-Nonetheless, an early review [@RN3213] suggested that cold stratification was not a requirement for the seed germination of several alpine species from different mountain ranges in the USA, a finding that has been reported by subsequent studies [@RN1484; @RN2995; @RN4721].
+A requirement for cold stratification to break dormancy allows seeds to sense the snow season, thereby postponing germination to a more favourable period for seedling survival and development.
+Thus, it can be expected that a cold stratification period is a requirement for seed germination in alpine plants.
+One of the first experiments on this topic was conducted by Söyrinki [-@RN4951], who showed how the storage of arctic-alpine species at winter temperatures (i.e.
+cold stratification) increased seed germination in most species.
+Nonetheless, Amen [-@RN3213] later suggested that cold stratification was not a requirement for the seed germination of alpine species from different mountain ranges in the USA, a finding that was supported by subsequent studies [@RN1484; @RN2995; @RN4721].
 More recently, research on Australian alpine plants found that a cold stratification period only increased germination in half of the species tested [@RN3703; @RN3285].
-In contrast, cold stratification increased seed germination over a range of temperatures in most of the species studied in the alpine zone of Japan [@RN707].
-Along an elevational gradient in the central Chilean Andes, cold stratification was an important requirement for seed germination in species from lower elevations, while species from higher sites did not germinate even after cold stratification [@RN4713].
+Cold stratification also increased seed germination over a range of temperatures in most of the species studied in the alpine zone of Japan [@RN707].
+Along an elevational gradient in the central Chilean Andes, cold stratification was an important requirement for seed germination in species from lower elevations, while species from higher sites did not germinate after a single period of cold stratification [@RN4713].
 Finally, in a comparative study on germination of unstratified seeds from different habitats of the Austrian Alps, species from alpine grasslands had the lowest final germination proportion, suggesting that they had the highest level of dormancy [@RN4930].
@@
 Seed dormancy is not the only trait that can prevent, in practice, autumn germination.
-Early studies on the germination ecology of alpine plants demonstrated that in most species recently dispersed seeds require relatively high temperatures for germination [@RN3258; @RN3213; @RN4712].
+Early studies on the germination ecology of alpine plants demonstrated that in most species recently dispersed seeds require relatively high temperatures for germination [@RN4951; @RN3258; @RN3213; @RN4712].
 Warm-cued germination has been considered as an adaptation to prevent seed germination at the time of seed dispersal (autumn) when temperatures are relatively low and there is a high risk of frost [@RN3393].
 Indeed, germination of alpine seeds tends to occur after winter, mainly in early summer [@RN2392; @RN2382].
-In recent times, an increasing number of studies reported that fresh seeds of some alpine plants also germinate at cool incubation temperatures or during cold stratification treatments [@RN2943; @RN3285; @RN2371; @RN4713].
+In recent times, an increasing number of studies reported that fresh seeds of some alpine plants can also germinate at cool incubation temperatures or during cold stratification treatments [@RN2943; @RN3285; @RN2371; @RN4713].
 This cold-cued germination could be an adaptation to germinate under snow or during snowmelt, which could presumably allow seedlings to develop a deeper root system before topsoil desiccation in summer [@RN4720], or to attain an optimal size for overwintering [@RN4712].
 Yet, the generality of warm-cued germination across different alpine regions and habitats has never been studied systematically, and the relative importance of cold stratification versus warm-cued germination remains unclear.
-A central question is whether physiological seed dormancy plays a relevant role in postponing alpine germination during autumn and winter, or the warm-cued germination is enough to prevent germination under cool autumn temperatures.
+A central question is whether physiological seed dormancy plays a relevant role in postponing alpine germination during autumn and winter, or whether the warm-cued germination is enough to prevent germination under cool autumn temperatures.
 The response to this question becomes pertinent under a warming climate, which could result in non-dormant seeds germinating readily in autumn.
@@
 Potentially, having a non-dormant or less dormant seed could be a better strategy for alpine plants to thrive under future climate [@RN3317; @RN2382].
-However, we might also expect that seedlings from these species will be exposed to frost damage if snow starts to melt earlier and insulation during winter disappears [@RN4944] and therefore a mechanism that enables diverse germination timing (i.e.
-low germination synchrony) should be preferred.
+However, we might also expect that seedlings from these species will be exposed to frost damage if snow starts to melt earlier and insulation during winter disappears [@RN4944] and therefore mechanisms that enable diverse germination timing (i.e.
+low germination synchrony) may be favoured.
 
 Besides patterns of seed dormancy and germination temperature, the response of alpine seeds to diurnal alternating temperatures and light can provide further information on their capacity to detect fine-scale environmental cues.
-Fluctuations of diurnal temperature decrease with increasing burial depths; below 10 cm soil temperature might be too stable to trigger germination in species requiring temperature alternation [@RN4910].
+Fluctuations of diurnal temperature decrease with increasing burial depths and the depth profile of this decrease depends on soil type, vegetation cover and snow [@RN1380; @RN4910; @RN4952].
 Solar irradiance in physiologically significant quantities penetrates only the first millimetres of the soil [@RN4909].
-Therefore, these two cues can indicate whether seeds are buried deep in the soil, in rock crevices or under snow.
+Therefore, the response to alternating temperatures and light can indicate seeds whether they are buried deep in the soil, in rock crevices or under snow.
 The depth of seed burial is crucial for seedling emergence, as small seeds lack the resources to survive until they reach the soil surface [@RN4898].
@@
 
-In this work, we make a global collection of primary germination data from laboratory experiments conducted with seeds of alpine plant species.
-The assembled dataset represents eight alpine regions belonging to four continents, and we use it to investigate the germination response of alpine plants to key environmental factors.
-These factors include dormancy-breaking treatments (cold stratification, GA~3~ and scarification), average germination temperature, diurnal alternating temperatures and light [@RN3368; @RN3214].
+In this work, we make a global collection of primary germination data from laboratory experiments conducted with seeds of alpine plant species from extratropical regions.
+The assembled dataset represents eight seasonal alpine regions belonging to four continents.
+We meta-analyse this dataset to investigate the germination response of alpine plants to key environmental factors: dormancy-breaking treatments (cold stratification, GA~3~ and scarification), average germination temperature, diurnal alternating temperatures and light [@RN3368; @RN3214].
 Given the heterogeneity of alpine species lineages adapted to different regions, we account for the shared evolutionary history using a phylogenetic comparative approach [@RN3081].
-We also analyse the relationship between germination ecology and two morphological traits which are key drivers of the seed ecological spectrum: seed mass and the embryo to seed ratio [@RN4655].
+We further analyse the relationships between germination ecology and two morphological traits which are key drivers of the seed ecological spectrum: seed mass and the embryo to seed ratio [@RN4655].
 Our main aim in conducting this work is to test the consistency of general assumptions about the influence of major environmental cues on seed germination of alpine plants.
 Specifically, we determine at the global scale whether alpine seeds: (1) have physiological dormancy; (2) require cold stratification to break dormancy; (3) need relatively warm temperatures to germinate; (4) do not require alternating temperatures for germination; and (5) respond positively to light.
-To understand the ecological constraints of our meta-analysis on alpine species pools, we analyse how the germination responses differ between strict alpine species (that occur exclusively above the treeline) and generalist species (that occur above the treeline but also in the lowlands).
+To understand the ecological constraints of our meta-analysis on alpine species pools, we compare the germination responses of strict alpine species (that occur exclusively above the treeline) and generalist species (that occur above the treeline but also at lower elevations).
 We can expect that the former group will show a more convergent alpine germination syndrome, while the latter will be more plastic in their response.
@@
 The submitted data had to report the studied species, the germination treatments, the number of seeds sown, and the number of seeds germinated.
-We received data (Table 1) from the Cantabrian Mountains of Spain, the European Alps, the Italian Apennines, the Northern Caucasus, the Qinghai-Tibet Plateau of China, the Chilean Andes, and the Australian Alps.
+We received data (Table 1) on species from the Cantabrian Mountains of Spain, the European Alps, the Italian Apennines, the Northern Caucasus, the Qinghai-Tibet Plateau of China, the Chilean Andes, and the Australian Alps.
 We also received data on European alpine species from *Enscobase*, the database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml).
@@
 
-Once received, we standardized the species names according to *The Plant List* [@RN2321] using the package *Taxonstand* [@RN4660] in *R*.
+We standardized the species names according to *The Plant List* [@RN2321] using the package *Taxonstand* [@RN4660] in *R*.
 For each species, we recorded its elevation range from local floras [@RN4741; @RN4739; @RN3243; @RN4742; @RN4740], catalogues [@RN4738] and herbaria (*Australian National Herbarium*, https://www.anbg.gov.au/cgi-bin/anhsir).
@@
 species that can grow above the treeline but also in the lowlands).
-For each species, we compiled three additional traits.
-First, seed dormancy class obtained from Baskin & Baskin [-@RN3214] and Rosbakh et al.
+For each species, we compiled three additional traits: seed dormancy class, seed mass and embryo:seed ratio.
+
+We obtained seed dormancy class from Baskin & Baskin [-@RN3214] and Rosbakh et al.
 [-@RN4743].
 We simplified the original dormancy classes to five categories: physiological (germination prevented by the seed’s internal balance of phytohormones), morphological (germination prevented by an embryo which is underdeveloped at the time of dispersal), morphophysiological (both physiological and morphological dormancy present), physical (germination prevented by a water-impermeable seed coat) or non-dormant.
-Second, seed mass as provided by the Royal Botanic Gardens Kew Seed Information Database [-@RN2987] and some additional sources [@RN4728; @RN4735; @RN3376].
-Third, the embryo to seed surface ratio.
-Embryo to surface area values were based on measurements of the surface area of the embryo and seed from drawings and photographs retrieved mainly from Martin [-@RN3244] and the Royal Botanic Gardens Kew Seed Information Database [-@RN2987], supplemented with own measurements (Vandelook, unpublished).
+We followed the concept by Vleeshouwers et al.
+[-@RN3254] to define dormancy as an inner seed property (sometimes called "endodormancy"), which should be separated from the absence of appropriate germination cues in the environment ("ecodormancy"); and the classification system of Baskin and Baskin [-@RN3261] based on previous work by Nikolaeva et al.
+[-@RN2877] which considers morphological and physical dormancy as classes separated from physiological dormancy.
+
+We collected mean seed mass for each species as provided by the Royal Botanic Gardens Kew Seed Information Database [-@RN2987] and additional sources [@RN4728; @RN4735; @RN3376].
+We recorded the embryo to seed surface area ratio for each species.
+Embryo to seed surface area values were calculated by dividing embryo area by the surface area of the seed, more specifically embryo plus endosperm and perisperm.
+Measurements were made using the *ImageJ* software [@RN4708] on drawings and photographs of seeds cut in half along the longitudinal axis, retrieved mainly from Martin [-@RN3244] and the Royal Botanic Gardens Kew Seed Information Database [-@RN2987], supplemented with own photographs (Vandelook, unpublished).
 For some species that were missing seed mass or embryo values, we calculated genus or family averages.
@@
 the number of seeds germinated out of seeds sown) at 8 weeks as the main response variable for the germination treatments.
+Eight weeks was a compromise deadline to unify datasets, as there was a small fraction of trials that had run for longer times.
 We further calculated the mean germination time and the germination uncertainty index with the package *GerminaR* [@RN4752].
+Equations for these two germination indices are available in Lozano-Isla et al.
+[-@RN4752].
 Mean germination time is the time that it takes for half of the seed lot to germinate, indicating the germination speed in each experimental condition.
-The germination uncertainty index estimates how scattered germination is through time, and gives an idea of whether germination is synchronous (i.e.
+The germination uncertainty index estimates how scattered germination is through time, giving an idea of whether germination is synchronous (i.e.
 most seeds germinate around the mean germination time) or asynchronous (i.e.
@@
 To calculate these indices, we used only records that met two conditions: (i) more than 50% final germination percentage; and (ii) a mean germination time under 4 weeks.
-We established these limits to prevent the results being altered by a modification of the seed’s dormancy status during the germination incubation (e.g.
+Within these limits, we were able to calculate mean germination time and uncertainty for `r MSfamdspp` species.
+We established this conservative approach to prevent the results being altered by a modification of the seed’s dormancy status during the germination incubation (e.g.
 a fraction of the seed lot is non-dormant at the beginning of the incubation and germinates shortly after the start of the experiment; while another fraction is originally dormant, loses dormancy in response to the incubation conditions, and germinates in a second wave).
@@
 We estimated the significance of model parameters by examining CIs, considering parameters with CIs overlapping with zero as non-significant.
-To estimate the phylogenetic signal of seed germination over all variables we used Pagels’s lambda (λ) [@RN4757], estimated simultaneously with the models by calculating the mean of the posterior distribution and the 95% CI of λ as indicated by De Villemereuil et al.
+To estimate the phylogenetic signal of seed germination over all variables, we used Pagels’s lambda (λ) [@RN4757], estimated simultaneously with the models by calculating the mean of the posterior distribution and the 95% CI of λ as indicated by De Villemereuil et al.
 [-@RN4756].
@@
 To do so, for each cue and species, we calculated a weighted average of the cue levels (in the case of temperature, cue levels were the temperature treatments; for the other cues the levels were 0 = absence and 1 = presence), weighting by the germination proportion at each level.
-This approach underrepresents the importance of the levels that were not tested for a given species, but can serve as a proxy of the response to the germination cues when visualized across the whole dataset; it must be stressed that this stage of the analysis serves only for visualization and not hypothesis testing.
+This approach underrepresents the importance of the levels that were not tested for a given species, but can serve as a proxy of the response to the germination cues when visualized across the whole dataset; it must be stressed that this stage of the analysis serves only for visualization and not for hypothesis testing.
 We also included germination timing traits by calculating single values for the mean germination time (the minimum, i.e.
 the time taken at the most favourable treatment) and the germination uncertainty index (the mean).
-We conducted the FAMD with only the `r MSfamdspp` species for which we had mean germination time and germination uncertainty index values.
+We conducted the FAMD with the `r MSfamdspp` species for which we had mean germination time and germination uncertainty index values.
 We finally included the continuous variables seed mass and embryo:seed, as well as the categorical variables dormancy class (physiological, morphophysiological, morphological, physical or non-dormant) and distribution (strict alpine vs.
@@
 The final dataset used in our meta-analysis contained `r MSrecords` records (i.e.
-germination proportions for a given seed lot of a species, recorded in a set of experimental conditions) contributed by `r MSgroups` research groups, representing `r MSfamilies` seed plant families and `r MSspecies` species (`r MSstrict` strict alpine species and `r MSgeneral` generalist species).
+germination proportions for a given seed lot of a species, recorded in a set of experimental conditions) contributed by `r MSgroups` research groups, representing `r MSfamilies` seed plant families and `r MSspecies` species.
+From these, `r MSstrict` were categorized as strict alpine species and `r MSgeneral` as generalist.
 Seed lots were originally sampled in `r MSregions` alpine regions of the world (Table 1), in both hemispheres.
+All regions correspond to middle latitudes.
 
@@
 \@ref(fig:fig1)a), but the frequency of the dormancy classes did not differ between generalist and strict alpine species.
-Among those species with dormant seeds, physiological dormancy was the most common dormancy class, followed by morpho-physiological and then physical.
+Among those species with dormant seeds, physiological dormancy was the most common dormancy class, followed by morphophysiological and then physical.
 Both strict alpine and generalist species had a similar range of values for seed mass (Fig.
@@
 The values of seed mass ranged from `r MSminMass` mg (*`r MSminMassSpp`*) to `r MSmaxMass` mg (*`r MSmaxMassSpp`*), with a median value of `r MSmedianMass` mg.
-Embryo:seed ratio encompassed the full range of potential values, from endospermic species with very small embryos (`r MSminEmbryo`) to non-endospermic seeds (`r MSmaxEmbryo`).
+Embryo:seed ratio encompassed the full range of potential values, from endospermic species with very small embryos (`r MSminEmbryo`) to non-endospermic seeds that store nutrients in the cotyledons (`r MSmaxEmbryo`).
 The median value was `r MSmedianEmbryo`.
@@
 an abrasion of the seed coat to allow water imbibition) was performed in `r MSscarY` records, and GA~3~ (a plant hormone that promotes embryo growth and stimulates germination) in `r MSga3` records.
-Since germination incubations ran for different periods of time, we established a cut-off time of 8 weeks to calculate the final germination percentages, to have a homogeneous comparison across records.
 
@@
 
-Temperature had a positive significant effect on the final germination proportion as shown by a credible interval not overlapping zero (Fig.
+Temperature had a positive significant effect on the final germination proportion (Fig.
 \@ref(fig:fig2)a), indicating a preference for warmer temperatures.
@@
 In other words, heavier and less endospermic seeds had a smaller increase in germination proportion and a smaller decrease in mean germination time when germinating in warmer temperatures.
-The germination uncertainty index did not show an interaction with seed mass or embryo:seed, as shown by the credible interval overlapping zero.
+The germination uncertainty index did not show an interaction with seed mass or embryo:seed.
 
@@
 Alternating temperatures had a positive effect on the final germination proportion (Fig.
-\@ref(fig:fig2)a), but not in the mean germination time (Fig.
+\@ref(fig:fig2)a), but not on the mean germination time (Fig.
 \@ref(fig:fig2)b) or the germination uncertainty index (Fig.
@@
 
-Random factors had a significant effect on the MCMCglmm models (Fig.
+Random factors had a significant effect in the MCMCglmm models (Fig.
 \@ref(fig:fig3)).
@@
 For the germination uncertainty index, it overlapped with zero, indicating a weak effect of phylogeny.
+The results for the germination uncertainty index indicate that it might have been more sensitive to the frequency of germination scoring, which was different in different labs.
 
@@
 
-The first FADM axis explained `r round(pcaAlpine$eig[1, 2], 0)`% of the variation.
+The first FAMD axis explained `r round(pcaAlpine$eig[1, 2], 0)`% of the variation (Fig.
+\@ref(fig:fig5)).
 The quantitative variables with the largest contribution to this first axis were germination temperature, germination uncertainty, GA~3~ and alternating temperatures.
 Among the qualitative factors, the strict alpine / generalist categories contributed to axis 1.
-In summary, the horizontal axis separated (i, left) generalist species with a positive response to alternating temperatures and more germination uncertainty from (ii, right) strict alpine species with a preference for warmer temperatures and a positive response to GA~3~ (Fig.
-\@ref(fig:fig5)).
+In summary, the horizontal axis separated (i, left) generalist species with a positive response to alternating temperatures and more germination uncertainty from (ii, right) strict alpine species with a preference for warmer temperatures and a positive response to GA~3~.
 Axis 2 explained `r round(pcaAlpine$eig[2, 2], 0)`% of the variability.
@@
 Our meta-analysis of primary germination data shows that alpine plants from different mountains of the world tend to show specialized seed germination characteristics.
-For most alpine species, both dormancy and warm-cued germination combine to ensure germination occurs in the season most conducive to survival: seeds tend to germinate when, after weeks of exposure to conditions that mimic winter, they are moved to mid-summer temperatures.
-Specifically, the germination ecology of strict alpine species is characterized by (1) physiological seed dormancy; (2) a strong need for cold stratification or GA~3~ to break dormancy; (3) warm-cued germination; (4) a positive response to alternating temperatures, although not so prominent as in generalist species; (5) a positive response to light; (6) slow and relatively synchronous germination; and (7) an interplay with seed and embryo size, with smaller and more endospermic seeds being more responsive to stratification, warmth, alternating temperatures and light.
-Here we discuss the ecological significance of the different aspects of the alpine seed germination spectrum.
+For most species thriving in alpine habitats, both dormancy and warm-cued germination combine to ensure that germination occurs in the season most conducive to survival.
+In particular, seeds tend to germinate when, after weeks of exposure to conditions that mimic winter, they experience mid-summer temperatures.
+The germination ecology of strict alpine species is therefore characterized by (1) physiological seed dormancy; (2) a strong need for cold stratification or GA~3~ to break dormancy; (3) warm-cued germination; (4) a positive response to alternating temperatures, although not so prominent as in generalist species; (5) a positive response to light; (6) slow and relatively synchronous germination; and (7) an interplay with seed and embryo size, with smaller and more endospermic seeds being more responsive to stratification, warmth, alternating temperatures and light.
+Here we discuss the ecological significance of these responses and their role to define the seed germination spectrum of alpine plants.
 
@@
 Our dataset indicates that seed dormancy is frequent across alpine species and plays an important role in postponing plant recruitment to favourable times.
-Furthermore, the frequency of dormancy is similar in strict alpine and generalist species, suggesting that this is a general trait of cold-adapted mountain species.
-Seeds with the two most frequent dormancy classes, physiological and morphophysiological, require a stratification period during which certain conditions (e.g.
+The frequency of dormancy classes displayed by strict alpine and generalist species is similar, suggesting that this is a general trait of cold-adapted mountain species.
+Seeds with the two most frequent dormancy classes, i.e.
+physiological and morphophysiological, require a stratification period during which certain conditions (e.g.
 cold and wet, although other combinations are possible) are met over a period of months [@RN3214; @RN4743].
-Indeed, our results show that cold stratification promotes germination.
-The phytohormone gibberellic acid (GA~3~) works as a general substitute of cold stratification to overcome the inherent physiological dormancy of alpine species.
+Indeed, our results show that cold stratification consistently promotes germination, while the phytohormone gibberellic acid (GA~3~) works as a general substitute of cold stratification to overcome the inherent physiological dormancy of alpine species.
 Additionally, we found some species with physical dormancy.
-This dormancy type can be broken by mechanical scarification via freezing-thawing cycles in spring or extreme temperature fluctuations during summer, but is also capable of responding to seasonal cycles of temperature and humidity [@RN780].
+This dormancy type can be broken by mechanical scarification via freezing-thawing cycles in spring or extreme temperature fluctuations during summer, but it is also capable of responding to seasonal cycles of temperature and humidity [@RN780].
 
@@
 Since global climate change is causing major changes in snow cover duration [@RN4917; @RN4918], it could disrupt natural cold stratification cycles in alpine environments, compromising population viability and indirectly favouring species with no stratification requirement [@RN3703].
-This risk will be partly alleviated by the plasticity of physiological seed dormancy in response to the seed maturation temperature [@RN2358 ; @RN2899], a mechanism of “seed memory” that facilitates acclimatisation to changing environments [@RN2249], and which could reduce the stratification requirements of seeds produced in a warmer climate.
+This risk will be partly alleviated by the plasticity of physiological seed dormancy in response to the seed maturation temperature [@RN2358; @RN2899], a mechanism of “seed memory” that facilitates acclimatisation to changing environments [@RN2249], reducing the stratification requirements of seeds produced in a warmer climate.
 
@@
 
-We found a general increase of seed germination with higher incubation temperatures.
+We also found a general increase of seed germination with higher incubation temperatures.
 This effect was more pronounced in strict alpine compared to generalist species, confirming the findings of local studies [@RN3266; @RN2371].
@@
 Moreover, strict alpine species may be less subjected to drought damage during establishment compared to generalist species: as elevation increases, plants face less drought risk because precipitation increases (except in Mediterranean and semi-arid mountains), but at the same time there is a higher probability of early or late season frosts [@RN4897; @RN2392; @RN4914].
-Together, these results indicate that germination patterns in alpine species are driven by an interplay of seed size and germination temperature to escape either frost or drought stress; small seeds are being selected for fast germination at warm temperatures, in order to escape unfavourable early-spring frost, and large seeds are being selected for slower germination under cooler conditions, which should maximize emergence during snowmelt (when water availability is highest) and gives seedlings adequate time to develop a root system to cope with desiccation risks in summer.
-Supporting this view, seedlings of alpine pioneer species have a low frost resistance [@RN4722], yet high heat tolerance up to 40–50°C [@RN4902].
+Together, these results indicate that germination patterns in alpine species are driven by an interplay of seed size and germination temperature to escape either frost or drought stress: small seeds are being selected for fast germination at warm temperatures, in order to escape unfavourable early-spring frost, and large seeds are being selected for slower germination under cooler conditions, which should maximize emergence during snowmelt (when water availability is highest) and gives seedlings adequate time to develop a root system to cope with desiccation risks in summer.
+Supporting this view, seedlings of alpine pioneer species have a low frost resistance [@RN4722], yet tolerate heat up to 40–50°C [@RN4902].
 
 In our analysis, increasing temperatures also decreased germination uncertainty, showing that germination synchrony is a plastic trait driven by temperature.
-While asynchronous germination is thought to be a form of adaptation to unpredictable alpine environments [@RN3734; @RN3690], our results indicate that asynchronous germination occurs when temperatures are relatively cool, likely as a bet-hedging strategy against the risk of early-spring frost.
+While asynchronous germination is thought to be a form of adaptation to unpredictable environments [@RN3734; @RN3690], our results indicate that asynchronous germination occurs when temperatures are relatively cool, likely as a bet-hedging strategy against the risk of early-spring frost.
 On the other hand, a fast and synchronised germination at warm conditions (i.e.
@@
 We found a positive germination response to alternating temperatures and light, conditions that mimic the environment in the upper soil and certain microsites during the snow-free season [@RN4712; @RN2392; @RN3029].
-Alternating temperatures had a slightly lower effect on strict alpine species, compared with generalist species, but still had an effect.
+Alternating temperatures had a slightly lower effect on strict alpine species, compared with generalist species, but still had some effect.
 This partly contradicts the findings of Liu et al.
@@
 When analysed at the global level, alternating temperatures generally increase germination of alpine plants, which corresponds well with the fact that most alpine regions have a strong day-night climatic variation during the growing season.
-It is also not surprising that we detected a negative interaction between seed mass and the germination response to light and alternating temperature regimes, since large seeds can afford to germinate at greater depths (where they are more protected from the alpine environment) and still reach the soil surface [@RN4904].
+It is also not surprising that we detected a negative interaction between seed mass and the germination response to light and alternating temperature regimes, since large seeds can afford to germinate at greater depths (where they are more protected from the alpine environment) and still reach the soil surface [@RN4956; @RN4904].
 
@@
 Using the whole set of traits compiled in this study we visualize the variation in the seed germination spectrum of alpine plants.
-On one end of the spectrum, strict alpine species tend to have slow and synchronous germination, require warm temperatures and light to germinate, and need stratification or GA~3~ to break physiological dormancy.
-On the other end, generalist species tend to have fast germination, show a bet-hedging strategy (high germination uncertainty) and germination traits suggesting possible fast germination after specific cues (scarification, alternating temperatures).
+On one end of the spectrum, strict alpine species tend to show slow and synchronous germination, require warm temperatures and light to germinate, and need stratification or GA~3~ to break physiological dormancy.
+On the other end, generalist species tend to have fast germination, showing a bet-hedging strategy (high germination uncertainty) and germination traits that suggest possible fast germination after specific cues (e.g.
+scarification or alternating temperatures).
 The dormancy classes without a physiological component (non-dormant, physical, and morphological) appear associated with generalist species.
@@
 This is in line with the ambiguity of relationships that have been found between seed mass and elevation for alpine species [@RN4912; @RN4903; @RN3685; @RN4896].
-The similarity in embryo:seed size between generalist and strict alpine species confirms the absence of a correlation between embryo-seed size and elevation found in the Apiaceae [@RN3685].
-In contrast with the supposed phylogenetic conservatism of these morphological traits, physiological germination responses are understood to be more plastic traits that can quickly acclimatize to new situations [@RN2249].
-While phylogenetic signal in seed germination is generally found, especially in highly selective environments [@RN3001; @RN2399; @RN2865], germination patterns can also be shared between phylogenetically distant species [@RN4911] and may have low or absent phylogenetic signal [@RN2365; @RN4900; @RN2371].
+The similarity in embryo:seed size between generalist and strict alpine species confirms the lack of correlation between embryo-seed size and elevation found in the Apiaceae [@RN3685].
+In contrast with the supposed phylogenetic conservatism of these morphological traits, physiological germination responses are understood to be more plastic traits that can quickly respond to new situations [@RN2249].
+While phylogenetic signal in seed germination is generally found, especially in highly selective environments [@RN3001; @RN2399; @RN2865], germination patterns can also be shared between phylogenetically distant species [@RN4911] and have low or absent phylogenetic signal [@RN2365; @RN4900; @RN2371].
 The phylogenetic signal found here indicates that evolutionary history cannot be neglected when studying alpine seed germination patterns.
@@
 Previous studies have shown that non-endospermic seeds are smaller and are likely to be non-dormant at the time of dispersal and thus germinate quicker than alpine endospermic seeds [@RN3703; @RN3285; @RN4907].
-Together, these results indicate that seed morphology and physiology are connected aspects.
+Together, these results indicate that seed morphology and physiology are connected.
 This connection, intertwined with their shared phylogenetic history, may constrict the potential variability of the seed ecological spectrum.
@@
 
-Alpine species from different regions of the world have specialised germination characteristics that support successful regeneration in challenging alpine environments.
-Our results point to an evolutionary convergence of germination patterns in alpine species.
-They also provide the first global assessment of the seed germination spectrum in relation with morphological seed traits of alpine plants, suggesting potential relationships and trade-offs between sets of regenerative traits that will need further research.
-New studies also will need to consider evolutionary history as an intrinsic constraint of the seed ecological spectrum.
-Overall, this study provides a global assessment of alpine seed germination that can be used as a reference in further studies focused on spatial or temporal variation in seed germination, their links with other plant traits, and the expected responses of alpine plants to climate change.
+Alpine plants from different regions of the world respond similarly to germination cues that are expected to support successful regeneration: cold stratification, warm and alternating temperatures, and light.
+These cues, particularly cold stratification and warm temperatures, are generally associated with cold-adapted herbaceous species in a broad sense [@RN3397; @RN3214; @RN2365].
+Nonetheless, our model shows that the response to stratification and temperature is more prominent in strict alpine species when comparing them with generalist plants that can grow opportunistically in the alpine belt, suggesting an evolutionary convergence of germination patterns in those species truly adapted to the alpine.
+We can conclude that, while alpine plants do not show a unique germination syndrome, they do show a more extreme or specialized version of the cold-adapted temperate germination syndrome.
+
+We were not able to analyse data from tropical alpine regions, highlighting how tropical alpine seeds are underrepresented in the literature.
+We recommend that tropical alpine regions be the focus of future research efforts, as it is possible that the cold-adapted germination syndrome is not prevalent there.
+Tropical alpine environments are relatively aseasonal and most plants have been reported as flowering throughout the year [@RN4957].
+Rather than being exposed to snow in winter and high temperatures in summer, tropical alpine plants face a year-round stress from night-time frost [@RN2392].
+That means that our study is valid for the temperate alpine system, but not necessarily for tropical mountain species pools, which are subjected to different selective pressures.
+
+Another aspect that calls for future efforts is a finer study of germination cues.
+To produce this meta-analysis, we had to assume some simplifications in order to merge disparate datasets: we considered coarse cues that are routinely recorded in germination tests, like the presence/absence of light, alternating temperatures or cold stratification.
+New research should aim to decompose these binary cues into finer quantitative variables which provide a more precise environmental signalling, such as the red:far red ratio [@RN604], the length of the photoperiod [@RN1883], the amplitude of the diurnal thermal oscillations [@RN2387] or the length and temperature of cold stratification [@RN2346].
+
+Overall, our first global assessment of the seed germination spectrum in relation with seed morphology supports the existence of trade-offs between sets of regenerative traits [@RN4655], such as the relationship we found between seed/embryo size, germination speed and germination responses to temperature and light.
+Future studies will need to consider, as we did, evolutionary history as an intrinsic constraint of the seed ecological spectrum.
 
@@
 EFP received financial support from the Government of Asturias and the FP7 – Marie Curie - COFUND programme of the European Commission (Grant ‘Clarín’ ACB17-19).
-EM is supported by the Future Leaders Fellowship - Diversity and Livelihoods, of the Royal Botanic Gardens, Kew.
-This research was supported by the NGSeeds lab virtual working group.
+BJA was funded by the Spanish Research Agency (PID2019-108636GA/ AEI/ 10.13039/501100011033).
+EM was supported by the Future Leaders Fellowship - Diversity and Livelihoods, of the Royal Botanic Gardens, Kew.
+This research was supported by NGSeeds, the Next Generation Seed Ecology, Evolution and Data Science virtual working group.
 
@@
 When the credible intervals overlap with the zero-effect line, the effect is not significant.
+Some effects that overlapped with zero (i.e.
+had no effect) and whose credible intervals were excessively wide are not shown for clarity's sake.
 The figure shows first the main effect, then the effects for the strict alpine and generalist groups, and then the interaction with seed mass and embryo:seed.
@@
 ```{r fig4, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Phylogenetic signal in the response to the germination environment.
-When lambda = 0, related taxa are no more similar than expected by chance; when lambda = 1, the trait is evolving following a constant variance random walk or Brownian motion model.", fig.align = "center", fig.width=7, fig.height=1.5}
+When lambda = 0, related taxa are no more similar than expected by chance; when lambda = 1, the trait is evolving following a constant variance random walk or Brownian motion model.
+MGT = mean germination time.
+UNC = germination uncertainty index.", fig.align = "center", fig.width=7, fig.height=1.5}
 cowplot::ggdraw(fig4)
@@
 
-```{r fig5, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "FADM ordination of the alpine seed germination spectrum.
+```{r fig5, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "FAMD ordination of the alpine seed germination spectrum.
 Each square is a species.
 Labels indicate the contribution of the variables to the axes.
+Panels A and B correspond to quantitative traits; panel C and D to qualitative traits.
+MGT = mean germination time.
+UNC = germination uncertainty index.
 Seed dormancy classes are abbreviated as follows: MD (morphological), MPD (morphophysiological), ND (non-dormant), PD (physiological), PY (physical).", fig.align = "center", fig.width=7, fig.height=7}

### 2020-10-27 src/manuscript-source.R

@@ -108,7 +108,7 @@ dataset %>%
   ggthemes::theme_tufte() +
   theme(legend.position = "right", axis.title.x = element_blank(),
         panel.background = element_rect(color = "grey96", fill = "grey96"),
-        axis.text.x = element_text(size = 7)) +
+        axis.text.x = element_text(size = 10)) +
   scale_fill_manual(values = c("olivedrab", "yellowgreen", "turquoise4", "gold", "purple")) -> fig1a
 
 dataset %>%
@@ -121,7 +121,7 @@ dataset %>%
   ggthemes::theme_tufte() +
   theme(legend.position = "none", axis.title.x = element_blank(),
         panel.background = element_rect(color = "grey96", fill = "grey96"),
-        axis.text.x = element_text(size = 7)) +
+        axis.text.x = element_text(size = 10)) +
   scale_fill_manual(values = c("yellowgreen", "gold")) -> fig1b
 
 dataset %>%
@@ -134,7 +134,7 @@ dataset %>%
   ggthemes::theme_tufte() +
   theme(legend.position = "none", axis.title.x = element_blank(),
         panel.background = element_rect(color = "grey96", fill = "grey96"),
-        axis.text.x = element_text(size = 7)) +
+        axis.text.x = element_text(size = 10)) +
   scale_fill_manual(values = c("yellowgreen", "gold")) -> fig1c
 
 gridExtra::arrangeGrob(fig1a, fig1b, fig1c, ncol = 3, widths = c(8, 7, 7)) -> fig1
@@ -543,7 +543,9 @@ ggplot(pcaInds, aes(x = Dim.1, y = Dim.2)) +
   scale_color_manual(values = c("yellowgreen", "gold")) +
   labs(title = "(D) Species distribution") -> fig5d
 
-gridExtra::arrangeGrob(fig5a, fig5b, fig5c, fig5d, ncol = 2) -> fig5
+gridExtra::arrangeGrob(fig5a, fig5b, fig5c, fig5d, ncol = 2,
+                       left = grid::textGrob("Second FAMD axis", rot = 90, gp = grid::gpar(fontfamily = "serif")),
+                       bottom = grid::textGrob("First FAMD axis", gp = grid::gpar(fontfamily = "serif"))) -> fig5
 
 # Save figures
 

### 2020-11-02 README.Rmd
readme pic 🌼🏔️
1st major revision 🍃

@@
 
-![Picos de Europa National Park, Cantabrian Mountains, Spain](data/fernandezpascual/Picos de Europa.JPG)
+![Picos de Europa National Park, Cantabrian Mountains, Spain](Picos de Europa.jpg)
 
@@
 
-This repository stores all the information related to the publication *The seed germination spectrum of alpine plants: a global meta-analysis*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.
+This repository stores all information related to the manuscript *The seed germination spectrum of alpine plants: a global meta-analysis*, including raw data, `R` scripts to process and analyse data, and `Rmarkdown` files to create the manuscript.
 
@@
 
-* `data` Original raw files including primary germination data and species traits.
+* `data` Primary germination data and species traits.
 * `doc` Files to create the manuscript using `Rmarkdown`.
-* `results` Output of the `R` scripts, including clean data and MCMCglmm analysis outputs.
-* `src` Scripts in `R` language used to clean the raw data and perform the analyses of the manuscript.
+* `results` Output of the `R` scripts, including MCMCglmm analysis outputs.
+* `src` Scripts in `R` language used to process data and perform the analyses of the manuscript.
 
@@
 
-Please cite the dataset as `CITATION`
+Please cite this repository as `The seed germination spectrum of alpine plants: a global meta-analysis. Manuscript in preparation.`

### 2020-11-02 doc/cover-letter.Rmd
1st major revision 🍃

@@
 date: "`r format(Sys.time(), '%d/%m/%Y')`"
-output: word_document
+output: pdf_document
 bibliography: refs.bib
@@
 
-We respond below to each of the issues pointed out by the referees.
+Thanks for the detailed reviews and the positive feedback.
+We respond below to all issues pointed out by the referees, grouping comments on similar topics on the same paragraph to facilitate the reading.
 We have adopted most of the proposed changes.
-Regarding light, we explain that this is an intrinsic limitation of the meta-analysis approach, and discuss the implications.
-As for the phylogenetic signal, we explain that this is explicitly taken into account by adopting a phylogenetic comparative statistical method such as the MCMCglmm, which considers phylogeny and phylogenetic signal within the analysis.
+Regarding the use of light, we explain with detail that this is an intrinsic limitation of the meta-analysis, mainly due to the quality of existing data, and discuss the implications.
+As for the phylogenetic signal, we explain that this is explicitly considered in a phylogenetic comparative statistical method such as the MCMCglmm we used, integrating phylogeny and phylogenetic signal within the analysis.
 
@@
 
-We have changed the introductory point to *Given the micro-climatic heterogeneity of alpine landscapes, and the wealth of local studies on alpine seed ecology, a current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants.*.
+We have changed the introductory point to a more general and stronger message: *Assumptions about the germination ecology of alpine plants are presently based on individual species and local studies.
+A current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants.*
 
@@
 They hamper the figure's visibility (i.e.
-they make all other effects appear too compressed), so we removed them from the plots.
+they make all other effects to appear compressed), so we removed them from the plots.
 We have included this clarification in the figure's caption.
@@
 We have moved the call to the section's beginning and labelled the axes.
-Indicating to which panel we are referring to in the text is not easy, because the different panels/traits are intertwined in the descriptions, therefore it would require to make many recurrent calls (and, in our opinion, that would burden the text too much.)
+Indicating to which panel we are referring to in the text is not easy, because the different panels/traits are intertwined in the descriptions, therefore it would require to make many recurrent calls (and, in our opinion, that would burden the text too much).
 
@@
 Yes, there is a knowledge gap for tropical mountains.
-We were unable to find any significant amount of data from the tropics (excluding single species studies).
-We have qualified *global* as excluding the tropics throughout the manuscript, and commented on the knowledge gap in the conclusions.
+We were unable to find any significant amount of data from the tropics.
+We have qualified *global* as excluding the tropics throughout the manuscript, and commented on the knowledge gap (and the need of further research in alpine plants of tropical regions) in a new *Future directions* section before the *Conclusions*.
 
@@
 The same can be said about other cues, like alternating temperature (amplitude of the alternation, duration of the cycles...) or cold stratification (actual temperature, length...).
-However, to conduct a meta-analysis, we need to find a compromise between data quality and quantity.
-Detailed descriptions of light spectral composition and photoperiod are only available for a small fraction of the whole dataset.
-Therefore, we must summarise cues into some variables (like light absence/presence) that are routinely recorded in the majority of trials.
-We have added these considerations to the conclusions.
+Unfortunately, most germination studies in the laboratory use this binary treatment, mainly due to the tradition of detecting major germination strategies.
+This is a limitation of the data that cannot be dealt with in a meta-analysis, thus we need to find a compromise between data quality and quantity.
+Detailed descriptions of light spectral composition and photoperiod are only available for a very small fraction of the whole dataset.
+Therefore, we must summarise cues into variables (like light absence/presence) that are routinely recorded in the majority of trials.
+We have added these considerations to a new *Future directions* section before the *Conclusions*.
 
@@
 To improve understanding, we have included a definition of non-endospermic seeds as *non-endospermic seeds that store nutrients in the cotyledons*.
-In any comparative biology analysis, all traits are always phylogenetically confounded.
+Nevertheless, we note that in any comparative biology analysis, all species traits are phylogenetically confounded.
 This is why we used a phylogenetic comparative analysis such as the MCMCglmm.
@@
 In this way, traits are analysed in relation to the underlying phylogenetic structure of the dataset.
-In our models, both the trait (seed:endosperm ratio) and phylogeny had significant effects.
-The use of phylogenetic comparative analysis is covered in the analysis section of the methods section.
+In our models, both the trait (seed:endosperm ratio) and phylogeny had significant effects, denoting the effect mentioned by the reviewer.
+The use of phylogenetic comparative analysis is covered in methods/analysis section.
 
@@
 
-In our manuscript we use the definition and classification of dormancy that is currently dominant in seed ecology and physiology.
+We agree on the conceptual distinction made by the reviewer, but we are using a different terminology for the same concepts.
+In particular, we use the definition and classification of dormancy that is currently most used in seed ecology and physiology.
 This framework is based on two main references: (i) the concept by Vleeshouwers et al.
 [-@RN3254] that dormancy is an inner seed property ("endodormancy"), which should be separated from the absence of appropriate germination cues in the environment ("ecodormancy"); and the classification system proposed by Baskin & Baskin [-@RN3261] based on previous work by Nikolaeva et al.
-[-@RN2877] and which considers morphological and physical dormancy separated from physiological.
-There are two practical reasons to use this framework: first, it will be readily understandable to the widest scope of readers; and second, it provides the largest amount of species trait data available [@RN3214; @RN4743].
-We have retained the definition and classification of dormancy originally used in our paper, however, we have also included explanation in the text of how this classification corresponds to the terminology preferred by the reviewer.
+[-@RN2877], which considers morphological and physical dormancy separated from physiological.
+There are two practical reasons to use this framework: first, it will be readily understandable to the widest scope of readers, including seed ecologists used to these terms; and second, this is the same framework used in the largest amount of seed trait data available [@RN3214; @RN4743].
+Although we have retained this definition and classification of dormancy, we also have explained how it corresponds to the “endodormancy” and “ecodormancy” terms suggested by the reviewer.
 
@@
 We have revised the entire conclusions paragraph to address these concerns, being more specific about both the conclusions and about the limitations of the data.
+In short, we have created a separate section of *Future directions* where we discuss (a) the need to account for phylogenetic constraints and interactions between seed morphology and physiology (citing specific traits/response) in future research; (b) that our results are valid for temperate regions, but little is known about the response of non-seasonal tropical mountains; and (c) that our meta-analysis shows that understanding seed germination of alpine plants still needs more detailed data for major environmental cues.
+In a final *Conclusions* section we now provide a more concrete take home message: (d) while alpine plants do not show a unique germination syndrome, they do show a more extreme or specialized version of the cold-adapted temperate herb germination syndrome (suggesting convergence).
 
@@
 
-The referee is right and we have screened the text for this class of words.
-*Change* is used to refer to a concept like climate change, or to encompass and summarise changes that can go in different directions (environmental changes, snow cover changes).
+The referee is right, and we have screened the text to detect these words.
+*Change* is used either to refer to a concept like climate change, or to encompass and summarise changes that can go in different directions (environmental changes, snow cover changes).
 *Interplay* and *intertwined* are used in places like the summary and conclusions, where we must summarise concepts that are detailed elsewhere.
@@
 We could not find *vary*.
-Regarding the specific point at L523, this is not a neutral statement: the direction of *interplay* is explained in the same sentence, after the *;*.
-To make this point clear, we have substituted the *;* with a *:*.
+Regarding the specific point at L523, this is not a neutral statement: the direction of *interplay* is explained in the same sentence, after the *;* (we have now substituted *;* with *:*).
 
@@
 
-We have removed wind.
+Thanks for this remark. We have removed wind.
 
@@
 
-We have removed the *soap*, and now the two introductory sentences link directly to former L73.
+Many of the co-authors generally agree with this view.
+In any case, this is not the aim of the study, so we have removed from the introduction our comments remarking the importance of regeneration for the general assessment of climate-warming effects on biodiversity.
+In the new version, the two introductory sentences link directly to former L73.
 
@@
 To the latter.
-To be more concrete, we have changed the sentence to *Given the high micro-climatic heterogeneity of alpine landscapes, and the wealth of local studies on alpine seed ecology (reviewed in the following paragraphs), a current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants*.
+To be more concrete, we have changed the sentence to *Despite the high micro-climatic heterogeneity of alpine landscapes, current assumptions about germination ecology of alpine plants are mainly based on local studies (reviewed in the following paragraphs)*.
 
@@
 We have changed the sentence to *Fluctuations of diurnal temperature decrease with increasing burial depths and the depth profile of this decrease depends on soil type, vegetation cover and snow*.
-We have read Söyrinki's work and cited it where appropriate, however, we did not find that he explicitly tested the effect of fluctuating temperatures.
+The German-speaking coauthors have revised Söyrinki's work and we have cited it where appropriate, however, we did not find that he explicitly tested the effect of fluctuating temperatures.
 
@@
 Measurements were made using the ImageJ software (Schneider et al.
-2012) on drawings and photographs of seeds cut in half along the longitudinal axis, retrieved mainly from Martin (1946) and the Royal Botanic Gardens Kew Seed Information Database (2017), supplemented with own photographs (Vandelook, unpublished).*.
-In our opinion this is the most straightforward method to obtain a good perception of the amount of nutrient reserves stored either inside or outside the embryo without going into more advanced 3D methods.
+2012) on drawings and photographs of seeds cut in half along the longitudinal axis, retrieved mainly from Martin (1946) and the Royal Botanic Gardens Kew Seed Information Database (2017), supplemented with own photographs (Vandelook, unpublished).* In our opinion this is the most straightforward method to obtain a good perception of the amount of nutrient reserves stored either inside or outside the embryo without going into more advanced 3D methods.
 It was approximately the same methodology as the one applied by Forbis et al.
@@
 These two variables are described later in the text, in the *Description of the dataset* section of the *Results*.
+The other comments of the reviewer are very interesting, we have mentioned a few of them in the *Future directions* section.
 
@@
 We have removed the adjective *high* and changed the sentence to *yet tolerate heat up to 40–50°C*.
+We are not aware of evidence suggesting that seedlings are more sensitive to high temperatures than adult tissue.
 
@@
 
-Indeed, this is not something exclusive of alpine environments. We have changed the sentence to *a form of adaptation to unpredictable environments*.
+Indeed, this is not exclusive of alpine environments. We have changed the sentence to *a form of adaptation to unpredictable environments*.
 

### 2020-11-02 doc/manuscript.Rmd
1st major revision 🍃

@@
 
-**Total word count for the main body of the text** 6,343; **Introduction** 1,968; **Materials and Methods** 1,342; **Results** 1,138; **Discussion** 1,896; **Number of figures** 5 in colour; **Tables** 1; **Supporting information** 0
+**Total word count for the main body of the text** 6,843; **Introduction** 1,888; **Materials and Methods** 1,517; **Results** 1,132; **Discussion** 2,306; **Number of figures** 5 in colour; **Tables** 1; **Supporting information** 0
 
@@
 
-* Given the micro-climatic heterogeneity of alpine landscapes, and the wealth of local studies on alpine seed ecology, a current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants.
-* We meta-analyse primary data from laboratory experiments conducted across four continents (excluding the tropics) and 661 species, to estimate the influence of six environmental cues on seed germination proportion, mean germination time and germination synchrony; accounting for possible effects of seed morphology (mass, embryo:seed ratio) and phylogeny.
-* Alpine plants show physiological seed dormancy, a strong need for cold stratification, warm-cued germination and positive germination responses to light and alternating temperatures.
-Species with a distribution limited to the alpine belt have a higher preference for warm temperatures and a stronger response to cold stratification than species whose distribution extends also to the lowlands.
+* Assumptions about the germination ecology of alpine plants are presently based on individual species and local studies.
+A current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants.
+* We meta-analyse primary data from laboratory experiments conducted across four continents (excluding the tropics) and 661 species, to estimate the influence of six environmental cues on germination proportion, mean germination time and germination synchrony; accounting for possible effects of seed morphology (mass, embryo:seed ratio) and phylogeny.
+* Most alpine plants show physiological seed dormancy, a strong need for cold stratification, warm-cued germination and positive germination responses to light and alternating temperatures.
+Species with a distribution limited to the alpine belt have a higher preference for warm temperatures and a stronger response to cold stratification than species whose distribution extends also below the treeline.
 Seed mass, embryo size and phylogeny have strong constraining effects on germination responses to the environment.
@@
 
-Alpine environments occur worldwide, above the maximum elevation at which trees can grow naturally [@RN3000; @RN3024; @RN4750].
+Alpine environments occur worldwide above the maximum elevation at which trees can grow naturally [@RN3000; @RN3024; @RN4750].
 These treeless habitats are characterized by low temperatures, unstable substrates and short growing seasons [@RN2392].
 Although most alpine plants can reproduce clonally [@RN2392], sexual reproduction by diaspores (i.e.
-seeds, fruits, or fruits with extra-ovarian structures attached; hereafter 'seeds') is the main strategy to maintain genetic diversity and colonize suitable new sites in response to environmental changes.
+seeds, fruits, or fruits with extra-ovarian structures attached; hereafter 'seeds') is the main strategy to maintain genetic diversity and to colonize suitable new sites in response to environmental changes.
 Plant regeneration by seed is a key process that determines whether species will persist in their current habitat or will be able to migrate to new sites [@RN3055]; and it is also fundamental for plant conservation [@RN3763] and seed-based ecological restoration [@RN3507; @RN2241].
@@
 The physiological process of seed germination is therefore an essential life stage that must be timed to occur when the environment is favourable for subsequent seedling survival and growth [@RN4691].
-In alpine habitats, seed germination will be a crucial test for plant populations coping with climate change, determining whether they will persist or go into extinction.
+In alpine habitats, seed germination will also be a crucial test for plant populations coping with global change, determining whether they will persist or go into extinction.
 
@@
 Thus, germination is mostly regulated by environmental cues related to water availability and soil temperature [@RN3368].
-The propagules of many plant species have also developed seed dormancy, by which germination is prevented during periods that are only ephemerally favourable, like a short warm spell in the middle of winter [@RN3214].
+The propagules of many plant species have also developed seed dormancy, by means of which germination is prevented during periods that are only ephemerally favourable, like a short warm spell in the middle of winter [@RN3214].
 Different degrees of dormancy within a seed population ensure the distribution of seedling emergence across time, a form of bet-hedging offspring survival against unpredictable environments [@RN3065].
 Other germination cues, such as the response to diurnal alternating temperatures [@RN1380] and the need for light or darkness [@RN4656], allow for a fine-scale detection of germination micro-niches and safe sites [@RN4719].
-Given the high micro-climatic heterogeneity of alpine landscapes, and the wealth of local studies on alpine seed ecology (reviewed in the following paragraphs), a current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants.
+Despite the high micro-climatic heterogeneity of alpine landscapes, current assumptions about germination ecology of alpine plants are mainly based on local studies (reviewed in the following paragraphs).
+A current challenge is thus to synthesize, at the global level, the seed ecological spectrum of alpine plants.
 The seed ecological spectrum [@RN4655] is defined as a set of seed traits that determines the ability of plants to disperse, persist, germinate, and establish in different habitats.
@@
 Regarding light and darkness, some local studies have found alpine species to have a preference for germinating in light [@RN4730; @RN4729].
-This trait could favour the creation of a persistent soil seed bank [@RN4718; @RN4929] which is advantageous in the temporally and spatially unpredictable alpine environments [@RN4731; @RN4732].
-Indeed, alpine soil seed banks are more frequent than previously supposed, as consistently reported in recent years [@RN4929; @RN4718; @RN4945].
-Nevertheless, other alpine species seem to prefer dark germination [@RN2943], which may be interpreted as a strategy for detecting safe sites in rock crevices [@RN3008].
+This trait is linked with the creation of a persistent soil seed bank [@RN4718; @RN4929] which may be advantageous in the temporally and spatially unpredictable alpine environments [@RN4731; @RN4732].
+Indeed, persistent alpine soil seed banks are more frequent than previously supposed, as consistently reported in recent years [@RN4929; @RN4718; @RN4945].
+Nevertheless, other alpine species seem to prefer germination in darkness [@RN2943], which may be interpreted as a strategy for detecting safe sites in rock crevices [@RN3008].
 Although some germination studies report laboratory experiments conducted with constant and alternating temperatures, or light and dark conditions, they are very much case-dependent, and therefore it is difficult to evaluate whether these factors show a consistent pattern across global alpine ecosystems.
@@
 The original data, plus R code for the analysis and creation of the manuscript using Rmarkdown, can be accessed at the GitHub repository https://github.com/efernandezpascual/alpineseeds.
-(The repository will be made open when the manuscript is accepted for publication, a zip version has been uploaded to https://drive.google.com/file/d/12Usqi7fOAbOoZ9Uu9wkJt2FANB629LFP/view?usp=sharing for peer-review).
 To gather data, we contacted research groups which have performed experiments on seed germination of alpine plants, asking them to provide raw data from laboratory germination experiments.
@@
 
-We obtained seed dormancy class from Baskin & Baskin [-@RN3214] and Rosbakh et al.
-[-@RN4743].
-We simplified the original dormancy classes to five categories: physiological (germination prevented by the seed’s internal balance of phytohormones), morphological (germination prevented by an embryo which is underdeveloped at the time of dispersal), morphophysiological (both physiological and morphological dormancy present), physical (germination prevented by a water-impermeable seed coat) or non-dormant.
+We classified the species in seed dormancy classes following Baskin & Baskin [-@RN3214] and Rosbakh et al.
+[-@RN4743], using five categories: physiological (germination prevented by the seed’s internal balance of phytohormones), morphological (germination prevented by an embryo which is underdeveloped at the time of dispersal), morphophysiological (both physiological and morphological dormancy present), physical (germination prevented by a water-impermeable seed coat) or non-dormant.
 We followed the concept by Vleeshouwers et al.
-[-@RN3254] to define dormancy as an inner seed property (sometimes called "endodormancy"), which should be separated from the absence of appropriate germination cues in the environment ("ecodormancy"); and the classification system of Baskin and Baskin [-@RN3261] based on previous work by Nikolaeva et al.
-[-@RN2877] which considers morphological and physical dormancy as classes separated from physiological dormancy.
+[-@RN3254] to define dormancy as an inner seed property (sometimes called "endodormancy"), which should be differentiated from the absence of appropriate germination cues in the environment ("ecodormancy"); and the classification system of Baskin and Baskin [-@RN3261] based on previous work by Nikolaeva et al.
+[-@RN2877] who consider morphological and physical dormancy as classes separated from physiological dormancy.
 
@@
 From these, `r MSstrict` were categorized as strict alpine species and `r MSgeneral` as generalist.
-Seed lots were originally sampled in `r MSregions` alpine regions of the world (Table 1), in both hemispheres.
-All regions correspond to middle latitudes.
+Seed lots were originally sampled in `r MSregions` alpine regions of the world (Table 1), at middle latitudes of both hemispheres.
 
@@
 
-Overall, these results confirm that overwintering is essential for promoting seed germination in alpine species, concurring with previous local studies [@RN3393; @RN2868; @RN707; @RN698; @RN3703; @RN2355; @RN3285; @RN2371; @RN4713].
+Overall, our results confirm that overwintering is essential for promoting seed germination in alpine species, concurring with previous local studies [@RN3393; @RN2868; @RN707; @RN698; @RN3703; @RN2355; @RN3285; @RN2371; @RN4713].
 Since global climate change is causing major changes in snow cover duration [@RN4917; @RN4918], it could disrupt natural cold stratification cycles in alpine environments, compromising population viability and indirectly favouring species with no stratification requirement [@RN3703].
-This risk will be partly alleviated by the plasticity of physiological seed dormancy in response to the seed maturation temperature [@RN2358; @RN2899], a mechanism of “seed memory” that facilitates acclimatisation to changing environments [@RN2249], reducing the stratification requirements of seeds produced in a warmer climate.
+This risk will be partly alleviated by the plasticity of physiological seed dormancy in response to the seed maturation temperature [@RN2358; @RN2899], a mechanism of “seed memory” via maternal effects that facilitates acclimatisation to changing environments [@RN2249], reducing the stratification requirements of seeds produced in a warmer climate.
 
@@
 We also found a general increase of seed germination with higher incubation temperatures.
-This effect was more pronounced in strict alpine compared to generalist species, confirming the findings of local studies [@RN3266; @RN2371].
+This effect was more pronounced in strict alpine compared to generalist species, confirming previous local studies [@RN3266; @RN2371].
 It was also more pronounced in small seeded and endospermic species.
@@
 staggering germination so that emergence occurs both before and after winter, is known in alpine species [@RN2392; @RN3285; @RN4907].
-Staggered germination is usually exhibited as a rapid first wave of germination followed by second wave several weeks or months later [@RN2392].
+Staggered germination is usually exhibited as a rapid first wave of germination followed by a second wave several weeks or months later [@RN2392].
 If autumn seedlings grow faster than spring seedlings [@RN4907], this will likely have cascading effects on individual fitness along the life of the plant [@RN3036; @RN2384].
@@
 When analysed at the global level, alternating temperatures generally increase germination of alpine plants, which corresponds well with the fact that most alpine regions have a strong day-night climatic variation during the growing season.
-It is also not surprising that we detected a negative interaction between seed mass and the germination response to light and alternating temperature regimes, since large seeds can afford to germinate at greater depths (where they are more protected from the alpine environment) and still reach the soil surface [@RN4956; @RN4904].
+It is also not surprising that we detected a negative interaction between seed mass and the germination response to light and alternating temperature regimes, since large seeds can afford to germinate at greater depths (where they are more protected from the alpine environment) and still reach the soil surface [@RN4898; @RN4904].
 
@@
 
-Using the whole set of traits compiled in this study we visualize the variation in the seed germination spectrum of alpine plants.
-On one end of the spectrum, strict alpine species tend to show slow and synchronous germination, require warm temperatures and light to germinate, and need stratification or GA~3~ to break physiological dormancy.
-On the other end, generalist species tend to have fast germination, showing a bet-hedging strategy (high germination uncertainty) and germination traits that suggest possible fast germination after specific cues (e.g.
+Using the whole set of traits compiled in this study we can visualize the variation in the seed germination spectrum of alpine plants.
+At one end of the spectrum, strict alpine species tend to show slow and synchronous germination, require warm temperatures and light to germinate, and need stratification or GA~3~ to break physiological dormancy.
+At the other end, generalist species tend to have fast germination, showing a bet-hedging strategy (high germination uncertainty) and germination traits that suggest possible fast germination after specific cues (e.g.
 scarification or alternating temperatures).
@@
 
-## Conclusions
+## Future directions
 
-Alpine plants from different regions of the world respond similarly to germination cues that are expected to support successful regeneration: cold stratification, warm and alternating temperatures, and light.
-These cues, particularly cold stratification and warm temperatures, are generally associated with cold-adapted herbaceous species in a broad sense [@RN3397; @RN3214; @RN2365].
-Nonetheless, our model shows that the response to stratification and temperature is more prominent in strict alpine species when comparing them with generalist plants that can grow opportunistically in the alpine belt, suggesting an evolutionary convergence of germination patterns in those species truly adapted to the alpine.
-We can conclude that, while alpine plants do not show a unique germination syndrome, they do show a more extreme or specialized version of the cold-adapted temperate germination syndrome.
+Our first global assessment of the seed germination spectrum in relation with seed morphology supports the existence of trade-offs between sets of regenerative traits [@RN4655], such as the relationship we found between seed/embryo size, germination speed and germination responses to temperature and light.
+Future studies will need to consider, as we did, evolutionary history and seed anatomy as intrinsic constraints of the seed ecological spectrum; while the seed germination spectrum we described here will need to be explored at the community level [@RN2352].
 
-We were not able to analyse data from tropical alpine regions, highlighting how tropical alpine seeds are underrepresented in the literature.
-We recommend that tropical alpine regions be the focus of future research efforts, as it is possible that the cold-adapted germination syndrome is not prevalent there.
+This study also highlights the general lack of germination studies on tropical alpine seeds.
+We were not able to analyse data from tropical alpine regions because they are underrepresented in the literature.
+Although there is an increasing interest in plant ecology of tropical alpine regions, we could not find seed germination data to be compared with other regions.
+We thus recommend to focus future research on seed ecology in high-mountain tropical areas, as it is possible that the cold-adapted germination syndrome is not prevalent there.
 Tropical alpine environments are relatively aseasonal and most plants have been reported as flowering throughout the year [@RN4957].
 Rather than being exposed to snow in winter and high temperatures in summer, tropical alpine plants face a year-round stress from night-time frost [@RN2392].
-That means that our study is valid for the temperate alpine system, but not necessarily for tropical mountain species pools, which are subjected to different selective pressures.
+That means that our study is valid for the temperate alpine system, where alpine habitats mainly occur [@RN4750], but not necessarily for tropical alpine systems, which are subjected to different selective pressures.
 
@@
 New research should aim to decompose these binary cues into finer quantitative variables which provide a more precise environmental signalling, such as the red:far red ratio [@RN604], the length of the photoperiod [@RN1883], the amplitude of the diurnal thermal oscillations [@RN2387] or the length and temperature of cold stratification [@RN2346].
+In addition, new field research is necessary to assess germination phenology in response to temperature and water availability on the upper soil layer.
 
-Overall, our first global assessment of the seed germination spectrum in relation with seed morphology supports the existence of trade-offs between sets of regenerative traits [@RN4655], such as the relationship we found between seed/embryo size, germination speed and germination responses to temperature and light.
-Future studies will need to consider, as we did, evolutionary history as an intrinsic constraint of the seed ecological spectrum.
+## Conclusions
+
+This study demonstrates how alpine plants from different regions of the world respond similarly to germination cues that support successful regeneration: cold stratification, warm and alternating temperatures, and light.
+These cues, particularly cold stratification and warm temperatures, are generally associated with cold-adapted herbaceous species in a broad sense [@RN3397; @RN3214; @RN2365].
+Nonetheless, our results show that the response to stratification and temperature is more prominent in strict alpine species when comparing them with generalist plants that can grow opportunistically in the alpine belt, suggesting an evolutionary convergence of germination patterns in those species truly adapted to alpine ecosystems.
+We can conclude that, while alpine plants do not show a unique germination syndrome, they do show a more extreme or specialized version of the cold-adapted temperate germination syndrome.
 
@@
 EFP received financial support from the Government of Asturias and the FP7 – Marie Curie - COFUND programme of the European Commission (Grant ‘Clarín’ ACB17-19).
-BJA was funded by the Spanish Research Agency (PID2019-108636GA/ AEI/ 10.13039/501100011033).
+BJA and EFP were funded by the Spanish Research Agency (PID2019-108636GA/ AEI/ 10.13039/501100011033).
 EM was supported by the Future Leaders Fellowship - Diversity and Livelihoods, of the Royal Botanic Gardens, Kew.
@@
 Upon publication, a version of record of the repository will be deposited in Zenodo.
-(The repository will be made open when the manuscript is accepted for publication, a zip version has been uploaded to https://drive.google.com/file/d/12Usqi7fOAbOoZ9Uu9wkJt2FANB629LFP/view?usp=sharing for peer-review)
 
@@
 
-```{r fig2, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the germination environment on germination, according to the MCMC meta-analysis of the primary data.
+```{r fig2, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the germination environment on germination, according to the MCMC meta-analysis of primary data.
 The dots indicate the posterior mean of the effect size, and the brackets its 95 % credible interval.
@@
 
-```{r fig3, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the random factors on germination, according to the MCMC meta-analysis of the primary data.
+```{r fig3, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the random factors on germination, according to the MCMC meta-analysis of primary data.
 The dots indicate the posterior mean of the effect size, and the brackets its 95 % credible interval.

### 2020-11-02 src/cleaning.R

@@ -1,1180 +1,40 @@
 library(tidyverse); library(readxl)
 
-# This script has the code to read the primary germination data from the different contributors
-# and merge it into a clean germination dataset
+# This script imports the primary data from my local drive
 
-# CLEAN GERMINATION DATA
+# Read germination data
 
-## Format Australian data (susanna Venn)
-## Read from original files with germination indices (S Rosbakh)
+read.csv("../#data/germination/results/alpineseedsDB.csv") -> germination
 
-read_excel("data/venn/VENN hand joined.xlsx",
-           sheet = 1, skip = 0) %>%
-  mutate(Region = "Australian Alps",
-         Tmin = ifelse(is.na(Tmin), Tmax, Tmin),
-         Tmean = (Tmax * Photoperiod + Tmin * (24 - Photoperiod)) / 24, 
-         Tdif = Tmax - Tmin, 
-         Alternating = as.factor(ifelse(Tdif != 0, "Y", "N"))) %>%
-  select(Taxon,
-         Source, 
-         Country, 
-         Region, 
-         Accession, 
-         Latitude, 
-         Longitude, 
-         Elevation, 
-         Dish, 
-         Scarification,
-         GA3,
-         Stratification_temperature, 
-         Stratification_days, 
-         Stratification, 
-         Light,
-         Photoperiod, 
-         Alternating,
-         Tmean,
-         Tmax, 
-         Tmin,
-         Tdif,
-         Sown, 
-         Germinated, 
-         Normal,
-         GRS:CVG) ->
-  AustraliaVenn
-
-## Format Australian data (Annisa Satyanti)
-## Read from original files with germination indices (S Rosbakh)
-
-read_excel("data/satyanti/UPDATE_Aust Alpine Germination timing 11MAY2020.xlsx",
-           sheet = 4, skip = 0) %>%
-  select(Species, Accession, Elevation, Treatment, Rep, seeds, GRS:CVG) -> satyanti1
-read_excel("data/satyanti/UPDATE_Aust Alpine Germination timing 11MAY2020.xlsx",
-           sheet = 8, skip = 0)  %>%
-  select(Species, Accession, Elevation, Treatment, Rep, seeds, GRS:CVG) -> satyanti2
-read_excel("data/satyanti/UPDATE_Aust Alpine Germination timing 11MAY2020.xlsx",
[…]

### 2020-11-02 src/manuscript-source.R

@@ -5,16 +5,8 @@ library(tidyverse)
 
 # Prepare dataset object to create tables, figures, etc for the manuscript
 
-read.csv("../data/fernandezpascual/Regions.csv") %>%
-  merge(read.csv("../results/germination.csv"), by = "Region", all = TRUE) %>%
-  merge(read.csv("../results/traits.csv"), by = "TPLName") %>%
-  mutate(Region = as.character(Region),
-         Region = ifelse(is.na(Macroregion), "Enscobase", Region),
-         Region = ifelse(Region %in% c("Northern Alps", "Southern Alps"), "Alps", Region),
-         Region = ifelse(Region %in% c("Central Andes", "Southern Andes"), "Andes", Region),
-         Region = ifelse(Region %in% c("Qinghai-Tibet Plateau"), "Qinghai-Tibet", Region),
-         Region = ifelse(Region %in% c("Cantabrian Mountains"), "Cantabrian Mts", Region),     
-         Region = as.factor(Region)) -> dataset
+read.csv("../data/germination.csv") %>%
+  merge(read.csv("../data/traits.csv"), by = "TPLName") -> dataset
 
 # Table 1
 
@@ -22,9 +14,6 @@ data.frame(
 merge(
 merge(
 dataset %>%
-  mutate(Source = fct_recode(Source, `Fernández-Pascual` = "Fernández-Pascual Plant Biol")) %>%
-  mutate(Region = fct_recode(Region, `European Alps` = "Alps")) %>%
-  mutate(Region = fct_recode(Region, `Europe` = "Enscobase")) %>%
   select(Source, Region, Country, TPLName, Alpine) %>%
   group_by(Source, Region) %>%
   tally %>%
@@ -32,9 +21,6 @@ dataset %>%
          `Coverage` = Region,
          Records = n),
 dataset %>%
-  mutate(Source = fct_recode(Source, `Fernández-Pascual` = "Fernández-Pascual Plant Biol")) %>%
-  mutate(Region = fct_recode(Region, `European Alps` = "Alps")) %>%
-  mutate(Region = fct_recode(Region, `Europe` = "Enscobase")) %>%
   select(Source, Region, Country, TPLName, Alpine) %>%
   unique %>%
   group_by(Source, Region) %>%
@@ -44,9 +30,6 @@ dataset %>%
          Species = n)),
 dataset %>%
   filter(Alpine == "Strict") %>%
[…]

### 2020-11-07 README.Rmd
minor revision 🌄

@@
 
-* `data` Primary germination data and species traits.
+* `data` Primary data, including the [germination](https://github.com/efernandezpascual/alpineseeds/blob/master/data/germination.csv) dataset, the species [traits](https://github.com/efernandezpascual/alpineseeds/blob/master/data/traits.csv) and the phylogenetic [tree](https://github.com/efernandezpascual/alpineseeds/blob/master/data/alpine.tree.pdf).
 * `doc` Files to create the manuscript using `Rmarkdown`.
-* `results` Output of the `R` scripts, including MCMCglmm analysis outputs.
+* `results` Output of the `R` scripts, including [MCMCglmm analysis](https://github.com/efernandezpascual/alpineseeds/tree/master/results/MCMCglmm) outputs.
 * `src` Scripts in `R` language used to process data and perform the analyses of the manuscript.

### 2020-11-07 doc/cover-letter.Rmd
minor revision 🌄

--- 2020-09-14 doc/cover-letter.Rmd
@@
 ---
-title: "Cover letter"
-author: "Eduardo Fernández Pascual"
-date: "14/09/2020"
-output: word_document
+title: "NPH-MS-2020-34477.R1 - Response to review"
+date: "`r format(Sys.time(), '%d/%m/%Y')`"
+output: pdf_document
 bibliography: refs.bib
@@
 always_allow_html: true
+urlcolor: blue
 ---
@@
 
-Dear Editorial Board,
+Dear Editor,
 
-I submit the manuscript *The seed germination spectrum of alpine plants: a global meta-analysis* to be considered for publication as a **full paper** in New Phytologist.
+Thank you for your positive comments and feedback.
+We submit a revised version of our manuscript, following your suggestions.
+In addition to the specific points commented below, we have revised the whole manuscript to improve its clarity.
 
-## What hypotheses or questions does this work address?
+> **1) Line 37: This is minor but I’m not in favour of using “meta-analyse” as a verb.
+Perhaps change to “We performed a meta-analysis…” I realize this uses a few more words but it reads better to my mind.**
 
-This is a global meta-analysis of primary germination data contributed by 12 research groups from 8 alpine regions in 4 continents.
-We used 9,799 records representing 62 seed plant families and 661 species to assess whether the germination ecological patterns detected by local studies are generalized at the global level.
+We have done the suggested change.
 
-## How does this work advance our current understanding of plant science?
+> **2) Line 270: You imply (although it is a little vague) that for “some” species you filled in their values based on the average values for the genus or family.
+Please state how many species this was done for (and how many were filled in with genus versus family info) and state clearly whether this is indicated in the GitHub dataset available which ones were filled in using this procedure.
+It matters greatly to how valuable this dataset becomes.**
 
-We describe a global alpine germination syndrome, information valuable for all scientists working in alpine environments.
-Our results indicate a convergence of the seed germination spectrum in strict alpine species.
-We also found interactive effects of phylogeny and seed morphology that point to trade-offs shaping the seed ecological spectrum.
+We have added to the text the number of species with values calculated from species-values, genus-averages or family-averages.
+We also added this information to the primary data available at the [GitHub repository](https://github.com/efernandezpascual/alpineseeds).
+We have indicated, in the relevant places of the text, which datasets ([germination](https://github.com/efernandezpascual/alpineseeds/blob/master/data/germination.csv), [species traits](https://github.com/efernandezpascual/alpineseeds/blob/master/data/traits.csv), and [phylogenetic tree](https://github.com/efernandezpascual/alpineseeds/blob/master/data/alpine.tree.pdf)) and results (detailed results of the [MCMCglmm analysis](https://github.com/efernandezpascual/alpineseeds/tree/master/results/MCMCglmm)) are available at GitHub.
 
-## Why is this work important and timely?
+> **3) Line 312: Is the phylogeny also available on Github?
+I’m guessing users would have to run the code to be able to generate it and it would be better if a visual was provided in the supplementary information of the paper itself.
+**
 
-This is the first global evaluation, for a coherent habitat, of the seed germination spectrum and its relationship with other aspects of the seed ecological spectrum.
-This was one of the objectives in the research agenda for trait-based seed ecology recently published by New Phytologist [@RN4655].
+The phylogeny is available at the [GitHub repository](https://github.com/efernandezpascual/alpineseeds) as a [.tree file](https://github.com/efernandezpascual/alpineseeds/blob/master/data/ALPINE.tree) that can be imported into R, or opened with software like [TreeGraph](http://treegraph.bioinfweb.info/).
+We have now also included a [pdf version of the tree](https://github.com/efernandezpascual/alpineseeds/blob/master/data/alpine.tree.pdf), so it can be visualized directly.
+We have indicated at the appropriate point of the text that the tree can be accessed at GitHub.
+We would prefer this solution to including it as a supplementary file, to avoid duplicating data sources.
 
-## References
+> **4) Lines 503-508: This is a very long sentence. Please divide into 2.**
+
+We have divided the sentence in two, as requested.
+
+> **5) Line 610: I think these are the sort of vague statements that Reviewer #2 was referring to regarding how phylogeny is confounding your models and the manuscript could still use some improvement here.
+For example, you provide phylogenetic signal in variables such as germination time on temperature (although I would argue that Figure 4 makes the values too difficult to distinguish in the 6 panel graphs), but I do not see values for the phylogenetic signal in whether a species is alpine vs.
+generalist, embryo-seed size, seed mass, etc.
+Please provide the phylogenetic signal values for all of these traits.
+Doing so will make statements regarding how “traits are connected and may be subjected to co-adaptation” much easier to interpret.
+Some improvements in Figure 4 would also be welcome (or perhaps the values can be provided instead in a table).**
+
+We have changed the sentence in former L610 to clarify that we are referring to the phylogenetic signal we found in the germination responses.
+
+Our study focuses on germination responses, and the MCMC models provide a value of phylogenetic signal in (i) the responses themselves [i.e.
+the main effects], and (2) the interaction between the responses and the other traits (i.e.
+the moderator effect of seed mass, embryo, strict/generalist).
+In Figure 4, we only show the signal in the main effects, because the signal in the interactions is almost identical to that in the corresponding main effects.
+However, the detailed values for each main effect and interaction can be found in the [MCMCglmm results](https://github.com/efernandezpascual/alpineseeds/tree/master/results/MCMCglmm) folder at GitHub (we now specify this in the text).
+
+We have calculated phylogenetic signal in the other traits.
+It was strong in both seed mass (lambda = 0.94) and the embryo:seed ratio (lambda = 0.99), corresponding to what can be expected from the literature.
+We also calculated phylogenetic signal in the strict alpine vs.
+generalist character using D [@RN4959], a metric for binary traits.
+The value of D was close to 1 (D = 0.92) indicating that the alpine vs.
+generalist character was randomly distributed across the phylogeny, this is, it had low phylogenetic signal.
+Nevertheless, we would rather not incorporate these values into the text, for two reasons.
+First, phylogenetic signal in other traits is not part of the story in this manuscript (rather, our focus would be in the interaction between other traits and germination responses).
+Second, since seed mass and embryo have been estimated from genus or family values for part of the species, estimates of phylogenetic signal in these traits can be misleading.
+There are many references focusing more explicitly in phylogenetic signal in seed mass and embryo size, and we discuss some of these in the discussion.
+
+We have revised fonts and sizes of all figures to improve clarity.
+
+> **6) Line 642: As Reviewer #2 mentioned with the previous submission, there is a tendency to state things in a vague way.
+Here again you mention that tropical systems are subjected to “different selective pressures”.
+Different how?**
+
+We have revised the entire paragraph, to make clear that we referred to the environmental conditions that we had described in previous sentences: absence of yearly climatic seasonality, strong night-day differences.
+
+# References

### 2020-11-07 doc/manuscript.Rmd
minor revision 🌄

@@
 
-**Total word count for the main body of the text** 6,843; **Introduction** 1,888; **Materials and Methods** 1,517; **Results** 1,132; **Discussion** 2,306; **Number of figures** 5 in colour; **Tables** 1; **Supporting information** 0
+**Total word count for the main body of the text** 6,837; **Introduction** 1,826; **Materials and Methods** 1,642; **Results** 1,155; **Discussion** 2,214; **Number of figures** 5 in colour; **Tables** 1; **Supporting information** 0
 
@@
 * Assumptions about the germination ecology of alpine plants are presently based on individual species and local studies.
-A current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants.
-* We meta-analyse primary data from laboratory experiments conducted across four continents (excluding the tropics) and 661 species, to estimate the influence of six environmental cues on germination proportion, mean germination time and germination synchrony; accounting for possible effects of seed morphology (mass, embryo:seed ratio) and phylogeny.
+A current challenge is to synthesize, at the global level, the alpine seed ecological spectrum.
+* We performed a meta-analysis of primary data from laboratory experiments conducted across four continents (excluding the tropics) and 661 species, to estimate the influence of six environmental cues on germination proportion, mean germination time and germination synchrony; accounting for possible effects of seed morphology (mass, embryo:seed ratio) and phylogeny.
 * Most alpine plants show physiological seed dormancy, a strong need for cold stratification, warm-cued germination and positive germination responses to light and alternating temperatures.
-Species with a distribution limited to the alpine belt have a higher preference for warm temperatures and a stronger response to cold stratification than species whose distribution extends also below the treeline.
+Species restricted to the alpine belt have a higher preference for warm temperatures and a stronger response to cold stratification than species whose distribution extends also below the treeline.
 Seed mass, embryo size and phylogeny have strong constraining effects on germination responses to the environment.
@@
 Although most alpine plants can reproduce clonally [@RN2392], sexual reproduction by diaspores (i.e.
-seeds, fruits, or fruits with extra-ovarian structures attached; hereafter 'seeds') is the main strategy to maintain genetic diversity and to colonize suitable new sites in response to environmental changes.
-Plant regeneration by seed is a key process that determines whether species will persist in their current habitat or will be able to migrate to new sites [@RN3055]; and it is also fundamental for plant conservation [@RN3763] and seed-based ecological restoration [@RN3507; @RN2241].
+seeds, fruits, or fruits with extra-ovarian structures attached; hereafter 'seeds') is the main strategy to maintain genetic diversity and to colonize suitable new sites in response to environmental changes [@RN3055].
 However, the short reproductive season of the alpine habitat constrains the phenological timing of flowering, pollination, seed development and seed dispersal.
@@
 Natural selection is expected to favour seed germination requirements that reduce the probability of facing environmental conditions which are not appropriate for seedling survival and growth [@RN3375].
-Thus, germination is mostly regulated by environmental cues related to water availability and soil temperature [@RN3368].
+Thus, germination is chiefly regulated by environmental cues related to water availability and soil temperature [@RN3368].
 The propagules of many plant species have also developed seed dormancy, by means of which germination is prevented during periods that are only ephemerally favourable, like a short warm spell in the middle of winter [@RN3214].
@@
 Other germination cues, such as the response to diurnal alternating temperatures [@RN1380] and the need for light or darkness [@RN4656], allow for a fine-scale detection of germination micro-niches and safe sites [@RN4719].
-Despite the high micro-climatic heterogeneity of alpine landscapes, current assumptions about germination ecology of alpine plants are mainly based on local studies (reviewed in the following paragraphs).
-A current challenge is thus to synthesize, at the global level, the seed ecological spectrum of alpine plants.
+Given that current assumptions about the germination ecology of alpine plants are mainly based on local studies, a current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants.
 The seed ecological spectrum [@RN4655] is defined as a set of seed traits that determines the ability of plants to disperse, persist, germinate, and establish in different habitats.
-A major goal for trait-based seed ecology is therefore to study the relationships and trade-offs between different aspects of the seed ecological spectrum [@RN4655], such as the physiological drivers of germination versus key morphological traits like seed mass [@RN4916] and the embryo to seed ratio [@RN3685].
+A major goal for trait-based seed ecology is to study the relationships and trade-offs between different aspects of the seed ecological spectrum [@RN4655], such as the physiological drivers of germination versus key morphological traits like seed mass [@RN4916] and the embryo to seed ratio [@RN3685].
 
@@
 cold stratification) increased seed germination in most species.
-Nonetheless, Amen [-@RN3213] later suggested that cold stratification was not a requirement for the seed germination of alpine species from different mountain ranges in the USA, a finding that was supported by subsequent studies [@RN1484; @RN2995; @RN4721].
+Nonetheless, Amen [-@RN3213] later suggested that cold stratification was not a requirement for seed germination of alpine species from different mountain ranges in the USA, a finding that was supported by subsequent studies [@RN1484; @RN2995; @RN4721].
 More recently, research on Australian alpine plants found that a cold stratification period only increased germination in half of the species tested [@RN3703; @RN3285].
-Cold stratification also increased seed germination over a range of temperatures in most of the species studied in the alpine zone of Japan [@RN707].
+Cold stratification increased seed germination over a range of temperatures in most of the species studied in the alpine zone of Japan [@RN707].
 Along an elevational gradient in the central Chilean Andes, cold stratification was an important requirement for seed germination in species from lower elevations, while species from higher sites did not germinate after a single period of cold stratification [@RN4713].
@@
 Seed dormancy is not the only trait that can prevent, in practice, autumn germination.
-Early studies on the germination ecology of alpine plants demonstrated that in most species recently dispersed seeds require relatively high temperatures for germination [@RN4951; @RN3258; @RN3213; @RN4712].
+Early studies on the germination ecology of alpine plants demonstrated that, in most species, recently dispersed seeds require relatively high temperatures for germination [@RN4951; @RN3258; @RN3213; @RN4712].
 Warm-cued germination has been considered as an adaptation to prevent seed germination at the time of seed dispersal (autumn) when temperatures are relatively low and there is a high risk of frost [@RN3393].
@@
 The assembled dataset represents eight seasonal alpine regions belonging to four continents.
-We meta-analyse this dataset to investigate the germination response of alpine plants to key environmental factors: dormancy-breaking treatments (cold stratification, GA~3~ and scarification), average germination temperature, diurnal alternating temperatures and light [@RN3368; @RN3214].
+We perform a meta-analysis of this dataset to investigate the germination response of alpine plants to key environmental factors: dormancy-breaking treatments (cold stratification, GA~3~ and scarification), average germination temperature, diurnal alternating temperatures and light [@RN3368; @RN3214].
 Given the heterogeneity of alpine species lineages adapted to different regions, we account for the shared evolutionary history using a phylogenetic comparative approach [@RN3081].
@@
 All data processing, analysis and manuscript production has been performed in *R* [@RN2315].
-The original data, plus R code for the analysis and creation of the manuscript using Rmarkdown, can be accessed at the GitHub repository https://github.com/efernandezpascual/alpineseeds.
+The original data, plus R code for the analysis and creation of the manuscript using Rmarkdown, can be accessed at the GitHub repository (see Data Availability Statement).
 To gather data, we contacted research groups which have performed experiments on seed germination of alpine plants, asking them to provide raw data from laboratory germination experiments.
@@
 We followed the concept by Vleeshouwers et al.
-[-@RN3254] to define dormancy as an inner seed property (sometimes called "endodormancy"), which should be differentiated from the absence of appropriate germination cues in the environment ("ecodormancy"); and the classification system of Baskin and Baskin [-@RN3261] based on previous work by Nikolaeva et al.
+[-@RN3254] to define dormancy as an inner seed property (sometimes called 'endodormancy'), which should be differentiated from the absence of appropriate germination cues in the environment ('ecodormancy'); and the classification system of Baskin and Baskin [-@RN3261] based on previous work by Nikolaeva et al.
 [-@RN2877] who consider morphological and physical dormancy as classes separated from physiological dormancy.
@@
 For some species that were missing seed mass or embryo values, we calculated genus or family averages.
+In the case of seed mass, we used species-level values for `r MSmassrankspp` species, and genus averages for another `r MSmassrankgen` species.
+For embryo mass, we used `r MSembrankspp` species-level values, `r MSembrankgen` genus averages, and `r MSembrankfam` family averages.
+The collected species traits, including information on which taxonomic rank was used to calculate the values for each species, is available in the data folder of the GitHub repository (see Data Availability Statement).
 
@@
 a fraction of the seed lot is non-dormant at the beginning of the incubation and germinates shortly after the start of the experiment; while another fraction is originally dormant, loses dormancy in response to the incubation conditions, and germinates in a second wave).
+The primary germination dataset is available in the data folder of the GitHub repository (see Data Availability Statement).
 
@@
 We placed taxa absent from the mega-tree at the genus-level basal node.
+The phylogenetic tree is available in the data folder of the GitHub repository (see Data Availability Statement).
 In all models, all variables were scaled so their contribution to the effect sizes could be compared.
@@
 When λ = 0, related taxa are no more similar than expected by chance, while when λ = 1, the trait is evolving following a constant variance random walk or Brownian motion model; intermediate values of λ indicate a phylogenetic correlation in trait evolution that does not fully follow a Brownian motion model [@RN4757].
+Detailed results of the MCMCglmms are available in the results folder of the GitHub repository (see Data Availability Statement).
 
@@
 The results for the germination uncertainty index indicate that it might have been more sensitive to the frequency of germination scoring, which was different in different labs.
+Detailed values of phylogenetic signal for all model interactions are available in the results folder of the GitHub repository (see Data Availability Statement).
 
@@
 
-Our meta-analysis of primary germination data shows that alpine plants from different mountains of the world tend to show specialized seed germination characteristics.
-For most species thriving in alpine habitats, both dormancy and warm-cued germination combine to ensure that germination occurs in the season most conducive to survival.
-In particular, seeds tend to germinate when, after weeks of exposure to conditions that mimic winter, they experience mid-summer temperatures.
-The germination ecology of strict alpine species is therefore characterized by (1) physiological seed dormancy; (2) a strong need for cold stratification or GA~3~ to break dormancy; (3) warm-cued germination; (4) a positive response to alternating temperatures, although not so prominent as in generalist species; (5) a positive response to light; (6) slow and relatively synchronous germination; and (7) an interplay with seed and embryo size, with smaller and more endospermic seeds being more responsive to stratification, warmth, alternating temperatures and light.
-Here we discuss the ecological significance of these responses and their role to define the seed germination spectrum of alpine plants.
+Our meta-analysis of primary germination data shows that alpine seeds tend to germinate when, after weeks of exposure to conditions that mimic winter, they experience mid-summer temperatures.
+The germination ecology of strict alpine species is characterized by (1) physiological seed dormancy; (2) a strong need for cold stratification or GA~3~ to break dormancy; (3) warm-cued germination; (4) a positive response to alternating temperatures, although not so prominent as in generalist species; (5) a positive response to light; (6) slow and relatively synchronous germination; and (7) an interplay with seed and embryo size, with smaller and more endospermic seeds being more responsive to stratification, warmth, alternating temperatures and light.
+Here we discuss the ecological significance of these responses and how they define the seed germination spectrum of alpine plants.
 
@@
 
-Our dataset indicates that seed dormancy is frequent across alpine species and plays an important role in postponing plant recruitment to favourable times.
+Our dataset shows that seed dormancy is frequent across alpine species and plays an important role in postponing plant recruitment to favourable times.
 The frequency of dormancy classes displayed by strict alpine and generalist species is similar, suggesting that this is a general trait of cold-adapted mountain species.
@@
 cold and wet, although other combinations are possible) are met over a period of months [@RN3214; @RN4743].
-Indeed, our results show that cold stratification consistently promotes germination, while the phytohormone gibberellic acid (GA~3~) works as a general substitute of cold stratification to overcome the inherent physiological dormancy of alpine species.
+Indeed, our results show that cold stratification consistently promotes germination, while the phytohormone gibberellic acid (GA~3~) works as a substitute of cold stratification to overcome physiological dormancy.
 Additionally, we found some species with physical dormancy.
@@
 
-Overall, our results confirm that overwintering is essential for promoting seed germination in alpine species, concurring with previous local studies [@RN3393; @RN2868; @RN707; @RN698; @RN3703; @RN2355; @RN3285; @RN2371; @RN4713].
+Together, these results confirm that overwintering is essential for promoting seed germination in alpine species, concurring with previous local studies [@RN3393; @RN2868; @RN707; @RN698; @RN3703; @RN2355; @RN3285; @RN2371; @RN4713].
 Since global climate change is causing major changes in snow cover duration [@RN4917; @RN4918], it could disrupt natural cold stratification cycles in alpine environments, compromising population viability and indirectly favouring species with no stratification requirement [@RN3703].
-This risk will be partly alleviated by the plasticity of physiological seed dormancy in response to the seed maturation temperature [@RN2358; @RN2899], a mechanism of “seed memory” via maternal effects that facilitates acclimatisation to changing environments [@RN2249], reducing the stratification requirements of seeds produced in a warmer climate.
+This risk will be partly alleviated by the plasticity of physiological seed dormancy in response to the seed maturation temperature [@RN2358; @RN2899].
+'Seed memory' via maternal effects facilitates acclimatisation to changing environments, and can reduce the stratification requirements of seeds produced in a warmer climate [@RN2249].
 
@@
 
-We also found a general increase of seed germination with higher incubation temperatures.
+We found a general increase of seed germination with higher incubation temperatures.
 This effect was more pronounced in strict alpine compared to generalist species, confirming previous local studies [@RN3266; @RN2371].
@@
 Moreover, strict alpine species may be less subjected to drought damage during establishment compared to generalist species: as elevation increases, plants face less drought risk because precipitation increases (except in Mediterranean and semi-arid mountains), but at the same time there is a higher probability of early or late season frosts [@RN4897; @RN2392; @RN4914].
-Together, these results indicate that germination patterns in alpine species are driven by an interplay of seed size and germination temperature to escape either frost or drought stress: small seeds are being selected for fast germination at warm temperatures, in order to escape unfavourable early-spring frost, and large seeds are being selected for slower germination under cooler conditions, which should maximize emergence during snowmelt (when water availability is highest) and gives seedlings adequate time to develop a root system to cope with desiccation risks in summer.
+Together, these results indicate that germination patterns in alpine species are driven by an interplay of seed size and germination temperature to escape either frost or drought stress: small seeds are being selected for fast germination at warm temperatures, in order to escape unfavourable early-spring frost; while large seeds are being selected for slower germination under cooler conditions, which should maximize emergence during snowmelt (when water availability is highest) giving seedlings adequate time to develop a root system to cope with desiccation risks in summer.
 Supporting this view, seedlings of alpine pioneer species have a low frost resistance [@RN4722], yet tolerate heat up to 40–50°C [@RN4902].
@@
 staggering germination so that emergence occurs both before and after winter, is known in alpine species [@RN2392; @RN3285; @RN4907].
-Staggered germination is usually exhibited as a rapid first wave of germination followed by a second wave several weeks or months later [@RN2392].
 If autumn seedlings grow faster than spring seedlings [@RN4907], this will likely have cascading effects on individual fitness along the life of the plant [@RN3036; @RN2384].
@@
 
-We found a positive germination response to alternating temperatures and light, conditions that mimic the environment in the upper soil and certain microsites during the snow-free season [@RN4712; @RN2392; @RN3029].
+We also found a positive germination response to alternating temperatures and light, conditions that mimic the environment in the upper soil and certain microsites during the snow-free season [@RN4712; @RN2392; @RN3029].
 Alternating temperatures had a slightly lower effect on strict alpine species, compared with generalist species, but still had some effect.
 This partly contradicts the findings of Liu et al.
-[-@RN3376], who reported a lack of a response to temperature fluctuation for the species distributed only at high elevations of the Qinghai-Tibet Plateau.
-Liu and co-workers argued that, in that area, high temperature fluctuations may occur in all months of the year and therefore fluctuating temperature alone may not be a reliable indicator of suitable conditions for seedling establishment and growth, while warm daily temperatures could be a cue that the short summer growing season has begun [@RN3376].
-When analysed at the global level, alternating temperatures generally increase germination of alpine plants, which corresponds well with the fact that most alpine regions have a strong day-night climatic variation during the growing season.
+[-@RN3376], who found no response to temperature fluctuations for species distributed only at high elevations of the Qinghai-Tibet Plateau.
+Liu and co-workers argued that, in that area, high temperature fluctuations may occur throughout the year, and therefore fluctuating temperature alone may not be a reliable indicator of suitable conditions for seedling establishment and growth [@RN3376].
+When analysed at the global level, alternating temperatures generally increase germination of alpine seeds, which corresponds well with the fact that most alpine regions have a strong day-night climatic variation during the growing season.
 It is also not surprising that we detected a negative interaction between seed mass and the germination response to light and alternating temperature regimes, since large seeds can afford to germinate at greater depths (where they are more protected from the alpine environment) and still reach the soil surface [@RN4898; @RN4904].
@@
 
-Using the whole set of traits compiled in this study we can visualize the variation in the seed germination spectrum of alpine plants.
-At one end of the spectrum, strict alpine species tend to show slow and synchronous germination, require warm temperatures and light to germinate, and need stratification or GA~3~ to break physiological dormancy.
-At the other end, generalist species tend to have fast germination, showing a bet-hedging strategy (high germination uncertainty) and germination traits that suggest possible fast germination after specific cues (e.g.
+Using the whole set of traits compiled in this study we can visualize the seed germination spectrum of alpine plants.
+At one end of the spectrum, strict alpine species tend to germinate slowly and synchronously, require warm temperatures and light to germinate, and need stratification or GA~3~ to break physiological dormancy.
+At the other end, generalist species tend to germinate fast, show a bet-hedging strategy (i.e.
+high germination uncertainty) and germination traits that suggest fast germination after specific cues (e.g.
 scarification or alternating temperatures).
@@
 This gradient of regenerative strategies mirrors prominent gradients of fast or slow population dynamics [@RN4922; @RN4921], fast or slow developing leaf traits [@RN4923], or large sets of plant ecological features [@RN2279].
-“Fast” regeneration can be understood as a plastic and opportunistic strategy to cope with frequently disturbed habitats such as avalanche ways, steep eroding slopes, regressing glaciers, and riverbeds [@RN4926; @RN4925].
-The “slow” regeneration of strict alpine species, on the other hand, seems to be a more specialized syndrome and might be the result of convergent evolution acting on those species that are truly restricted to the alpine vegetation belt.
+'Fast' regeneration can be understood as a plastic and opportunistic strategy to cope with frequently disturbed habitats such as avalanche ways, steep eroding slopes, regressing glaciers, and riverbeds [@RN4926; @RN4925].
+The 'slow' regeneration of strict alpine species, on the other hand, seems to be a more specialized syndrome and might be the result of convergent evolution acting on those species that are truly restricted to the alpine vegetation belt.
 
 A general assumption of seed trait ecology is that seed mass and embryo:seed ratios are relatively conserved across related species [@RN4916; @RN3685].
-In our alpine dataset, both traits showed no differences between strict alpine and generalist species.
-This is in line with the ambiguity of relationships that have been found between seed mass and elevation for alpine species [@RN4912; @RN4903; @RN3685; @RN4896].
-The similarity in embryo:seed size between generalist and strict alpine species confirms the lack of correlation between embryo-seed size and elevation found in the Apiaceae [@RN3685].
+In our alpine dataset, both traits show no differences between strict alpine and generalist species.
+This is in line with the ambiguity of relationships between seed mass and elevation that have been found for alpine species, including examples of negative [@RN4912], positive [@RN4903] or absent [@RN3685; @RN4896] correlations.
+The similarity in embryo:seed ratio between generalist and strict alpine species also confirms the lack of correlation between embryo-seed ratio and elevation found in the Apiaceae [@RN3685].
 In contrast with the supposed phylogenetic conservatism of these morphological traits, physiological germination responses are understood to be more plastic traits that can quickly respond to new situations [@RN2249].
 While phylogenetic signal in seed germination is generally found, especially in highly selective environments [@RN3001; @RN2399; @RN2865], germination patterns can also be shared between phylogenetically distant species [@RN4911] and have low or absent phylogenetic signal [@RN2365; @RN4900; @RN2371].
-The phylogenetic signal found here indicates that evolutionary history cannot be neglected when studying alpine seed germination patterns.
-Furthermore, we detected close relationships between seed-embryo size and physiological responses of seeds, suggesting that both sets of traits are connected and may be subjected to co-adaptation.
+The phylogenetic signal in germination responses that we found here indicates that evolutionary history cannot be neglected when studying seed germination patterns.
+Furthermore, we detected that seed-embryo size influences germination responses to environmental cues, with smaller and more endospermic seeds being more responsive to warmth, light and alternating temperatures.
 Previous studies have shown that non-endospermic seeds are smaller and are likely to be non-dormant at the time of dispersal and thus germinate quicker than alpine endospermic seeds [@RN3703; @RN3285; @RN4907].
-Together, these results indicate that seed morphology and physiology are connected.
+Together, these results indicate that seed morphology and physiology are connected and may be subjected to co-adaptation.
 This connection, intertwined with their shared phylogenetic history, may constrict the potential variability of the seed ecological spectrum.
@@
 Our first global assessment of the seed germination spectrum in relation with seed morphology supports the existence of trade-offs between sets of regenerative traits [@RN4655], such as the relationship we found between seed/embryo size, germination speed and germination responses to temperature and light.
-Future studies will need to consider, as we did, evolutionary history and seed anatomy as intrinsic constraints of the seed ecological spectrum; while the seed germination spectrum we described here will need to be explored at the community level [@RN2352].
+Future studies will need to consider, as we did, evolutionary history and seed morphology as intrinsic constraints of the seed ecological spectrum; while the seed germination spectrum we described here will need to be explored at the community level [@RN2352].
 
-This study also highlights the general lack of germination studies on tropical alpine seeds.
-We were not able to analyse data from tropical alpine regions because they are underrepresented in the literature.
-Although there is an increasing interest in plant ecology of tropical alpine regions, we could not find seed germination data to be compared with other regions.
-We thus recommend to focus future research on seed ecology in high-mountain tropical areas, as it is possible that the cold-adapted germination syndrome is not prevalent there.
-Tropical alpine environments are relatively aseasonal and most plants have been reported as flowering throughout the year [@RN4957].
+This study also highlights the general lack of seed germination studies on tropical alpine species.
+Although there is an increasing interest in the ecology of plants from tropical alpine regions, we could not find seed germination data to include in our meta-analysis.
+Compared with temperate alpine ecosystems, tropical alpine environments are relatively aseasonal and most plants have been reported as flowering throughout the year [@RN4957].
 Rather than being exposed to snow in winter and high temperatures in summer, tropical alpine plants face a year-round stress from night-time frost [@RN2392].
-That means that our study is valid for the temperate alpine system, where alpine habitats mainly occur [@RN4750], but not necessarily for tropical alpine systems, which are subjected to different selective pressures.
+These different selective pressures suggest that the cold-adapted germination syndrome of temperate alpine seeds may not be prevalent in the tropical alpine flora.
+That means that our study is valid for temperate alpine regions, where alpine habitats mainly occur [@RN4750], but not necessarily for tropical alpine systems.
+We thus recommend to focus future research on the seed ecology of high-mountain tropical areas.
 
@@
 These cues, particularly cold stratification and warm temperatures, are generally associated with cold-adapted herbaceous species in a broad sense [@RN3397; @RN3214; @RN2365].
-Nonetheless, our results show that the response to stratification and temperature is more prominent in strict alpine species when comparing them with generalist plants that can grow opportunistically in the alpine belt, suggesting an evolutionary convergence of germination patterns in those species truly adapted to alpine ecosystems.
+Nonetheless, our results show that the response to stratification and temperature is more prominent in strict alpine species when comparing them with generalist plants that can grow opportunistically in the alpine belt.
+This suggests an evolutionary convergence of germination patterns in those species truly adapted to alpine ecosystems.
 We can conclude that, while alpine plants do not show a unique germination syndrome, they do show a more extreme or specialized version of the cold-adapted temperate germination syndrome.
@@
 ```{r fig1, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Seed dormancy classes, seed mass and embryo to seed ratio in the subsets of strict alpine and generalist species analysed in this study.
+Strict species are restricted to the alpine belt, generalist species can be found in the alpine belt but also below the treeline.
 Seed dormancy classes are abbreviated as follows: MD (morphological), MPD (morphophysiological), ND (non-dormant), PD (physiological), PY (physical).", fig.align = "center", fig.width=7, fig.height=3}
@@
 When lambda = 0, related taxa are no more similar than expected by chance; when lambda = 1, the trait is evolving following a constant variance random walk or Brownian motion model.
+FGP = final germination proportion.
 MGT = mean germination time.
@@
 UNC = germination uncertainty index.
-Seed dormancy classes are abbreviated as follows: MD (morphological), MPD (morphophysiological), ND (non-dormant), PD (physiological), PY (physical).", fig.align = "center", fig.width=7, fig.height=7}
+Seed dormancy classes are abbreviated as follows: MD (morphological), MPD (morphophysiological), ND (non-dormant), PD (physiological), PY (physical).
+Strict species are restricted to the alpine belt, generalist species can be found in the alpine belt but also below the treeline.", fig.align = "center", fig.width=7, fig.height=7}
 cowplot::ggdraw(fig5)

### 2020-11-07 src/cleaning.R

@@ -14,6 +14,7 @@ read_excel("../#data/lifeforms/data/alpineseeds/Traits_to_complete_PP_20200510.x
 
 ## Read and prepare seed mass
 
+read.csv("../#data/seedmass/results/Alpine_spp_genus_level_values.csv") -> genus.mass
 read.csv("../#data/seedmass/results/seedmass.csv") -> seed.mass
 
 ## Embryo trait (Filip)
@@ -32,7 +33,9 @@ read.csv("../#data/tpl/results/TPLNames.csv") %>%
   merge(seed.mass, all.x = TRUE) %>% # Merge seed mass
   merge(read.csv("../#data/baskin/results/dormancy.csv"), all.x = TRUE) %>%
   merge(filip, all.x = TRUE) %>%
-  select(TPLName, Family, Alpine, `Plant category`, `Lifespan`, Seed.mass, Embryo, Dormancy) -> traits
+  merge(read.csv("../#data/embryos/results/Alpine_spp_genus_level_values.csv")) %>% # Add embryo values rank
+  mutate(Seed.mass.rank = ifelse(TPLName %in% genus.mass$TPLName, "Genus", "Species")) %>% # Add seed mass values rank
+  select(TPLName, Family, Alpine, Seed.mass, Seed.mass.rank, Embryo, Embryo.rank, Dormancy) -> traits
 
 # SAVE CLEAN FILES
 

### 2020-11-07 src/manuscript-source.R

@@ -45,6 +45,12 @@ dataset %>%
 
 # Numbers for the main text
 
+read.csv("../data/traits.csv") %>% filter(Embryo.rank == "Species") %>% pull(TPLName) %>% unique %>% length -> MSembrankspp
+read.csv("../data/traits.csv") %>% filter(Embryo.rank == "Genus") %>% pull(TPLName) %>% unique %>% length -> MSembrankgen
+read.csv("../data/traits.csv") %>% filter(Embryo.rank == "Family") %>% pull(TPLName) %>% unique %>% length -> MSembrankfam
+
+read.csv("../data/traits.csv") %>% filter(Seed.mass.rank == "Species") %>% pull(TPLName) %>% unique %>% length -> MSmassrankspp
+read.csv("../data/traits.csv") %>% filter(Seed.mass.rank == "Genus") %>% pull(TPLName) %>% unique %>% length -> MSmassrankgen
 dataset %>% summary
 dataset %>% pull(Region) %>% unique %>% length -> MSregions
 dataset %>% tally() -> MSrecords
@@ -87,7 +93,8 @@ dataset %>%
   ggthemes::theme_tufte() +
   theme(legend.position = "right", axis.title.x = element_blank(),
         panel.background = element_rect(color = "grey96", fill = "grey96"),
-        axis.text.x = element_text(size = 10)) +
+        axis.title = element_text(size = 14),
+        axis.text = element_text(size = 11.5, color = "black")) +
   scale_fill_manual(values = c("olivedrab", "yellowgreen", "turquoise4", "gold", "purple")) -> fig1a
 
 dataset %>%
@@ -100,7 +107,8 @@ dataset %>%
   ggthemes::theme_tufte() +
   theme(legend.position = "none", axis.title.x = element_blank(),
         panel.background = element_rect(color = "grey96", fill = "grey96"),
-        axis.text.x = element_text(size = 10)) +
+        axis.title = element_text(size = 14),
+        axis.text = element_text(size = 11.5, color = "black")) +
   scale_fill_manual(values = c("yellowgreen", "gold")) -> fig1b
 
 dataset %>%
@@ -113,10 +121,11 @@ dataset %>%
   ggthemes::theme_tufte() +
   theme(legend.position = "none", axis.title.x = element_blank(),
         panel.background = element_rect(color = "grey96", fill = "grey96"),
-        axis.text.x = element_text(size = 10)) +
+        axis.title = element_text(size = 14),
[…]

### 2020-11-07 README.Rmd
tree link 📖

--- 2020-11-02 README.Rmd
@@
 
-* `data` Primary germination data and species traits.
+* `data` Primary data, including the [germination](https://github.com/efernandezpascual/alpineseeds/blob/master/data/germination.csv) dataset, the species [traits](https://github.com/efernandezpascual/alpineseeds/blob/master/data/traits.csv) and the phylogenetic [tree](https://github.com/efernandezpascual/alpineseeds/blob/master/data/ALPINE.tree.pdf).
 * `doc` Files to create the manuscript using `Rmarkdown`.
-* `results` Output of the `R` scripts, including MCMCglmm analysis outputs.
+* `results` Output of the `R` scripts, including [MCMCglmm analysis](https://github.com/efernandezpascual/alpineseeds/tree/master/results/MCMCglmm) outputs.
 * `src` Scripts in `R` language used to process data and perform the analyses of the manuscript.

### 2020-11-07 doc/cover-letter.Rmd
tree link 📖

@@
 We also added this information to the primary data available at the [GitHub repository](https://github.com/efernandezpascual/alpineseeds).
-We have indicated, in the relevant places of the text, which datasets ([germination](https://github.com/efernandezpascual/alpineseeds/blob/master/data/germination.csv), [species traits](https://github.com/efernandezpascual/alpineseeds/blob/master/data/traits.csv), and [phylogenetic tree](https://github.com/efernandezpascual/alpineseeds/blob/master/data/alpine.tree.pdf)) and results (detailed results of the [MCMCglmm analysis](https://github.com/efernandezpascual/alpineseeds/tree/master/results/MCMCglmm)) are available at GitHub.
+We have indicated, in the relevant places of the text, which datasets ([germination](https://github.com/efernandezpascual/alpineseeds/blob/master/data/germination.csv), [species traits](https://github.com/efernandezpascual/alpineseeds/blob/master/data/traits.csv), and [phylogenetic tree](https://github.com/efernandezpascual/alpineseeds/blob/master/data/ALPINE.tree.pdf)) and results (detailed results of the [MCMCglmm analysis](https://github.com/efernandezpascual/alpineseeds/tree/master/results/MCMCglmm)) are available at GitHub.
 
@@
 The phylogeny is available at the [GitHub repository](https://github.com/efernandezpascual/alpineseeds) as a [.tree file](https://github.com/efernandezpascual/alpineseeds/blob/master/data/ALPINE.tree) that can be imported into R, or opened with software like [TreeGraph](http://treegraph.bioinfweb.info/).
-We have now also included a [pdf version of the tree](https://github.com/efernandezpascual/alpineseeds/blob/master/data/alpine.tree.pdf), so it can be visualized directly.
+We have now also included a [pdf version of the tree](https://github.com/efernandezpascual/alpineseeds/blob/master/data/ALPINE.tree.pdf), so it can be visualized directly.
 We have indicated at the appropriate point of the text that the tree can be accessed at GitHub.

### 2020-11-12 README.Rmd
accepted version 🏔️

@@
 
-Please cite this repository as `The seed germination spectrum of alpine plants: a global meta-analysis. Manuscript in preparation.`
+Please cite this repository as:
+
+> Fernández-Pascual E, Carta A, Mondoni A, Cavieres L, Rosbakh S, Venn S, Satyanti A, Guja L, Briceño VF, Vandelook F, Mattana E, Saatkamp A, Bu H, Sommerville K, Poschlod P, Liu K, Nicotra A, Jiménez-Alfaro B (2020) The seed germination spectrum of alpine plants: a global meta-analysis.
+*New Phytologist*.

### 2020-11-12 doc/manuscript.Rmd
github ms 📰

@@
 title: "The seed germination spectrum of alpine plants: a global meta-analysis"
-output: 
-  bookdown::word_document2:
-    number_sections: false
-    reference_docx: template.docx
+output: github_document
 bibliography: refs.bib
@@
 ```{r message = FALSE, echo = FALSE, warning = FALSE}
-library(tidyverse); library(flextable)
+library(tidyverse)
 source("../src/manuscript-source.R", encoding = "UTF-8")
@@
 Telephone: +34985104787.
-
-**Total word count for the main body of the text** 6,837; **Introduction** 1,826; **Materials and Methods** 1,642; **Results** 1,155; **Discussion** 2,214; **Number of figures** 5 in colour; **Tables** 1; **Supporting information** 0
 
@@
 
+```{r table1, echo = FALSE}
+Table1 %>% kableExtra::kable(caption = "**Table 1** *Description of the datasets contributing primary data to this meta-analysis.
+The number of germination records, number of species and number of strict alpine species are given.*", format = "pandoc")
+```
+
 ---
@@
 
-Most species in the dataset were reported as having dormant seeds (Fig.
-\@ref(fig:fig1)a), but the frequency of the dormancy classes did not differ between generalist and strict alpine species.
+Most species in the dataset were reported as having dormant seeds (**Fig.
+1a**), but the frequency of the dormancy classes did not differ between generalist and strict alpine species.
 Among those species with dormant seeds, physiological dormancy was the most common dormancy class, followed by morphophysiological and then physical.
-Both strict alpine and generalist species had a similar range of values for seed mass (Fig.
-\@ref(fig:fig1)b) and embryo:seed ratio (Fig.
-\@ref(fig:fig1)c).
+Both strict alpine and generalist species had a similar range of values for seed mass (**Fig.
+1b**) and embryo:seed ratio (**Fig.
+1c**).
 The values of seed mass ranged from `r MSminMass` mg (*`r MSminMassSpp`*) to `r MSmaxMass` mg (*`r MSmaxMassSpp`*), with a median value of `r MSmedianMass` mg.
@@
 The median value was `r MSmedianEmbryo`.
+
+```{r fig1, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "**Figure 1** *Seed dormancy classes, seed mass and embryo to seed ratio in the subsets of strict alpine and generalist species analysed in this study.
+Strict species are restricted to the alpine belt, generalist species can be found in the alpine belt but also below the treeline.
+Seed dormancy classes are abbreviated as follows: MD (morphological), MPD (morphophysiological), ND (non-dormant), PD (physiological), PY (physical).*", fig.align = "center", fig.width=7, fig.height=3}
+cowplot::ggdraw(fig1)
+```
 
@@
 
-In general, stratification significantly increased the final germination proportion (Fig.
-\@ref(fig:fig2)a), as shown by a credible interval not overlapping zero.
+In general, stratification significantly increased the final germination proportion (**Fig.
+2a**), as shown by a credible interval not overlapping zero.
 This positive effect occurred in both generalist and strict alpine species, although the effect size was larger in the strict alpine group.
-Stratification slightly reduced the mean germination time (Fig.
-\@ref(fig:fig2)b), which was evident in the generalist species, but not in the strict alpine species.
-Further, stratification reduced the germination uncertainty index as a general response (Fig.
-\@ref(fig:fig2)c), and slightly more so in generalist species.
+Stratification slightly reduced the mean germination time (**Fig.
+2b**), which was evident in the generalist species, but not in the strict alpine species.
+Further, stratification reduced the germination uncertainty index as a general response (**Fig.
+2c**), and slightly more so in generalist species.
 Seed mass did not interact with stratification, but species with larger embryos had a smaller response to stratification in their final germination proportion and reduced less markedly their mean germination time and germination uncertainty index.
@@
 
-Scarification promoted the germination proportion (Fig.
-\@ref(fig:fig2)a) and reduced the mean germination time (Fig.
-\@ref(fig:fig2)b), but it did not affect the germination uncertainty index (Fig.
-\@ref(fig:fig2)c).
+Scarification promoted the germination proportion (**Fig.
+2a**) and reduced the mean germination time (**Fig.
+2b**), but it did not affect the germination uncertainty index (**Fig.
+2c**).
 Strict alpine species were slightly more responsive to scarification than generalist species.
@@
 
+```{r fig2, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "**Figure 2** *Effect of the germination environment on germination, according to the MCMC meta-analysis of primary data.
+The dots indicate the posterior mean of the effect size, and the brackets its 95 % credible interval.
+The line of zero effect is shown.
+When the credible intervals overlap with the zero-effect line, the effect is not significant.
+Some effects that overlapped with zero (i.e.
+had no effect) and whose credible intervals were excessively wide are not shown for clarity's sake.
+The figure shows first the main effect, then the effects for the strict alpine and generalist groups, and then the interaction with seed mass and embryo:seed.
+A negative interaction (e.g.
+seed mass and light, in the case of the germination proportion) indicates that, as the interacting factor increases (e.g.
+seed mass) the response to the germination cue (e.g.
+light) decreases, in comparison with the main effect.*", fig.align = "center", fig.width=7, fig.height=5.5}
+cowplot::ggdraw(fig2)
+```
+
 ## Germination temperature
 
-Temperature had a positive significant effect on the final germination proportion (Fig.
-\@ref(fig:fig2)a), indicating a preference for warmer temperatures.
-Warmer temperatures also significantly decreased mean germination time (Fig.
-\@ref(fig:fig2)b) and the germination uncertainty index (Fig.
-\@ref(fig:fig2)c).
+Temperature had a positive significant effect on the final germination proportion (**Fig.
+2a**), indicating a preference for warmer temperatures.
+Warmer temperatures also significantly decreased mean germination time (**Fig.
+2b**) and the germination uncertainty index (**Fig.
+2c**).
 The effect of warmer temperatures was slightly higher in strict alpine than in generalist species, although the credible intervals of the two groups of species do overlap.
@@
 
-Alternating temperatures had a positive effect on the final germination proportion (Fig.
-\@ref(fig:fig2)a), but not on the mean germination time (Fig.
-\@ref(fig:fig2)b) or the germination uncertainty index (Fig.
-\@ref(fig:fig2)c).
+Alternating temperatures had a positive effect on the final germination proportion (**Fig.
+2a**), but not on the mean germination time (**Fig.
+2b**) or the germination uncertainty index (**Fig.
+2c**).
 The positive effect on the germination proportion was slightly higher in the generalist species.
@@
 
-Light increased the final germination proportion (Fig.
-\@ref(fig:fig2)a), with no differences between strict alpine and generalist species.
-It did not affect the mean germination time (Fig.
-\@ref(fig:fig2)b), but it did result in more variable germination (increased the germination uncertainty index) of generalist species (Fig.
-\@ref(fig:fig2)c).
+Light increased the final germination proportion (**Fig.
+2a**), with no differences between strict alpine and generalist species.
+It did not affect the mean germination time (**Fig.
+2b**), but it did result in more variable germination (increased the germination uncertainty index) of generalist species (**Fig.
+2c**).
 Heavier and less endospermic seeds responded less to light.
@@
 
-Random factors had a significant effect in the MCMCglmm models (Fig.
-\@ref(fig:fig3)).
+Random factors had a significant effect in the MCMCglmm models (**Fig.
+3**).
 The final germination proportion and the mean germination time were affected more strongly, on average, by the seed lot and the species identity, suggesting that these traits are subject to intra- and interspecific variability, respectively.
@@
 The germination uncertainty index showed the opposite pattern, with a stronger effect of lab and alpine region.
-The phylogenetic signal in the response to all germination parameters (Fig.
-\@ref(fig:fig4)) was significantly positive for the final germination proportion and the mean germination time.
+The phylogenetic signal in the response to all germination parameters (**Fig.
+4**) was significantly positive for the final germination proportion and the mean germination time.
 For the germination uncertainty index, it overlapped with zero, indicating a weak effect of phylogeny.
@@
 
+```{r fig3, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "**Figure 3** *Effect of the random factors on germination, according to the MCMC meta-analysis of primary data.
+The dots indicate the posterior mean of the effect size, and the brackets its 95 % credible interval.
+The line of zero effect is shown.*", fig.align = "center", fig.width=7, fig.height=5}
+cowplot::ggdraw(fig3)
+```
+
+```{r fig4, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "**Figure 4** *Phylogenetic signal in the response to the germination environment.
+When lambda = 0, related taxa are no more similar than expected by chance; when lambda = 1, the trait is evolving following a constant variance random walk or Brownian motion model.
+FGP = final germination proportion.
+MGT = mean germination time.
+UNC = germination uncertainty index.*", fig.align = "center", fig.width=7, fig.height=1.5}
+cowplot::ggdraw(fig4)
+```
+
 ## FAMD ordination
 
-The first FAMD axis explained `r round(pcaAlpine$eig[1, 2], 0)`% of the variation (Fig.
-\@ref(fig:fig5)).
+The first FAMD axis explained `r round(pcaAlpine$eig[1, 2], 0)`% of the variation (**Fig.
+5**).
 The quantitative variables with the largest contribution to this first axis were germination temperature, germination uncertainty, GA~3~ and alternating temperatures.
@@
 This axis separated (iii, bottom) morphophysiologically dormant species with longer mean time to germination from (iv, top) physically dormant species that responded to scarification and had heavier, less endospermic seeds.
+
+```{r fig5, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "**Figure 5** *FAMD ordination of the alpine seed germination spectrum.
+Each square is a species.
+Labels indicate the contribution of the variables to the axes.
+Panels A and B correspond to quantitative traits; panel C and D to qualitative traits.
+MGT = mean germination time.
+UNC = germination uncertainty index.
+Seed dormancy classes are abbreviated as follows: MD (morphological), MPD (morphophysiological), ND (non-dormant), PD (physiological), PY (physical).
+Strict species are restricted to the alpine belt, generalist species can be found in the alpine belt but also below the treeline.*", fig.align = "center", fig.width=7, fig.height=7}
+cowplot::ggdraw(fig5)
+```
 
@@
 EM was supported by the Future Leaders Fellowship - Diversity and Livelihoods, of the Royal Botanic Gardens, Kew.
+Authors thank Amelia Martyn Yenson for help in experimental work.
 This research was supported by NGSeeds, the Next Generation Seed Ecology, Evolution and Data Science virtual working group.
@@
 BJA conceived the original idea.
-EFP, AC, AM, LC, SR, SV, AnS, LG, VB, FV, HB, KS, PP contributed data.
+EFP, AC, AM, LC, SR, SV, AnS, LG, VB, FV, HB, KS, PP, KL, AN contributed data.
 EFP assembled and managed the database.
@@
 The original data, R code for the analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/alpineseeds.
-Upon publication, a version of record of the repository will be deposited in Zenodo.
+A version of record of the repository can be found at https://doi.org/10.5281/zenodo.4270344.
 
@@
 :::
-
-# Figures
-
-```{r fig1, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Seed dormancy classes, seed mass and embryo to seed ratio in the subsets of strict alpine and generalist species analysed in this study.
-Strict species are restricted to the alpine belt, generalist species can be found in the alpine belt but also below the treeline.
-Seed dormancy classes are abbreviated as follows: MD (morphological), MPD (morphophysiological), ND (non-dormant), PD (physiological), PY (physical).", fig.align = "center", fig.width=7, fig.height=3}
-cowplot::ggdraw(fig1)
-```
-
-```{r fig2, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the germination environment on germination, according to the MCMC meta-analysis of primary data.
-The dots indicate the posterior mean of the effect size, and the brackets its 95 % credible interval.
-The line of zero effect is shown.
-When the credible intervals overlap with the zero-effect line, the effect is not significant.
-Some effects that overlapped with zero (i.e.
-had no effect) and whose credible intervals were excessively wide are not shown for clarity's sake.
-The figure shows first the main effect, then the effects for the strict alpine and generalist groups, and then the interaction with seed mass and embryo:seed.
-A negative interaction (e.g.
-seed mass and light, in the case of the germination proportion) indicates that, as the interacting factor increases (e.g.
-seed mass) the response to the germination cue (e.g.
-light) decreases, in comparison with the main effect.", fig.align = "center", fig.width=7, fig.height=5.5}
-cowplot::ggdraw(fig2)
-```
-
-```{r fig3, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the random factors on germination, according to the MCMC meta-analysis of primary data.
-The dots indicate the posterior mean of the effect size, and the brackets its 95 % credible interval.
-The line of zero effect is shown.", fig.align = "center", fig.width=7, fig.height=5}
-cowplot::ggdraw(fig3)
-```
-
-```{r fig4, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Phylogenetic signal in the response to the germination environment.
-When lambda = 0, related taxa are no more similar than expected by chance; when lambda = 1, the trait is evolving following a constant variance random walk or Brownian motion model.
-FGP = final germination proportion.
-MGT = mean germination time.
-UNC = germination uncertainty index.", fig.align = "center", fig.width=7, fig.height=1.5}
-cowplot::ggdraw(fig4)
-```
-
-```{r fig5, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "FAMD ordination of the alpine seed germination spectrum.
-Each square is a species.
-Labels indicate the contribution of the variables to the axes.
-Panels A and B correspond to quantitative traits; panel C and D to qualitative traits.
-MGT = mean germination time.
-UNC = germination uncertainty index.
-Seed dormancy classes are abbreviated as follows: MD (morphological), MPD (morphophysiological), ND (non-dormant), PD (physiological), PY (physical).
-Strict species are restricted to the alpine belt, generalist species can be found in the alpine belt but also below the treeline.", fig.align = "center", fig.width=7, fig.height=7}
-cowplot::ggdraw(fig5)
-```
-
-# Tables
-
-```{r table1, echo = FALSE}
-flextable(Table1) %>% width(width = c(3.3, 1, 0.8, 0.7, 0.6)) %>% fontsize(size = 8) %>% padding(padding = 0) %>% set_caption("Table 1 Description of the datasets contributing primary data to this meta-analysis.
-The number of germination records, number of species and number of strict alpine species are given.")
-```

### 2020-11-12 src/manuscript-source.R

@@ -552,19 +552,19 @@ gridExtra::arrangeGrob(fig5a, fig5b, fig5c, fig5d, ncol = 2,
                                                                                   fontfamily = "serif"))) -> fig5
 
 # Save figures
-
-ggsave(fig1, file = "../results/Figure_1.tiff", 
-       path = NULL, scale = 1, width = 180, height = 73, units = "mm", dpi = 600)
-
-ggsave(fig2, file = "../results/Figure_2.tiff", 
-       path = NULL, scale = 1, width = 180, height = 145, units = "mm", dpi = 600)
-
-ggsave(fig3, file = "../results/Figure_3.tiff", 
-       path = NULL, scale = 1, width = 180, height = 135, units = "mm", dpi = 600)
-
-ggsave(fig4, file = "../results/Figure_4.tiff", 
-       path = NULL, scale = 1, width = 180, height = 45, units = "mm", dpi = 600)
-
-ggsave(fig5, file = "../results/Figure_5.tiff", 
-       path = NULL, scale = 1, width = 180, height = 180, units = "mm", dpi = 600)
-
+# 
+# ggsave(fig1, file = "../results/Figure_1.tiff", 
+#        path = NULL, scale = 1, width = 180, height = 73, units = "mm", dpi = 600)
+# 
+# ggsave(fig2, file = "../results/Figure_2.tiff", 
+#        path = NULL, scale = 1, width = 180, height = 145, units = "mm", dpi = 600)
+# 
+# ggsave(fig3, file = "../results/Figure_3.tiff", 
+#        path = NULL, scale = 1, width = 180, height = 135, units = "mm", dpi = 600)
+# 
+# ggsave(fig4, file = "../results/Figure_4.tiff", 
+#        path = NULL, scale = 1, width = 180, height = 45, units = "mm", dpi = 600)
+# 
+# ggsave(fig5, file = "../results/Figure_5.tiff", 
+#        path = NULL, scale = 1, width = 180, height = 180, units = "mm", dpi = 600)
+# 

### 2021-05-16 README.Rmd
github ms 📰
journal issue 🌅

--- 2020-11-07 README.Rmd
@@
 
+[![DOI](https://zenodo.org/badge/254951190.svg)](https://zenodo.org/badge/latestdoi/254951190)
+
 ![Picos de Europa National Park, Cantabrian Mountains, Spain](Picos de Europa.jpg)
@@
 
-This repository stores all information related to the manuscript *The seed germination spectrum of alpine plants: a global meta-analysis*, including raw data, `R` scripts to process and analyse data, and `Rmarkdown` files to create the manuscript.
+This repository stores all information related to the publication *The seed germination spectrum of alpine plants: a global meta-analysis*, including raw data, `R` scripts to process and analyse data, and `Rmarkdown` files to create the manuscript.
 
@@
 * `data` Primary data, including the [germination](https://github.com/efernandezpascual/alpineseeds/blob/master/data/germination.csv) dataset, the species [traits](https://github.com/efernandezpascual/alpineseeds/blob/master/data/traits.csv) and the phylogenetic [tree](https://github.com/efernandezpascual/alpineseeds/blob/master/data/ALPINE.tree.pdf).
-* `doc` Files to create the manuscript using `Rmarkdown`.
+* `doc` Files to create the [manuscript](https://github.com/efernandezpascual/alpineseeds/blob/master/doc/manuscript.md) using `Rmarkdown`.
 * `results` Output of the `R` scripts, including [MCMCglmm analysis](https://github.com/efernandezpascual/alpineseeds/tree/master/results/MCMCglmm) outputs.
@@
 
-Please cite this repository as `The seed germination spectrum of alpine plants: a global meta-analysis. Manuscript in preparation.`
+Please cite this repository as:
+
+> Fernández-Pascual E, Carta A, Mondoni A, Cavieres L, Rosbakh S, Venn S, Satyanti A, Guja L, Briceño VF, Vandelook F, Mattana E, Saatkamp A, Bu H, Sommerville K, Poschlod P, Liu K, Nicotra A, Jiménez-Alfaro B (2020) [The seed germination spectrum of alpine plants: a global meta-analysis.](https://doi.org/10.1111/nph.17086) *New Phytologist* 229, 3573-3586.

### 2021-05-16 src/manuscript-source.R

@@ -552,19 +552,19 @@ gridExtra::arrangeGrob(fig5a, fig5b, fig5c, fig5d, ncol = 2,
                                                                                   fontfamily = "serif"))) -> fig5
 
 # Save figures
-# 
-# ggsave(fig1, file = "../results/Figure_1.tiff", 
-#        path = NULL, scale = 1, width = 180, height = 73, units = "mm", dpi = 600)
-# 
-# ggsave(fig2, file = "../results/Figure_2.tiff", 
-#        path = NULL, scale = 1, width = 180, height = 145, units = "mm", dpi = 600)
-# 
-# ggsave(fig3, file = "../results/Figure_3.tiff", 
-#        path = NULL, scale = 1, width = 180, height = 135, units = "mm", dpi = 600)
-# 
-# ggsave(fig4, file = "../results/Figure_4.tiff", 
-#        path = NULL, scale = 1, width = 180, height = 45, units = "mm", dpi = 600)
-# 
-# ggsave(fig5, file = "../results/Figure_5.tiff", 
-#        path = NULL, scale = 1, width = 180, height = 180, units = "mm", dpi = 600)
-# 
+
+ggsave(fig1, file = "../results/Figure_1.png",
+       path = NULL, scale = 1, width = 180, height = 73, units = "mm", dpi = 600)
+
+ggsave(fig2, file = "../results/Figure_2.png",
+       path = NULL, scale = 1, width = 180, height = 145, units = "mm", dpi = 600)
+
+ggsave(fig3, file = "../results/Figure_3.png",
+       path = NULL, scale = 1, width = 180, height = 135, units = "mm", dpi = 600)
+
+ggsave(fig4, file = "../results/Figure_4.png",
+       path = NULL, scale = 1, width = 180, height = 45, units = "mm", dpi = 600)
+
+ggsave(fig5, file = "../results/Figure_5.png",
+       path = NULL, scale = 1, width = 180, height = 180, units = "mm", dpi = 600)
+
